### 2021-07-27 paper/main.tex
WIP paper

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{apacite}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{pxfonts}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage[left]{lineno}
\usepackage{soul}
\linenumbers

\title{High-order dynamic neural correlations reflect naturalistic processing
in humans} \author{Lucy
  L. W. Owen$^1$, and\
  Jeremy R. Manning\textsuperscript{$1, \dagger$}\\
  [0.1in]$^1$Department of Psychological and Brain
  Sciences,\\Dartmouth
  College, Hanover, NH\\
  \textsuperscript{$\dagger$}Address correspondence to
  jeremy.r.manning@dartmouth.edu}

\begin{document}
\maketitle

\begin{abstract}
We applied dimensionality reduction algorithms to the activity patterns in each experimental condition.  Specifically, we sought to understand the “dimensionality” of the neural patterns that were sufficient to decode participants’ listening times (or approach was similar to that of Mack et al. 2017).  We found that even low-dimensional embeddings of the data were sufficient to accurately decode listening times from the intact story recording, whereas finer temporal scramblings of the story required higher-dimensional embeddings of the data to reach peak decoding accuracy.
\end{abstract}

\doublespacing

\section*{Introduction}

We're interested in the complexity of brain patterns that underly
different types of thoughts. To explore this question space, we take
brain patterns recorded under different experimental conditions and
project them into lower dimensional spaces using principle components
analysis. We can then ask how well those low-dimensional embeddings of
the data retain cognitively relevant information like when in a story
someone is listening to. 
Then we can explore decoding accuracy as a
function of the number of components, or dimensions, in the
low-dimensional space under different cognitive conditions.

You can imagine two reasonable predictions of
how cognition is reflected in brain patterns.  The first is as our
thoughts become more complex, they are supported by more complex brain
patterns, and require more components to decode.  The second is that
when thoughts are deeper and more complicated, the units of neural
activity would carry more information, and would require therefore
fewer components to decode.

This idea can be explored in this visual analogy for neural
compression.  Here there are two images of pies, the top pie is more
complex than the bottom.  On the left we're illustrating that it takes
fewer componenets to reach the same 95 percent variance explained in
the less complex pie, which corresponds to higher compression.
However, on the right with very few components similar
variance is explaining both pies.

We investigated the dimensionality of neural patterns by training
classifiers using more and more principle components. Or, in other
words, we used less and less compression to decode.
We applied the approach to
a neuroimaging dataset comprising data collected as participants
listened to a story varying in cognitive richness~\citep{SimoEtal16}.

\section*{Methods}
A major challenge to studying such patterns is that typically neither
the correlations nor the hierarchical organizations of those
correlations may be directly observed.  Rather, these fundamental
properties must be inferred indirectly by examining the observable
parts of the system-- e.g., the behaviors of the individual units of
that system.  Here we propose a series of mathematical operations that may be used to approximate dynamic correlations at a range of scales (i.e., orders of interaction).

There are two basic steps to our approach (Fig.~\ref{fig:methods_fig}).  In the first step, we take
a number-of-timepoints ($T$) by number-of-features ($F$)
\textit{matrix of observations} ($\mathbf{X}$) and we return a $T$ by
$\frac{F^2 - F}{2}$ \textit{matrix of dynamic correlations}
($\mathbf{Y}$).  Here $\mathbf{Y_0}$ describes, at each moment, how
all of the features (columns of $\mathbf{X}$) are inferred to be
interacting.  (Since the interactions are assumed to be non-recurrent
and symmetric, only the upper triangle of the full correlation matrix
is computed.)  In the second step, we project $\mathbf{Y_0}$ onto an
$F$-dimensional space, resulting in a new $T$ by $F$ matrix
$\mathbf{Y_1}$.  Note that $\mathbf{Y_1}$ contains information about
the correlation dynamics present in $\mathbf{X}$, but represented in a
compressed number of dimensions.  By repeatedly applying these two
steps in sequence, we can examine and explore higher order dynamic
correlations in $\mathbf{X}$.

\subsection*{Dynamic correlations}
Given a matrix of observations, we can compute the (static)
correlations between any pair of observations, $\mathbf{X}_i$ and
$\mathbf{X}_j$ using:
\begin{align}
  \mathrm{corr}(\mathbf{X}_i, \mathbf{X}_j) &= \frac{\sum_{t=1}^T \left(\mathbf{X}_i(t)
                                              -
                                              \bar{\mathbf{X}_i}\right) \left(\mathbf{X}_j(t)
                                              -
                                              \bar{\mathbf{X}_j}\right)}{\sqrt{\sum_{t=1}^T
                                              \sigma^2_{\mathbf{X}_i} 
                                              \sigma^2_{\mathbf{X}_j}}},~\mathrm{where}\\\label{eqn:corr}
  \bar{\mathbf{X}_k} &= \sum_{t=1}^T
                       \mathbf{X}_k(t),~\mathrm{and}\\
  \sigma^2_{\mathbf{X}_k} &= \sum_{t=1}^T \left( \mathbf{X}_k -
                            \bar{\mathbf{X}_k} \right)^2 
\end{align}

We can generalize this formula to compute time-varying correlations by
incorporating a \textit{weight function} that takes a time $t$ as input,
and returns how much the observed data every timepoint (including $t$)
contribute to the correlations at time $t$ (Fig.~\ref{fig:weights}).
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/kernels}
  \caption{\textbf{Examples of time-varying weights.  Each panel
      displays per-timepoint weights at $t = 50$, evaluated for 100
      timepoints ($1, ..., 100$).  a. Uniform weights.} The weights
    are timepoint-invariant; observations at all timepoints are
    weighted equally, and do not change as a function of $t$.  This is
    a special case of weight function that reduces dynamic
    correlations to static correlations.  \textbf{b. Dirac delta
      function.} Only the observation at timepoint $t$ is given weight
    (of 1), and weights for observations at all other timepoints are
    set to 0.  \textbf{c. Gaussian weights.} Each observation's
    weights fall off in time according to a Gaussian probability
    density function centered on $\mu = t$.  Weights derived using
    several different example variance parameters ($\sigma^2$ are
    displayed.  \textbf{d. Laplace weights.}  Each observation's
    weights fall off in time according to a Laplace probability
    density function centered on $\mu = t$.  Weights derived using
    several different example scale parameters ($b$) are displayed.
    \textbf{e. Mexican hat (Ricker wavelet) weights.}  Each
    observation's weights fall off in time according to a Ricker
    wavelet centered on $t$.  This function highlights the
    \textit{contrasts} between local versus surrounding activity
    patterns in estimating dynamic correlations. Weights derived using
    several different example width parameters ($\sigma$) are
    displayed.  }
  \label{fig:weights}
\end{figure}

Given a weight function $w(t)$ for timepoint $t$, evaluated at
timepoints in the interval $\left[ 1, ..., T \right]$, we can extend the static correlation formula
in Equation~\ref{eqn:corr} to reflect an \textit{instantaneous
  correlation} at timepoint $t$:
\begin{align}
  \mathrm{timecorr}(\mathbf{X}_i, \mathbf{X}_j, t) &= \frac{\sum_{t=1}^T \left(\mathbf{X}_i(t)
                                              -
                                              \widetilde{\mathbf{X}}_i(t)\right) \left(\mathbf{X}_j(t)
                                              -
                                              \widetilde{\mathbf{X}}_j(t)\right)}{\sqrt{\sum_{t=1}^T
                                              \widetilde{\sigma}^2_{\mathbf{X}_i}(t) 
                                              \widetilde{\sigma}^2_{\mathbf{X}_j}}(t)},~\mathrm{where}\\\label{eqn:timecorr}
  \widetilde{\mathbf{X}}_k(t) &= \sum_{i=1}^T
                       w(t, i)\mathbf{X}_k(i),\\
  \widetilde{\sigma}^2_{\mathbf{X}_k}(t) &= \sum_{i=1}^T \left( \mathbf{X}_k(i) -
                            \widetilde{\mathbf{X}}_k(t) \right)^2,
\end{align}
and $w(t, i)$ is shorthand for $w(t)$ evaluated at timepoint $i$.
Equation~\ref{eqn:timecorr} may be used to estimate the instantaneous
correlations between every pair of observations, at each timepoint
(i.e., $\mathbf{Y}$).

\subsubsection*{Inter-subject dynamic correlations}
Equation~\ref{eqn:timecorr} provides a means of taking a single
observation matrix, $\mathbf{X}$ and estimating the dynamic
correlations from moment to moment, $\mathbf{Y}$.  Suppose that one
has access to a set of multiple observation matrices that reflect the
same phenomenon.  For example, one might collect neuroimaging data
from several experimental participants, as each participant performs
the same task (or sequence of tasks).  Let $\mathbf{X}_1$,
$\mathbf{X}_2$, ..., $\mathbf{X}_P$ reflect the $T$ by $F$ observation
matrices for each of $P$ participants in an experiment.  We can use
\textit{inter-subject functional connectivity}~\citep[ISFC;
][]{SimoEtal16} to compute the degree of stimulus-driven correlations
reflected in the multi-participant dataset at a given timepoint $t$
using:
\begin{align}
\bar{\mathbf{C}}(t) = M\left(R\left(\frac{1}{2P} \sum_{i=1}^P
  Z\left(Y_i(t)\right)^\mathrm{T} + Z\left(Y_i,(t)\right)\right)\right),
\end{align}
where $M$ extracts and vectorizes the diagonal and upper triangle of a symmetric
matrix, $Z$ is the Fisher $z$-transformation~\citep{Zar10}:
\begin{align}
Z(r) = \frac{\log(1+r) - \log(1-r)}{2}
\end{align}
$R$ is the inverse of $Z$:
\begin{align}
R(z) = \frac{\exp(2z - 1)}{\exp(2z + 1)},
\end{align}
and $\mathbf{Y}_i(t)$ denotes the correlation matrix
(Eqn.~\ref{eqn:corr}) between each column of $\mathbf{X}_i$ and each
column of the average observations from all \textit{other}
participants, $\bar{\mathbf{X}}_{ \backslash i}$:
\begin{align}
  \bar{\mathbf{X}}_{ \backslash i} = R\left(\frac{1}{P-1}\sum_{i \in
  \backslash i} Z\left( \mathbf{X}_i \right) \right),
\end{align}
where $ \backslash i$ denotes the set of all participants other than
participant $i$. In this way, the $T$ by $\left( \frac{F^2 - F}{2}
  + F \right)$
matrix $\bar{\mathbf{C}}$ is the time-varying extension of the ISFC
approach developed by \cite{SimoEtal16}.

\subsection*{Higher-order correlations}
Given a timeseries of dynamic correlations (e.g., obtained using
Eqn.~\ref{eqn:timecorr}), higher-order correlations reflect the
dynamic correlations between columns of $\mathbf{Y}$.  Given unlimited
computing resources, one could use repeated applications of
Equation~\ref{eqn:timecorr} to estimate these higher-order
correlations (i.e., substituting in the previous output, $\mathbf{Y}$,
for the input, $\mathbf{X}$ in the equation).  However, because each
output $\mathbf{Y}$ has $\mathcal{O}(F^2)$ columns relative to $F$ columns in
the input $\mathbf{X}$, the output of Equation~\ref{eqn:timecorr}
grows with the square of the number of repeated applications (total
cost of computing $n$\textsuperscript{th} order correlations is
$\mathcal{O}(F^{2n})$ for $n \in \mathcal{J}, n > 0$).  When $F$ or $n$ is large,
this approach quickly becomes intractable.

To make progress in computing $\mathbf{Y_{n+1}}$, we can approximate
$\mathbf{Y}_n$ by computing an $\mathcal{O}(F)$-dimensional embedding of
$\mathbf{Y}_n$, termed $\hat{\mathbf{Y}}_n$, and then we can apply
Equation~\ref{eqn:timecorr} to $\hat{\mathbf{Y}}_n$ rather than
directly to $\mathbf{Y}_n$.  This enables us to maintain $\mathcal{O}(n)$
scaling with respect to $n$, rather than exponential scaling via the
direct approach.

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{figs/methods_fig}
  \caption{\textbf{Computing higher order correlations. } Dynamic
    correlations
  are computed then approximated to the same size as original data
  using dimensionality reduction.
  This process is repeated, and can be used to compute up to any
  arbitrary order with compuations scaling linearly (as opposed to exponentially) with order. }
  \label{fig:methods_fig}
\end{figure}

There are many possible methods for computing $\hat{\mathbf{Y}}_n$
from $\mathbf{Y}_n$, including traditional dimensionality reduction
approaches and graph theory based approaches as described next.  In the
\textit{Discussion} section we elaborate on other potential approaches.

\subsubsection*{Dimensionality reduction-based approaches to computing
  $\hat{\mathbf{Y}}_n$}

Commonly used dimensionality reduction algorithms include Principal
Components Analysis~\citep[PCA; ][]{Pear01}, Probabilistic
PCA~\citep[PPCA; ][]{TippBish99}, Exploratory Factor
Analysis~\citep[EFA; ][]{Spea04}, Independent Components
Analysis~\citep[ICA; ][]{JuttHera91, ComoEtal91}, $t$-Stochastic
Neighbor Embedding~\citep[$t$-SNE; ][]{MaatHint08}, Uniform Manifold
Approximation and Projection~\citep[UMAP; ][]{McInHeal18},
non-negative matrix factorization~\citep[NMF; ][]{LeeSeun99},
Topographic Factor Analysis (TFA)~\cite{MannEtal14b}, Hierarchical
Topographic Factor analysis (HTFA)~\cite{MannEtal18}, Topographic
Latent Source Analysis (TLSA)~\cite{GersEtal11}, Dictionary
learning~\citep{MairEtal09a, MairEtal09b}, deep
autoencoders~\citep{HintSala06}, among others.  While complete
characterizations of each of these algorithms is beyond the scope of
the present manuscript, the general intuition driving these
approaches is to compute the $\hat{\mathbf{Y}}$ with $i$ columns that
is closest to the original $\mathbf{Y}$ with $j$ columns, and where
(typically) $i \ll j$.  The different approaches place different
constraints on what properties $\hat{\mathbf{Y}}$ must satisfy and
which aspects of the data are compared (and how) to characterize the
match between $\hat{\mathbf{Y}}$ and  $\mathbf{Y}$.

Applying dimensionality reduction algorithms to $\mathbf{Y}$ yields a
$\hat{\mathbf{Y}}$ whose columns reflect weighted combinations (or
nonlinear transformations) of the original columns of $\mathbf{Y}$.
This has two main consequences.  First, with each repeated
dimensionality reduction, the resulting $\hat{\mathbf{Y}}_n$ has lower
and lower fidelity (with respect to what the ``true'' $\mathbf{Y}_n$
might have looked like without using dimensionality reduction to
maintain scalability).  In other words, computing $\hat{\mathbf{Y}}_n$
is a lossy operation.  Second, whereas the columns of $\mathbf{Y}_n$
may be mapped directly onto pairs of columns of $\mathbf{Y}_{n-1}$,
that mapping either becomes less cleanly defined in
$\hat{\mathbf{Y}}_n$ due to the reweightings and/or nonlinear
transformations.

\subsubsection*{Graph theory-based approaches to computing
  $\hat{\mathbf{Y}}_n$}

Graph theoretic measures take as input a matrix of interactions (e.g.,
using the above notation, an $F \times F$ correlation matrix or
binarized correlation matrix reconstituded from a single timepoint's
row of $\mathbf{Y}$) and return as output a set of $F$ measures
describing how each node (feature) sits within that interactions
matrix with respect to the rest of the population.  Common measures
include betweeness centrality~\citep[the proportion of shortest paths
between each pair of nodes in the population that involves the given
node in question; e.g., ][]{Newm05, OpsaEtal10, Bart04, GeisEtal08,
  Free77}; diversity and dissimilarity~\citep[characterizations of how
differently connected a given node is from others in the population;
e.g., ][]{Rao82, Lin09, RicoSzei06}; Eigenvector centrality and
pagerank centrality~\citep[measures of how influential a given node is
within the broader network; e.g., ][]{Newm08, Bona07, LohmEtal10,
  HaluEtal13}; transfer entropy and flow coefficients~\citep[a measure
of how much information is flowing from a given node to other nodes in
the network; e.g., ][]{HoneEtal07, Schr00}; $k$-coreness
centrality~\citep[a measure of the connectivity of a node within its
local sub-graph; e.g., ][]{AlvaEtal05, ChriFowl10}; within-module
degree~\citep[a measure of how many connections a node has to its
close neighbors in the network; e.g., ][]{RubiSpor10}; participation
coefficient~\citep[a measure of the diversity of a node's connections
to different sub-graphs in the network; e.g., ][]{RubiSpor10}; and
sub-graph centrality~\citep[a measure of a node's participation in all
of the network's sub-graphs; e.g., ][]{EstrRodr05}.

As an alternative to the above dimensionality reduction approach to
embedding $\mathbf{Y}_n$ in a lower-dimensional space, but still
allowing for scalable explorations of higher-order structure in the
data, we also explore using the above graph theoretic measures as a
means of obtaining $\hat{\mathbf{Y}}_n$.  In particular: for a given
graph theoretic measure, $\eta: \mathcal{R}^{F \times F} \rightarrow
\mathcal{R}^F$, we can use $\eta$ to tranform each row of
$\mathbf{Y}_n$ in a way that characterizes the corresponding
graph-theoretic properties of each column.  Whereas the dimensionality
reduction approach to computing $\hat{\mathbf{Y}}_n$ is lossy,
the graph-theory approach is lossless.  However, whereas the
dimensionality reduction approach maintains ties (direct or indirect)
to the original activity patterns reflected in $\mathbf{Y}_{n-1}$, the
graph-theory approach does not.  Instead, the graph-theory
characterizes the nature and timecourse of each feature's
\textit{participation} in the network.

\subsection*{Evaluation metrics}
We evaluate our approach to extracting dynamic correlations and
higher-order correlations using several metrics detailed next.  First,
we generated synthetic data using known time-varying correlations, and
then we evaluated the fidelity with which Equation~\ref{eqn:timecorr}
could recover those correlations (for synthetic datasets with
different properties, and using different kernels to define the
weights; Fig.~\ref{fig:weights}).  We then turned to a series of
analyses on a (real) neuroimaging dataset where the ground truth
correlations were \textit{not} known.  We evaluated whether the
recovered correlations could be used to accurately label held-out
neuroimaging data with the time at which it was collected.  We used this latter evaluations (using timepoint
decoding) as a proxy for
gauging how much explanatory power the recovered correlations held
with respect to the observed data.

\subsubsection*{Generating synthetic data}
To explore recovery of a constant covariance (Fig.~\ref{fig:synthetic_data},  a.), we generated syntheic
data sampled from a constant covariance matrix. To do this, we created one random
covariance matrix, $K$, with 50 features, and for each of the 300 timepoints  we
sampled from a Gaussian distribution centered on $K$.  Similarly, we generated synthetic data
sampled from a random covariance matrix (Fig.~\ref{fig:synthetic_data}, b.) by creating a new random
covariance matrix $K(t)$, for each of the 300 timepoints and sampled from a
Gaussian distribution centered on $K(t)$.

To generate synthetic data from a dynamically changing covariance
matrix (Fig.~\ref{fig:synthetic_data},  c.),
we generated two random covariance
matrices, $K_{1}$ and $K_{2}$.  We then computed a weighted average covariance matrix
for each of the 300 timepoint, $K(t)$, 
by taking the linearly spaced weights ($w$) of the two random matrices, 
\begin{align}
K(t) = w(t) * K_{1} + (1 - w(t)) *  K_{2}, \\\label{eqn:ramping}
\end{align}
and for each of the 300 timepoints sampled from a
Gaussian distribution centered on $K(t)$.

Lastly, for the synthetic data containing block structure (Fig.~\ref{fig:synthetic_data},  d.), we followed the same
process of creating synthetic data sampled from a constant covariance
matrix (see above) but sampled from a new random covariance matrix
after 60 consecutive timepoints.  We then pieced the blocks together
to create a sythetic dataset with 300 total timepoints but drawn from
5 separate covariance matrices.

\subsubsection*{Recovery of ground truth parameters from synthetic
  data}

We applied timecorr, using delta and gaussian (width = 10) kernels
Fig.~\ref{fig:weights}) to each of these 
synthetic datasets, then correlated each recovered
correlation matrix with the ground truth.  We repeated this process 10
times and explored how recovery varies
with the kernel and the specific structure of the data. For the
ramping synthetic dataset (Fig.~\ref{fig:synthetic_data},  c.)  and for the
block synthetic dataset (Fig.~\ref{fig:synthetic_data},  d.)  we made further
comparisons of the timecorr recovered correlation matrices. We
compared the ramping recovered correlation matrices to only the first random covariance matrix $K_{1}$
(First, Fig.~\ref{fig:synthetic_data},  c.) and to only the last
random covariance matrix $K_{2}$ (Last, Fig.~\ref{fig:synthetic_data},
c.) from Equation~\ref{eqn:ramping}. We also compared the block recovered correlation matrices in to
the block specific covariance matrix (Block 1-5,
Fig.~\ref{fig:synthetic_data},  d.).

\subsubsection*{Timepoint decoding}

To explore how higher-order structure varies with stimulus structure
and complexity, we used a previous neuroimaging dataset
~\cite{SimoEtal16}  in which participants listened to an audio recording of a story; 36 particpants listen to an intact version of the story, 17 participants listen to time-scrambled recordings of the same story where paragraphs were scrambled, 36 particpants listen to word-scrambled version and 36 participants lay in rest condition.

Prior work has shown participants share similar neural responses to
richly structured stimuli when compared to stimuli with less
structure.  To assess whether the moment-by-moment higher order
correlations were reliably preserved across participants, we used
inter-subject functional connectivity (ISFC) to isolate the
time-varying correlational structure (functional connectivity patterns
that were specifically driven by the story participants listened to.
Following the analyses conducted by (HTFA)~\cite{MannEtal18}, we first
applied \textit{hierarchical topographic factor analysis} (HTFA) to
the fMRI datasets to obtain a time series of 700 node activities for
every participant.  We then computed the dynamic weighted ISFC using a
range of kernels and widths.  Specifically, we used Gaussian, Laplace,
and mexican hat kernels, as well as widths of 5, 10, 20, and 50.  We
then approximated these dynamic correlation using two reduction measures, PCA and eigenvector centrality, and computed the dynamic weighted ISFC on the approximations.  We repeated this process up to 10th order approximated correlations.

To assess decoding accuracy, we randomly divided participants for each
stimulus into training and testing groups. For the zeroth order, we
computed the mean factor activity for each group.  For all subsequent
orders up to the tenth order, we computed the mean approximated
dynamic ISFC of factor activity for each group. To assess how
additional higher-order correlations contribute to decoding accuracy,
for each order we included a weighted-mixutre (described below) of the activity patterns of all
previous orders.  For each group of participants in turn, we compared these activity patterns (using Pearson correlations) to estimate the story times each pattern corresponded to. Specifically, we asked, for each timepoint: what are the correlations
between the first group's and second group's activity patterns at each
order. We note that the decoding test we used is a conservative in which we count a timepoint label as incorrect if it is not an exact match.

For each order we obtained the weighted-mixture of the correlation
matrices for the current order and all previous orders using mixing parameter, $\phi$, where $0 <\phi< 1$ reflects a
weighted mixture of order based decoding Fig.~\ref{fig:decoding_level}
Panel c. ). We calculated  $\phi$, by
subdividing the training group and using the quasi-Newton method of
Broyden, Fletcher, Goldfarb, and Shanno (BFGS ~\citep{NoceWrig06}) for optimization. We
repeated this cross-validation process 10 times for each parameter set.

\section*{Results}
\subsubsection*{Synthetic data}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/synthetic_data.pdf}
  \caption{\textbf{Dynamic correlation recovery with synthetic data. } Using synthetic data containing different
    underlying correlational structure,
   we test how well we can recover dynamic correlation matrices using two kernels (delta and gaussian,
    width = 10) when compared to groud truth.  We plot recovery using of datasets containing the
    following underlying structure:
    \textbf{a. Constant. b. Random. c. Ramping.  d. Block.}}
  \label{fig:synthetic_data}
\end{figure}

To assess the performance of dynamic correlation recovery using
timecorr, we varied width the kernel and the specific structure of the
data. We applied timecorr, using delta and gaussian kernels
Fig.~\ref{fig:weights}) to each of the following  
synthetic datasets: constant, random, ramping, and block.  We then correlated each recovered
correlation matrix with the ground truth.

For the constant synthetic dataset, a gaussian kernel (width=10)
outperformed the delta kernel (Fig.~\ref{fig:synthetic_data},  a.).  This is in contrast with the random
synthetic dataset, for which the delta kernel best captures the
rapidly changing structure (Fig.~\ref{fig:synthetic_data},  b.). For
the ramping synthetic dataset, the slow changing strucutre within the
data is best
captured by the gaussian kernel and the best recovery occurs in the
middle (Ramping, Fig.~\ref{fig:synthetic_data},
c.). In addition to comparing the timecorr recovered correlation
matrices to the ground truth, we
further compared the ramping recovered correlation matrices to only the first random covariance matrix $K_{1}$
(First, Fig.~\ref{fig:synthetic_data},  c.) and to only the last
random covariance matrix $K_{2}$ (Last, Fig.~\ref{fig:synthetic_data},
c.), both of which perform best at the beginning and end respectively.

Similary for the block sythetic dataset, we compared the timecorr
recovered correlation matrices to the ground truth as well as to each
block-specific covariance matrix (Block 1-5,
Fig.~\ref{fig:synthetic_data},  d.).  Although the structure is
changing by block, the gaussian kernel once again outperforms the
delta kernel.  Performance does however drop near even boundaries for
when using the gaussian kernel.

\subsubsection*{Neuroimaging dataset~\citep{SimoEtal16}}
For our decoding analysis, we used HTFA-derived node activities
~\cite{MannEtal18} from fMRI data collected as participants listened
to an audio recording of a story (intact condition; 36 participants),
listened to time scrambled recordings of the same story (17
participants in the paragraph-scrambled condition listened to the
paragraphs in a randomized order and 36 in the word-scrambled
condition listened to the words in a randomized order), or lay resting
with their eyes open in the scanner (rest condition; 36
participants). We sought to demonstrate how higher-order correlations
may be used to examine dynamic interactions of brain patterns in
(real) multi-subject fMRI datasets. This story listening dataset was
collected as part of a separate study, where the full imaging
parameters, image preprocessing methods, and experimental details may
be found~\citep{SimoEtal16}. The dataset is available at
\href{url}{http://arks. princeton.edu/ark:/88435/dsp015d86p269k}.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level.pdf}
  \caption{\textbf{Decoding by order.
      a.\&d. Relative decoding accuracy by order.} Ribbons of each
    color display cross-validated decoding performance averaged over
    all parameters for each condition (intact, paragraph, word,
    and rest) using PCA (a.) or eigenvector centrality (d.) to
    approximate correlations. Decoders were trained using increasingly more
    higher-order information and this ribbons are displayed relative
    to chance at 0. The red dots indicates
    maximum decoding accuracy for each condition. \textbf{b.\&e. Z-transformed decoding accuracy
      by order.}  We Z-transformed the decoding accuracy by order to better
    visualize the order with the maximum decoding accuracy for each
    condition using PCA (b.) or eigenvector centrality (e.) to
    approximate correlations. \textbf{c.\&f. Optimized weights.} Bar heights indicate the
    optimized mixing paramete $\phi$ of each contributing order up to
    and including the order with the maximum
    decoding accuracy for each contributing order using PCA (c.) or eigenvector centrality (f.) to
    approximate correlations.
  For the order with maximum decoding accuracy by condition, we show
  barplots of the optimized weights $\phi$  for each contributing
  order.}
  \label{fig:decoding_level}
\end{figure}

We next evaluated if our model of high-order correlations in brain activity can capture cognitively relevant brain patterns. We performed a decoding analysis, using cross validation to estimate (using other participants’ data) which parts of the story each weighted-mixture of higher-order brain activity pattern corresponded to (see \textit{Materials and methods}). We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.

Separately for each experimental condition, we divided participants
into two groups. For the zeroth order, we computed the mean factor
activity for each group.  For all subsequent orders up to the tenth
order, we computed the mean approximated dynamic ISFC of factor
activity for each group (see \textit{Materials and methods}), and
combined in a weighted mixutre with all previous orders
(i.e. cross-validation for the second
order contained a weighted-mixture of zeroth, first, and second order
(Fig.~\ref{fig:decoding_level}, c.\&f.).  For
each order, we correlated the group 1 activity patterns with group 2
activity patterns.  We then subdivided the group 1 to obtain an
optimal weighting parameter for each order’s correlation matrix using
the same cross validation method. We used the optimal weighting
parameters to obtain a weighted-mixture (see \textit{Materials and methods}) of each order’s correlation
matrix. Using these correlations, we labeled the group 1 timepoints
using the group 2 timepoints with which they were most highly
correlated; we then computed the proportion of correctly labeled group
1 timepoints. (We also performed the symmetric analysis whereby we
labeled the group 2 timepoints using the group 1 timepoints as a
template.) We repeated this procedure 100 times (randomly re-assigning
participants to the two groups each time) to obtain a distribution of
decoding accuracies for each experimental condition. There were 272
timepoints for paragraph condition, 300 timepoints for intact and word
conditions, and 400 timepoints for rest condition,  so chance
performance on this decoding test is was $\frac{1}{272}$,
$\frac{1}{300}$, and $\frac{1}{400}$ respectively.

We repeated this process for each set of parameters, varying kernel
type and width, and averaged over the reduction technique used to
approximate the higher-order correlations (PCA Fig.~\ref{fig:decoding_level},  a.-c. and eigenvector
centrality Fig.~\ref{fig:decoding_level},  d.-f.).  Since there is no ground truth in these analyses, and we did not know
which parameters best capture the data, we instead report a robustness
search by averaging over the parameters and reporting which results
consistently showed up across all parameters.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/brain_plots.pdf}
  \caption{\textbf{Average correlations by order for the intact listening condition.} Using eigenvector
    centrality to approximate higher-order correlations for the intact
    listening condition, we plot the
    strongest 25\% absolute value
    mean correlation for order: \textbf{a. 1. b. 2. c. 3.,
      and d. 4}, representing the degree of agreement by
    location pair over time.}
  \label{fig:brain_plots}
\end{figure}

The two methods used to approximate the higher-order correlations (PCA Fig.~\ref{fig:decoding_level},  a.-c. and eigenvector
centrality Fig.~\ref{fig:decoding_level},  d.-f.) capture different
facets of the activity patterns.  Using PCA, the higher-order
correlations are all linked to the original activity patterns, whereas
eigenvectory centrality breaks the immediate link with specific brain
areas and instead characterizes the position of the nodes in the
network that are similar over time.

We found for both PCA and eigenvector centrality, during the intact condition in the
experiment, classifiers that incorporated higher-order correlations
yielded consistently higher accuracy than classifiers trained only on
lower-order patterns (Fig.~\ref{fig:decoding_level},  a.\&d.).  We plot
the average correlations for up to the fourth order for the intact
condition (Fig.~\ref{fig:brain_plots}) representing the degree of
agreement by location pair over time.  By
contrast, we found that incorporating higher-order (greater than first
order) correlations did
not further improve decoding accuracy for the other listening
conditions or rest condition.  This suggests
that the cognitive processing that supported the most cognitively rich
condition
involved higher-order network dynamics.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/direction_of_field.pdf}
  \caption{\textbf{Direction of the field (adapted from~\citep{Turk13}).} The evolution of fMRI analyses started with
    univariate activation, which refers to
    the average amplitude of BOLD activity evoked by events of an
    experimental condition. Next, multivariate classifiers are trained
    on patterns of activation across voxels to decode distributed
    representations for specific events. The next, resting
    connectivity, is the temporal correlation of one or more seed
    regions with the remainder of the brain during rest. Additionally,
    task-based connectivity examines how these correlations differ by
    cognitive state. Following this increasing trajectory of
    increasing complexity, full connectivity considers all pairwise
    correlations in the brain, most commonly at rest.  Next, dynamic
    full connectivity considers how full connectivity changes over
    time. Continuing this line of reasoning, we expect higher-order network dynamics might provide even richer insights into the neural basis of cognition.}
  \label{fig:direction_of_field}
\end{figure}

Based on prior work ~\citep{Deme19} and following the direction of the field ~\citep{Turk13} we think our thoughts might be encoded in
dynamic network patterns, and possibly higher order network
patterns (Fig.~\ref{fig:direction_of_field}). We sought to test this hypothesis by developing an approach
to inferring high-order network dynamics from timeseries data.

One challenge in studying dynamic interactions is the
computational resources required to calculate higher-order correlations. 
We developed a computationally tractable model of network dynamics (Fig.~\ref{fig:methods_fig}) that takes in a feature
timeseries and outputs approximated first-order dynamics (i.e.,
dynamic functional correlations), second-order dynamics
(reflecting homologous networks that dynamically form and disperse),
and higher-order network dynamics (up to tenth-order dynamic
correlations).

We first validated our model using synthetic data, and explored how
recovery varied with different underlying data structures and kernels.   We then 
applied the approach to an fMRI dataset
~\citep{SimoEtal16} in which participants listened to an audio
recording of a story, as well as scrambled versions of the same story
(where the scrambling was applied at different temporal scales).  We
trained classifiers to take the output of the model and decode the
timepoint in the story (or scrambled story) that the participants were
listening to. We found that, during the intact listening condition in the
experiment, classifiers that incorporated higher-order correlations
yielded consistently higher accuracy than classifiers trained only on
lower-order patterns (Fig.~\ref{fig:decoding_level},  a.\&d.).  By contrast, these
higher-order correlations were not necessary to support decoding the other
listening conditions and (minimally
above chance) during a control rest condition.  This suggests
that the cognitive processing that supported the most cogntively rich listening conditions
involved second-order (or higher) network dynamics.

Although we found decoding accuracy was best when incorporating
higher-order network dynamics for all but rest
  condition, it is unclear if this is a product of the brain or the
  data collection technique.  It could be that the brain is
  second-order or it could be that fMRI can
  only reliably give second-order interactions. Exploring this method
  with other data collection technique will be important to
  disentangle this question.

  \subsection*{Concluding remarks}

How can we better understand how brain patterns change over
time? How can we quantify the potential network dynamics that might be
driving these changes? One way to judge the techniques of the future is
to look at the trajectory of the fMRI field so far has taken so far
(Fig.~\ref{fig:methods_fig}).  The field started with 
univariate activation, measuring the average activity for each voxel.
Analyses of multivariate activation followed, looking at spatial patterns of
activity over voxels. Next, correlations of activity were explored, first
with measures like resting connectivity that take temporal correlation
between a seed voxel and all other voxels then with full connectivty
that measure all pairwise correlations.  Additionally, this path of increasing
complexity also moved from static to dynamic measurements.  One
logical next step in this trajectory would be dynamic higher-order
correlations. We have created a method 
to support these calculations by scalably approximating dynamic higher-order
correlations.

\section*{Acknowledgements}
We acknowledge discussions with Luke Chang, Hany Farid, Paxton
Fitzpatrick, Andrew Heusser, Eshin Jolly, Qiang Liu, Matthijs van der
Meer, Judith Mildner, Gina Notaro, Stephen Satterthwaite, Emily
Whitaker, Weizhen Xie, and Kirsten Ziman. Our work was supported in
part by NSF EPSCoR Award Number 1632738 to J.R.M. and by a sub-award
of DARPA RAM Cooperative Agreement N66001-14-2-4-032 to J.R.M.  The
content is solely the responsibility of the authors and does not
necessarily represent the official views of our supporting
organizations.

\section*{Author contributions}
Concept: J.R.M.  Implementation: T.H.C., L.L.W.O., and
J.R.M.  Analyses: L.L.W.O and J.R.M.

\bibliographystyle{apacite}
\bibliography{CDL-bibliography/memlab}

\end{document}

### 2021-07-27 code/scripts/config.py
redo analysis

@@ -6,9 +6,9 @@ config = dict()
 config['template'] = 'run_job.sh'
 
 # ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-if (socket.gethostname() == 'Lucys-MacBook-Pro-3.local') or (socket.gethostname() == 'vertex.kiewit.dartmouth.edu') or (socket.gethostname() == 'vertex.local')or (socket.gethostname() == 'vpn-two-factor-general-230-139-212.dartmouth.edu'):
-    config['datadir'] = '/Users/lucyowen/Desktop/pca_env/pca_paper/pieman/data'
-    config['workingdir'] = '/Users/lucyowen/Desktop/pca_env/pca_paper/pieman'
+if (socket.gethostname() == 'Lucys-MacBook-Pro.local') or (socket.gethostname() == 'vertex.kiewit.dartmouth.edu') or (socket.gethostname() == 'vertex.local')or (socket.gethostname() == 'vpn-investment-office-231-132-37.dartmouth.edu'):
+    config['datadir'] = '/Users/lucyowen/repos/pca_paper-1/data'
+    config['workingdir'] = '/Users/lucyowen/repos/pca_paper-1'
     config['startdir'] = os.path.dirname(os.path.dirname(os.path.realpath(__file__)))  # directory to start the job in
     config['template'] = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'run_job_local.sh')
 else:
@@ -16,6 +16,11 @@ else:
     config['workingdir'] = '/dartfs/rc/lab/D/DBIC/CDL/f002s72/pca_paper/pieman'
     config['startdir'] = '/dartfs/rc/lab/D/DBIC/CDL/f002s72'
     config['template'] = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'run_job.sh')
+
+# config['datadir'] = '/opt/project/data'
+# config['workingdir'] = '/opt/project/'
+# config['startdir'] = os.path.dirname(os.path.dirname(os.path.realpath(__file__)))  # directory to start the job in
+# config['template'] = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'run_job_local.sh')
 
 # job creation options
 config['scriptdir'] = os.path.join(config['workingdir'], 'scripts')

### 2021-07-27 code/scripts/pieman_pca_analysis.py

@@ -1,6 +1,10 @@
 
-import timecorr as tc
-from timecorr.helpers import isfc, wisfc, mean_combine, corrmean_combine
+import sys
+import os
+sys.path.append(sys.path.append(os.path.dirname(os.path.dirname(os.path.abspath(__file__)))))
+
+#sys.path.append(os.path.abspath('..'))
+from decode_helpers.helpers import pca_decoder
 from scipy.io import loadmat
 import numpy as np
 import sys
@@ -12,11 +16,8 @@ import hypertools as hyp
 cond = sys.argv[1]
 chunk = sys.argv[2]
 reps = sys.argv[3]
-cfun = sys.argv[4]
-rfun = sys.argv[5]
-width = int(sys.argv[6])
-wp = sys.argv[7]
-ndims = sys.argv[8]
+rfun = sys.argv[4]
+ndims = sys.argv[5]
 
 if len(sys.argv) < 10:
     debug = False
@@ -24,6 +25,7 @@ else:
     debug = eval(sys.argv[9])
 
 result_name = 'pca_decode'
+
 
 if debug:
     results_dir = os.path.join(config['resultsdir'], result_name, rfun + '_debug', 'ndims_'+ ndims)
@@ -38,12 +40,6 @@ except OSError as err:
    print(err)
 
 
-
-laplace = {'name': 'Laplace', 'weights': tc.laplace_weights, 'params': {'scale': width}}
-delta = {'name': '$\delta$', 'weights': tc.eye_weights, 'params': tc.eye_params}
-gaussian = {'name': 'Gaussian', 'weights': tc.gaussian_weights, 'params': {'var': width}}
-mexican_hat = {'name': 'Mexican hat', 'weights': tc.mexican_hat_weights, 'params': {'sigma': width}}
-
 factors = 700
 
 if factors == 100:
@@ -55,7 +51,6 @@ pieman_data = loadmat(os.path.join(config['datadir'], pieman_name))
 pieman_conds = ['intact', 'paragraph', 'word', 'rest']
 
 
-weights_paramter = eval(wp)
 
 if debug:
     data = []
@@ -90,12 +85,7 @@ append_iter = pd.DataFrame()
 
 pca_data = np.asarray(hyp.reduce(list(data[conds == cond]), ndims=int(ndims)))
 
-iter_results = tc.helpers.pca_decoder(data[conds == cond], nfolds=2, dims=int(ndims),
-                                    combine=mean_combine,
-                                    cfun=eval(cfun),
-                                    rfun=None,
[…]

### 2021-07-27 code/scripts/pieman_pca_analysis_submit.py

@@ -24,20 +24,11 @@ cond_type = ['intact', 'paragraph', 'word', 'rest']
 
 
 # options for reps: integer
-reps =  str('100')
-
-# options for reps: cfuns
-cfuns =  [str('isfc')]
+reps =  str('10')
 
 # options for reps: rfuns
-#rfuns =  [str('eigenvector_centrality'), str('pagerank_centrality'), str('strength')]
 rfuns =  [str('PCA')]
 
-# options for widths: integer
-widths = [str(10)]
-
-# options for weight functions: laplace, gaussian, mexican hat, delta
-weights = ['gaussian']
 
 # options for dims: integers
 dims = ['100']
@@ -46,13 +37,13 @@ dims = ['100']
 debug = str('False')
 
 job_commands = list(np.array([list(map(lambda x: x[0]+" "+str(x[1])+" "+'unchunked'+" "+str(r)+
-                                                 " "+cfuns[0]+" "+rfuns[0]+" "+widths[0]+" "+weights[0]+" "+ dims[0]+ " "+debug,
+                                                 " "+rfuns[0]+" "+ dims[0]+ " "+debug,
                                        zip([job_script]*len(cond_type), cond_type)))
                               for r in range(int(reps))]).flat)
 
 # job_names should specify the file name of each script (as a list, of the same length as job_commands)
 job_names = list(np.array([list(map(lambda x: os.path.basename(os.path.splitext(x)[0])+'_'+'unchunked'+'_'+str(r)+'_'
-                                              +cfuns[0]+'_'+rfuns[0]+'_'+widths[0]+'_'+weights[0]+'_'+ dims[0] + '_'+debug+'.sh', cond_type))
+                                              '_'+rfuns[0]+'_'+ dims[0] + '_'+debug+'.sh', cond_type))
                            for r in range(int(reps))]).flat)
 
 

### 2021-07-27 paper/main.tex
WIP paper

@@
 
-This idea can be explored in this visual analogy for neural
+This idea can be explored in this visual analogy (Fig.~\ref{fig: pie_example}) for neural
 compression.  Here there are two images of pies, the top pie is more
@@
 variance is explaining both pies.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{figs/pie_example.pdf}
+  \caption{\textbf{Illustration of compression.} Visual analogy for
+    neural compression. Here are 2 images of pies, one more complex
+    thantheother. \textbf{a.}It takes fewer components to reach the same percent
+    variance explained in the less complex pie, which corresponds to
+    higher compression. \textbf{b.}However, with very few components, similar variance is explained in both pies.
+\textbf{c.} Plots the cumulative explained variance for more and more components.}
+  \label{fig:pie_example}
+\end{figure}
 
@@
 \section*{Methods}
-A major challenge to studying such patterns is that typically neither
-the correlations nor the hierarchical organizations of those
-correlations may be directly observed.  Rather, these fundamental
-properties must be inferred indirectly by examining the observable
-parts of the system-- e.g., the behaviors of the individual units of
-that system.  Here we propose a series of mathematical operations that may be used to approximate dynamic correlations at a range of scales (i.e., orders of interaction).
 
-There are two basic steps to our approach (Fig.~\ref{fig:methods_fig}).  In the first step, we take
-a number-of-timepoints ($T$) by number-of-features ($F$)
-\textit{matrix of observations} ($\mathbf{X}$) and we return a $T$ by
-$\frac{F^2 - F}{2}$ \textit{matrix of dynamic correlations}
-($\mathbf{Y}$).  Here $\mathbf{Y_0}$ describes, at each moment, how
-all of the features (columns of $\mathbf{X}$) are inferred to be
-interacting.  (Since the interactions are assumed to be non-recurrent
-and symmetric, only the upper triangle of the full correlation matrix
-is computed.)  In the second step, we project $\mathbf{Y_0}$ onto an
-$F$-dimensional space, resulting in a new $T$ by $F$ matrix
-$\mathbf{Y_1}$.  Note that $\mathbf{Y_1}$ contains information about
-the correlation dynamics present in $\mathbf{X}$, but represented in a
-compressed number of dimensions.  By repeatedly applying these two
-steps in sequence, we can examine and explore higher order dynamic
-correlations in $\mathbf{X}$.
+\subsubsection*{Dimensionality reduction}
 
-\subsection*{Dynamic correlations}
-Given a matrix of observations, we can compute the (static)
-correlations between any pair of observations, $\mathbf{X}_i$ and
-$\mathbf{X}_j$ using:
-\begin{align}
-  \mathrm{corr}(\mathbf{X}_i, \mathbf{X}_j) &= \frac{\sum_{t=1}^T \left(\mathbf{X}_i(t)
-                                              -
-                                              \bar{\mathbf{X}_i}\right) \left(\mathbf{X}_j(t)
-                                              -
-                                              \bar{\mathbf{X}_j}\right)}{\sqrt{\sum_{t=1}^T
-                                              \sigma^2_{\mathbf{X}_i} 
-                                              \sigma^2_{\mathbf{X}_j}}},~\mathrm{where}\\\label{eqn:corr}
-  \bar{\mathbf{X}_k} &= \sum_{t=1}^T
-                       \mathbf{X}_k(t),~\mathrm{and}\\
-  \sigma^2_{\mathbf{X}_k} &= \sum_{t=1}^T \left( \mathbf{X}_k -
-                            \bar{\mathbf{X}_k} \right)^2 
-\end{align}
-
-We can generalize this formula to compute time-varying correlations by
-incorporating a \textit{weight function} that takes a time $t$ as input,
-and returns how much the observed data every timepoint (including $t$)
-contribute to the correlations at time $t$ (Fig.~\ref{fig:weights}).
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/kernels}
-  \caption{\textbf{Examples of time-varying weights.  Each panel
-      displays per-timepoint weights at $t = 50$, evaluated for 100
-      timepoints ($1, ..., 100$).  a. Uniform weights.} The weights
-    are timepoint-invariant; observations at all timepoints are
-    weighted equally, and do not change as a function of $t$.  This is
-    a special case of weight function that reduces dynamic
-    correlations to static correlations.  \textbf{b. Dirac delta
-      function.} Only the observation at timepoint $t$ is given weight
-    (of 1), and weights for observations at all other timepoints are
-    set to 0.  \textbf{c. Gaussian weights.} Each observation's
-    weights fall off in time according to a Gaussian probability
-    density function centered on $\mu = t$.  Weights derived using
-    several different example variance parameters ($\sigma^2$ are
-    displayed.  \textbf{d. Laplace weights.}  Each observation's
-    weights fall off in time according to a Laplace probability
-    density function centered on $\mu = t$.  Weights derived using
-    several different example scale parameters ($b$) are displayed.
-    \textbf{e. Mexican hat (Ricker wavelet) weights.}  Each
-    observation's weights fall off in time according to a Ricker
-    wavelet centered on $t$.  This function highlights the
-    \textit{contrasts} between local versus surrounding activity
-    patterns in estimating dynamic correlations. Weights derived using
-    several different example width parameters ($\sigma$) are
-    displayed.  }
-  \label{fig:weights}
-\end{figure}
-
-Given a weight function $w(t)$ for timepoint $t$, evaluated at
-timepoints in the interval $\left[ 1, ..., T \right]$, we can extend the static correlation formula
-in Equation~\ref{eqn:corr} to reflect an \textit{instantaneous
-  correlation} at timepoint $t$:
-\begin{align}
-  \mathrm{timecorr}(\mathbf{X}_i, \mathbf{X}_j, t) &= \frac{\sum_{t=1}^T \left(\mathbf{X}_i(t)
-                                              -
-                                              \widetilde{\mathbf{X}}_i(t)\right) \left(\mathbf{X}_j(t)
-                                              -
-                                              \widetilde{\mathbf{X}}_j(t)\right)}{\sqrt{\sum_{t=1}^T
-                                              \widetilde{\sigma}^2_{\mathbf{X}_i}(t) 
-                                              \widetilde{\sigma}^2_{\mathbf{X}_j}}(t)},~\mathrm{where}\\\label{eqn:timecorr}
-  \widetilde{\mathbf{X}}_k(t) &= \sum_{i=1}^T
-                       w(t, i)\mathbf{X}_k(i),\\
-  \widetilde{\sigma}^2_{\mathbf{X}_k}(t) &= \sum_{i=1}^T \left( \mathbf{X}_k(i) -
-                            \widetilde{\mathbf{X}}_k(t) \right)^2,
-\end{align}
-and $w(t, i)$ is shorthand for $w(t)$ evaluated at timepoint $i$.
-Equation~\ref{eqn:timecorr} may be used to estimate the instantaneous
-correlations between every pair of observations, at each timepoint
-(i.e., $\mathbf{Y}$).
-
-\subsubsection*{Inter-subject dynamic correlations}
-Equation~\ref{eqn:timecorr} provides a means of taking a single
-observation matrix, $\mathbf{X}$ and estimating the dynamic
-correlations from moment to moment, $\mathbf{Y}$.  Suppose that one
-has access to a set of multiple observation matrices that reflect the
-same phenomenon.  For example, one might collect neuroimaging data
-from several experimental participants, as each participant performs
-the same task (or sequence of tasks).  Let $\mathbf{X}_1$,
-$\mathbf{X}_2$, ..., $\mathbf{X}_P$ reflect the $T$ by $F$ observation
-matrices for each of $P$ participants in an experiment.  We can use
-\textit{inter-subject functional connectivity}~\citep[ISFC;
-][]{SimoEtal16} to compute the degree of stimulus-driven correlations
-reflected in the multi-participant dataset at a given timepoint $t$
-using:
-\begin{align}
-\bar{\mathbf{C}}(t) = M\left(R\left(\frac{1}{2P} \sum_{i=1}^P
-  Z\left(Y_i(t)\right)^\mathrm{T} + Z\left(Y_i,(t)\right)\right)\right),
-\end{align}
-where $M$ extracts and vectorizes the diagonal and upper triangle of a symmetric
-matrix, $Z$ is the Fisher $z$-transformation~\citep{Zar10}:
-\begin{align}
-Z(r) = \frac{\log(1+r) - \log(1-r)}{2}
-\end{align}
-$R$ is the inverse of $Z$:
-\begin{align}
-R(z) = \frac{\exp(2z - 1)}{\exp(2z + 1)},
-\end{align}
-and $\mathbf{Y}_i(t)$ denotes the correlation matrix
-(Eqn.~\ref{eqn:corr}) between each column of $\mathbf{X}_i$ and each
-column of the average observations from all \textit{other}
-participants, $\bar{\mathbf{X}}_{ \backslash i}$:
-\begin{align}
-  \bar{\mathbf{X}}_{ \backslash i} = R\left(\frac{1}{P-1}\sum_{i \in
-  \backslash i} Z\left( \mathbf{X}_i \right) \right),
-\end{align}
-where $ \backslash i$ denotes the set of all participants other than
-participant $i$. In this way, the $T$ by $\left( \frac{F^2 - F}{2}
-  + F \right)$
-matrix $\bar{\mathbf{C}}$ is the time-varying extension of the ISFC
-approach developed by \cite{SimoEtal16}.
-
-\subsection*{Higher-order correlations}
-Given a timeseries of dynamic correlations (e.g., obtained using
-Eqn.~\ref{eqn:timecorr}), higher-order correlations reflect the
-dynamic correlations between columns of $\mathbf{Y}$.  Given unlimited
-computing resources, one could use repeated applications of
-Equation~\ref{eqn:timecorr} to estimate these higher-order
-correlations (i.e., substituting in the previous output, $\mathbf{Y}$,
-for the input, $\mathbf{X}$ in the equation).  However, because each
-output $\mathbf{Y}$ has $\mathcal{O}(F^2)$ columns relative to $F$ columns in
-the input $\mathbf{X}$, the output of Equation~\ref{eqn:timecorr}
-grows with the square of the number of repeated applications (total
-cost of computing $n$\textsuperscript{th} order correlations is
-$\mathcal{O}(F^{2n})$ for $n \in \mathcal{J}, n > 0$).  When $F$ or $n$ is large,
-this approach quickly becomes intractable.
-
-To make progress in computing $\mathbf{Y_{n+1}}$, we can approximate
-$\mathbf{Y}_n$ by computing an $\mathcal{O}(F)$-dimensional embedding of
-$\mathbf{Y}_n$, termed $\hat{\mathbf{Y}}_n$, and then we can apply
-Equation~\ref{eqn:timecorr} to $\hat{\mathbf{Y}}_n$ rather than
-directly to $\mathbf{Y}_n$.  This enables us to maintain $\mathcal{O}(n)$
-scaling with respect to $n$, rather than exponential scaling via the
-direct approach.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.5\textwidth]{figs/methods_fig}
-  \caption{\textbf{Computing higher order correlations. } Dynamic
-    correlations
-  are computed then approximated to the same size as original data
-  using dimensionality reduction.
-  This process is repeated, and can be used to compute up to any
-  arbitrary order with compuations scaling linearly (as opposed to exponentially) with order. }
-  \label{fig:methods_fig}
-\end{figure}
-
-There are many possible methods for computing $\hat{\mathbf{Y}}_n$
-from $\mathbf{Y}_n$, including traditional dimensionality reduction
-approaches and graph theory based approaches as described next.  In the
-\textit{Discussion} section we elaborate on other potential approaches.
-
-\subsubsection*{Dimensionality reduction-based approaches to computing
-  $\hat{\mathbf{Y}}_n$}
-
-Commonly used dimensionality reduction algorithms include Principal
-Components Analysis~\citep[PCA; ][]{Pear01}, Probabilistic
-PCA~\citep[PPCA; ][]{TippBish99}, Exploratory Factor
-Analysis~\citep[EFA; ][]{Spea04}, Independent Components
-Analysis~\citep[ICA; ][]{JuttHera91, ComoEtal91}, $t$-Stochastic
-Neighbor Embedding~\citep[$t$-SNE; ][]{MaatHint08}, Uniform Manifold
-Approximation and Projection~\citep[UMAP; ][]{McInHeal18},
-non-negative matrix factorization~\citep[NMF; ][]{LeeSeun99},
-Topographic Factor Analysis (TFA)~\cite{MannEtal14b}, Hierarchical
-Topographic Factor analysis (HTFA)~\cite{MannEtal18}, Topographic
-Latent Source Analysis (TLSA)~\cite{GersEtal11}, Dictionary
-learning~\citep{MairEtal09a, MairEtal09b}, deep
-autoencoders~\citep{HintSala06}, among others.  While complete
-characterizations of each of these algorithms is beyond the scope of
-the present manuscript, the general intuition driving these
-approaches is to compute the $\hat{\mathbf{Y}}$ with $i$ columns that
-is closest to the original $\mathbf{Y}$ with $j$ columns, and where
-(typically) $i \ll j$.  The different approaches place different
-constraints on what properties $\hat{\mathbf{Y}}$ must satisfy and
-which aspects of the data are compared (and how) to characterize the
-match between $\hat{\mathbf{Y}}$ and  $\mathbf{Y}$.
-
-Applying dimensionality reduction algorithms to $\mathbf{Y}$ yields a
-$\hat{\mathbf{Y}}$ whose columns reflect weighted combinations (or
-nonlinear transformations) of the original columns of $\mathbf{Y}$.
-This has two main consequences.  First, with each repeated
-dimensionality reduction, the resulting $\hat{\mathbf{Y}}_n$ has lower
-and lower fidelity (with respect to what the ``true'' $\mathbf{Y}_n$
-might have looked like without using dimensionality reduction to
-maintain scalability).  In other words, computing $\hat{\mathbf{Y}}_n$
-is a lossy operation.  Second, whereas the columns of $\mathbf{Y}_n$
-may be mapped directly onto pairs of columns of $\mathbf{Y}_{n-1}$,
-that mapping either becomes less cleanly defined in
-$\hat{\mathbf{Y}}_n$ due to the reweightings and/or nonlinear
-transformations.
-
-\subsubsection*{Graph theory-based approaches to computing
-  $\hat{\mathbf{Y}}_n$}
-
-Graph theoretic measures take as input a matrix of interactions (e.g.,
-using the above notation, an $F \times F$ correlation matrix or
-binarized correlation matrix reconstituded from a single timepoint's
-row of $\mathbf{Y}$) and return as output a set of $F$ measures
-describing how each node (feature) sits within that interactions
-matrix with respect to the rest of the population.  Common measures
-include betweeness centrality~\citep[the proportion of shortest paths
-between each pair of nodes in the population that involves the given
-node in question; e.g., ][]{Newm05, OpsaEtal10, Bart04, GeisEtal08,
-  Free77}; diversity and dissimilarity~\citep[characterizations of how
-differently connected a given node is from others in the population;
-e.g., ][]{Rao82, Lin09, RicoSzei06}; Eigenvector centrality and
-pagerank centrality~\citep[measures of how influential a given node is
-within the broader network; e.g., ][]{Newm08, Bona07, LohmEtal10,
-  HaluEtal13}; transfer entropy and flow coefficients~\citep[a measure
-of how much information is flowing from a given node to other nodes in
-the network; e.g., ][]{HoneEtal07, Schr00}; $k$-coreness
-centrality~\citep[a measure of the connectivity of a node within its
-local sub-graph; e.g., ][]{AlvaEtal05, ChriFowl10}; within-module
-degree~\citep[a measure of how many connections a node has to its
-close neighbors in the network; e.g., ][]{RubiSpor10}; participation
-coefficient~\citep[a measure of the diversity of a node's connections
-to different sub-graphs in the network; e.g., ][]{RubiSpor10}; and
-sub-graph centrality~\citep[a measure of a node's participation in all
-of the network's sub-graphs; e.g., ][]{EstrRodr05}.
-
-As an alternative to the above dimensionality reduction approach to
-embedding $\mathbf{Y}_n$ in a lower-dimensional space, but still
-allowing for scalable explorations of higher-order structure in the
-data, we also explore using the above graph theoretic measures as a
-means of obtaining $\hat{\mathbf{Y}}_n$.  In particular: for a given
-graph theoretic measure, $\eta: \mathcal{R}^{F \times F} \rightarrow
-\mathcal{R}^F$, we can use $\eta$ to tranform each row of
-$\mathbf{Y}_n$ in a way that characterizes the corresponding
-graph-theoretic properties of each column.  Whereas the dimensionality
-reduction approach to computing $\hat{\mathbf{Y}}_n$ is lossy,
-the graph-theory approach is lossless.  However, whereas the
-dimensionality reduction approach maintains ties (direct or indirect)
-to the original activity patterns reflected in $\mathbf{Y}_{n-1}$, the
-graph-theory approach does not.  Instead, the graph-theory
-characterizes the nature and timecourse of each feature's
-\textit{participation} in the network.
+Explain PCA
 
 \subsection*{Evaluation metrics}
-We evaluate our approach to extracting dynamic correlations and
-higher-order correlations using several metrics detailed next.  First,
-we generated synthetic data using known time-varying correlations, and
-then we evaluated the fidelity with which Equation~\ref{eqn:timecorr}
-could recover those correlations (for synthetic datasets with
-different properties, and using different kernels to define the
-weights; Fig.~\ref{fig:weights}).  We then turned to a series of
-analyses on a (real) neuroimaging dataset where the ground truth
-correlations were \textit{not} known.  We evaluated whether the
-recovered correlations could be used to accurately label held-out
+ We evaluated the degree of
+compression of held-out
 neuroimaging data with the time at which it was collected.  We used this latter evaluations (using timepoint
 decoding) as a proxy for
-gauging how much explanatory power the recovered correlations held
+gauging how much explanatory power the compressed data held
 with respect to the observed data.
-
-\subsubsection*{Generating synthetic data}
-To explore recovery of a constant covariance (Fig.~\ref{fig:synthetic_data},  a.), we generated syntheic
-data sampled from a constant covariance matrix. To do this, we created one random
-covariance matrix, $K$, with 50 features, and for each of the 300 timepoints  we
-sampled from a Gaussian distribution centered on $K$.  Similarly, we generated synthetic data
-sampled from a random covariance matrix (Fig.~\ref{fig:synthetic_data}, b.) by creating a new random
-covariance matrix $K(t)$, for each of the 300 timepoints and sampled from a
-Gaussian distribution centered on $K(t)$.
-
-To generate synthetic data from a dynamically changing covariance
-matrix (Fig.~\ref{fig:synthetic_data},  c.),
-we generated two random covariance
-matrices, $K_{1}$ and $K_{2}$.  We then computed a weighted average covariance matrix
-for each of the 300 timepoint, $K(t)$, 
-by taking the linearly spaced weights ($w$) of the two random matrices, 
-\begin{align}
-K(t) = w(t) * K_{1} + (1 - w(t)) *  K_{2}, \\\label{eqn:ramping}
-\end{align}
-and for each of the 300 timepoints sampled from a
-Gaussian distribution centered on $K(t)$.
-
-Lastly, for the synthetic data containing block structure (Fig.~\ref{fig:synthetic_data},  d.), we followed the same
-process of creating synthetic data sampled from a constant covariance
-matrix (see above) but sampled from a new random covariance matrix
-after 60 consecutive timepoints.  We then pieced the blocks together
-to create a sythetic dataset with 300 total timepoints but drawn from
-5 separate covariance matrices.
-
-\subsubsection*{Recovery of ground truth parameters from synthetic
-  data}
-
-We applied timecorr, using delta and gaussian (width = 10) kernels
-Fig.~\ref{fig:weights}) to each of these 
-synthetic datasets, then correlated each recovered
-correlation matrix with the ground truth.  We repeated this process 10
-times and explored how recovery varies
-with the kernel and the specific structure of the data. For the
-ramping synthetic dataset (Fig.~\ref{fig:synthetic_data},  c.)  and for the
-block synthetic dataset (Fig.~\ref{fig:synthetic_data},  d.)  we made further
-comparisons of the timecorr recovered correlation matrices. We
-compared the ramping recovered correlation matrices to only the first random covariance matrix $K_{1}$
-(First, Fig.~\ref{fig:synthetic_data},  c.) and to only the last
-random covariance matrix $K_{2}$ (Last, Fig.~\ref{fig:synthetic_data},
-c.) from Equation~\ref{eqn:ramping}. We also compared the block recovered correlation matrices in to
-the block specific covariance matrix (Block 1-5,
-Fig.~\ref{fig:synthetic_data},  d.).
 
@@
 the fMRI datasets to obtain a time series of 700 node activities for
-every participant.  We then computed the dynamic weighted ISFC using a
-range of kernels and widths.  Specifically, we used Gaussian, Laplace,
-and mexican hat kernels, as well as widths of 5, 10, 20, and 50.  We
-then approximated these dynamic correlation using two reduction measures, PCA and eigenvector centrality, and computed the dynamic weighted ISFC on the approximations.
-We repeated this process up to 10th order approximated correlations.
+every participant.  We then applied dimensionality reduction
+(Incremental PCA) for each group.
+
+We then compared the groups’ activity patterns (using Pearson
+correlations) to estimate the story times each corresponding pattern
+using more and more principle components.
 
 To assess decoding accuracy, we randomly divided participants for each
-stimulus into training and testing groups. For the zeroth order, we
-computed the mean factor activity for each group.  For all subsequent
-orders up to the tenth order, we computed the mean approximated
-dynamic ISFC of factor activity for each group. To assess how
-additional higher-order correlations contribute to decoding accuracy,
-for each order we included a weighted-mixutre (described below) of the activity patterns of all
-previous orders.
-For each group of participants in turn, we compared these activity patterns (using Pearson correlations) to estimate the story times each pattern corresponded to.
-Specifically, we asked, for each timepoint: what are the correlations
+stimulus into training and testing groups. We then compared the
+groups’ activity patterns (using Pearson correlations) to estimate the
+story times each corresponding pattern using more and more principle
+components (as the data became less compressed). Specifically, we asked, for each timepoint: what are the correlations
 between the first group's and second group's activity patterns at each
@@
 
-For each order we obtained the weighted-mixture of the correlation
-matrices for the current order and all previous orders using mixing parameter, $\phi$, where $0 <\phi< 1$ reflects a
-weighted mixture of order based decoding Fig.~\ref{fig:decoding_level}
-Panel c. ). We calculated  $\phi$, by
-subdividing the training group and using the quasi-Newton method of
-Broyden, Fletcher, Goldfarb, and Shanno (BFGS ~\citep{NoceWrig06}) for optimization. We
-repeated this cross-validation process 10 times for each parameter set.
+\section*{Results}
 
-\section*{Results}
-\subsubsection*{Synthetic data}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/synthetic_data.pdf}
-  \caption{\textbf{Dynamic correlation recovery with synthetic data. } Using synthetic data containing different
-    underlying correlational structure,
-   we test how well we can recover dynamic correlation matrices using two kernels (delta and gaussian,
-    width = 10) when compared to groud truth.  We plot recovery using of datasets containing the
-    following underlying structure:
-    \textbf{a. Constant. b. Random. c. Ramping.  d. Block.}}
-  \label{fig:synthetic_data}
-\end{figure}
-
-To assess the performance of dynamic correlation recovery using
-timecorr, we varied width the kernel and the specific structure of the
-data. We applied timecorr, using delta and gaussian kernels
-Fig.~\ref{fig:weights}) to each of the following  
-synthetic datasets: constant, random, ramping, and block.  We then correlated each recovered
-correlation matrix with the ground truth.
-
-For the constant synthetic dataset, a gaussian kernel (width=10)
-outperformed the delta kernel (Fig.~\ref{fig:synthetic_data},  a.).  This is in contrast with the random
-synthetic dataset, for which the delta kernel best captures the
-rapidly changing structure (Fig.~\ref{fig:synthetic_data},  b.). For
-the ramping synthetic dataset, the slow changing strucutre within the
-data is best
-captured by the gaussian kernel and the best recovery occurs in the
-middle (Ramping, Fig.~\ref{fig:synthetic_data},
-c.). In addition to comparing the timecorr recovered correlation
-matrices to the ground truth, we
-further compared the ramping recovered correlation matrices to only the first random covariance matrix $K_{1}$
-(First, Fig.~\ref{fig:synthetic_data},  c.) and to only the last
-random covariance matrix $K_{2}$ (Last, Fig.~\ref{fig:synthetic_data},
-c.), both of which perform best at the beginning and end respectively.
-
-Similary for the block sythetic dataset, we compared the timecorr
-recovered correlation matrices to the ground truth as well as to each
-block-specific covariance matrix (Block 1-5,
-Fig.~\ref{fig:synthetic_data},  d.).  Although the structure is
-changing by block, the gaussian kernel once again outperforms the
-delta kernel.  Performance does however drop near even boundaries for
-when using the gaussian kernel.
-
-\subsubsection*{Neuroimaging dataset~\citep{SimoEtal16}}
 For our decoding analysis, we used HTFA-derived node activities
@@
 with their eyes open in the scanner (rest condition; 36
-participants). We sought to demonstrate how higher-order correlations
-may be used to examine dynamic interactions of brain patterns in
-(real) multi-subject fMRI datasets. This story listening dataset was
+participants). We sought to understand the 'dimensionality' of the
+neural patterns by using more and more principle
+components to decode a multi-subject fMRI datasets. This story listening dataset was
 collected as part of a separate study, where the full imaging
@@
   \centering
-  \includegraphics[width=\textwidth]{figs/decode_level.pdf}
-  \caption{\textbf{Decoding by order.
-      a.\&d. Relative decoding accuracy by order.} Ribbons of each
-    color display cross-validated decoding performance averaged over
-    all parameters for each condition (intact, paragraph, word,
-    and rest) using PCA (a.) or eigenvector centrality (d.) to
-    approximate correlations. Decoders were trained using increasingly more
-    higher-order information and this ribbons are displayed relative
-    to chance at 0. The red dots indicates
-    maximum decoding accuracy for each condition. \textbf{b.\&e. Z-transformed decoding accuracy
-      by order.}  We Z-transformed the decoding accuracy by order to better
-    visualize the order with the maximum decoding accuracy for each
-    condition using PCA (b.) or eigenvector centrality (e.) to
-    approximate correlations. \textbf{c.\&f. Optimized weights.} Bar heights indicate the
-    optimized mixing paramete $\phi$ of each contributing order up to
-    and including the order with the maximum
-    decoding accuracy for each contributing order using PCA (c.) or eigenvector centrality (f.) to
-    approximate correlations.
-  For the order with maximum decoding accuracy by condition, we show
-  barplots of the optimized weights $\phi$  for each contributing
-  order.}
-  \label{fig:decoding_level}
+  \includegraphics[width=\textwidth]{figs/decode_interpret.pdf}
+  \caption{\textbf{Decoding accuracy.} \textbf{a. Decoding accuracy by
+      number of components.} Ribbons of each color display
+    cross-validated decoding performance for each condition (intact,
+    paragraph, word, and rest). Decoders were trained using
+    increasingly more principle components and displayed relative to
+    chance (red line). \textbf{b. Fixed decoding accuracy by number of
+      components.} We zoom in on the plot shown in \textbf{a.} and add
+    a line denoting fixed decoding accuracy (.05). We plot where the
+    intact, paragraph, and word conditions intersect.
+    \textbf{c.
+Explanation of inflection metric.} First the we fit a sigmoid function to the decoding accuracy by number of components.
+Second, we found where the second derivate is both positive and less than .0001.
+Last, we then plot that inflection point as a single metric to capture the slope and asymptote of the curve.}
+  \label{fig:decode_interpret}
 \end{figure}
 
-We next evaluated if our model of high-order correlations in brain activity can capture cognitively relevant brain patterns.
-We performed a decoding analysis, using cross validation to estimate (using other participants’ data) which parts of the story each weighted-mixture of higher-order brain activity pattern corresponded to (see \textit{Materials and methods}).
+We performed a decoding analysis, using cross validation to estimate
+(using other participants’ data) which parts of the story the additional added
+principle component corresponded to (see \textit{Materials and methods}).
 We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.
@@
 Separately for each experimental condition, we divided participants
-into two groups. For the zeroth order, we computed the mean factor
-activity for each group.  For all subsequent orders up to the tenth
-order, we computed the mean approximated dynamic ISFC of factor
-activity for each group (see \textit{Materials and methods}), and
-combined in a weighted mixutre with all previous orders
-(i.e. cross-validation for the second
-order contained a weighted-mixture of zeroth, first, and second order
-(Fig.~\ref{fig:decoding_level}, c.\&f.).  For
-each order, we correlated the group 1 activity patterns with group 2
-activity patterns.  We then subdivided the group 1 to obtain an
-optimal weighting parameter for each order’s correlation matrix using
-the same cross validation method. We used the optimal weighting
-parameters to obtain a weighted-mixture (see \textit{Materials and methods}) of each order’s correlation
-matrix. Using these correlations, we labeled the group 1 timepoints
-using the group 2 timepoints with which they were most highly
-correlated; we then computed the proportion of correctly labeled group
-1 timepoints. (We also performed the symmetric analysis whereby we
-labeled the group 2 timepoints using the group 1 timepoints as a
-template.) We repeated this procedure 100 times (randomly re-assigning
-participants to the two groups each time) to obtain a distribution of
-decoding accuracies for each experimental condition. There were 272
+into two groups. Starting with 1 principle comonent, for
+each dimension we added another principle component, and we correlated the group 1 activity patterns with group 2
+activity patterns.  There were 272
 timepoints for paragraph condition, 300 timepoints for intact and word
@@
 
-We repeated this process for each set of parameters, varying kernel
-type and width, and averaged over the reduction technique used to
-approximate the higher-order correlations (PCA Fig.~\ref{fig:decoding_level},  a.-c. and eigenvector
-centrality Fig.~\ref{fig:decoding_level},  d.-f.).  Since there is no ground truth in these analyses, and we did not know
-which parameters best capture the data, we instead report a robustness
-search by averaging over the parameters and reporting which results
-consistently showed up across all parameters.
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/brain_plots.pdf}
-  \caption{\textbf{Average correlations by order for the intact listening condition.} Using eigenvector
-    centrality to approximate higher-order correlations for the intact
-    listening condition, we plot the
-    strongest 25\% absolute value
-    mean correlation for order: \textbf{a. 1. b. 2. c. 3.,
-      and d. 4}, representing the degree of agreement by
-    location pair over time.}
-  \label{fig:brain_plots}
-\end{figure}
-
-The two methods used to approximate the higher-order correlations (PCA Fig.~\ref{fig:decoding_level},  a.-c. and eigenvector
-centrality Fig.~\ref{fig:decoding_level},  d.-f.) capture different
-facets of the activity patterns.  Using PCA, the higher-order
-correlations are all linked to the original activity patterns, whereas
-eigenvectory centrality breaks the immediate link with specific brain
-areas and instead characterizes the position of the nodes in the
-network that are similar over time.
-
-We found for both PCA and eigenvector centrality, during the intact condition in the
-experiment, classifiers that incorporated higher-order correlations
-yielded consistently higher accuracy than classifiers trained only on
-lower-order patterns (Fig.~\ref{fig:decoding_level},  a.\&d.).  We plot
-the average correlations for up to the fourth order for the intact
-condition (Fig.~\ref{fig:brain_plots}) representing the degree of
-agreement by location pair over time.  By
-contrast, we found that incorporating higher-order (greater than first
-order) correlations did
-not further improve decoding accuracy for the other listening
-conditions or rest condition.  This suggests
-that the cognitive processing that supported the most cognitively rich
-condition
-involved higher-order network dynamics.
+- As complexity of the stimulus increases, decoding accuracy
+increases (Fig.~\ref{fig:decode_interpret},  a.).
 

### 2023-02-13 code/decode_helpers/decode_helpers/helpers.py
push whatever was here

@@ -137,7 +137,7 @@ class CorrelationClassifier(BaseEstimator):
 
         full_corrs = 1 - pdist(X, metric='correlation')
         for i in range(X.shape[1]):
-            next_X = np.delete(X.copy(), i, axis=1)
+            next_X = np.delete(X.copy(), i, axis=1) # This isn't actually deleting anything at this point
             next_corrs = 1 - pdist(next_X, metric='correlation')
             self.feature_importances_[i] = pearsonr(full_corrs, next_corrs)[0]
 
@@ -162,6 +162,12 @@ def decoding(in_data, out_data, n_features, **kwargs):
 
     acc = np.sum(y == predictions.round().astype(int))/in_data.shape[0]
 
+
+    try_it = (1 - sd.cdist(np.atleast_2d(y).T, np.atleast_2d(predictions).T, 'euclidean'))
+
+    import seaborn as sns
+    sns.heatmap(try_it)
+    plt.show()
     return acc
 
 def find_features(data, dims=10):
@@ -221,7 +227,7 @@ def pca_decoder_rfe(data, nfolds=2, dims=10):
             results_pd_full = pd.concat([results_pd_full, results_pd])
 
 
-    return try_results_pd_full
+    return results_pd_full
 
 
 def pca_decoder(data, nfolds=2, dims=10):

### 2023-02-13 paper/main.tex

@@
 
-\title{High-order dynamic neural correlations reflect naturalistic processing
+\title{Understanding brain complexity in naturalistic processing
 in humans} \author{Lucy
@@
 - As complexity of the stimulus increases, decoding accuracy
-increases (Fig.~\ref{fig:decode_interpret},  a.).
+increases (Fig.~\ref{fig:decode_interpret},  a.).  Replication of
+Simony findings.
+
+- As complexity of the stimuli increases, we need fewer components to
+decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
+
+- As complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{figs/decode_pcs_network.pdf}
+  \caption{\textbf{Inflection points by network.} \textbf{a.}
+    Inflection point was calculated as explained in
+    Fig.~\ref{fig:decode_interpret}, b. Analyses were limited by the
+    brain networks (using the Yeoetal11 network parcellation) and
+    arranged in increasing order relative to the intact condition.
+    \textbf{b.
+and c.} For the total time in the intact condition, we are plotting the relative inflection points (\textbf{b.}) and corresponding number of components (\textbf{c.}) by network.
+\textbf{d.} The network parcellation defined by Yeo et al.
+(2011) is displayed on the inflated brain maps.
+The colors and network labels serve as a legend for \textbf{a.} and \textbf{d.}}
+  \label{fig:decode_pcs_network}
+\end{figure}
+
+As complexity of the stimuli increases, decoding accuracy increases
+with higher cognitive areas. (Fig.~\ref{fig: decode_pcs_network}).
+
+ If, there is some understanding of the narrative that accumulates over time, we should be able to see that change.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{figs/decode_pcs.pdf}
+  \caption{\textbf{Inflection points by thirds.} \textbf{a.}
+    Decoding accuracy by number of components not broken into thirds
+    (Fig.~\ref{fig:decode_interpret} a.). \textbf{b. and c.}
+    Quantifying changes in decoding accuracy across
+time. \textbf{b.} Slope of decoding accuracy was calculated by fitting
+a regression line for each component/condition for each
+third. \textbf{c.}  We also repeated the analysis
+(Fig.~\ref{fig:decode_interpret}, b.) to obtain the inflection point
+for each condition and for each third.
+\textbf{d.}  Decoding accuracy by number of components for each third of the scan time.
+We repeated the same analysis in Fig.~\ref{fig:decode_interpret} a.
+but breaking the scan time for each condition into 3 intervals.
+}
+  \label{fig: decode_pcs_thirds}
+\end{figure}
+
+- Increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
+- Decreases in decoding accuracy for the word-scrambled and rest
+condition.
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/direction_of_field.pdf}
-  \caption{\textbf{Direction of the field (adapted from~\citep{Turk13}).} The evolution of fMRI analyses started with
-    univariate activation, which refers to
-    the average amplitude of BOLD activity evoked by events of an
-    experimental condition. Next, multivariate classifiers are trained
-    on patterns of activation across voxels to decode distributed
-    representations for specific events. The next, resting
-    connectivity, is the temporal correlation of one or more seed
-    regions with the remainder of the brain during rest. Additionally,
-    task-based connectivity examines how these correlations differ by
-    cognitive state. Following this increasing trajectory of
-    increasing complexity, full connectivity considers all pairwise
-    correlations in the brain, most commonly at rest.  Next, dynamic
-    full connectivity considers how full connectivity changes over
-    time. Continuing this line of reasoning, we expect higher-order network dynamics might provide even richer insights into the neural basis of cognition.}
-  \label{fig:direction_of_field}
-\end{figure}
+- We trained classifiers using more and more principle components to decode, and compared across condi- tions with varying degrees of cognitive richness.
+-We found that as listening conditions become more cognitively rich,
+decoding accuracy increased.
+-Also, decoding accuracy increased as understanding of the narrative
+accumulated over time, in more complex listening conditions.
+- Decoding accuracy also increased in higher cognitive areas, in more complex listening conditions.
+-We found that as story listening conditions become more complex, more
+components are required to decode.
+-We also found we could decode better with more impoverished data when
+there is the underlying structure of the narrative providing more cognitive richness.
+-We posit that as the complexity of our thoughts increases, neural
+compression decreases. However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
 

### 2023-02-13 paper/main.tex

@@
 We're interested in the complexity of brain patterns that underly
-different types of thoughts. To explore this question space, we take
-brain patterns recorded under different experimental conditions and
+different types of thoughts. To explore this question space, we will take
+brain patterns recorded under different experimental conditions used in Aim 2, and
 project them into lower dimensional spaces using principle components
@@
 someone is listening to.
-Then we can explore decoding accuracy as a
-function of the number of components, or dimensions, in the
-low-dimensional space under different cognitive conditions.
+
+This work has been inspired, in part, by~\cite{MackEtal20}. In this paper, 
+they investigated the role of the prefrontal cortex in filtering out 
+irrelevant content.  Specifically, they looked at if the vmPFC performs 
+data reduction on incoming information through compression.  
+This was motivated, in part, by orbital frontal cortex (OFC) 
+compression in rats~\citep{ZhouEtal19}.
+They studied this using a learning paradigm in which participants had to classify insects based on different numbers of feature dimensions.
+The idea was that participants in some learning blocks, participants could identify the insects based on one feature (low complexity) or several features (high complexity), but importantly the stimuli remained the same across all learning problems.
+They found that complexity and compression had an inverse relationship; the lower complexity of a conceptual space, the higher the degree of compression.
+Building on this idea, we wonder if varying degrees of compression is performed throughout the brain.
+We also want to test this idea, but using varying levels of engagement listening to a naturalistic stimuli.
+
+To understand the degree of compression throughout the brain during cognition, we will use the same fMRI data from Aim 2, collected while participants listened to a story in different scrambling conditions.
+We will measure the degree that multivoxel activation patterns are compressed during story listening using principle components analysis (PCA) a method for low-rank approximation of multidimensional data~\citep{EckaYoun36}.
+We will explore this using decoding accuracy as a function of the number of components, or dimensions, in the low-dimensional space under different cognitive conditions.
 
@@
 
-This idea can be explored in this visual analogy (Fig.~\ref{fig: pie_example}) for neural
-compression.  Here there are two images of pies, the top pie is more
-complex than the bottom.  On the left we're illustrating that it takes
-fewer componenets to reach the same 95 percent variance explained in
+This idea can be explored in this visual analogy (Fig.~\ref{fig: pie_example}) for neural compression.
+Here there are two images of pies, the top pie is more complex than the bottom.
+On the left we're illustrating that it takes
+fewer components to reach the same 95 percent variance explained in
 the less complex pie, which corresponds to higher compression.
@@
     neural compression. Here are 2 images of pies, one more complex
-    thantheother. \textbf{a.}It takes fewer components to reach the same percent
+    than the other. \textbf{a.} It takes fewer components to reach the same percent
     variance explained in the less complex pie, which corresponds to
-    higher compression. \textbf{b.}However, with very few components, similar variance is explained in both pies.
+    higher compression. \textbf{b.} However, with very few components, similar variance is explained in both pies.
 \textbf{c.} Plots the cumulative explained variance for more and more components.}
@@
 
-\section*{Methods}
+% PCA, or the closely related factor analysis and singular value decomposition (SVD) (Hastie et al., 2009), is widely used in the study of individual differences and aids estimating how many latent components, or “factors”, underlie a pattern of (item) responses within or across participants, as for instance in the context of intelligence (Spearman, 1904) or personality tests (Cattell, 1947).
+In the context of neuroimaging, PCA has been used to identify brain networks (Huth et al., 2012; Friston et al., 1993).
 
-\subsubsection*{Dimensionality reduction}
-
-Explain PCA
+ % add citation:
+% S.M.Smith,A.Hyva ̈rinen,G.Varoquaux,K.L.Miller,and C. F. Beckmann, “Group-PCA for very large fMRI datasets,” NeuroImage, vol. 101, pp. 738, 2014.
 
 \subsection*{Evaluation metrics}
- We evaluated the degree of
+ We will evaluate the degree of
 compression of held-out
-neuroimaging data with the time at which it was collected.  We used this latter evaluations (using timepoint
+neuroimaging data by assessing the time at which it was collected.  We will use this evaluation (timepoint
 decoding) as a proxy for
@@
 
-To explore how higher-order structure varies with stimulus structure
-and complexity, we used a previous neuroimaging dataset
+To explore how compression varies with complexity, we will use a previous neuroimaging dataset
 ~\cite{SimoEtal16}  in which participants listened to an audio recording of a story; 36 particpants listen to an intact version of the story, 17 participants listen to time-scrambled recordings of the same story where paragraphs were scrambled, 36 particpants listen to word-scrambled version and 36 participants lay in rest condition.
 
-Prior work has shown participants share similar neural responses to
-richly structured stimuli when compared to stimuli with less
-structure.  To assess whether the moment-by-moment higher order
-correlations were reliably preserved across participants, we used
-inter-subject functional connectivity (ISFC) to isolate the
-time-varying correlational structure (functional connectivity patterns
-that were specifically driven by the story participants listened to.
 Following the analyses conducted by (HTFA)~\cite{MannEtal18}, we first
-applied \textit{hierarchical topographic factor analysis} (HTFA) to
+apply \textit{hierarchical topographic factor analysis} (HTFA) to
 the fMRI datasets to obtain a time series of 700 node activities for
-every participant.  We then applied dimensionality reduction
+every participant.  We then apply dimensionality reduction
 (Incremental PCA) for each group.
 
-We then compared the groups’ activity patterns (using Pearson
+We then compare the groups’ activity patterns (using Pearson
 correlations) to estimate the story times each corresponding pattern
@@
 
-To assess decoding accuracy, we randomly divided participants for each
-stimulus into training and testing groups. We then compared the
+To assess decoding accuracy, we randomly divide participants for each
+stimulus into training and testing groups. We then compare the
 groups’ activity patterns (using Pearson correlations) to estimate the
 story times each corresponding pattern using more and more principle
-components (as the data became less compressed). Specifically, we asked, for each timepoint: what are the correlations
+components (as the data became less compressed). Specifically, we ask, for each timepoint: what are the correlations
 between the first group's and second group's activity patterns at each
@@
 
+% \subsection*{Evaluation metrics}
+%  We evaluated the degree of
+% compression of held-out
+% neuroimaging data with the time at which it was collected.  We used this latter evaluations (using timepoint
+% decoding) as a proxy for
+% gauging how much explanatory power the compressed data held
+% with respect to the observed data.
+
+% \subsubsection*{Timepoint decoding}
+
+% To explore how higher-order structure varies with stimulus structure
+% and complexity, we used a previous neuroimaging dataset
+% ~\cite{SimoEtal16}  in which participants listened to an audio recording of a story; 36 particpants listen to an intact version of the story, 17 participants listen to time-scrambled recordings of the same story where paragraphs were scrambled, 36 particpants listen to word-scrambled version and 36 participants lay in rest condition.
+
+% Prior work has shown participants share similar neural responses to
+% richly structured stimuli when compared to stimuli with less
+% structure.  To assess whether the moment-by-moment higher order
+% correlations were reliably preserved across participants, we used
+% inter-subject functional connectivity (ISFC) to isolate the
+% time-varying correlational structure (functional connectivity patterns
+% that were specifically driven by the story participants listened to.
+% Following the analyses conducted by (HTFA)~\cite{MannEtal18}, we first
+% applied \textit{hierarchical topographic factor analysis} (HTFA) to
+% the fMRI datasets to obtain a time series of 700 node activities for
+% every participant.  We then applied dimensionality reduction
+% (Incremental PCA) for each group.
+
+% We then compared the groups’ activity patterns (using Pearson
+% correlations) to estimate the story times each corresponding pattern
+% using more and more principle components.
+
+% To assess decoding accuracy, we randomly divided participants for each
+% stimulus into training and testing groups. We then compared the
+% groups’ activity patterns (using Pearson correlations) to estimate the
+% story times each corresponding pattern using more and more principle
+% components (as the data became less compressed). Specifically, we asked, for each timepoint: what are the correlations
+% between the first group's and second group's activity patterns at each
+% order. We note that the decoding test we used is a conservative in which we count a timepoint label as incorrect if it is not an exact match.
+
 \section*{Results}
-
-For our decoding analysis, we used HTFA-derived node activities
-~\cite{MannEtal18} from fMRI data collected as participants listened
-to an audio recording of a story (intact condition; 36 participants),
-listened to time scrambled recordings of the same story (17
-participants in the paragraph-scrambled condition listened to the
-paragraphs in a randomized order and 36 in the word-scrambled
-condition listened to the words in a randomized order), or lay resting
-with their eyes open in the scanner (rest condition; 36
-participants). We sought to understand the 'dimensionality' of the
-neural patterns by using more and more principle
-components to decode a multi-subject fMRI datasets. This story listening dataset was
-collected as part of a separate study, where the full imaging
-parameters, image preprocessing methods, and experimental details may
-be found~\citep{SimoEtal16}. The dataset is available at
-\href{url}{http://arks. princeton.edu/ark:/88435/dsp015d86p269k}.
+By training classifiers using more and more principle components to decode, and comparing across conditions with varying degrees of cognitive richness, we can assess the explanatory power of the compressed data held with respect to the observed data (see \textit{Methods}).
+We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.
 
@@
 Explanation of inflection metric.} First the we fit a sigmoid function to the decoding accuracy by number of components.
-Second, we found where the second derivate is both positive and less than .0001.
+Second, we found where the second derivative is both positive and less than .0001.
 Last, we then plot that inflection point as a single metric to capture the slope and asymptote of the curve.}
@@
 
-We performed a decoding analysis, using cross validation to estimate
-(using other participants’ data) which parts of the story the additional added
-principle component corresponded to (see \textit{Materials and methods}).
+Prior work has shown participants share similar neural responses to
+richly structured stimuli when compared to stimuli with less
+structure \cite{SimoEtal16}.  We replicate this finding, showing as complexity of the stimulus increases, decoding accuracy
+increases (Fig.~\ref{fig:decode_interpret},  a.).  
+Additionally, we found that as complexity of the stimuli increases, we need fewer components to decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
+However, we also found that as complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
+We posit that as the complexity of our thoughts increases, neural
+compression decreases. However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
+
+% We performed a decoding analysis, using cross validation to estimate
+% (using other participants’ data) which parts of the story the additional added
+% principle component corresponded to (see \textit{Materials and methods}).
 We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.
 
-Separately for each experimental condition, we divided participants
-into two groups. Starting with 1 principle comonent, for
-each dimension we added another principle component, and we correlated the group 1 activity patterns with group 2
-activity patterns.  There were 272
-timepoints for paragraph condition, 300 timepoints for intact and word
-conditions, and 400 timepoints for rest condition,  so chance
-performance on this decoding test is was $\frac{1}{272}$,
-$\frac{1}{300}$, and $\frac{1}{400}$ respectively.
+% Separately for each experimental condition, we divided participants
+% into two groups. Starting with 1 principle comonent, for
+% each dimension we added another principle component, and we correlated the group 1 activity patterns with group 2
+% activity patterns.  There were 272
+% timepoints for paragraph condition, 300 timepoints for intact and word
+% conditions, and 400 timepoints for rest condition,  so chance
+% performance on this decoding test is was $\frac{1}{272}$,
+% $\frac{1}{300}$, and $\frac{1}{400}$ respectively.
 
-- As complexity of the stimulus increases, decoding accuracy
-increases (Fig.~\ref{fig:decode_interpret},  a.).  Replication of
-Simony findings.
+% - As complexity of the stimulus increases, decoding accuracy
+% increases (Fig.~\ref{fig:decode_interpret},  a.).  Replication of
+% Simony findings.
 
-- As complexity of the stimuli increases, we need fewer components to
-decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
+% - As complexity of the stimuli increases, we need fewer components to
+% decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
 
-- As complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
+% - As complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
 
@@
     Fig.~\ref{fig:decode_interpret}, b. Analyses were limited by the
-    brain networks (using the Yeoetal11 network parcellation) and
+    brain networks (using the \cite{YeoEtal11} network parcellation) and
     arranged in increasing order relative to the intact condition.
@@
 and c.} For the total time in the intact condition, we are plotting the relative inflection points (\textbf{b.}) and corresponding number of components (\textbf{c.}) by network.
-\textbf{d.} The network parcellation defined by Yeo et al.
-(2011) is displayed on the inflated brain maps.
+\textbf{d.} The network parcellation defined by \cite{YeoEtal11} is displayed on the inflated brain maps.
 The colors and network labels serve as a legend for \textbf{a.} and \textbf{d.}}
@@
 
-As complexity of the stimuli increases, decoding accuracy increases
-with higher cognitive areas. (Fig.~\ref{fig: decode_pcs_network}).
+We also wondered how this compression would change across brain regions.
+We repeated the analysis but limited the brain hubs to 7 networks using the Yeo et al.
+(2011) network parcellation shown here in the inflated brain (Fig.~\ref{fig:decode_pcs_network}, d.).
+We found that as complexity of the stimuli increases, decoding accuracy increases
+with higher cognitive areas. (Fig.~\ref{fig:decode_pcs_network}).
 
- If, there is some understanding of the narrative that accumulates over time, we should be able to see that change.
+% As complexity of the stimuli increases, decoding accuracy increases
+% with higher cognitive areas. (Fig.~\ref{fig: decode_pcs_network}).
+
+%  If, there is some understanding of the narrative that accumulates over time, we should be able to see that change.
 
@@
 }
-  \label{fig: decode_pcs_thirds}
+  \label{fig:decode_pcs_thirds}
 \end{figure}
 
-- Increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
-- Decreases in decoding accuracy for the word-scrambled and rest
+We were also curious how compression would change across time.
+If, there is some understanding of the narrative that accumulates over time, we should be able to see that difference.
+We found increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
+We also found decreases in decoding accuracy for the word-scrambled and rest
 condition.
+
+Overall, we found that as story listening conditions become more complex, more components are required to decode. We also found we could decode better with more impoverished data when
+there is the underlying structure of the narrative providing more cognitive richness.
+We posit that as the complexity of our thoughts increases, neural compression decreases.
+However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
+
+% - Increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
+% - Decreases in decoding accuracy for the word-scrambled and rest
+% condition.
 
@@
 \section*{Author contributions}
-Concept: J.R.M.  Implementation: T.H.C., L.L.W.O., and
+Concept: J.R.M. and L.L.W.O. Implementation: L.L.W.O., and
 J.R.M.  Analyses: L.L.W.O and J.R.M.
@@
 \bibliographystyle{apacite}
-\bibliography{CDL-bibliography/memlab}
+\bibliography{cdl.bib}
 

### 2023-02-20 paper/main.tex
checking  in some thoughts, minor file reorganization

@@
 
-\title{Understanding brain complexity in naturalistic processing
-in humans} \author{Lucy
-  L. W. Owen$^1$, and\
-  Jeremy R. Manning\textsuperscript{$1, \dagger$}\\
-  [0.1in]$^1$Department of Psychological and Brain
-  Sciences,\\Dartmouth
-  College, Hanover, NH\\
-  \textsuperscript{$\dagger$}Address correspondence to
+\title{High-order cognition is supported by complex but compressible brain activity patterns}
+
+\author{Lucy L. W. Owen$^1, 2$, and Jeremy R. Manning\textsuperscript{$1,
+*$}\\$^1$Department of Psychological and Brain Sciences,\\Dartmouth College,
+Hanover, NH\\[0.1cm]$^2$Carney Institute for Brain Sciences,\\Brown University,
+Providence, RI\\[0.1cm] \textsuperscript{*}Address correspondence to
 jeremy.r.manning@dartmouth.edu}
@@
 \begin{abstract}
-We applied dimensionality reduction algorithms to the activity patterns in each experimental condition.
-Specifically, we sought to understand the “dimensionality” of the neural patterns that were sufficient to decode participants’ listening times (or approach was similar to that of Mack et al.
-2017).
-We found that even low-dimensional embeddings of the data were sufficient to accurately decode listening times from the intact story recording, whereas finer temporal scramblings of the story required higher-dimensional embeddings of the data to reach peak decoding accuracy.
+
+We applied dimensionality reduction algorithms to the activity
+patterns in each experimental condition. Specifically, we sought to understand
+the “dimensionality” of the neural patterns that were sufficient to decode
+participants’ listening times (or approach was similar to that of Mack et al.
+2017). We found that even low-dimensional embeddings of the data were
+sufficient to accurately decode listening times from the intact story
+recording, whereas finer temporal scramblings of the story required
+higher-dimensional embeddings of the data to reach peak decoding accuracy.
+
 \end{abstract}
@@
 \section*{Introduction}
+
+Large-scale networks, including the human brain, may be conceptualized as
+occupying one or more positions along on a continuum. At one extreme, every
+node is fully independent of every other node. At the other extreme, all nodes
+are fully coupled and behave identically. Each extreme optimizes key properties
+of how the network functions. When every node is independent, the network is
+maximally \textit{expressive}: if we define the network's ``state'' as the
+total set of activity patterns across nodes, then every state is equally
+reachable by a network with fully independent nodes. On the other hand, a fully
+coupled network optimizes \textit{robustness}: any subset of nodes, other than
+the entire network, may be removed from the network without any loss of
+function or expressive power. Note that a given set of nodes might reconfigure
+its connections or behaviors under different circumstances to change its
+position along this continuum according to the needs at hand. Presumably, most
+systems tend to occupy positions between the above extremes. We wondered: might
+the human brain reconfigure itself to be more flexible or more robust according
+to ongoing demands?
 

### 2023-02-20 paper/main.tex
draft of abstract

@@
 
-We applied dimensionality reduction algorithms to the activity
-patterns in each experimental condition. Specifically, we sought to understand
-the “dimensionality” of the neural patterns that were sufficient to decode
-participants’ listening times (or approach was similar to that of Mack et al.
-2017). We found that even low-dimensional embeddings of the data were
-sufficient to accurately decode listening times from the intact story
-recording, whereas finer temporal scramblings of the story required
-higher-dimensional embeddings of the data to reach peak decoding accuracy.
+We applied dimensionality reduction algorithms and pattern classifiers to
+functional neuroimaging data collected as participants listened to a story,
+temporally scrambled versions of the story, or underwent a resting state
+scanning session. These experimental conditions were intended to require
+different depths of processing and insprire different levels of engagement. We
+considered two primary aspects of the data. First, we treated the number of
+features (components) required to achieve a threshold decoding accuracy as a
+proxy for the ``compressibility'' of the neural patterns (where fewer
+components indicate higher compressibility). Second, we treated the maximum
+achievable decoding accuracy across participants as an indicator of the
+``stability'' of the recorded patterns. Overall, we found that neural patterns
+recorded as participants listened to the intact story required fewer features
+to achieve comparable classification accuracy to the other experimental
+conditions. However, the peak decoding accuracy (achievable with more features)
+was also highest during intact story listening. Taken together, our work
+suggests that our brain networks flexibly reconfigure according to ongoing task
+demands, and that the activity patterns associated with higher-order cognition
+and high engagement are both more complex and more compressible than the
+activity patterns associated with lower-order tasks and lower levels of
+engagement.
 
@@
 
-We're interested in the complexity of brain patterns that underly
-different types of thoughts. To explore this question space, we will take
-brain patterns recorded under different experimental conditions used in Aim 2, and
-project them into lower dimensional spaces using principle components
-analysis. We can then ask how well those low-dimensional embeddings of
-the data retain cognitively relevant information like when in a story
-someone is listening to.
+We're interested in the complexity of brain patterns that underly different
+types of thoughts. To explore this question space, we will take brain patterns
+recorded under different experimental conditions used in Aim 2, and project
+them into lower dimensional spaces using principle components analysis. We can
+then ask how well those low-dimensional embeddings of the data retain
+cognitively relevant information like when in a story someone is listening to.
 
-This work has been inspired, in part, by~\cite{MackEtal20}. In this paper, 
-they investigated the role of the prefrontal cortex in filtering out 
-irrelevant content.  Specifically, they looked at if the vmPFC performs 
-data reduction on incoming information through compression.  
-This was motivated, in part, by orbital frontal cortex (OFC) 
-compression in rats~\citep{ZhouEtal19}.
-They studied this using a learning paradigm in which participants had to classify insects based on different numbers of feature dimensions.
-The idea was that participants in some learning blocks, participants could identify the insects based on one feature (low complexity) or several features (high complexity), but importantly the stimuli remained the same across all learning problems.
-They found that complexity and compression had an inverse relationship; the lower complexity of a conceptual space, the higher the degree of compression.
-Building on this idea, we wonder if varying degrees of compression is performed throughout the brain.
-We also want to test this idea, but using varying levels of engagement listening to a naturalistic stimuli.
+This work has been inspired, in part, by~\cite{MackEtal20}. In this paper, they
+investigated the role of the prefrontal cortex in filtering out irrelevant
+content. Specifically, they looked at if the vmPFC performs data reduction on
+incoming information through compression. This was motivated, in part, by
+orbital frontal cortex (OFC) compression in rats~\citep{ZhouEtal19}. They
+studied this using a learning paradigm in which participants had to classify
+insects based on different numbers of feature dimensions. The idea was that
+participants in some learning blocks, participants could identify the insects
+based on one feature (low complexity) or several features (high complexity),
+but importantly the stimuli remained the same across all learning problems.
+They found that complexity and compression had an inverse relationship; the
+lower complexity of a conceptual space, the higher the degree of compression.
+Building on this idea, we wonder if varying degrees of compression is performed
+throughout the brain. We also want to test this idea, but using varying levels
+of engagement listening to a naturalistic stimuli.
 
-To understand the degree of compression throughout the brain during cognition, we will use the same fMRI data from Aim 2, collected while participants listened to a story in different scrambling conditions.
-We will measure the degree that multivoxel activation patterns are compressed during story listening using principle components analysis (PCA) a method for low-rank approximation of multidimensional data~\citep{EckaYoun36}.
-We will explore this using decoding accuracy as a function of the number of components, or dimensions, in the low-dimensional space under different cognitive conditions.
+To understand the degree of compression throughout the brain during cognition,
+we will use the same fMRI data from Aim 2, collected while participants
+listened to a story in different scrambling conditions. We will measure the
+degree that multivoxel activation patterns are compressed during story
+listening using principle components analysis (PCA) a method for low-rank
+approximation of multidimensional data~\citep{EckaYoun36}. We will explore this
+using decoding accuracy as a function of the number of components, or
+dimensions, in the low-dimensional space under different cognitive conditions.
 
-You can imagine two reasonable predictions of
-how cognition is reflected in brain patterns.  The first is as our
-thoughts become more complex, they are supported by more complex brain
-patterns, and require more components to decode.  The second is that
-when thoughts are deeper and more complicated, the units of neural
-activity would carry more information, and would require therefore
-fewer components to decode.
+You can imagine two reasonable predictions of how cognition is reflected in
+brain patterns. The first is as our thoughts become more complex, they are
+supported by more complex brain patterns, and require more components to
+decode. The second is that when thoughts are deeper and more complicated, the
+units of neural activity would carry more information, and would require
+therefore fewer components to decode.
 
-This idea can be explored in this visual analogy (Fig.~\ref{fig: pie_example}) for neural compression.
-Here there are two images of pies, the top pie is more complex than the bottom.
-On the left we're illustrating that it takes
-fewer components to reach the same 95 percent variance explained in
-the less complex pie, which corresponds to higher compression.
-However, on the right with very few components similar
-variance is explaining both pies.
+This idea can be explored in this visual analogy (Fig.~\ref{fig: pie_example})
+for neural compression. Here there are two images of pies, the top pie is more
+complex than the bottom. On the left we're illustrating that it takes fewer
+components to reach the same 95 percent variance explained in the less complex
+pie, which corresponds to higher compression. However, on the right with very
+few components similar variance is explaining both pies.
 

### 2023-02-20 paper/main.tex
starting on intro, additional file reorganization
typo fixes

@@
 
-\author{Lucy L. W. Owen$^1, 2$, and Jeremy R. Manning\textsuperscript{$1,
-*$}\\$^1$Department of Psychological and Brain Sciences,\\Dartmouth College,
-Hanover, NH\\[0.1cm]$^2$Carney Institute for Brain Sciences,\\Brown University,
+\author{Lucy L. W. Owen\textsuperscript{1, 2}, and Jeremy R. Manning\textsuperscript{1,
+*}\\\textsuperscript{1}Department of Psychological and Brain Sciences,\\Dartmouth College,
+Hanover, NH\\[0.1cm]\textsuperscript{2}Carney Institute for Brain Sciences,\\Brown University,
 Providence, RI\\[0.1cm] \textsuperscript{*}Address correspondence to
@@
 \bibliographystyle{apacite}
-\bibliography{cdl.bib}
+\bibliography{CDL-bibliography/cdl}
 

### 2023-02-20 paper/main.tex
added some notes
working on introduction (backing up progress)
added a paragraph about information to the intro

@@
 position along this continuum according to the needs at hand. Presumably, most
-systems tend to occupy positions between the above extremes. We wondered: might
+systems tend to occupy positions between these extremes. We wondered: might
 the human brain reconfigure itself to be more flexible or more robust according
 to ongoing demands?
+
+% compression (redundancy) vs. complexity -- appear to balance each other, butnot necessarily.
+
+Closely related to the above notions of expressiveness versus robustness are
+measures of how much \textit{information}~\citep{Shan48} is contained in a
+given signal or pattern. Formally, information is defined as the amount of
+uncertainty (i.e., entropy) in a given variables' outcomes, measured in
+\textit{bits}, or the optimal number of yes/no questions needed to reduce
+uncertainty about the variable's outcomes to zero. The \textit{redundancy} of a
+signal denotes the difference how expressive the signal could be (i.e.,
+proportional to the number of unique states or symbols used to transmit the
+signal) and the actual information rate (i.e., the entropy of each individual
+state or symbol). If a brain network's nodes are fully independent, then the
+number of bits required to express a single activity pattern is proportional to
+the number of nodes. The network would also be minimally redundant, since the
+status of every node would be needed to fully express a single brain activity
+pattern. If a brain network's nodes are fully coupled and identical, then the
+number of bits required to express a single activity pattern is proportional to
+the number of unique states or values any individual node can take on. Such a
+network would be highly redundant, since knowing any individual node's state
+would be sufficient to recover the full-brain activity pattern.
+
+%   - e.g., you could have something that is highly compressible (just a few components are sufficient to achieve some threshold information level) but
+%     *also* very rich-- e.g., the maximum number of informative components is large
+% compression and task demands -- MackEtal20, ZhouEtal19
+% superEEG -- prediction of held-out data; implies redundancy
+% timecorr -- higher order network patterns during higher order cognition implies more complexity
+% cognitive richness and scrambling-- various Hasson lab papers
 

### 2023-02-21 paper/main.tex
i seem to be making backwards progress...but i'm reading some related papers to help frame things and flesh out the intro

@@
 node is fully independent of every other node. At the other extreme, all nodes
-are fully coupled and behave identically. Each extreme optimizes key properties
-of how the network functions. When every node is independent, the network is
-maximally \textit{expressive}: if we define the network's ``state'' as the
-total set of activity patterns across nodes, then every state is equally
-reachable by a network with fully independent nodes. On the other hand, a fully
-coupled network optimizes \textit{robustness}: any subset of nodes, other than
-the entire network, may be removed from the network without any loss of
-function or expressive power. Note that a given set of nodes might reconfigure
-its connections or behaviors under different circumstances to change its
-position along this continuum according to the needs at hand. Presumably, most
-systems tend to occupy positions between these extremes. We wondered: might
-the human brain reconfigure itself to be more flexible or more robust according
-to ongoing demands?
+behave identically. Each extreme optimizes key properties of how the network
+functions. When every node is independent, the network is maximally
+\textit{expressive}: if we define the network's ``state'' as the activity
+pattern across its nodes, then every state is equally reachable by a network
+with fully independent nodes. On the other hand, a network of identically
+behaved nodes optimizes \textit{robustness}: any subset of nodes may be removed
+from the network without any loss of function or expressive power, as long as
+any single node remains. Presumably, most natural systems tend to occupy
+positions between these extremes. We wondered: might the human brain
+reconfigure itself to be more flexible or more robust according to ongoing
+demands? In other words, might the brain reconfigure its connections or
+behaviors under different circumstances to change its position along this
+continuum?
+
+We take as a given that brain activity is highly flexible: our brains can
+exhibit nearly infinite activity patterns. However, those activity patterns are
+also highly structured. For example, full-brain correlation matrices are stable
+within~\citep{FinnEtal15} and across~\citep{YeoEtal11} individuals. This
+stability implies that our brains are not
+
+For example, despite stable individual differences in
+full-brain correlations~\citep{FinnEtal15},
+
+A number of prior studies suggest that
 
@@
 Closely related to the above notions of expressiveness versus robustness are
-measures of how much \textit{information}~\citep{Shan48} is contained in a
-given signal or pattern. Formally, information is defined as the amount of
-uncertainty (i.e., entropy) in a given variables' outcomes, measured in
-\textit{bits}, or the optimal number of yes/no questions needed to reduce
-uncertainty about the variable's outcomes to zero. The \textit{redundancy} of a
-signal denotes the difference how expressive the signal could be (i.e.,
-proportional to the number of unique states or symbols used to transmit the
-signal) and the actual information rate (i.e., the entropy of each individual
-state or symbol). If a brain network's nodes are fully independent, then the
-number of bits required to express a single activity pattern is proportional to
-the number of nodes. The network would also be minimally redundant, since the
-status of every node would be needed to fully express a single brain activity
-pattern. If a brain network's nodes are fully coupled and identical, then the
-number of bits required to express a single activity pattern is proportional to
-the number of unique states or values any individual node can take on. Such a
-network would be highly redundant, since knowing any individual node's state
-would be sufficient to recover the full-brain activity pattern.
+measures of how much \textit{information} is contained in a given signal or
+pattern, and how \textit{redundant} a signal is. Formally, information is
+defined as the amount of uncertainty about a given variables' outcomes (i.e.,
+entropy), measured in \textit{bits}, or the optimal number of yes/no questions
+needed to reduce uncertainty about the variable's outcomes to zero. Highly
+complex systems with many degrees of freedom (i.e., high flexibility and
+expressiveness), are more information-rich than simpler or more constrained
+systems. The redundancy of a signal denotes the difference how expressive the
+signal \textit{could} be (i.e., proportional to the number of unique states or
+symbols used to transmit the signal) and the actual information rate (i.e., the
+entropy of each individual state or symbol). If a brain network's nodes are
+fully independent, then the number of bits required to express a single
+activity pattern is proportional to the number of nodes. The network would also
+be minimally redundant, since the status of every node would be needed to fully
+express a single brain activity pattern. If a brain network's nodes are fully
+coupled and identical, then the number of bits required to express a single
+activity pattern is proportional to the number of unique states or values any
+individual node can take on. Such a network would be highly redundant, since
+knowing any individual node's state would be sufficient to recover the
+full-brain activity pattern. Highly redundant systems are also robust, since
+there is little information loss from losing any given observation.
+
+% Some papers to add in:
+
+% https://www.nature.com/articles/s41467-019-13930-8
+% https://onlinelibrary.wiley.com/doi/full/10.1002/jnr.24316?casa_token=PjnffsBh3vgAAAAA%3AynPc_easF6EXPrEK7ihnQ6Jc8QeWq6x0u-ytDsMwxQ4a0jnIvXhg8xBQDDVnAoMbYhUWiwUtlj4vTA
+% https://www.sciencedirect.com/science/article/pii/S1053811918305238?casa_token=xSH5EIN1rZYAAAAA:R2Zf9Lu_np3v0t5-lNZNal0JaJLk1JyYe4QxtmHnba9_lMGJUJgKNK4lHoybuITemQqYUIcuNA
+% https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3174820/
+% https://www.sciencedirect.com/science/article/pii/S0896627314004000
+% https://www.pnas.org/doi/abs/10.1073/pnas.0905267106
+% https://www.sciencedirect.com/science/article/pii/S0896627318302411
+% https://www.liebertpub.com/doi/full/10.1089/brain.2011.0068
+% https://www.sciencedirect.com/science/article/pii/S1053811917302872#f0015
+% https://www.sciencedirect.com/science/article/abs/pii/S1053811919303027
+% https://pubmed.ncbi.nlm.nih.gov/34625348/
+
+If brain activity patterns contain rich task-relevant information, we should be
+able to use the activity patterns to accurately differentiate between different
+aspects of the task~\citep[e.g., using pattern classifiers;][]{NormEtal06b}.
+For example, prior work has shown a direct correspondence between
+classification accuracy and the information content of a signal~\citep{Alva02}.
+If brain activity patterns are
 

### 2023-02-22 code/scripts/config.py
updates on the cluster from long ago

@@ -5,7 +5,7 @@ config = dict()
 
 config['template'] = 'run_job.sh'
 
-local = True
+local = False
 
 # ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
 
@@ -59,7 +59,7 @@ config['jobname'] = "pca_analysis"  # default job name
 config['q'] = "default"  # options: default, testing, largeq
 config['nnodes'] = 1  # how many nodes to use for this one job
 config['ppn'] = 4  # how many processors to use for this one job (assume 4GB of RAM per processor)
-config['walltime'] = '3:00:00'  # maximum runtime, in h:MM:SS
+config['walltime'] = '24:00:00'  # maximum runtime, in h:MM:SS
 #config['startdir'] = '/ihome/lowen/repos/supereeg/examples'  # directory to start the job in
 config['cmd_wrapper'] = "python3"  # replace with actual command wrapper (e.g. matlab, python, etc.)
 config['modules'] = "(\"python/3.6\")"  # separate each module with a space and enclose in (escaped) double quotes

### 2023-02-22 code/scripts/pieman_pca_elimination_analysis_submit.py

@@ -37,13 +37,13 @@ dims = ['700']
 # options for debug: True or False
 debug = str('False')
 
-job_commands = list(np.array([list(map(lambda x: x[0]+" "+str(x[1])+" "+'unchunked'+" "+str(r)+
+job_commands = list(np.array([list(map(lambda x: x[0]+" "+str(x[1])+" "+'elim_unchunked'+" "+str(r)+
                                                  " "+rfuns[0]+" "+ dims[0]+ " "+debug,
                                        zip([job_script]*len(cond_type), cond_type)))
                               for r in range(int(reps))]).flat)
 
 # job_names should specify the file name of each script (as a list, of the same length as job_commands)
-job_names = list(np.array([list(map(lambda x: os.path.basename(os.path.splitext(x)[0])+'_'+'unchunked'+'_'+str(r)+'_'
+job_names = list(np.array([list(map(lambda x: os.path.basename(os.path.splitext(x)[0])+'_'+'elim_unchunked'+'_'+str(r)+'_'
                                               '_'+rfuns[0]+'_'+ dims[0] + '_'+debug+'.sh', cond_type))
                            for r in range(int(reps))]).flat)
 

### 2023-02-22 paper/main.tex
working on intro, pie figure, added information and compressibility figure

@@
 
-We take as a given that brain activity is highly flexible: our brains can
-exhibit nearly infinite activity patterns. However, those activity patterns are
-also highly structured. For example, full-brain correlation matrices are stable
-within~\citep{FinnEtal15} and across~\citep{YeoEtal11} individuals. This
-stability implies that our brains are not
-
-For example, despite stable individual differences in
-full-brain correlations~\citep{FinnEtal15},
-
-A number of prior studies suggest that
-
 % compression (redundancy) vs. complexity -- appear to balance each other, butnot necessarily.
@@
 measures of how much \textit{information} is contained in a given signal or
-pattern, and how \textit{redundant} a signal is. Formally, information is
-defined as the amount of uncertainty about a given variables' outcomes (i.e.,
-entropy), measured in \textit{bits}, or the optimal number of yes/no questions
-needed to reduce uncertainty about the variable's outcomes to zero. Highly
-complex systems with many degrees of freedom (i.e., high flexibility and
-expressiveness), are more information-rich than simpler or more constrained
-systems. The redundancy of a signal denotes the difference how expressive the
-signal \textit{could} be (i.e., proportional to the number of unique states or
-symbols used to transmit the signal) and the actual information rate (i.e., the
-entropy of each individual state or symbol). If a brain network's nodes are
-fully independent, then the number of bits required to express a single
-activity pattern is proportional to the number of nodes. The network would also
-be minimally redundant, since the status of every node would be needed to fully
-express a single brain activity pattern. If a brain network's nodes are fully
-coupled and identical, then the number of bits required to express a single
-activity pattern is proportional to the number of unique states or values any
-individual node can take on. Such a network would be highly redundant, since
-knowing any individual node's state would be sufficient to recover the
-full-brain activity pattern. Highly redundant systems are also robust, since
-there is little information loss from losing any given observation.
+pattern, and how \textit{redundant} a signal is~\citep{Shan48}. Formally,
+information is defined as the amount of uncertainty about a given variables'
+outcomes (i.e., entropy), measured in \textit{bits}, or the optimal number of
+yes/no questions needed to reduce uncertainty about the variable's outcomes to
+zero. Highly complex systems with many degrees of freedom (i.e., high
+flexibility and expressiveness), are more information-rich than simpler or more
+constrained systems. The redundancy of a signal denotes the difference how
+expressive the signal \textit{could} be (i.e., proportional to the number of
+unique states or symbols used to transmit the signal) and the actual
+information rate (i.e., the entropy of each individual state or symbol). If a
+brain network's nodes are fully independent, then the number of bits required
+to express a single activity pattern is proportional to the number of nodes.
+The network would also be minimally redundant, since the status of every node
+would be needed to fully express a single brain activity pattern. If a brain
+network's nodes are fully coupled and identical, then the number of bits
+required to express a single activity pattern is proportional to the number of
+unique states or values any individual node can take on. Such a network would
+be highly redundant, since knowing any individual node's state would be
+sufficient to recover the full-brain activity pattern. Highly redundant systems
+are also robust, since there is little information loss from losing any given
+observation.
 
-% Some papers to add in:
+We take as a given that brain activity is highly flexible: our brains can
+exhibit nearly infinite activity patterns. This flexibility implies that our
+brains activity patterns are highly information rich. However, brain activity
+patterns are also highly structured. For example, full-brain correlation
+matrices are stable within~\citep{FinnEtal15, FinnEtal17, GratEtal18} and
+across~\citep{YeoEtal11, GlerEtal12, GratEtal18, ColeEtal14} individuals. This
+stability suggests that our brains' activity patterns are at least partially
+constrained, for example by anatomical, external, or internal factors.
+Constraints on brain activity that limit its flexibility decrease
+expressiveness (i.e., its information rate). However, constraints on brain
+activity also increase its robustness to noise (e.g., ``missing'' or corrupted
+signals may be partially recovered). For example, recent work has shown that
+full-brain activity patterns may be reliably recovered from only a relatively
+small number of implanted electrodes~\citep{OwenEtal20, ScanEtal21}. This
+robustness property suggests that the relevant signal (e.g., underlying factors
+that have some influence over brain activity patterns) are compressible.
 
-% https://www.nature.com/articles/s41467-019-13930-8
-% https://onlinelibrary.wiley.com/doi/full/10.1002/jnr.24316?casa_token=PjnffsBh3vgAAAAA%3AynPc_easF6EXPrEK7ihnQ6Jc8QeWq6x0u-ytDsMwxQ4a0jnIvXhg8xBQDDVnAoMbYhUWiwUtlj4vTA
-% https://www.sciencedirect.com/science/article/pii/S1053811918305238?casa_token=xSH5EIN1rZYAAAAA:R2Zf9Lu_np3v0t5-lNZNal0JaJLk1JyYe4QxtmHnba9_lMGJUJgKNK4lHoybuITemQqYUIcuNA
-% https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3174820/
-% https://www.sciencedirect.com/science/article/pii/S0896627314004000
-% https://www.pnas.org/doi/abs/10.1073/pnas.0905267106
-% https://www.sciencedirect.com/science/article/pii/S0896627318302411
-% https://www.liebertpub.com/doi/full/10.1089/brain.2011.0068
-% https://www.sciencedirect.com/science/article/pii/S1053811917302872#f0015
-% https://www.sciencedirect.com/science/article/abs/pii/S1053811919303027
-% https://pubmed.ncbi.nlm.nih.gov/34625348/
+To the extent that brain activity patterns contain rich task-relevant
+information, we should be able to use the activity patterns to accurately
+differentiate between different aspects of the task~\citep[e.g., using pattern
+classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
+correspondence between classification accuracy and the information content of a
+signal~\citep{Alva02}. To the extent that brain activity patterns are
+compressible, we should be able to generate simplified (e.g., lower
+dimensional) representations of the data while still preserving the relevant or
+important aspects of the original signal. In general, information content and
+compressibility are related but are partially dissociable
+(Fig.~\ref{fig:information--compression}).
 
-If brain activity patterns contain rich task-relevant information, we should be
-able to use the activity patterns to accurately differentiate between different
-aspects of the task~\citep[e.g., using pattern classifiers;][]{NormEtal06b}.
-For example, prior work has shown a direct correspondence between
-classification accuracy and the information content of a signal~\citep{Alva02}.
-If brain activity patterns are
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{figs/pie_example.pdf}
+  \caption{\textbf{Illustration of compression.} Visual analogy for
+    neural compression. Here are 2 images of pies, one more complex
+    than the other. \textbf{a.} It takes fewer components to reach the same percent
+    variance explained in the less complex pie, which corresponds to
+    higher compression. \textbf{b.} However, with very few components, similar variance is explained in both pies.
+\textbf{c.} Plots the cumulative explained variance for more and more components.}
+  \label{fig:information-compression}
+\end{figure}
+
+Specifically, if a given signal
+(e.g., a representation of brain activity patterns) contains more information
+about ongoing cognitive processes, then the peak decoding accuracy should be
+high. And if the signal is compressible, a low-dimensional embedding of the
+signal will be similarly informative to the original signal. A signal can be
+informative but not compressible, informative but
 

### 2023-02-23 paper/main.tex

@@
 compressibility are related but are partially dissociable
-(Fig.~\ref{fig:information--compression}).
+(Fig.~\ref{fig:information-compression}).
 
-\begin{figure}
+\begin{figure}[tp]
+
 \centering
-  \includegraphics[width=\textwidth]{figs/pie_example.pdf}
-  \caption{\textbf{Illustration of compression.} Visual analogy for
-    neural compression. Here are 2 images of pies, one more complex
-    than the other. \textbf{a.} It takes fewer components to reach the same percent
-    variance explained in the less complex pie, which corresponds to
-    higher compression. \textbf{b.} However, with very few components, similar variance is explained in both pies.
-\textbf{c.} Plots the cumulative explained variance for more and more components.}
+\includegraphics[width=\textwidth]{figs/information_and_compressibility}
+
+\caption{\textbf{Information content and compressibility.} \textbf{A. Variance
+explained for two images.} We applied principal components analysis to a
+photograph and drawing, treating each row of the images as ``observations.''
+Across columns, we identified the number of components required to explain
+100\%, 95\%, 75\%, or 50\% of the cumulative variance in each image (the 100\%
+columns denote the original images). The numbers of components are indicate in
+parentheses, and the resulting ``compressed'' images are displayed. \textbf{B.
+Representing two images with different numbers of components.} Using the same
+principal component decompositions as in Panel A, we computed the cumulative
+proportion of variance explained with 380 (original images), 20, 4, or 2
+components. \textbf{C. Cumulative variance explained versus number of
+components.} For the images displayed in Panels A and B, we plot the cumulative
+proportion of variance explained as a function of the number of components used
+to represent each image. \textbf{D. Information rate and compressibility.}
+Across multiple images, the information rate (i.e., the amount of information
+contained in each image; horizontal axis) is high when each individual pixel
+provides information that cannot be inferred from other pixels.
+High-information rate images tend to be high-resolution, and low-information
+rate images tend to be low-resolution. Compressibility is related to the
+difference between the information required to specify the original versus
+compressed images (vertical axis). Highly compressible images often contain
+predictable structure (redundancies) that can be leveraged to represent the
+images much more efficiently than in their original feature spaces.}
+
 \label{fig:information-compression} 
@@
 
-Specifically, if a given signal
-(e.g., a representation of brain activity patterns) contains more information
-about ongoing cognitive processes, then the peak decoding accuracy should be
-high. And if the signal is compressible, a low-dimensional embedding of the
-signal will be similarly informative to the original signal. A signal can be
-informative but not compressible, informative but
+% Specifically, if a given signal
+% (e.g., a representation of brain activity patterns) contains more information
+% about ongoing cognitive processes, then the peak decoding accuracy should be
+% high. And if the signal is compressible, a low-dimensional embedding of the
+% signal will be similarly informative to the original signal. A signal can be
+% informative but not compressible, informative but
 
-%   - e.g., you could have something that is highly compressible (just a few components are sufficient to achieve some threshold information level) but
-%     *also* very rich-- e.g., the maximum number of informative components is large
-% compression and task demands -- MackEtal20, ZhouEtal19
-% superEEG -- prediction of held-out data; implies redundancy
-% timecorr -- higher order network patterns during higher order cognition implies more complexity
-% cognitive richness and scrambling-- various Hasson lab papers
+% %   - e.g., you could have something that is highly compressible (just a few components are sufficient to achieve some threshold information level) but
+% %     *also* very rich-- e.g., the maximum number of informative components is large
+% % compression and task demands -- MackEtal20, ZhouEtal19
+% % superEEG -- prediction of held-out data; implies redundancy
+% % timecorr -- higher order network patterns during higher order cognition implies more complexity
+% % cognitive richness and scrambling-- various Hasson lab papers
 
-We're interested in the complexity of brain patterns that underly different
-types of thoughts. To explore this question space, we will take brain patterns
-recorded under different experimental conditions used in Aim 2, and project
-them into lower dimensional spaces using principle components analysis. We can
-then ask how well those low-dimensional embeddings of the data retain
-cognitively relevant information like when in a story someone is listening to.
+% We're interested in the complexity of brain patterns that underly different
+% types of thoughts. To explore this question space, we will take brain patterns
+% recorded under different experimental conditions used in Aim 2, and project
+% them into lower dimensional spaces using principle components analysis. We can
+% then ask how well those low-dimensional embeddings of the data retain
+% cognitively relevant information like when in a story someone is listening to.
 
-This work has been inspired, in part, by~\cite{MackEtal20}. In this paper, they
-investigated the role of the prefrontal cortex in filtering out irrelevant
-content. Specifically, they looked at if the vmPFC performs data reduction on
-incoming information through compression. This was motivated, in part, by
-orbital frontal cortex (OFC) compression in rats~\citep{ZhouEtal19}. They
-studied this using a learning paradigm in which participants had to classify
-insects based on different numbers of feature dimensions. The idea was that
-participants in some learning blocks, participants could identify the insects
-based on one feature (low complexity) or several features (high complexity),
-but importantly the stimuli remained the same across all learning problems.
-They found that complexity and compression had an inverse relationship; the
-lower complexity of a conceptual space, the higher the degree of compression.
-Building on this idea, we wonder if varying degrees of compression is performed
-throughout the brain. We also want to test this idea, but using varying levels
-of engagement listening to a naturalistic stimuli.
+% This work has been inspired, in part, by~\cite{MackEtal20}. In this paper, they
+% investigated the role of the prefrontal cortex in filtering out irrelevant
+% content. Specifically, they looked at if the vmPFC performs data reduction on
+% incoming information through compression. This was motivated, in part, by
+% orbital frontal cortex (OFC) compression in rats~\citep{ZhouEtal19}. They
+% studied this using a learning paradigm in which participants had to classify
+% insects based on different numbers of feature dimensions. The idea was that
+% participants in some learning blocks, participants could identify the insects
+% based on one feature (low complexity) or several features (high complexity),
+% but importantly the stimuli remained the same across all learning problems.
+% They found that complexity and compression had an inverse relationship; the
+% lower complexity of a conceptual space, the higher the degree of compression.
+% Building on this idea, we wonder if varying degrees of compression is performed
+% throughout the brain. We also want to test this idea, but using varying levels
+% of engagement listening to a naturalistic stimuli.
 
-To understand the degree of compression throughout the brain during cognition,
-we will use the same fMRI data from Aim 2, collected while participants
-listened to a story in different scrambling conditions. We will measure the
-degree that multivoxel activation patterns are compressed during story
-listening using principle components analysis (PCA) a method for low-rank
-approximation of multidimensional data~\citep{EckaYoun36}. We will explore this
-using decoding accuracy as a function of the number of components, or
-dimensions, in the low-dimensional space under different cognitive conditions.
+% To understand the degree of compression throughout the brain during cognition,
+% we will use the same fMRI data from Aim 2, collected while participants
+% listened to a story in different scrambling conditions. We will measure the
+% degree that multivoxel activation patterns are compressed during story
+% listening using principle components analysis (PCA) a method for low-rank
+% approximation of multidimensional data~\citep{EckaYoun36}. We will explore this
+% using decoding accuracy as a function of the number of components, or
+% dimensions, in the low-dimensional space under different cognitive conditions.
 
-You can imagine two reasonable predictions of how cognition is reflected in
-brain patterns. The first is as our thoughts become more complex, they are
-supported by more complex brain patterns, and require more components to
-decode. The second is that when thoughts are deeper and more complicated, the
-units of neural activity would carry more information, and would require
-therefore fewer components to decode.
+% You can imagine two reasonable predictions of how cognition is reflected in
+% brain patterns. The first is as our thoughts become more complex, they are
+% supported by more complex brain patterns, and require more components to
+% decode. The second is that when thoughts are deeper and more complicated, the
+% units of neural activity would carry more information, and would require
+% therefore fewer components to decode.
 
-This idea can be explored in this visual analogy (Fig.~\ref{fig: pie_example})
-for neural compression. Here there are two images of pies, the top pie is more
-complex than the bottom. On the left we're illustrating that it takes fewer
-components to reach the same 95 percent variance explained in the less complex
-pie, which corresponds to higher compression. However, on the right with very
-few components similar variance is explaining both pies.
+% This idea can be explored in this visual analogy (Fig.~\ref{fig: pie_example})
+% for neural compression. Here there are two images of pies, the top pie is more
+% complex than the bottom. On the left we're illustrating that it takes fewer
+% components to reach the same 95 percent variance explained in the less complex
+% pie, which corresponds to higher compression. However, on the right with very
+% few components similar variance is explaining both pies.
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/pie_example.pdf}
-  \caption{\textbf{Illustration of compression.} Visual analogy for
-    neural compression. Here are 2 images of pies, one more complex
-    than the other. \textbf{a.} It takes fewer components to reach the same percent
-    variance explained in the less complex pie, which corresponds to
-    higher compression. \textbf{b.} However, with very few components, similar variance is explained in both pies.
-\textbf{c.} Plots the cumulative explained variance for more and more components.}
-  \label{fig:pie_example}
-\end{figure}
+% We investigated the dimensionality of neural patterns by training
+% classifiers using more and more principle components. Or, in other
+% words, we used less and less compression to decode.
+% We applied the approach to
+% a neuroimaging dataset comprising data collected as participants
+% listened to a story varying in cognitive richness~\citep{SimoEtal16}.
 
-We investigated the dimensionality of neural patterns by training
-classifiers using more and more principle components. Or, in other
-words, we used less and less compression to decode.
-We applied the approach to
-a neuroimaging dataset comprising data collected as participants
-listened to a story varying in cognitive richness~\citep{SimoEtal16}.
-
-% PCA, or the closely related factor analysis and singular value decomposition (SVD) (Hastie et al., 2009), is widely used in the study of individual differences and aids estimating how many latent components, or “factors”, underlie a pattern of (item) responses within or across participants, as for instance in the context of intelligence (Spearman, 1904) or personality tests (Cattell, 1947).
+% % PCA, or the closely related factor analysis and singular value decomposition (SVD) (Hastie et al., 2009), is widely used in the study of individual differences and aids estimating how many latent components, or “factors”, underlie a pattern of (item) responses within or across participants, as for instance in the context of intelligence (Spearman, 1904) or personality tests (Cattell, 1947).
 In the context of neuroimaging, PCA has been used to identify brain networks (Huth et al., 2012; Friston et al., 1993).
 
- % add citation:
-% S.M.Smith,A.Hyva ̈rinen,G.Varoquaux,K.L.Miller,and C. F. Beckmann, “Group-PCA for very large fMRI datasets,” NeuroImage, vol. 101, pp. 738, 2014.
+%  % add citation:
+% % S.M.Smith,A.Hyva ̈rinen,G.Varoquaux,K.L.Miller,and C. F. Beckmann, “Group-PCA for very large fMRI datasets,” NeuroImage, vol. 101, pp. 738, 2014.
 
-\subsection*{Evaluation metrics}
- We will evaluate the degree of
-compression of held-out
-neuroimaging data by assessing the time at which it was collected.  We will use this evaluation (timepoint
-decoding) as a proxy for
-gauging how much explanatory power the compressed data held
-with respect to the observed data.
+% \subsection*{Evaluation metrics}
+%  We will evaluate the degree of
+% compression of held-out
+% neuroimaging data by assessing the time at which it was collected.  We will use this evaluation (timepoint
+% decoding) as a proxy for
+% gauging how much explanatory power the compressed data held
+% with respect to the observed data.
 
-\subsubsection*{Timepoint decoding}
+% \subsubsection*{Timepoint decoding}
 
-To explore how compression varies with complexity, we will use a previous neuroimaging dataset
-~\cite{SimoEtal16}  in which participants listened to an audio recording of a story; 36 particpants listen to an intact version of the story, 17 participants listen to time-scrambled recordings of the same story where paragraphs were scrambled, 36 particpants listen to word-scrambled version and 36 participants lay in rest condition.
+% To explore how compression varies with complexity, we will use a previous neuroimaging dataset
+% ~\cite{SimoEtal16}  in which participants listened to an audio recording of a story; 36 particpants listen to an intact version of the story, 17 participants listen to time-scrambled recordings of the same story where paragraphs were scrambled, 36 particpants listen to word-scrambled version and 36 participants lay in rest condition.
 
-Following the analyses conducted by (HTFA)~\cite{MannEtal18}, we first
-apply \textit{hierarchical topographic factor analysis} (HTFA) to
-the fMRI datasets to obtain a time series of 700 node activities for
-every participant.  We then apply dimensionality reduction
-(Incremental PCA) for each group.
+% Following the analyses conducted by (HTFA)~\cite{MannEtal18}, we first
+% apply \textit{hierarchical topographic factor analysis} (HTFA) to
+% the fMRI datasets to obtain a time series of 700 node activities for
+% every participant.  We then apply dimensionality reduction
+% (Incremental PCA) for each group.
 
-We then compare the groups’ activity patterns (using Pearson
-correlations) to estimate the story times each corresponding pattern
-using more and more principle components.
+% We then compare the groups’ activity patterns (using Pearson
+% correlations) to estimate the story times each corresponding pattern
+% using more and more principle components.
 
-To assess decoding accuracy, we randomly divide participants for each
-stimulus into training and testing groups. We then compare the
-groups’ activity patterns (using Pearson correlations) to estimate the
-story times each corresponding pattern using more and more principle
-components (as the data became less compressed). Specifically, we ask, for each timepoint: what are the correlations
-between the first group's and second group's activity patterns at each
-order. We note that the decoding test we used is a conservative in which we count a timepoint label as incorrect if it is not an exact match.
+% To assess decoding accuracy, we randomly divide participants for each
+% stimulus into training and testing groups. We then compare the
+% groups’ activity patterns (using Pearson correlations) to estimate the
+% story times each corresponding pattern using more and more principle
+% components (as the data became less compressed). Specifically, we ask, for each timepoint: what are the correlations
+% between the first group's and second group's activity patterns at each
+% order. We note that the decoding test we used is a conservative in which we count a timepoint label as incorrect if it is not an exact match.
 

### 2023-02-23 paper/main.tex
working on intro

@@
 compressibility are related but are partially dissociable
-(Fig.~\ref{fig:information-compression}).
+(Fig.~\ref{fig:information-compression}). If a given signal (e.g., a
+representation of brain activity patterns) contains more information about
+ongoing cognitive processes, then the peak decoding accuracy should be high.
+And if the signal is compressible, a low-dimensional embedding of the signal
+will be similarly informative to the original signal
+(Fig.~\ref{fig:information-compression}D).
 
@@
 
-% Specifically, if a given signal
-% (e.g., a representation of brain activity patterns) contains more information
-% about ongoing cognitive processes, then the peak decoding accuracy should be
-% high. And if the signal is compressible, a low-dimensional embedding of the
-% signal will be similarly informative to the original signal. A signal can be
-% informative but not compressible, informative but
+Several recent studies suggest that the complexity of brain activity is
+task-dependent, whereby simpler tasks with lower cognitive demands are
+reflected by simpler and more compressible brain activity patterns, and more
+complex tasks with higher cognitive demands are reflected by more complex and
+less compressible brain activity patterns~\citep{MackEtal20, OwenEtal21}. These
+findings complement other work suggesting that functional connectivity
+(correlation) patterns are task-dependent~\citep{FinnEtal17, OwenEtal20,
+ColeEtal14}, although see~\cite{GratEtal18}. Higher-order cognitive processing
+of a common stimulus also appears to drive more stereotyped task-related
+activity and functional connectivity across individuals~\citep{HassEtal08,
+LernEtal11, SimoChan20, SimoEtal16}.
 
-% %   - e.g., you could have something that is highly compressible (just a few components are sufficient to achieve some threshold information level) but
-% %     *also* very rich-- e.g., the maximum number of informative components is large
 % % compression and task demands -- MackEtal20, ZhouEtal19
-% % superEEG -- prediction of held-out data; implies redundancy
 % % timecorr -- higher order network patterns during higher order cognition implies more complexity

### 2023-02-23 paper/main.tex
checking in a draft of the introduction

@@
 
-To the extent that brain activity patterns contain rich task-relevant
-information, we should be able to use the activity patterns to accurately
-differentiate between different aspects of the task~\citep[e.g., using pattern
-classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
-correspondence between classification accuracy and the information content of a
-signal~\citep{Alva02}. To the extent that brain activity patterns are
-compressible, we should be able to generate simplified (e.g., lower
-dimensional) representations of the data while still preserving the relevant or
-important aspects of the original signal. In general, information content and
-compressibility are related but are partially dissociable
-(Fig.~\ref{fig:information-compression}). If a given signal (e.g., a
-representation of brain activity patterns) contains more information about
-ongoing cognitive processes, then the peak decoding accuracy should be high.
-And if the signal is compressible, a low-dimensional embedding of the signal
-will be similarly informative to the original signal
-(Fig.~\ref{fig:information-compression}D).
-
 \begin{figure}[tp]
@@
 
+To the extent that brain activity patterns contain rich task-relevant
+information, we should be able to use the activity patterns to accurately
+differentiate between different aspects of the task~\citep[e.g., using pattern
+classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
+correspondence between classification accuracy and the information content of a
+signal~\citep{Alva02}. To the extent that brain activity patterns are
+compressible, we should be able to generate simplified (e.g., lower
+dimensional) representations of the data while still preserving the relevant or
+important aspects of the original signal. In general, information content and
+compressibility are related but are partially dissociable
+(Fig.~\ref{fig:information-compression}). If a given signal (e.g., a
+representation of brain activity patterns) contains more information about
+ongoing cognitive processes, then the peak decoding accuracy should be high.
+And if the signal is compressible, a low-dimensional embedding of the signal
+will be similarly informative to the original signal
+(Fig.~\ref{fig:information-compression}D).
+
 Several recent studies suggest that the complexity of brain activity is
@@
 
-% % compression and task demands -- MackEtal20, ZhouEtal19
-% % timecorr -- higher order network patterns during higher order cognition implies more complexity
-% % cognitive richness and scrambling-- various Hasson lab papers
+The above prior studies are consistent with two potential descriptions of how
+cognitive processes are reflects in brain activity patterns. One possibility is
+that the information rate of brain activity increases during more complex or
+higher-level cognitive processing. If so, then the ability to reliably decode
+cognitive states from brain activity patterns should improve with task
+complexity or with the level (or ``depth'') of cognitive processing. A second
+possibility is that the compressibility of brain activity patterns increases
+during more complex or higher-level cognitive processing. If so, then
+individual features of brain recordings, or compressed representations of brain
+recordings, should carry more information during complex or high-level (versus
+simple or low-level) cognitive tasks.
 
-% We're interested in the complexity of brain patterns that underly different
-% types of thoughts. To explore this question space, we will take brain patterns
-% recorded under different experimental conditions used in Aim 2, and project
-% them into lower dimensional spaces using principle components analysis. We can
-% then ask how well those low-dimensional embeddings of the data retain
-% cognitively relevant information like when in a story someone is listening to.
-
-% This work has been inspired, in part, by~\cite{MackEtal20}. In this paper, they
-% investigated the role of the prefrontal cortex in filtering out irrelevant
-% content. Specifically, they looked at if the vmPFC performs data reduction on
-% incoming information through compression. This was motivated, in part, by
-% orbital frontal cortex (OFC) compression in rats~\citep{ZhouEtal19}. They
-% studied this using a learning paradigm in which participants had to classify
-% insects based on different numbers of feature dimensions. The idea was that
-% participants in some learning blocks, participants could identify the insects
-% based on one feature (low complexity) or several features (high complexity),
-% but importantly the stimuli remained the same across all learning problems.
-% They found that complexity and compression had an inverse relationship; the
-% lower complexity of a conceptual space, the higher the degree of compression.
-% Building on this idea, we wonder if varying degrees of compression is performed
-% throughout the brain. We also want to test this idea, but using varying levels
-% of engagement listening to a naturalistic stimuli.
-
-% To understand the degree of compression throughout the brain during cognition,
-% we will use the same fMRI data from Aim 2, collected while participants
-% listened to a story in different scrambling conditions. We will measure the
-% degree that multivoxel activation patterns are compressed during story
-% listening using principle components analysis (PCA) a method for low-rank
-% approximation of multidimensional data~\citep{EckaYoun36}. We will explore this
-% using decoding accuracy as a function of the number of components, or
-% dimensions, in the low-dimensional space under different cognitive conditions.
-
-% You can imagine two reasonable predictions of how cognition is reflected in
-% brain patterns. The first is as our thoughts become more complex, they are
-% supported by more complex brain patterns, and require more components to
-% decode. The second is that when thoughts are deeper and more complicated, the
-% units of neural activity would carry more information, and would require
-% therefore fewer components to decode.
-
-% This idea can be explored in this visual analogy (Fig.~\ref{fig: pie_example})
-% for neural compression. Here there are two images of pies, the top pie is more
-% complex than the bottom. On the left we're illustrating that it takes fewer
-% components to reach the same 95 percent variance explained in the less complex
-% pie, which corresponds to higher compression. However, on the right with very
-% few components similar variance is explaining both pies.
-
-% We investigated the dimensionality of neural patterns by training
-% classifiers using more and more principle components. Or, in other
-% words, we used less and less compression to decode.
-% We applied the approach to
-% a neuroimaging dataset comprising data collected as participants
-% listened to a story varying in cognitive richness~\citep{SimoEtal16}.
+We used a previously collected neuroimaging dataset to estimate the extent to
+which each of these two possibilities might hold. The dataset we examined
+comprised functional magnetic resonance imaging (fMRI) data collected as
+participants listened to an audio recording of a 10-minute story, temporally
+scrambled recordings of the story, or underwent a resting state
+scan~\citep{SimoEtal16}. Each of these experimental conditions evokes different
+depths of cognitive processing~\citep{SimoEtal16,LernEtal11,
+HassEtal08,OwenEtal21}. We used across-participant classifiers to decode
+listening times in each condition, as a proxy for how ``informative'' the
+task-specific activity patterns were~\citep{SimoChan20}. We also use principle
+components analysis to generate lower-dimensional representations of the
+activity patterns. We then repeated the classification analyses after
+preserving different numbers of components and examined how classification
+accuracy changed across the different experimental conditions.
 

### 2023-02-24 paper/boneyard.tex
backing up progress

--- 2023-02-23 paper/main.tex
@@
-\documentclass[english]{article}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{hyperref}
-\usepackage{setspace}
-\usepackage{apacite}
-\usepackage{hyperref}
-\usepackage{natbib}
-\usepackage{pxfonts}
-\usepackage[utf8]{inputenc}
-\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
-\usepackage[left]{lineno}
-\usepackage{soul}
-\linenumbers
+By training classifiers using more and more principle components to decode, and comparing across conditions with varying degrees of cognitive richness, we can assess the explanatory power of the compressed data held with respect to the observed data (see \textit{Methods}).
+We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.
 
-\title{High-order cognition is supported by complex but compressible brain activity patterns}
-
-\author{Lucy L. W. Owen\textsuperscript{1, 2}, and Jeremy R. Manning\textsuperscript{1,
-*}\\\textsuperscript{1}Department of Psychological and Brain Sciences,\\Dartmouth College,
-Hanover, NH\\[0.1cm]\textsuperscript{2}Carney Institute for Brain Sciences,\\Brown University,
-Providence, RI\\[0.1cm] \textsuperscript{*}Address correspondence to
-jeremy.r.manning@dartmouth.edu}
-
-\begin{document}
-\maketitle
-
-\begin{abstract}
-
-We applied dimensionality reduction algorithms and pattern classifiers to
-functional neuroimaging data collected as participants listened to a story,
-temporally scrambled versions of the story, or underwent a resting state
-scanning session. These experimental conditions were intended to require
-different depths of processing and insprire different levels of engagement. We
-considered two primary aspects of the data. First, we treated the number of
-features (components) required to achieve a threshold decoding accuracy as a
-proxy for the ``compressibility'' of the neural patterns (where fewer
-components indicate higher compressibility). Second, we treated the maximum
-achievable decoding accuracy across participants as an indicator of the
-``stability'' of the recorded patterns. Overall, we found that neural patterns
-recorded as participants listened to the intact story required fewer features
-to achieve comparable classification accuracy to the other experimental
-conditions. However, the peak decoding accuracy (achievable with more features)
-was also highest during intact story listening. Taken together, our work
-suggests that our brain networks flexibly reconfigure according to ongoing task
-demands, and that the activity patterns associated with higher-order cognition
-and high engagement are both more complex and more compressible than the
-activity patterns associated with lower-order tasks and lower levels of
-engagement.
-
-\end{abstract}
-
-\doublespacing
-
-\section*{Introduction}
-
-Large-scale networks, including the human brain, may be conceptualized as
-occupying one or more positions along on a continuum. At one extreme, every
-node is fully independent of every other node. At the other extreme, all nodes
-behave identically. Each extreme optimizes key properties of how the network
-functions. When every node is independent, the network is maximally
-\textit{expressive}: if we define the network's ``state'' as the activity
-pattern across its nodes, then every state is equally reachable by a network
-with fully independent nodes. On the other hand, a network of identically
-behaved nodes optimizes \textit{robustness}: any subset of nodes may be removed
-from the network without any loss of function or expressive power, as long as
-any single node remains. Presumably, most natural systems tend to occupy
-positions between these extremes. We wondered: might the human brain
-reconfigure itself to be more flexible or more robust according to ongoing
-demands? In other words, might the brain reconfigure its connections or
-behaviors under different circumstances to change its position along this
-continuum?
-
-% compression (redundancy) vs. complexity -- appear to balance each other, butnot necessarily.
-
-Closely related to the above notions of expressiveness versus robustness are
-measures of how much \textit{information} is contained in a given signal or
-pattern, and how \textit{redundant} a signal is~\citep{Shan48}. Formally,
-information is defined as the amount of uncertainty about a given variables'
-outcomes (i.e., entropy), measured in \textit{bits}, or the optimal number of
-yes/no questions needed to reduce uncertainty about the variable's outcomes to
-zero. Highly complex systems with many degrees of freedom (i.e., high
-flexibility and expressiveness), are more information-rich than simpler or more
-constrained systems. The redundancy of a signal denotes the difference how
-expressive the signal \textit{could} be (i.e., proportional to the number of
-unique states or symbols used to transmit the signal) and the actual
-information rate (i.e., the entropy of each individual state or symbol). If a
-brain network's nodes are fully independent, then the number of bits required
-to express a single activity pattern is proportional to the number of nodes.
-The network would also be minimally redundant, since the status of every node
-would be needed to fully express a single brain activity pattern. If a brain
-network's nodes are fully coupled and identical, then the number of bits
-required to express a single activity pattern is proportional to the number of
-unique states or values any individual node can take on. Such a network would
-be highly redundant, since knowing any individual node's state would be
-sufficient to recover the full-brain activity pattern. Highly redundant systems
-are also robust, since there is little information loss from losing any given
-observation.
-
-We take as a given that brain activity is highly flexible: our brains can
-exhibit nearly infinite activity patterns. This flexibility implies that our
-brains activity patterns are highly information rich. However, brain activity
-patterns are also highly structured. For example, full-brain correlation
-matrices are stable within~\citep{FinnEtal15, FinnEtal17, GratEtal18} and
-across~\citep{YeoEtal11, GlerEtal12, GratEtal18, ColeEtal14} individuals. This
-stability suggests that our brains' activity patterns are at least partially
-constrained, for example by anatomical, external, or internal factors.
-Constraints on brain activity that limit its flexibility decrease
-expressiveness (i.e., its information rate). However, constraints on brain
-activity also increase its robustness to noise (e.g., ``missing'' or corrupted
-signals may be partially recovered). For example, recent work has shown that
-full-brain activity patterns may be reliably recovered from only a relatively
-small number of implanted electrodes~\citep{OwenEtal20, ScanEtal21}. This
-robustness property suggests that the relevant signal (e.g., underlying factors
-that have some influence over brain activity patterns) are compressible.
-
-\begin{figure}[tp]
-
+\begin{figure}
   \centering
-  \includegraphics[width=\textwidth]{figs/information_and_compressibility}
-
-  \caption{\textbf{Information content and compressibility.} \textbf{A. Variance
-  explained for two images.} We applied principal components analysis to a
-  photograph and drawing, treating each row of the images as ``observations.''
-  Across columns, we identified the number of components required to explain
-  100\%, 95\%, 75\%, or 50\% of the cumulative variance in each image (the 100\%
-  columns denote the original images). The numbers of components are indicate in
-  parentheses, and the resulting ``compressed'' images are displayed. \textbf{B.
-  Representing two images with different numbers of components.} Using the same
-  principal component decompositions as in Panel A, we computed the cumulative
-  proportion of variance explained with 380 (original images), 20, 4, or 2
-  components. \textbf{C. Cumulative variance explained versus number of
-  components.} For the images displayed in Panels A and B, we plot the cumulative
-  proportion of variance explained as a function of the number of components used
-  to represent each image. \textbf{D. Information rate and compressibility.}
-  Across multiple images, the information rate (i.e., the amount of information
-  contained in each image; horizontal axis) is high when each individual pixel
-  provides information that cannot be inferred from other pixels.
-  High-information rate images tend to be high-resolution, and low-information
-  rate images tend to be low-resolution. Compressibility is related to the
-  difference between the information required to specify the original versus
-  compressed images (vertical axis). Highly compressible images often contain
-  predictable structure (redundancies) that can be leveraged to represent the
-  images much more efficiently than in their original feature spaces.}
-
-  \label{fig:information-compression} 
+  \includegraphics[width=\textwidth]{figs/decode_interpret.pdf}
+  \caption{\textbf{Decoding accuracy.} \textbf{a. Decoding accuracy by
+      number of components.} Ribbons of each color display
+    cross-validated decoding performance for each condition (intact,
+    paragraph, word, and rest). Decoders were trained using
+    increasingly more principle components and displayed relative to
+    chance (red line). \textbf{b. Fixed decoding accuracy by number of
+      components.} We zoom in on the plot shown in \textbf{a.} and add
+    a line denoting fixed decoding accuracy (.05). We plot where the
+    intact, paragraph, and word conditions intersect.
+    \textbf{c.
+Explanation of inflection metric.} First the we fit a sigmoid function to the decoding accuracy by number of components.
+Second, we found where the second derivative is both positive and less than .0001.
+Last, we then plot that inflection point as a single metric to capture the slope and asymptote of the curve.}
+    \label{fig:decode_interpret}
   \end{figure}
 
-To the extent that brain activity patterns contain rich task-relevant
-information, we should be able to use the activity patterns to accurately
-differentiate between different aspects of the task~\citep[e.g., using pattern
-classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
-correspondence between classification accuracy and the information content of a
-signal~\citep{Alva02}. To the extent that brain activity patterns are
-compressible, we should be able to generate simplified (e.g., lower
-dimensional) representations of the data while still preserving the relevant or
-important aspects of the original signal. In general, information content and
-compressibility are related but are partially dissociable
-(Fig.~\ref{fig:information-compression}). If a given signal (e.g., a
-representation of brain activity patterns) contains more information about
-ongoing cognitive processes, then the peak decoding accuracy should be high.
-And if the signal is compressible, a low-dimensional embedding of the signal
-will be similarly informative to the original signal
-(Fig.~\ref{fig:information-compression}D).
+% We performed a decoding analysis, using cross validation to estimate
+% (using other participants’ data) which parts of the story the additional added
+% principle component corresponded to (see \textit{Materials and methods}).
+We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.
 
-Several recent studies suggest that the complexity of brain activity is
-task-dependent, whereby simpler tasks with lower cognitive demands are
-reflected by simpler and more compressible brain activity patterns, and more
-complex tasks with higher cognitive demands are reflected by more complex and
-less compressible brain activity patterns~\citep{MackEtal20, OwenEtal21}. These
-findings complement other work suggesting that functional connectivity
-(correlation) patterns are task-dependent~\citep{FinnEtal17, OwenEtal20,
-ColeEtal14}, although see~\cite{GratEtal18}. Higher-order cognitive processing
-of a common stimulus also appears to drive more stereotyped task-related
-activity and functional connectivity across individuals~\citep{HassEtal08,
-LernEtal11, SimoChan20, SimoEtal16}.
+% Separately for each experimental condition, we divided participants
+% into two groups. Starting with 1 principle comonent, for
+% each dimension we added another principle component, and we correlated the group 1 activity patterns with group 2
+% activity patterns.  There were 272
+% timepoints for paragraph condition, 300 timepoints for intact and word
+% conditions, and 400 timepoints for rest condition,  so chance
+% performance on this decoding test is was $\frac{1}{272}$,
+% $\frac{1}{300}$, and $\frac{1}{400}$ respectively.
 
-The above prior studies are consistent with two potential descriptions of how
-cognitive processes are reflects in brain activity patterns. One possibility is
-that the information rate of brain activity increases during more complex or
-higher-level cognitive processing. If so, then the ability to reliably decode
-cognitive states from brain activity patterns should improve with task
-complexity or with the level (or ``depth'') of cognitive processing. A second
-possibility is that the compressibility of brain activity patterns increases
-during more complex or higher-level cognitive processing. If so, then
-individual features of brain recordings, or compressed representations of brain
-recordings, should carry more information during complex or high-level (versus
-simple or low-level) cognitive tasks.
+% - As complexity of the stimulus increases, decoding accuracy
+% increases (Fig.~\ref{fig:decode_interpret},  a.).  Replication of
+% Simony findings.
 
-We used a previously collected neuroimaging dataset to estimate the extent to
-which each of these two possibilities might hold. The dataset we examined
-comprised functional magnetic resonance imaging (fMRI) data collected as
-participants listened to an audio recording of a 10-minute story, temporally
-scrambled recordings of the story, or underwent a resting state
-scan~\citep{SimoEtal16}. Each of these experimental conditions evokes different
-depths of cognitive processing~\citep{SimoEtal16,LernEtal11,
-HassEtal08,OwenEtal21}. We used across-participant classifiers to decode
-listening times in each condition, as a proxy for how ``informative'' the
-task-specific activity patterns were~\citep{SimoChan20}. We also use principle
-components analysis to generate lower-dimensional representations of the
-activity patterns. We then repeated the classification analyses after
-preserving different numbers of components and examined how classification
-accuracy changed across the different experimental conditions.
+% - As complexity of the stimuli increases, we need fewer components to
+% decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
+
+% - As complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
+
+Prior work has shown participants share similar neural responses to
+richly structured stimuli when compared to stimuli with less
+structure \cite{SimoEtal16}.  We replicate this finding, showing as complexity of the stimulus increases, decoding accuracy
+increases (Fig.~\ref{fig:decode_interpret},  a.).  
+Additionally, we found that as complexity of the stimuli increases, we need fewer components to decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
+However, we also found that as complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
+We posit that as the complexity of our thoughts increases, neural
+compression decreases. However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
 
@@
 
-\section*{Results}
-By training classifiers using more and more principle components to decode, and comparing across conditions with varying degrees of cognitive richness, we can assess the explanatory power of the compressed data held with respect to the observed data (see \textit{Methods}).
-We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.
+  Most complex systems reflect dynamic interactions between myriad
+  evolving components (e.g., interacting molecules, interacting brain
+  systems, interacting individuals within a social network or
+  ecological system, coordinated components within a mechanical or
+  digital device, etc.).  Despite that these interactions are central
+  to the full system's behavior (e.g., removing a component from the
+  full system can change the entire system's behavior), dynamic
+  interactions cannot typically be directly measured.  Rather, the
+  interactions must be inferred through their hypothesized role in
+  guiding the dynamics of system components.  Here we use a
+  model-based approach to inferring dynamic interactions from
+  timeseries data.  In addition to examining first-order interactions
+  (e.g., between pairs of components) we also examine higher-order
+  interactions (e.g., that characterize mirrored structure in the
+  patterns of interaction dynamics displayed by different subsets of
+  components).  We apply our approach to two datasets.  First, we use
+  a synthetic dataset, for which the underlying dynamic interactions
+  are known, to show that our model recovers those ground-truth
+  dynamic interactions.  We also apply our model to a neuroimaging
+  dataset and show that the high-order dynamic interactions exhibited
+  by brain data vary meaningfully as a function of the cognitive
+  ``richness'' of the stimulus people are experiencing.
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/decode_interpret.pdf}
-  \caption{\textbf{Decoding accuracy.} \textbf{a. Decoding accuracy by
-      number of components.} Ribbons of each color display
-    cross-validated decoding performance for each condition (intact,
-    paragraph, word, and rest). Decoders were trained using
-    increasingly more principle components and displayed relative to
-    chance (red line). \textbf{b. Fixed decoding accuracy by number of
-      components.} We zoom in on the plot shown in \textbf{a.} and add
-    a line denoting fixed decoding accuracy (.05). We plot where the
-    intact, paragraph, and word conditions intersect.
-    \textbf{c.
-Explanation of inflection metric.} First the we fit a sigmoid function to the decoding accuracy by number of components.
-Second, we found where the second derivative is both positive and less than .0001.
-Last, we then plot that inflection point as a single metric to capture the slope and asymptote of the curve.}
-    \label{fig:decode_interpret}
-  \end{figure}
+%should we introduce the name levelup in the introduction? ie. "in the second step, levelup, we use..." -ecw
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Methods}
+%%%%%%%%%%%%%%%%%%%%%%%%
 
-Prior work has shown participants share similar neural responses to
-richly structured stimuli when compared to stimuli with less
-structure \cite{SimoEtal16}.  We replicate this finding, showing as complexity of the stimulus increases, decoding accuracy
-increases (Fig.~\ref{fig:decode_interpret},  a.).  
-Additionally, we found that as complexity of the stimuli increases, we need fewer components to decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
-However, we also found that as complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
-We posit that as the complexity of our thoughts increases, neural
-compression decreases. However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
+The mathematical theories behind the timecorr approach for calculating the moment-by-moment dynamic correlations in systems are described here.
+The function takes a number-of-timepoint by number-of-features matrix as well as a weight function to determine a moment-by-moment correlations matrix.
+To calculate the correlations, expectation values are found using equation 1:
 
-% We performed a decoding analysis, using cross validation to estimate
-% (using other participants’ data) which parts of the story the additional added
-% principle component corresponded to (see \textit{Materials and methods}).
-We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.
+\begin{equation}
+\mathrm{corr}_{x,y}(t) = \frac{1}{T} \sum_{t} \frac{\left(x(t)-\mathbb{E}[x]\right)\left(y(t)-\mathbb{E}[y]\right)}{\sigma_{x}(t)\sigma_{y}(t)}
+\end{equation}
 
-% Separately for each experimental condition, we divided participants
-% into two groups. Starting with 1 principle comonent, for
-% each dimension we added another principle component, and we correlated the group 1 activity patterns with group 2
-% activity patterns.  There were 272
-% timepoints for paragraph condition, 300 timepoints for intact and word
-% conditions, and 400 timepoints for rest condition,  so chance
-% performance on this decoding test is was $\frac{1}{272}$,
-% $\frac{1}{300}$, and $\frac{1}{400}$ respectively.
+%should the variables in this equation/description change to match the description in the intro?
+Here, x is the timepoint component of the matrix and y is the feature component.
+Timepoint by timepoint, x by x, at each given moment (t) in the matrix is subtracted from the expectation value of the full timepoint matrix.
+The same process occurs for the features points (y) at each timepoint (t).
+The product of these are then divided by the standard deviation at that moment for both the timepoint and the feature.
+Only the upper triangle of the matrix is calculated as the interactions are assumed to be both symmetric and non-recurrent.
 
-% - As complexity of the stimulus increases, decoding accuracy
-% increases (Fig.~\ref{fig:decode_interpret},  a.).  Replication of
-% Simony findings.
+Two weight parameters have been used to calculate the connectivity of the data are a Gaussian weight parameter, equation 2, and a LaPlace weights parameter, equation 3.
 
-% - As complexity of the stimuli increases, we need fewer components to
-% decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
+%label as Gaussian? 
+\begin{equation}
+\mathcal{N}\left( x | \mu, \sigma \right) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left(-\frac{|| x - \mu ||^2}{2\sigma^2}\right)
+\end{equation}
+where sigma squared is the variance and mu is the expected value.
 
-% - As complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
+%label as LaPlace? 
+\begin{equation}
+\mathcal{L}\left( x | \mu, \lambda \right) = \frac{1}{2 \lambda}\left( x | \mu, \lambda \right) = \exp\left(-\frac{|x - \mu|}{\lambda}\right)
+\end{equation}
+where mu is the expectation values, lambda is a defined decay constant.
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/decode_pcs_network.pdf}
-  \caption{\textbf{Inflection points by network.} \textbf{a.}
-    Inflection point was calculated as explained in
-    Fig.~\ref{fig:decode_interpret}, b. Analyses were limited by the
-    brain networks (using the \cite{YeoEtal11} network parcellation) and
-    arranged in increasing order relative to the intact condition.
-    \textbf{b.
-and c.} For the total time in the intact condition, we are plotting the relative inflection points (\textbf{b.}) and corresponding number of components (\textbf{c.}) by network.
-\textbf{d.} The network parcellation defined by \cite{YeoEtal11} is displayed on the inflated brain maps.
-The colors and network labels serve as a legend for \textbf{a.} and \textbf{d.}}
-  \label{fig:decode_pcs_network}
-\end{figure}
+%figure showing the difference b/w two weights in same data set? Using the synthetic data?
 
-We also wondered how this compression would change across brain regions.
-We repeated the analysis but limited the brain hubs to 7 networks using the Yeo et al.
-(2011) network parcellation shown here in the inflated brain (Fig.~\ref{fig:decode_pcs_network}, d.).
-We found that as complexity of the stimuli increases, decoding accuracy increases
-with higher cognitive areas. (Fig.~\ref{fig:decode_pcs_network}).
+When using the Gaussian weights parameter timepoints are more closely correlated to timepoints further away in the matrix than when using then the LaPlace weight function.
 
-% As complexity of the stimuli increases, decoding accuracy increases
-% with higher cognitive areas. (Fig.~\ref{fig: decode_pcs_network}).
+%Plot of kernels as Fig 1?
 
-%  If, there is some understanding of the narrative that accumulates over time, we should be able to see that change.
+% New section? Mathematical Methods of Levelup? 
+In order to understand how the data is interconnected we use the function levelup. Levelup uses the same input as the timecorr function, a timepoint-by-function matrix.
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/decode_pcs.pdf}
-  \caption{\textbf{Inflection points by thirds.} \textbf{a.}
-    Decoding accuracy by number of components not broken into thirds
-    (Fig.~\ref{fig:decode_interpret} a.). \textbf{b. and c.}
-    Quantifying changes in decoding accuracy across
-time. \textbf{b.} Slope of decoding accuracy was calculated by fitting
-a regression line for each component/condition for each
-third. \textbf{c.}  We also repeated the analysis
-(Fig.~\ref{fig:decode_interpret}, b.) to obtain the inflection point
-for each condition and for each third.
-\textbf{d.}  Decoding accuracy by number of components for each third of the scan time.
-We repeated the same analysis in Fig.~\ref{fig:decode_interpret} a.
-but breaking the scan time for each condition into 3 intervals.
-}
-  \label{fig:decode_pcs_thirds}
-\end{figure}
+%Diagram of input-->timecorr-->PCaA= levelup?
 
-We were also curious how compression would change across time.
-If, there is some understanding of the narrative that accumulates over time, we should be able to see that difference.
-We found increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
-We also found decreases in decoding accuracy for the word-scrambled and rest
-condition.
+In the levelup method the the data first goes through the timecorr process and then is reduced using Principal Component Analysis (PCA).
+By utilizing PCA the output data maintains the same features as the input method.
+Not only does this allow for a sanity check, but also it allows for repeated use of the function.
+Using PCA or other reduction techniques is needed as without it the output data would get exponentially large.
+On the flip side using PCA, or any dimensionality reduction technique, makes the output data less accurate as information is lost.
 
-Overall, we found that as story listening conditions become more complex, more components are required to decode. We also found we could decode better with more impoverished data when
-there is the underlying structure of the narrative providing more cognitive richness.
-We posit that as the complexity of our thoughts increases, neural compression decreases.
-However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
+%PCA equation?
+%figure/ chart of how many levels we "go up" before data gets too muddled- depends on the correlation?
 
-% - Increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
-% - Decreases in decoding accuracy for the word-scrambled and rest
-% condition.
+Timepoints that share high similarity across all subjects have a high probability of being stimulus driven.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Discussion}
+To compute the correlations of multiple subjects the across mode of both timecorr and levelup are used.
+The across function compares one voxel in a sample to the averages of that voxel in the rest of the dataframe.
+This occurs throughout every voxel in the dataset.
+The output of the across mode is a correlation matrix that is a number of timepoints by features matrix.
 
-- We trained classifiers using more and more principle components to decode, and compared across condi- tions with varying degrees of cognitive richness.
--We found that as listening conditions become more cognitively rich,
-decoding accuracy increased.
--Also, decoding accuracy increased as understanding of the narrative
-accumulated over time, in more complex listening conditions.
-- Decoding accuracy also increased in higher cognitive areas, in more complex listening conditions.
--We found that as story listening conditions become more complex, more
-components are required to decode.
--We also found we could decode better with more impoverished data when
-there is the underlying structure of the narrative providing more cognitive richness.
--We posit that as the complexity of our thoughts increases, neural
-compression decreases. However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
-
-Based on prior work ~\citep{Deme19} and following the direction of the field ~\citep{Turk13} we think our thoughts might be encoded in
-dynamic network patterns, and possibly higher order network
-patterns (Fig.~\ref{fig:direction_of_field}). We sought to test this hypothesis by developing an approach
-to inferring high-order network dynamics from timeseries data.
-
-One challenge in studying dynamic interactions is the
-computational resources required to calculate higher-order correlations. 
-We developed a computationally tractable model of network dynamics (Fig.~\ref{fig:methods_fig}) that takes in a feature
-timeseries and outputs approximated first-order dynamics (i.e.,
-dynamic functional correlations), second-order dynamics
-(reflecting homologous networks that dynamically form and disperse),
-and higher-order network dynamics (up to tenth-order dynamic
-correlations).
-
-We first validated our model using synthetic data, and explored how
-recovery varied with different underlying data structures and kernels.   We then 
-applied the approach to an fMRI dataset
-~\citep{SimoEtal16} in which participants listened to an audio
-recording of a story, as well as scrambled versions of the same story
-(where the scrambling was applied at different temporal scales).  We
-trained classifiers to take the output of the model and decode the
-timepoint in the story (or scrambled story) that the participants were
-listening to. We found that, during the intact listening condition in the
-experiment, classifiers that incorporated higher-order correlations
-yielded consistently higher accuracy than classifiers trained only on
-lower-order patterns (Fig.~\ref{fig:decoding_level},  a.\&d.).  By contrast, these
-higher-order correlations were not necessary to support decoding the other
-listening conditions and (minimally
-above chance) during a control rest condition.  This suggests
-that the cognitive processing that supported the most cogntively rich listening conditions
-involved second-order (or higher) network dynamics.
-
-Although we found decoding accuracy was best when incorporating
-higher-order network dynamics for all but rest
-  condition, it is unclear if this is a product of the brain or the
-  data collection technique.  It could be that the brain is
-  second-order or it could be that fMRI can
-  only reliably give second-order interactions. Exploring this method
-  with other data collection technique will be important to
-  disentangle this question.
-
-  \subsection*{Concluding remarks}
-
-How can we better understand how brain patterns change over
-time? How can we quantify the potential network dynamics that might be
-driving these changes? One way to judge the techniques of the future is
-to look at the trajectory of the fMRI field so far has taken so far
-(Fig.~\ref{fig:methods_fig}).  The field started with 
-univariate activation, measuring the average activity for each voxel.
-Analyses of multivariate activation followed, looking at spatial patterns of
-activity over voxels. Next, correlations of activity were explored, first
-with measures like resting connectivity that take temporal correlation
-between a seed voxel and all other voxels then with full connectivty
-that measure all pairwise correlations.  Additionally, this path of increasing
-complexity also moved from static to dynamic measurements.  One
-logical next step in this trajectory would be dynamic higher-order
-correlations. We have created a method 
-to support these calculations by scalably approximating dynamic higher-order
-correlations.
-
-\section*{Acknowledgements}
-We acknowledge discussions with Luke Chang, Hany Farid, Paxton
-Fitzpatrick, Andrew Heusser, Eshin Jolly, Qiang Liu, Matthijs van der
-Meer, Judith Mildner, Gina Notaro, Stephen Satterthwaite, Emily
-Whitaker, Weizhen Xie, and Kirsten Ziman. Our work was supported in
-part by NSF EPSCoR Award Number 1632738 to J.R.M. and by a sub-award
-of DARPA RAM Cooperative Agreement N66001-14-2-4-032 to J.R.M.  The
-content is solely the responsibility of the authors and does not
-necessarily represent the official views of our supporting
-organizations.
-
-\section*{Author contributions}
-Concept: J.R.M. and L.L.W.O. Implementation: L.L.W.O., and
-J.R.M.  Analyses: L.L.W.O and J.R.M.
-
-\bibliographystyle{apacite}
-\bibliography{CDL-bibliography/cdl}
-
-\end{document}
+%F^2-F/2 +F

### 2023-02-24 paper/main.tex
minor typo fix
backing up progress

@@
 
-\author{Lucy L. W. Owen\textsuperscript{1, 2}, and Jeremy R. Manning\textsuperscript{1,
+\author{Lucy L. W. Owen\textsuperscript{1, 2} and Jeremy R. Manning\textsuperscript{1,
 *}\\\textsuperscript{1}Department of Psychological and Brain Sciences,\\Dartmouth College,
@@
 scanning session. These experimental conditions were intended to require
-different depths of processing and insprire different levels of engagement. We
-considered two primary aspects of the data. First, we treated the number of
-features (components) required to achieve a threshold decoding accuracy as a
-proxy for the ``compressibility'' of the neural patterns (where fewer
-components indicate higher compressibility). Second, we treated the maximum
-achievable decoding accuracy across participants as an indicator of the
+different depths of processing and inspire different levels of cognitive
+engagement. We considered two primary aspects of the data. First, we treated
+the number of features (components) required to achieve a threshold decoding
+accuracy as a proxy for the ``compressibility'' of the neural patterns (where
+fewer components indicate greater compression). Second, we treated the
+maximum achievable decoding accuracy across participants as an indicator of the
 ``stability'' of the recorded patterns. Overall, we found that neural patterns
@@
 
-% % PCA, or the closely related factor analysis and singular value decomposition (SVD) (Hastie et al., 2009), is widely used in the study of individual differences and aids estimating how many latent components, or “factors”, underlie a pattern of (item) responses within or across participants, as for instance in the context of intelligence (Spearman, 1904) or personality tests (Cattell, 1947).
-In the context of neuroimaging, PCA has been used to identify brain networks (Huth et al., 2012; Friston et al., 1993).
-
-%  % add citation:
-% % S.M.Smith,A.Hyva ̈rinen,G.Varoquaux,K.L.Miller,and C. F. Beckmann, “Group-PCA for very large fMRI datasets,” NeuroImage, vol. 101, pp. 738, 2014.
-
-% \subsection*{Evaluation metrics}
-%  We will evaluate the degree of
-% compression of held-out
-% neuroimaging data by assessing the time at which it was collected.  We will use this evaluation (timepoint
-% decoding) as a proxy for
-% gauging how much explanatory power the compressed data held
-% with respect to the observed data.
-
-% \subsubsection*{Timepoint decoding}
-
-% To explore how compression varies with complexity, we will use a previous neuroimaging dataset
-% ~\cite{SimoEtal16}  in which participants listened to an audio recording of a story; 36 particpants listen to an intact version of the story, 17 participants listen to time-scrambled recordings of the same story where paragraphs were scrambled, 36 particpants listen to word-scrambled version and 36 participants lay in rest condition.
-
-% Following the analyses conducted by (HTFA)~\cite{MannEtal18}, we first
-% apply \textit{hierarchical topographic factor analysis} (HTFA) to
-% the fMRI datasets to obtain a time series of 700 node activities for
-% every participant.  We then apply dimensionality reduction
-% (Incremental PCA) for each group.
-
-% We then compare the groups’ activity patterns (using Pearson
-% correlations) to estimate the story times each corresponding pattern
-% using more and more principle components.
-
-% To assess decoding accuracy, we randomly divide participants for each
-% stimulus into training and testing groups. We then compare the
-% groups’ activity patterns (using Pearson correlations) to estimate the
-% story times each corresponding pattern using more and more principle
-% components (as the data became less compressed). Specifically, we ask, for each timepoint: what are the correlations
-% between the first group's and second group's activity patterns at each
-% order. We note that the decoding test we used is a conservative in which we count a timepoint label as incorrect if it is not an exact match.
-
-% \subsection*{Evaluation metrics}
-%  We evaluated the degree of
-% compression of held-out
-% neuroimaging data with the time at which it was collected.  We used this latter evaluations (using timepoint
-% decoding) as a proxy for
-% gauging how much explanatory power the compressed data held
-% with respect to the observed data.
-
-% \subsubsection*{Timepoint decoding}
-
-% To explore how higher-order structure varies with stimulus structure
-% and complexity, we used a previous neuroimaging dataset
-% ~\cite{SimoEtal16}  in which participants listened to an audio recording of a story; 36 particpants listen to an intact version of the story, 17 participants listen to time-scrambled recordings of the same story where paragraphs were scrambled, 36 particpants listen to word-scrambled version and 36 participants lay in rest condition.
-
-% Prior work has shown participants share similar neural responses to
-% richly structured stimuli when compared to stimuli with less
-% structure.  To assess whether the moment-by-moment higher order
-% correlations were reliably preserved across participants, we used
-% inter-subject functional connectivity (ISFC) to isolate the
-% time-varying correlational structure (functional connectivity patterns
-% that were specifically driven by the story participants listened to.
-% Following the analyses conducted by (HTFA)~\cite{MannEtal18}, we first
-% applied \textit{hierarchical topographic factor analysis} (HTFA) to
-% the fMRI datasets to obtain a time series of 700 node activities for
-% every participant.  We then applied dimensionality reduction
-% (Incremental PCA) for each group.
-
-% We then compared the groups’ activity patterns (using Pearson
-% correlations) to estimate the story times each corresponding pattern
-% using more and more principle components.
-
-% To assess decoding accuracy, we randomly divided participants for each
-% stimulus into training and testing groups. We then compared the
-% groups’ activity patterns (using Pearson correlations) to estimate the
-% story times each corresponding pattern using more and more principle
-% components (as the data became less compressed). Specifically, we asked, for each timepoint: what are the correlations
-% between the first group's and second group's activity patterns at each
-% order. We note that the decoding test we used is a conservative in which we count a timepoint label as incorrect if it is not an exact match.
-
 \section*{Results}
-By training classifiers using more and more principle components to decode, and comparing across conditions with varying degrees of cognitive richness, we can assess the explanatory power of the compressed data held with respect to the observed data (see \textit{Methods}).
-We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figs/decode_interpret.pdf}
-  \caption{\textbf{Decoding accuracy.} \textbf{a. Decoding accuracy by
+
+  \includegraphics[width=0.8\textwidth]{figs/decoding_and_inflection}
+
+  \caption{\textbf{Decoding accuracy and compression.} \textbf{A. Decoding accuracy by
       number of components.} Ribbons of each color display
     cross-validated decoding performance for each condition (intact,
-    paragraph, word, and rest). Decoders were trained using
-    increasingly more principle components and displayed relative to
-    chance (red line). \textbf{b. Fixed decoding accuracy by number of
-      components.} We zoom in on the plot shown in \textbf{a.} and add
-    a line denoting fixed decoding accuracy (.05). We plot where the
-    intact, paragraph, and word conditions intersect.
-    \textbf{c.
-Explanation of inflection metric.} First the we fit a sigmoid function to the decoding accuracy by number of components.
-Second, we found where the second derivative is both positive and less than .0001.
-Last, we then plot that inflection point as a single metric to capture the slope and asymptote of the curve.}
-    \label{fig:decode_interpret}
+    paragraph, word, and rest), as a function of the number of components (features) used to
+    represent the data.  The horizontal red line denotes chance performance, and the horizontal black
+    line denotes 5\% decoding accuracy (used as a reference in Panel B).  \textbf{B. Numbers of components
+    required to reach a fixed decoding accuracy threshold, by condition.}  The panel displays a zoomed-in
+    view of the inset in Panel A.  Intersections between each condition's decoding accuracy curve
+    and the 5\% decoding accuracy reference line are marked by arrows.  \textbf{C. Estimating
+    inflection points.}  We sought to identify an ``inflection point'' for each decoding curve, denoting
+    the number of components at which the decoding curve asymptotes.  We fit sigmoid functions to each
+    decoding curve (left sub-panel) and then computed the minimum number of components where the
+    second derivative of the sigmoid was both positive and less than a threshold value of 0.0001.
+    \textbf{D. Inflection points by condition.}  Each dot displays the number of components ($x$-axis)
+    and decoding accuracy ($y$-axis) at one condition's inflection point.}
+
+    \label{fig:decoding-inflection}
   \end{figure}
-
-Prior work has shown participants share similar neural responses to
-richly structured stimuli when compared to stimuli with less
-structure \cite{SimoEtal16}.  We replicate this finding, showing as complexity of the stimulus increases, decoding accuracy
-increases (Fig.~\ref{fig:decode_interpret},  a.).  
-Additionally, we found that as complexity of the stimuli increases, we need fewer components to decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
-However, we also found that as complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
-We posit that as the complexity of our thoughts increases, neural
-compression decreases. However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
-
-% We performed a decoding analysis, using cross validation to estimate
-% (using other participants’ data) which parts of the story the additional added
-% principle component corresponded to (see \textit{Materials and methods}).
-We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.
-
-% Separately for each experimental condition, we divided participants
-% into two groups. Starting with 1 principle comonent, for
-% each dimension we added another principle component, and we correlated the group 1 activity patterns with group 2
-% activity patterns.  There were 272
-% timepoints for paragraph condition, 300 timepoints for intact and word
-% conditions, and 400 timepoints for rest condition,  so chance
-% performance on this decoding test is was $\frac{1}{272}$,
-% $\frac{1}{300}$, and $\frac{1}{400}$ respectively.
-
-% - As complexity of the stimulus increases, decoding accuracy
-% increases (Fig.~\ref{fig:decode_interpret},  a.).  Replication of
-% Simony findings.
-
-% - As complexity of the stimuli increases, we need fewer components to
-% decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
-
-% - As complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
 

### 2023-02-27 paper/boneyard.tex
updating figures (captions and layouts)

@@
   \end{figure}
+
+  We also wondered how this compression would change across brain regions.
+We repeated the analysis but limited the brain hubs to 7 networks using the Yeo et al.
+(2011) network parcellation shown here in the inflated brain (Fig.~\ref{fig:decode_pcs_network}, d.).
+We found that as complexity of the stimuli increases, decoding accuracy increases
+  with higher cognitive areas. (Fig.~\ref{fig:decode_pcs_network}).
+
+  % As complexity of the stimuli increases, decoding accuracy increases
+  % with higher cognitive areas. (Fig.~\ref{fig: decode_pcs_network}).
+
+  %  If, there is some understanding of the narrative that accumulates over time, we should be able to see that change.
 

### 2023-02-27 paper/main.tex
updated network figure
updating figures (captions and layouts)

@@
 
-  \caption{\textbf{Decoding accuracy and compression.} \textbf{A. Decoding accuracy by
-      number of components.} Ribbons of each color display
-    cross-validated decoding performance for each condition (intact,
-    paragraph, word, and rest), as a function of the number of components (features) used to
-    represent the data.  The horizontal red line denotes chance performance, and the horizontal black
-    line denotes 5\% decoding accuracy (used as a reference in Panel B).  \textbf{B. Numbers of components
-    required to reach a fixed decoding accuracy threshold, by condition.}  The panel displays a zoomed-in
-    view of the inset in Panel A.  Intersections between each condition's decoding accuracy curve
-    and the 5\% decoding accuracy reference line are marked by arrows.  \textbf{C. Estimating
-    inflection points.}  We sought to identify an ``inflection point'' for each decoding curve, denoting
-    the number of components at which the decoding curve asymptotes.  We fit sigmoid functions to each
-    decoding curve (left sub-panel) and then computed the minimum number of components where the
-    second derivative of the sigmoid was both positive and less than a threshold value of 0.0001.
-    \textbf{D. Inflection points by condition.}  Each dot displays the number of components ($x$-axis)
-    and decoding accuracy ($y$-axis) at one condition's inflection point.}
+\caption{\textbf{Decoding accuracy and compression.} \textbf{A. Decoding
+accuracy by number of components.} Ribbons of each color display
+cross-validated decoding performance for each condition (intact, paragraph,
+word, and rest), as a function of the number of components (features) used to
+represent the data. The horizontal red line denotes chance performance, and the
+horizontal black line denotes 5\% decoding accuracy (used as a reference in
+Panel B). \textbf{B. Numbers of components required to reach a fixed decoding
+accuracy threshold, by condition.} The panel displays a zoomed-in view of the
+inset in Panel A. Intersections between each condition's decoding accuracy
+curve and the 5\% decoding accuracy reference line are marked by arrows.
+\textbf{C. Estimating inflection points.} We sought to identify an ``inflection
+point'' for each decoding curve, denoting the number of components at which the
+decoding curve asymptotes. We fit sigmoid functions to each decoding curve
+(left sub-panel) and then computed the minimum number of components where the
+second derivative of the sigmoid was both positive and less than a threshold
+value of 0.0001. \textbf{D. Inflection points by condition.} Each dot displays
+the number of components ($x$-axis) and decoding accuracy ($y$-axis) at one
+condition's inflection point. All error ribbons denote bootstrap-estimated 95\%
+confidence intervals.}
 
-    \label{fig:decoding-inflection}
+    \label{fig:inflection}
   \end{figure}
@@
   \centering
-  \includegraphics[width=\textwidth]{figs/decode_pcs_network.pdf}
-  \caption{\textbf{Inflection points by network.} \textbf{a.}
-    Inflection point was calculated as explained in
-    Fig.~\ref{fig:decode_interpret}, b. Analyses were limited by the
-    brain networks (using the \cite{YeoEtal11} network parcellation) and
-    arranged in increasing order relative to the intact condition.
-    \textbf{b.
-and c.} For the total time in the intact condition, we are plotting the relative inflection points (\textbf{b.}) and corresponding number of components (\textbf{c.}) by network.
-\textbf{d.} The network parcellation defined by \cite{YeoEtal11} is displayed on the inflated brain maps.
-The colors and network labels serve as a legend for \textbf{a.} and \textbf{d.}}
-  \label{fig:decode_pcs_network}
+  \includegraphics[width=0.8\textwidth]{figs/decode_pcs_network}
+
+  \caption{\textbf{Network-specific decoding accuracy and compression.}
+  \textbf{A. Decoding accuracy and number of components for network-specific
+  inflection points.} We considered the seven networks identified by
+  \cite{YeoEtal11}. We computed each network's inflection point, for each
+  experimental condition, using the procedure described in
+  Figure~\ref{fig:inflection}C. \textbf{B. Network-specific decoding
+  accuracy.} Each of the seven networks are colored according to the decoding
+  accuracy at the network's inflection point for the ``intact'' experimental
+  condition (corresponding to the dark blue curve in Panel A). \textbf{C.
+  Network-specific compression.} Each of the seven networks are colored
+  according to the number of components at the network's inflection point for
+  the intact experimental condition. Larger numbers of components reflect lower
+  compressibility.}
+
+  \label{fig:networks}
 \end{figure}
-
-We also wondered how this compression would change across brain regions.
-We repeated the analysis but limited the brain hubs to 7 networks using the Yeo et al.
-(2011) network parcellation shown here in the inflated brain (Fig.~\ref{fig:decode_pcs_network}, d.).
-We found that as complexity of the stimuli increases, decoding accuracy increases
-with higher cognitive areas. (Fig.~\ref{fig:decode_pcs_network}).
-
-% As complexity of the stimuli increases, decoding accuracy increases
-% with higher cognitive areas. (Fig.~\ref{fig: decode_pcs_network}).
-
-%  If, there is some understanding of the narrative that accumulates over time, we should be able to see that change.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figs/decode_pcs.pdf}
-  \caption{\textbf{Inflection points by thirds.} \textbf{a.}
-    Decoding accuracy by number of components not broken into thirds
-    (Fig.~\ref{fig:decode_interpret} a.). \textbf{b. and c.}
-    Quantifying changes in decoding accuracy across
-time. \textbf{b.} Slope of decoding accuracy was calculated by fitting
-a regression line for each component/condition for each
-third. \textbf{c.}  We also repeated the analysis
-(Fig.~\ref{fig:decode_interpret}, b.) to obtain the inflection point
-for each condition and for each third.
-\textbf{d.}  Decoding accuracy by number of components for each third of the scan time.
-We repeated the same analysis in Fig.~\ref{fig:decode_interpret} a.
-but breaking the scan time for each condition into 3 intervals.
-}
-  \label{fig:decode_pcs_thirds}
+  \includegraphics[width=0.8\textwidth]{figs/information_and_compression_over_time.pdf}
+
+\caption{\textbf{Changes in decoding accuracy and compression over time.}
+\textbf{A. Decoding accuracy by number of components, by story segment.} Each
+family of curves is plotted in the same format as Figure~\ref{fig:inflection}A
+but reflects data only from one third of the dataset. \textbf{B. Inflection
+points by condition and segment.} The dots re-plot the inflection points from
+Figure~\ref{fig:inflection}D for reference. The numbers denote the inflection
+points for each third of the dataset (1: first third; 2: second third; 3: third
+third; colors denote experimental conditions). \textbf{C. Change in decoding
+accuracy over time, by number of components.} For each number of components
+($x$-axis) and condition (color), we fit a regression line to the decoding
+accuracies obtained for the first, second, and third thirds of the dataset
+(corresponding to the left, middle, and right columns of Panel A,
+respectively). The $y$-axis denotes the slopes of the regression lines. All
+error ribbons denote bootstrap-estimated 95\% confidence intervals.}
+
+\label{fig:inflection-thirds}
+
 \end{figure}

### 2023-02-27 paper/boneyard.tex
finished integrating figures, including drafts of all captions.  also did some basic cleanup.

@@
+- We trained classifiers using more and more principle components to decode, and compared across condi- tions with varying degrees of cognitive richness.
+-We found that as listening conditions become more cognitively rich,
+decoding accuracy increased.
+-Also, decoding accuracy increased as understanding of the narrative
+accumulated over time, in more complex listening conditions.
+- Decoding accuracy also increased in higher cognitive areas, in more complex listening conditions.
+-We found that as story listening conditions become more complex, more
+components are required to decode.
+-We also found we could decode better with more impoverished data when
+there is the underlying structure of the narrative providing more cognitive richness.
+-We posit that as the complexity of our thoughts increases, neural
+compression decreases. However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
+
+Based on prior work ~\citep{Deme19} and following the direction of the field ~\citep{Turk13} we think our thoughts might be encoded in
+dynamic network patterns, and possibly higher order network
+patterns (Fig.~\ref{fig:direction_of_field}). We sought to test this hypothesis by developing an approach
+to inferring high-order network dynamics from timeseries data.
+
+One challenge in studying dynamic interactions is the
+computational resources required to calculate higher-order correlations. 
+We developed a computationally tractable model of network dynamics (Fig.~\ref{fig:methods_fig}) that takes in a feature
+timeseries and outputs approximated first-order dynamics (i.e.,
+dynamic functional correlations), second-order dynamics
+(reflecting homologous networks that dynamically form and disperse),
+and higher-order network dynamics (up to tenth-order dynamic
+correlations).
+
+We first validated our model using synthetic data, and explored how
+recovery varied with different underlying data structures and kernels.   We then 
+applied the approach to an fMRI dataset
+~\citep{SimoEtal16} in which participants listened to an audio
+recording of a story, as well as scrambled versions of the same story
+(where the scrambling was applied at different temporal scales).  We
+trained classifiers to take the output of the model and decode the
+timepoint in the story (or scrambled story) that the participants were
+listening to. We found that, during the intact listening condition in the
+experiment, classifiers that incorporated higher-order correlations
+yielded consistently higher accuracy than classifiers trained only on
+lower-order patterns (Fig.~\ref{fig:decoding_level},  a.\&d.).  By contrast, these
+higher-order correlations were not necessary to support decoding the other
+listening conditions and (minimally
+above chance) during a control rest condition.  This suggests
+that the cognitive processing that supported the most cogntively rich listening conditions
+involved second-order (or higher) network dynamics.
+
+Although we found decoding accuracy was best when incorporating
+higher-order network dynamics for all but rest
+  condition, it is unclear if this is a product of the brain or the
+  data collection technique.  It could be that the brain is
+  second-order or it could be that fMRI can
+  only reliably give second-order interactions. Exploring this method
+  with other data collection technique will be important to
+  disentangle this question.
+
+  \subsection*{Concluding remarks}
+
+How can we better understand how brain patterns change over
+time? How can we quantify the potential network dynamics that might be
+driving these changes? One way to judge the techniques of the future is
+to look at the trajectory of the fMRI field so far has taken so far
+(Fig.~\ref{fig:methods_fig}).  The field started with 
+univariate activation, measuring the average activity for each voxel.
+Analyses of multivariate activation followed, looking at spatial patterns of
+activity over voxels. Next, correlations of activity were explored, first
+with measures like resting connectivity that take temporal correlation
+between a seed voxel and all other voxels then with full connectivty
+that measure all pairwise correlations.  Additionally, this path of increasing
+complexity also moved from static to dynamic measurements.  One
+logical next step in this trajectory would be dynamic higher-order
+correlations. We have created a method 
+to support these calculations by scalably approximating dynamic higher-order
+correlations.
+
 By training classifiers using more and more principle components to decode, and comparing across conditions with varying degrees of cognitive richness, we can assess the explanatory power of the compressed data held with respect to the observed data (see \textit{Methods}).
@@
   with higher cognitive areas. (Fig.~\ref{fig:decode_pcs_network}).
+
+We were also curious how compression would change across time.
+If, there is some understanding of the narrative that accumulates over time, we should be able to see that difference.
+We found increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
+We also found decreases in decoding accuracy for the word-scrambled and rest
+condition.
+
+Overall, we found that as story listening conditions become more complex, more components are required to decode. We also found we could decode better with more impoverished data when
+there is the underlying structure of the narrative providing more cognitive richness.
+We posit that as the complexity of our thoughts increases, neural compression decreases.
+However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
+
+% - Increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
+% - Decreases in decoding accuracy for the word-scrambled and rest
+% condition.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2023-02-27 paper/main.tex
finished integrating figures, including drafts of all captions.  also did some basic cleanup.

@@
 \linenumbers
+
+\newcommand{\synthThirds}{S1}
 
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{figs/information_and_compression_over_time.pdf}
+  \includegraphics[width=0.8\textwidth]{figs/information_and_compression_over_time}
 
@@
 
-We were also curious how compression would change across time.
-If, there is some understanding of the narrative that accumulates over time, we should be able to see that difference.
-We found increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
-We also found decreases in decoding accuracy for the word-scrambled and rest
-condition.
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{figs/pca_neurosynth}
 
-Overall, we found that as story listening conditions become more complex, more components are required to decode. We also found we could decode better with more impoverished data when
-there is the underlying structure of the narrative providing more cognitive richness.
-We posit that as the complexity of our thoughts increases, neural compression decreases.
-However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
+\caption{\textbf{Top terms associated with the highest-weighted components by
+condition.} Each row corresponds to an experimental condition, and the colors
+correspond to the component number (ranked by proportion of variance
+explained). The inflated brain plots display the spatial distributions of each
+components' hubs (see \textit{Topographic Factor Analysis}). The lists on the
+right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
+each components' brain map. Analogous maps computed separately for each story
+segment may be found in Figure~\synthThirds.}
 
-% - Increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
-% - Decreases in decoding accuracy for the word-scrambled and rest
-% condition.
+\label{fig:neurosynth-pca}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{figs/pca_changes_neurosynth}
+
+\caption{\textbf{Changes in decoding accuracy and compression over time.} Each
+row corresponds to an experimental condition, and the colors correspond to
+change direction (positive: red; negative: blue). The saturation denotes the
+component number (columns on the right).  The lists on the
+right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
+each components' brain map.}
+
+\label{fig:neurosynth-changes}
+
+\end{figure}
+
 \section*{Discussion}
 
-- We trained classifiers using more and more principle components to decode, and compared across condi- tions with varying degrees of cognitive richness.
--We found that as listening conditions become more cognitively rich,
-decoding accuracy increased.
--Also, decoding accuracy increased as understanding of the narrative
-accumulated over time, in more complex listening conditions.
-- Decoding accuracy also increased in higher cognitive areas, in more complex listening conditions.
--We found that as story listening conditions become more complex, more
-components are required to decode.
--We also found we could decode better with more impoverished data when
-there is the underlying structure of the narrative providing more cognitive richness.
--We posit that as the complexity of our thoughts increases, neural
-compression decreases. However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
-
-Based on prior work ~\citep{Deme19} and following the direction of the field ~\citep{Turk13} we think our thoughts might be encoded in
-dynamic network patterns, and possibly higher order network
-patterns (Fig.~\ref{fig:direction_of_field}). We sought to test this hypothesis by developing an approach
-to inferring high-order network dynamics from timeseries data.
-
-One challenge in studying dynamic interactions is the
-computational resources required to calculate higher-order correlations. 
-We developed a computationally tractable model of network dynamics (Fig.~\ref{fig:methods_fig}) that takes in a feature
-timeseries and outputs approximated first-order dynamics (i.e.,
-dynamic functional correlations), second-order dynamics
-(reflecting homologous networks that dynamically form and disperse),
-and higher-order network dynamics (up to tenth-order dynamic
-correlations).
-
-We first validated our model using synthetic data, and explored how
-recovery varied with different underlying data structures and kernels.   We then 
-applied the approach to an fMRI dataset
-~\citep{SimoEtal16} in which participants listened to an audio
-recording of a story, as well as scrambled versions of the same story
-(where the scrambling was applied at different temporal scales).  We
-trained classifiers to take the output of the model and decode the
-timepoint in the story (or scrambled story) that the participants were
-listening to. We found that, during the intact listening condition in the
-experiment, classifiers that incorporated higher-order correlations
-yielded consistently higher accuracy than classifiers trained only on
-lower-order patterns (Fig.~\ref{fig:decoding_level},  a.\&d.).  By contrast, these
-higher-order correlations were not necessary to support decoding the other
-listening conditions and (minimally
-above chance) during a control rest condition.  This suggests
-that the cognitive processing that supported the most cogntively rich listening conditions
-involved second-order (or higher) network dynamics.
-
-Although we found decoding accuracy was best when incorporating
-higher-order network dynamics for all but rest
-  condition, it is unclear if this is a product of the brain or the
-  data collection technique.  It could be that the brain is
-  second-order or it could be that fMRI can
-  only reliably give second-order interactions. Exploring this method
-  with other data collection technique will be important to
-  disentangle this question.
-
-  \subsection*{Concluding remarks}
-
-How can we better understand how brain patterns change over
-time? How can we quantify the potential network dynamics that might be
-driving these changes? One way to judge the techniques of the future is
-to look at the trajectory of the fMRI field so far has taken so far
-(Fig.~\ref{fig:methods_fig}).  The field started with 
-univariate activation, measuring the average activity for each voxel.
-Analyses of multivariate activation followed, looking at spatial patterns of
-activity over voxels. Next, correlations of activity were explored, first
-with measures like resting connectivity that take temporal correlation
-between a seed voxel and all other voxels then with full connectivty
-that measure all pairwise correlations.  Additionally, this path of increasing
-complexity also moved from static to dynamic measurements.  One
-logical next step in this trajectory would be dynamic higher-order
-correlations. We have created a method 
-to support these calculations by scalably approximating dynamic higher-order
-correlations.
+\section*{Methods}
 
 \section*{Acknowledgements}
-We acknowledge discussions with Luke Chang, Hany Farid, Paxton
-Fitzpatrick, Andrew Heusser, Eshin Jolly, Qiang Liu, Matthijs van der
-Meer, Judith Mildner, Gina Notaro, Stephen Satterthwaite, Emily
-Whitaker, Weizhen Xie, and Kirsten Ziman. Our work was supported in
-part by NSF EPSCoR Award Number 1632738 to J.R.M. and by a sub-award
-of DARPA RAM Cooperative Agreement N66001-14-2-4-032 to J.R.M.  The
-content is solely the responsibility of the authors and does not
-necessarily represent the official views of our supporting
-organizations.
+
+We acknowledge discussions with Rick Betzel, Emily Finn, and Jim Haxby. Our
+work was supported in part by NSF CAREER Award Number 2145172 to J.R.M. The
+content is solely the responsibility of the authors and does not necessarily
+represent the official views of our supporting organizations. The funders had
+no role in study design, data collection and analysis, decision to publish, or
+preparation of the manuscript.
 
 \section*{Author contributions}
-Concept: J.R.M. and L.L.W.O. Implementation: L.L.W.O., and
-J.R.M.  Analyses: L.L.W.O and J.R.M.
+
+Conceptulization: J.R.M. and L.L.W.O. Methodology: J.R.M. and L.L.W.O.
+Implementation: L.L.W.O. Analysis: L.L.W.O. Writing, Reviewing, and Editing:
+J.R.M. and L.L.W.O. Supervision: J.R.M.
 

### 2023-02-27 paper/boneyard.tex
finished draft of methods and supplemental materials

@@
-- We trained classifiers using more and more principle components to decode, and compared across condi- tions with varying degrees of cognitive richness.
--We found that as listening conditions become more cognitively rich,
-decoding accuracy increased.
+- We trained classifiers using more and more principle components to decode,
+and compared across condi- tions with varying degrees of cognitive richness.
+
+-We found that as listening conditions become more cognitively rich, decoding
+accuracy increased.
+
 -Also, decoding accuracy increased as understanding of the narrative
 accumulated over time, in more complex listening conditions.
-- Decoding accuracy also increased in higher cognitive areas, in more complex listening conditions.
--We found that as story listening conditions become more complex, more
-components are required to decode.
--We also found we could decode better with more impoverished data when
-there is the underlying structure of the narrative providing more cognitive richness.
--We posit that as the complexity of our thoughts increases, neural
-compression decreases. However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
 
-Based on prior work ~\citep{Deme19} and following the direction of the field ~\citep{Turk13} we think our thoughts might be encoded in
-dynamic network patterns, and possibly higher order network
-patterns (Fig.~\ref{fig:direction_of_field}). We sought to test this hypothesis by developing an approach
-to inferring high-order network dynamics from timeseries data.
+- Decoding accuracy also increased in higher cognitive areas, in more complex
+listening conditions. -We found that as story listening conditions become more
+complex, more components are required to decode.
 
-One challenge in studying dynamic interactions is the
-computational resources required to calculate higher-order correlations. 
-We developed a computationally tractable model of network dynamics (Fig.~\ref{fig:methods_fig}) that takes in a feature
-timeseries and outputs approximated first-order dynamics (i.e.,
-dynamic functional correlations), second-order dynamics
-(reflecting homologous networks that dynamically form and disperse),
-and higher-order network dynamics (up to tenth-order dynamic
-correlations).
+-We also found we could decode better with more impoverished data when there is
+the underlying structure of the narrative providing more cognitive richness.
 
-We first validated our model using synthetic data, and explored how
-recovery varied with different underlying data structures and kernels.   We then 
-applied the approach to an fMRI dataset
-~\citep{SimoEtal16} in which participants listened to an audio
-recording of a story, as well as scrambled versions of the same story
-(where the scrambling was applied at different temporal scales).  We
-trained classifiers to take the output of the model and decode the
-timepoint in the story (or scrambled story) that the participants were
-listening to. We found that, during the intact listening condition in the
-experiment, classifiers that incorporated higher-order correlations
-yielded consistently higher accuracy than classifiers trained only on
-lower-order patterns (Fig.~\ref{fig:decoding_level},  a.\&d.).  By contrast, these
-higher-order correlations were not necessary to support decoding the other
-listening conditions and (minimally
-above chance) during a control rest condition.  This suggests
-that the cognitive processing that supported the most cogntively rich listening conditions
-involved second-order (or higher) network dynamics.
+-We posit that as the complexity of our thoughts increases, neural compression
+decreases. However, as our thoughts become deeper and richer, more reliable
+information is available at higher neural compression.
 
-Although we found decoding accuracy was best when incorporating
-higher-order network dynamics for all but rest
-  condition, it is unclear if this is a product of the brain or the
-  data collection technique.  It could be that the brain is
-  second-order or it could be that fMRI can
-  only reliably give second-order interactions. Exploring this method
-  with other data collection technique will be important to
-  disentangle this question.
+Based on prior work ~\citep{Deme19} and following the direction of the field
+~\citep{Turk13} we think our thoughts might be encoded in dynamic network
+patterns, and possibly higher order network patterns
+(Fig.~\ref{fig:direction_of_field}). We sought to test this hypothesis by
+developing an approach to inferring high-order network dynamics from timeseries
+data.
+
+One challenge in studying dynamic interactions is the computational resources
+required to calculate higher-order correlations. We developed a computationally
+tractable model of network dynamics (Fig.~\ref{fig:methods_fig}) that takes in
+a feature timeseries and outputs approximated first-order dynamics (i.e.,
+dynamic functional correlations), second-order dynamics (reflecting homologous
+networks that dynamically form and disperse), and higher-order network dynamics
+(up to tenth-order dynamic correlations).
+
+We first validated our model using synthetic data, and explored how recovery
+varied with different underlying data structures and kernels. We then applied
+the approach to an fMRI dataset ~\citep{SimoEtal16} in which participants
+listened to an audio recording of a story, as well as scrambled versions of the
+same story (where the scrambling was applied at different temporal scales). We
+trained classifiers to take the output of the model and decode the timepoint in
+the story (or scrambled story) that the participants were listening to. We
+found that, during the intact listening condition in the experiment,
+classifiers that incorporated higher-order correlations yielded consistently
+higher accuracy than classifiers trained only on lower-order patterns
+(Fig.~\ref{fig:decoding_level}, a.\&d.). By contrast, these higher-order
+correlations were not necessary to support decoding the other listening
+conditions and (minimally above chance) during a control rest condition. This
+suggests that the cognitive processing that supported the most cogntively rich
+listening conditions involved second-order (or higher) network dynamics.
+
+Although we found decoding accuracy was best when incorporating higher-order
+network dynamics for all but rest condition, it is unclear if this is a product
+of the brain or the data collection technique. It could be that the brain is
+second-order or it could be that fMRI can only reliably give second-order
+interactions. Exploring this method with other data collection technique will
+be important to disentangle this question.
 
@@
 
-By training classifiers using more and more principle components to decode, and comparing across conditions with varying degrees of cognitive richness, we can assess the explanatory power of the compressed data held with respect to the observed data (see \textit{Methods}).
-We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.
+By training classifiers using more and more principle components to decode, and
+comparing across conditions with varying degrees of cognitive richness, we can
+assess the explanatory power of the compressed data held with respect to the
+observed data (see \textit{Methods}). We note that our primary goal was not to
+achieve perfect decoding accuracy, but rather to use decoding accuracy as a
+benchmark for assessing whether different neural features specifically capture
+cognitively relevant brain patterns.
 
-\begin{figure}
-  \centering
+\begin{figure} \centering
 \includegraphics[width=\textwidth]{figs/decode_interpret.pdf}
-  \caption{\textbf{Decoding accuracy.} \textbf{a. Decoding accuracy by
-      number of components.} Ribbons of each color display
-    cross-validated decoding performance for each condition (intact,
-    paragraph, word, and rest). Decoders were trained using
-    increasingly more principle components and displayed relative to
-    chance (red line). \textbf{b. Fixed decoding accuracy by number of
-      components.} We zoom in on the plot shown in \textbf{a.} and add
-    a line denoting fixed decoding accuracy (.05). We plot where the
-    intact, paragraph, and word conditions intersect.
-    \textbf{c.
-Explanation of inflection metric.} First the we fit a sigmoid function to the decoding accuracy by number of components.
-Second, we found where the second derivative is both positive and less than .0001.
-Last, we then plot that inflection point as a single metric to capture the slope and asymptote of the curve.}
-    \label{fig:decode_interpret}
-  \end{figure}
+\caption{\textbf{Decoding accuracy.} \textbf{a. Decoding accuracy by number of
+components.} Ribbons of each color display cross-validated decoding performance
+for each condition (intact, paragraph, word, and rest). Decoders were trained
+using increasingly more principle components and displayed relative to chance
+(red line). \textbf{b. Fixed decoding accuracy by number of components.} We
+zoom in on the plot shown in \textbf{a.} and add a line denoting fixed decoding
+accuracy (.05). We plot where the intact, paragraph, and word conditions
+intersect. \textbf{c. Explanation of inflection metric.} First the we fit a
+sigmoid function to the decoding accuracy by number of components. Second, we
+found where the second derivative is both positive and less than .0001. Last,
+we then plot that inflection point as a single metric to capture the slope and
+asymptote of the curve.} \label{fig:decode_interpret} \end{figure}
 
-  We also wondered how this compression would change across brain regions.
-We repeated the analysis but limited the brain hubs to 7 networks using the Yeo et al.
-(2011) network parcellation shown here in the inflated brain (Fig.~\ref{fig:decode_pcs_network}, d.).
-We found that as complexity of the stimuli increases, decoding accuracy increases
-  with higher cognitive areas. (Fig.~\ref{fig:decode_pcs_network}).
+  We also wondered how this compression would change across brain regions. We
+  repeated the analysis but limited the brain hubs to 7 networks using the Yeo
+  et al. (2011) network parcellation shown here in the inflated brain
+  (Fig.~\ref{fig:decode_pcs_network}, d.). We found that as complexity of the
+  stimuli increases, decoding accuracy increases with higher cognitive areas.
+  (Fig.~\ref{fig:decode_pcs_network}).
 
-We were also curious how compression would change across time.
-If, there is some understanding of the narrative that accumulates over time, we should be able to see that difference.
-We found increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
+We were also curious how compression would change across time. If, there is
+some understanding of the narrative that accumulates over time, we should be
+able to see that difference. We found increases in decoding accuracy with the
+same number or fewer components for more complex, cognitively rich, conditions.
 We also found decreases in decoding accuracy for the word-scrambled and rest
@@
 
-Overall, we found that as story listening conditions become more complex, more components are required to decode. We also found we could decode better with more impoverished data when
-there is the underlying structure of the narrative providing more cognitive richness.
-We posit that as the complexity of our thoughts increases, neural compression decreases.
-However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
+Overall, we found that as story listening conditions become more complex, more
+components are required to decode. We also found we could decode better with
+more impoverished data when there is the underlying structure of the narrative
+providing more cognitive richness. We posit that as the complexity of our
+thoughts increases, neural compression decreases. However, as our thoughts
+become deeper and richer, more reliable information is available at higher
+neural compression.
 
-% - Increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
-% - Decreases in decoding accuracy for the word-scrambled and rest
-% condition.
+% - Increases in decoding accuracy with the same number or fewer components for
+more complex, cognitively rich, conditions. % - Decreases in decoding accuracy
+for the word-scrambled and rest % condition.
 
@@
 
-  %  If, there is some understanding of the narrative that accumulates over time, we should be able to see that change.
+  % If, there is some understanding of the narrative that accumulates over
+  time, we should be able to see that change.
 
-% We performed a decoding analysis, using cross validation to estimate
-% (using other participants’ data) which parts of the story the additional added
-% principle component corresponded to (see \textit{Materials and methods}).
-We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.
+% We performed a decoding analysis, using cross validation to estimate % (using
+other participants’ data) which parts of the story the additional added %
+principle component corresponded to (see \textit{Materials and methods}). We
+note that our primary goal was not to achieve perfect decoding accuracy, but
+rather to use decoding accuracy as a benchmark for assessing whether different
+neural features specifically capture cognitively relevant brain patterns.
 
-% Separately for each experimental condition, we divided participants
-% into two groups. Starting with 1 principle comonent, for
-% each dimension we added another principle component, and we correlated the group 1 activity patterns with group 2
-% activity patterns.  There were 272
-% timepoints for paragraph condition, 300 timepoints for intact and word
-% conditions, and 400 timepoints for rest condition,  so chance
-% performance on this decoding test is was $\frac{1}{272}$,
-% $\frac{1}{300}$, and $\frac{1}{400}$ respectively.
+% Separately for each experimental condition, we divided participants % into
+two groups. Starting with 1 principle comonent, for % each dimension we added
+another principle component, and we correlated the group 1 activity patterns
+with group 2 % activity patterns. There were 272 % timepoints for paragraph
+condition, 300 timepoints for intact and word % conditions, and 400 timepoints
+for rest condition, so chance % performance on this decoding test is was
+$\frac{1}{272}$, % $\frac{1}{300}$, and $\frac{1}{400}$ respectively.
 
-% - As complexity of the stimulus increases, decoding accuracy
-% increases (Fig.~\ref{fig:decode_interpret},  a.).  Replication of
-% Simony findings.
+% - As complexity of the stimulus increases, decoding accuracy % increases
+(Fig.~\ref{fig:decode_interpret}, a.). Replication of % Simony findings.
 
-% - As complexity of the stimuli increases, we need fewer components to
-% decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
+% - As complexity of the stimuli increases, we need fewer components to %
+decode the same amount (Fig.~\ref{fig:decode_interpret}, b.).
 
-% - As complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
+% - As complexity of the stimuli increases, more components are required to
+reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret}, c.).
 
-Prior work has shown participants share similar neural responses to
-richly structured stimuli when compared to stimuli with less
-structure \cite{SimoEtal16}.  We replicate this finding, showing as complexity of the stimulus increases, decoding accuracy
-increases (Fig.~\ref{fig:decode_interpret},  a.).  
-Additionally, we found that as complexity of the stimuli increases, we need fewer components to decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).
-However, we also found that as complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).
-We posit that as the complexity of our thoughts increases, neural
-compression decreases. However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
+Prior work has shown participants share similar neural responses to richly
+structured stimuli when compared to stimuli with less structure
+\cite{SimoEtal16}. We replicate this finding, showing as complexity of the
+stimulus increases, decoding accuracy increases
+(Fig.~\ref{fig:decode_interpret}, a.). Additionally, we found that as
+complexity of the stimuli increases, we need fewer components to decode the
+same amount (Fig.~\ref{fig:decode_interpret}, b.). However, we also found that
+as complexity of the stimuli increases, more components are required to reach
+peak decoding accuracy (Fig.~\ref{fig:decode_interpret}, c.). We posit that as
+the complexity of our thoughts increases, neural compression decreases.
+However, as our thoughts become deeper and richer, more reliable information is
+available at higher neural compression.
 
-% % PCA, or the closely related factor analysis and singular value decomposition (SVD) (Hastie et al., 2009), is widely used in the study of individual differences and aids estimating how many latent components, or “factors”, underlie a pattern of (item) responses within or across participants, as for instance in the context of intelligence (Spearman, 1904) or personality tests (Cattell, 1947).
-In the context of neuroimaging, PCA has been used to identify brain networks (Huth et al., 2012; Friston et al., 1993).
+% % PCA, or the closely related factor analysis and singular value
+decomposition (SVD) (Hastie et al., 2009), is widely used in the study of
+individual differences and aids estimating how many latent components, or
+“factors”, underlie a pattern of (item) responses within or across
+participants, as for instance in the context of intelligence (Spearman, 1904)
+or personality tests (Cattell, 1947). In the context of neuroimaging, PCA has
+been used to identify brain networks (Huth et al., 2012; Friston et al., 1993).
 
-%  % add citation:
-% % S.M.Smith,A.Hyva ̈rinen,G.Varoquaux,K.L.Miller,and C. F. Beckmann, “Group-PCA for very large fMRI datasets,” NeuroImage, vol. 101, pp. 738, 2014.
+% % add citation: % % S.M.Smith,A.Hyva ̈rinen,G.Varoquaux,K.L.Miller,and C. F.
+Beckmann, “Group-PCA for very large fMRI datasets,” NeuroImage, vol. 101, pp.
+738, 2014.
 
-% \subsection*{Evaluation metrics}
-%  We will evaluate the degree of
-% compression of held-out
-% neuroimaging data by assessing the time at which it was collected.  We will use this evaluation (timepoint
-% decoding) as a proxy for
-% gauging how much explanatory power the compressed data held
-% with respect to the observed data.
+% \subsection*{Evaluation metrics} % We will evaluate the degree of %
+compression of held-out % neuroimaging data by assessing the time at which it
+was collected. We will use this evaluation (timepoint % decoding) as a proxy
+for % gauging how much explanatory power the compressed data held % with
+respect to the observed data.
 

### 2023-02-27 paper/main.tex
finished draft of methods and supplemental materials

@@
 correspond to the component number (ranked by proportion of variance
-explained). The inflated brain plots display the spatial distributions of each
-components' hubs (see \textit{Topographic Factor Analysis}). The lists on the
-right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
+explained). The inflated brain plots display the top 20 highest-weighted hubs
+(see \textit{Topographic Factor Analysis}) for each components'. The lists on
+the right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
 each components' brain map. Analogous maps computed separately for each story
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/pca_changes_neurosynth}
-
-\caption{\textbf{Changes in decoding accuracy and compression over time.} Each
-row corresponds to an experimental condition, and the colors correspond to
-change direction (positive: red; negative: blue). The saturation denotes the
-component number (columns on the right).  The lists on the
-right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
-each components' brain map.}
-
-\label{fig:neurosynth-changes}
-
-\end{figure}
-
 \section*{Discussion}
@@
 \section*{Methods}
+
+We measured properties of recorded neuroimaging data under different task
+conditions that varied systematically in cognitive engagement and depth of
+processing. We were especially interested in how \textit{informative} and
+\textit{compressible} the activity patterns were under these different
+conditions (Fig.~\ref{fig:information-compression}).
+
+\subsection*{Functional neuroimaging data collected during story
+  listening}
+
+We examined an fMRI dataset collected by \cite{SimoEtal16} that the authors
+have made publicly available at
+\href{http://arks.princeton.edu/ark:/88435/dsp015d86p269k}{arks.princeton.edu/ark:/88435/dsp015d86p269k}.
+The dataset comprises neuroimaging data collected as participants listened to
+an audio recording of a story (intact condition; 36 participants), listened to
+temporally scrambled recordings of the same story (17 participants in the
+paragraph-scrambled condition listened to the paragraphs in a randomized order
+and 36 in the word-scrambled condition listened to the words in a randomized
+order), or lay resting with their eyes open in the scanner (rest condition; 36
+participants). Full neuroimaging details may be found in the original paper for
+which the data were collected~\cite{SimoEtal16}. Procedures were approved by
+the Princeton University Committee on Activities Involving Human Subjects, and
+by the Western Institutional Review Board (Puyallup, WA). All subjects were
+native English speakers with normal hearing and provided written informed
+consent.
+
+\subsection*{Hierarchical topographic factor analysis (HTFA)}
+
+Following our prior related work, we used HTFA~\cite{MannEtal18} to derive a
+compact representation of the neuroimaging data. In brief, this approach
+approximates the timeseries of voxel activations (44,415 voxels) using a much
+smaller number of radial basis function (RBF) nodes~\citep[in this case, 700
+nodes, as determined by an optimization procedure;][]{MannEtal18}). This
+provides a convenient representation for examining full-brain activity patterns
+and network dynamics. All of the analyses we carried out on the neuroimaging
+dataset were performed in this lower-dimensional space. In other words, each
+participant's data matrix, $\mathbf{X}$, was a number-of-timepoints ($T$) by 700
+matrix of HTFA-derived factor weights (where the row and column labels were
+matched across participants). Code for carrying out HTFA on fMRI data may be
+found as part of the BrainIAK toolbox~\cite{CapoEtal17, KumaEtal21}, which may
+be downloaded at \href{https://brainiak.org/}{brainiak.org}.
+
+\subsection*{Principal components analysis (PCA)}
+
+We applied group PCA~\citep{SmitEtal14} separately to the HTFA-derived
+representations of the data (i.e., factor loadings) from each experimental
+condition. Specifically, for each condition, we considered the set of all
+participants' $T$ by 700 factor weight matrices. We used group PCA to project
+these 700-dimensional matrices into a series of $k$-dimensional spaces, for $k
+\in \{3, 4, 5, ..., 700\}$. This yielded a set of number-of-participants
+matrices, each with $T$ rows and $k$ columns.
+
+\subsection*{Temporal decoding}
+
+We sought to identify neural patterns that reflected participants' ongoing
+cognitive processing of incoming stimulus information. As reviewed by
+\cite{SimoEtal16}, one way of homing in on these stimulus-driven neural
+patterns is to compare activity patterns across individuals. In particular,
+neural patterns will be similar across individuals to the extent that the
+neural patterns under consideration are stimulus-driven, and to the extent that
+the corresponding cognitive representations are reflected in similar spatial
+patterns across people~\cite{SimoChan20}. Following this logic, we used an
+across-participant temporal decoding test developed by \cite{MannEtal18} to
+assess the degree to which different neural patterns reflected ongoing
+stimulus-driven cognitive processing across people. The approach entails using
+a subset of the data to train a classifier to decode stimulus timepoints (i.e.,
+moments in the story participants listened to) from neural patterns. We use
+decoding (forward inference) accuracy on held-out data, from held-out
+participants, as a proxy for the extent to which the inputted neural patterns
+reflected stimulus-driven cognitive processing in a similar way across
+individuals.
+
+\subsubsection*{Forward inference and decoding accuracy}
+
+We used an across-participant correlation-based classifier to decode which
+stimulus timepoint matched each timepoint's neural pattern. For a given value
+of $k$ (i.e., number of principal components), we first used group PCA to
+project the data from each condition into a $k$-dimensional space. Next, we
+divided the participants into two groups: a template group,
+$\mathcal{G}_{\mathrm{template}}$ (i.e., training data), and a to-be-decoded
+group, $\mathcal{G}_{\mathrm{decode}}$ (i.e., test data). We averaged the
+projected data within each group to obtain a single $T$ by $k$ matrix for each
+group. Next, we correlated the rows of the two averaged matrices to form a $T$
+by $T$ decoding matrix, $\mathbf{\Lambda}$. In this way, the rows of
+$\mathbf{\Lambda}$ reflected timepoints from the template group, while the
+columns reflected timepoints from the to-be-decoded group. We used
+$\mathbf{\Lambda}$ to assign temporal labels to each timepoint (row) from the
+test group's matrix, using the row of the training group's matrix with which it
+was most highly correlated. We repeated this decoding procedure, but using
+$\mathcal{G}_{\mathrm{decode}}$ as the template group and
+$\mathcal{G}_{\mathrm{template}}$ as the to-be-decoded group. Given the true
+timepoint labels (for each group), we defined the decoding accuracy as the
+average proportion of correctly decoded timepoints, across both groups (where
+chance perfomance is $\frac{1}{T}$). In Figures~\ref{fig:inflection}
+and~\ref{fig:inflection-thirds} we report the decoding accuracy for each
+condition and value of $k$, averaged across $n = 1,000$ cross validation folds.
+
+\subsection*{Reverse inference}
+
+To help interpret the brain activity patterns we found within the contexts of
+other studies, we created summary maps of each principal component, for each
+experimental condition, by summing together the 20 HTFA-derived RBF nodes (see
+\textit{Hierarchical Topographic Factor Analysis}) with the highest absolute
+value weights for each of the top 5 components (Figs.~\ref{fig:neurosynth-pca},
+\synthThirds). We then carried out a meta analysis using
+Neurosynth~\cite{RubiEtal17} to identify the 5 terms most commonly associated
+with the given map.
+
+\section*{Data and code availability}
+
+All of the code used to produce the figures and results in this manuscript,
+along with links to the corresponding datasets, may be found at 
+\href{https://github.com/ContextLab/pca_paper}{github.com/ContextLab/pca\_paper}.
 

### 2023-02-27 paper/supplement.tex
finished draft of methods and supplemental materials

--- 2023-02-27 paper/main.tex
@@
 
-\newcommand{\synthThirds}{S1}
+\newcommand{\synth}{1}
+\renewcommand{\thefigure}{S\arabic{figure}}
 
-\title{High-order cognition is supported by complex but compressible brain activity patterns}
+\title{\textit{Supplemental materials for:} High-order cognition is supported by complex but compressible brain activity patterns}
 
@@
 
-\begin{abstract}
-
-We applied dimensionality reduction algorithms and pattern classifiers to
-functional neuroimaging data collected as participants listened to a story,
-temporally scrambled versions of the story, or underwent a resting state
-scanning session. These experimental conditions were intended to require
-different depths of processing and inspire different levels of cognitive
-engagement. We considered two primary aspects of the data. First, we treated
-the number of features (components) required to achieve a threshold decoding
-accuracy as a proxy for the ``compressibility'' of the neural patterns (where
-fewer components indicate greater compression). Second, we treated the
-maximum achievable decoding accuracy across participants as an indicator of the
-``stability'' of the recorded patterns. Overall, we found that neural patterns
-recorded as participants listened to the intact story required fewer features
-to achieve comparable classification accuracy to the other experimental
-conditions. However, the peak decoding accuracy (achievable with more features)
-was also highest during intact story listening. Taken together, our work
-suggests that our brain networks flexibly reconfigure according to ongoing task
-demands, and that the activity patterns associated with higher-order cognition
-and high engagement are both more complex and more compressible than the
-activity patterns associated with lower-order tasks and lower levels of
-engagement.
-
-\end{abstract}
-
-\doublespacing
-
-\section*{Introduction}
-
-Large-scale networks, including the human brain, may be conceptualized as
-occupying one or more positions along on a continuum. At one extreme, every
-node is fully independent of every other node. At the other extreme, all nodes
-behave identically. Each extreme optimizes key properties of how the network
-functions. When every node is independent, the network is maximally
-\textit{expressive}: if we define the network's ``state'' as the activity
-pattern across its nodes, then every state is equally reachable by a network
-with fully independent nodes. On the other hand, a network of identically
-behaved nodes optimizes \textit{robustness}: any subset of nodes may be removed
-from the network without any loss of function or expressive power, as long as
-any single node remains. Presumably, most natural systems tend to occupy
-positions between these extremes. We wondered: might the human brain
-reconfigure itself to be more flexible or more robust according to ongoing
-demands? In other words, might the brain reconfigure its connections or
-behaviors under different circumstances to change its position along this
-continuum?
-
-% compression (redundancy) vs. complexity -- appear to balance each other, butnot necessarily.
-
-Closely related to the above notions of expressiveness versus robustness are
-measures of how much \textit{information} is contained in a given signal or
-pattern, and how \textit{redundant} a signal is~\citep{Shan48}. Formally,
-information is defined as the amount of uncertainty about a given variables'
-outcomes (i.e., entropy), measured in \textit{bits}, or the optimal number of
-yes/no questions needed to reduce uncertainty about the variable's outcomes to
-zero. Highly complex systems with many degrees of freedom (i.e., high
-flexibility and expressiveness), are more information-rich than simpler or more
-constrained systems. The redundancy of a signal denotes the difference how
-expressive the signal \textit{could} be (i.e., proportional to the number of
-unique states or symbols used to transmit the signal) and the actual
-information rate (i.e., the entropy of each individual state or symbol). If a
-brain network's nodes are fully independent, then the number of bits required
-to express a single activity pattern is proportional to the number of nodes.
-The network would also be minimally redundant, since the status of every node
-would be needed to fully express a single brain activity pattern. If a brain
-network's nodes are fully coupled and identical, then the number of bits
-required to express a single activity pattern is proportional to the number of
-unique states or values any individual node can take on. Such a network would
-be highly redundant, since knowing any individual node's state would be
-sufficient to recover the full-brain activity pattern. Highly redundant systems
-are also robust, since there is little information loss from losing any given
-observation.
-
-We take as a given that brain activity is highly flexible: our brains can
-exhibit nearly infinite activity patterns. This flexibility implies that our
-brains activity patterns are highly information rich. However, brain activity
-patterns are also highly structured. For example, full-brain correlation
-matrices are stable within~\citep{FinnEtal15, FinnEtal17, GratEtal18} and
-across~\citep{YeoEtal11, GlerEtal12, GratEtal18, ColeEtal14} individuals. This
-stability suggests that our brains' activity patterns are at least partially
-constrained, for example by anatomical, external, or internal factors.
-Constraints on brain activity that limit its flexibility decrease
-expressiveness (i.e., its information rate). However, constraints on brain
-activity also increase its robustness to noise (e.g., ``missing'' or corrupted
-signals may be partially recovered). For example, recent work has shown that
-full-brain activity patterns may be reliably recovered from only a relatively
-small number of implanted electrodes~\citep{OwenEtal20, ScanEtal21}. This
-robustness property suggests that the relevant signal (e.g., underlying factors
-that have some influence over brain activity patterns) are compressible.
-
-\begin{figure}[tp]
-
-  \centering
-  \includegraphics[width=\textwidth]{figs/information_and_compressibility}
-
-  \caption{\textbf{Information content and compressibility.} \textbf{A. Variance
-  explained for two images.} We applied principal components analysis to a
-  photograph and drawing, treating each row of the images as ``observations.''
-  Across columns, we identified the number of components required to explain
-  100\%, 95\%, 75\%, or 50\% of the cumulative variance in each image (the 100\%
-  columns denote the original images). The numbers of components are indicate in
-  parentheses, and the resulting ``compressed'' images are displayed. \textbf{B.
-  Representing two images with different numbers of components.} Using the same
-  principal component decompositions as in Panel A, we computed the cumulative
-  proportion of variance explained with 380 (original images), 20, 4, or 2
-  components. \textbf{C. Cumulative variance explained versus number of
-  components.} For the images displayed in Panels A and B, we plot the cumulative
-  proportion of variance explained as a function of the number of components used
-  to represent each image. \textbf{D. Information rate and compressibility.}
-  Across multiple images, the information rate (i.e., the amount of information
-  contained in each image; horizontal axis) is high when each individual pixel
-  provides information that cannot be inferred from other pixels.
-  High-information rate images tend to be high-resolution, and low-information
-  rate images tend to be low-resolution. Compressibility is related to the
-  difference between the information required to specify the original versus
-  compressed images (vertical axis). Highly compressible images often contain
-  predictable structure (redundancies) that can be leveraged to represent the
-  images much more efficiently than in their original feature spaces.}
-
-  \label{fig:information-compression} 
-  \end{figure}
-
-To the extent that brain activity patterns contain rich task-relevant
-information, we should be able to use the activity patterns to accurately
-differentiate between different aspects of the task~\citep[e.g., using pattern
-classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
-correspondence between classification accuracy and the information content of a
-signal~\citep{Alva02}. To the extent that brain activity patterns are
-compressible, we should be able to generate simplified (e.g., lower
-dimensional) representations of the data while still preserving the relevant or
-important aspects of the original signal. In general, information content and
-compressibility are related but are partially dissociable
-(Fig.~\ref{fig:information-compression}). If a given signal (e.g., a
-representation of brain activity patterns) contains more information about
-ongoing cognitive processes, then the peak decoding accuracy should be high.
-And if the signal is compressible, a low-dimensional embedding of the signal
-will be similarly informative to the original signal
-(Fig.~\ref{fig:information-compression}D).
-
-Several recent studies suggest that the complexity of brain activity is
-task-dependent, whereby simpler tasks with lower cognitive demands are
-reflected by simpler and more compressible brain activity patterns, and more
-complex tasks with higher cognitive demands are reflected by more complex and
-less compressible brain activity patterns~\citep{MackEtal20, OwenEtal21}. These
-findings complement other work suggesting that functional connectivity
-(correlation) patterns are task-dependent~\citep{FinnEtal17, OwenEtal20,
-ColeEtal14}, although see~\cite{GratEtal18}. Higher-order cognitive processing
-of a common stimulus also appears to drive more stereotyped task-related
-activity and functional connectivity across individuals~\citep{HassEtal08,
-LernEtal11, SimoChan20, SimoEtal16}.
-
-The above prior studies are consistent with two potential descriptions of how
-cognitive processes are reflects in brain activity patterns. One possibility is
-that the information rate of brain activity increases during more complex or
-higher-level cognitive processing. If so, then the ability to reliably decode
-cognitive states from brain activity patterns should improve with task
-complexity or with the level (or ``depth'') of cognitive processing. A second
-possibility is that the compressibility of brain activity patterns increases
-during more complex or higher-level cognitive processing. If so, then
-individual features of brain recordings, or compressed representations of brain
-recordings, should carry more information during complex or high-level (versus
-simple or low-level) cognitive tasks.
-
-We used a previously collected neuroimaging dataset to estimate the extent to
-which each of these two possibilities might hold. The dataset we examined
-comprised functional magnetic resonance imaging (fMRI) data collected as
-participants listened to an audio recording of a 10-minute story, temporally
-scrambled recordings of the story, or underwent a resting state
-scan~\citep{SimoEtal16}. Each of these experimental conditions evokes different
-depths of cognitive processing~\citep{SimoEtal16,LernEtal11,
-HassEtal08,OwenEtal21}. We used across-participant classifiers to decode
-listening times in each condition, as a proxy for how ``informative'' the
-task-specific activity patterns were~\citep{SimoChan20}. We also use principle
-components analysis to generate lower-dimensional representations of the
-activity patterns. We then repeated the classification analyses after
-preserving different numbers of components and examined how classification
-accuracy changed across the different experimental conditions.
-
-\section*{Results}
-
 \begin{figure}
   \centering
+  \includegraphics[width=\textwidth]{figs/pca_neurosynth_thirds}
 
-  \includegraphics[width=0.8\textwidth]{figs/decoding_and_inflection}
+\caption{\textbf{Top terms associated with the highest-weighted components by
+condition, broken down by story segment.} Each group of three rows corresponds
+to an experimental condition, and the colors correspond to the component number
+(ranked by proportion of variance explained). The rows' numbering denotes the
+story segment used to compute each map (1: first third; 2: second third; 3:
+third third). The inflated brain plots display the top 20 highest-weighted hubs
+(see \textit{Topographic Factor Analysis}) for each components. The lists on
+the right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
+each components' brain map. Analogous maps computed for the entire story may be
+found in Figure~\synth.}
 
-\caption{\textbf{Decoding accuracy and compression.} \textbf{A. Decoding
-accuracy by number of components.} Ribbons of each color display
-cross-validated decoding performance for each condition (intact, paragraph,
-word, and rest), as a function of the number of components (features) used to
-represent the data. The horizontal red line denotes chance performance, and the
-horizontal black line denotes 5\% decoding accuracy (used as a reference in
-Panel B). \textbf{B. Numbers of components required to reach a fixed decoding
-accuracy threshold, by condition.} The panel displays a zoomed-in view of the
-inset in Panel A. Intersections between each condition's decoding accuracy
-curve and the 5\% decoding accuracy reference line are marked by arrows.
-\textbf{C. Estimating inflection points.} We sought to identify an ``inflection
-point'' for each decoding curve, denoting the number of components at which the
-decoding curve asymptotes. We fit sigmoid functions to each decoding curve
-(left sub-panel) and then computed the minimum number of components where the
-second derivative of the sigmoid was both positive and less than a threshold
-value of 0.0001. \textbf{D. Inflection points by condition.} Each dot displays
-the number of components ($x$-axis) and decoding accuracy ($y$-axis) at one
-condition's inflection point. All error ribbons denote bootstrap-estimated 95\%
-confidence intervals.}
-
-    \label{fig:inflection}
-  \end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{figs/decode_pcs_network}
-
-  \caption{\textbf{Network-specific decoding accuracy and compression.}
-  \textbf{A. Decoding accuracy and number of components for network-specific
-  inflection points.} We considered the seven networks identified by
-  \cite{YeoEtal11}. We computed each network's inflection point, for each
-  experimental condition, using the procedure described in
-  Figure~\ref{fig:inflection}C. \textbf{B. Network-specific decoding
-  accuracy.} Each of the seven networks are colored according to the decoding
-  accuracy at the network's inflection point for the ``intact'' experimental
-  condition (corresponding to the dark blue curve in Panel A). \textbf{C.
-  Network-specific compression.} Each of the seven networks are colored
-  according to the number of components at the network's inflection point for
-  the intact experimental condition. Larger numbers of components reflect lower
-  compressibility.}
-
-  \label{fig:networks}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{figs/information_and_compression_over_time}
-
-\caption{\textbf{Changes in decoding accuracy and compression over time.}
-\textbf{A. Decoding accuracy by number of components, by story segment.} Each
-family of curves is plotted in the same format as Figure~\ref{fig:inflection}A
-but reflects data only from one third of the dataset. \textbf{B. Inflection
-points by condition and segment.} The dots re-plot the inflection points from
-Figure~\ref{fig:inflection}D for reference. The numbers denote the inflection
-points for each third of the dataset (1: first third; 2: second third; 3: third
-third; colors denote experimental conditions). \textbf{C. Change in decoding
-accuracy over time, by number of components.} For each number of components
-($x$-axis) and condition (color), we fit a regression line to the decoding
-accuracies obtained for the first, second, and third thirds of the dataset
-(corresponding to the left, middle, and right columns of Panel A,
-respectively). The $y$-axis denotes the slopes of the regression lines. All
-error ribbons denote bootstrap-estimated 95\% confidence intervals.}
-
-\label{fig:inflection-thirds}
+\label{fig:neurosynth-thirds}
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/pca_neurosynth}
-
-\caption{\textbf{Top terms associated with the highest-weighted components by
-condition.} Each row corresponds to an experimental condition, and the colors
-correspond to the component number (ranked by proportion of variance
-explained). The inflated brain plots display the spatial distributions of each
-components' hubs (see \textit{Topographic Factor Analysis}). The lists on the
-right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
-each components' brain map. Analogous maps computed separately for each story
-segment may be found in Figure~\synthThirds.}
-
-\label{fig:neurosynth-pca}
-
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/pca_changes_neurosynth}
-
-\caption{\textbf{Changes in decoding accuracy and compression over time.} Each
-row corresponds to an experimental condition, and the colors correspond to
-change direction (positive: red; negative: blue). The saturation denotes the
-component number (columns on the right).  The lists on the
-right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
-each components' brain map.}
-
-\label{fig:neurosynth-changes}
-
-\end{figure}
-
-\section*{Discussion}
-
-\section*{Methods}
-
-\section*{Acknowledgements}
-
-We acknowledge discussions with Rick Betzel, Emily Finn, and Jim Haxby. Our
-work was supported in part by NSF CAREER Award Number 2145172 to J.R.M. The
-content is solely the responsibility of the authors and does not necessarily
-represent the official views of our supporting organizations. The funders had
-no role in study design, data collection and analysis, decision to publish, or
-preparation of the manuscript.
-
-\section*{Author contributions}
-
-Conceptulization: J.R.M. and L.L.W.O. Methodology: J.R.M. and L.L.W.O.
-Implementation: L.L.W.O. Analysis: L.L.W.O. Writing, Reviewing, and Editing:
-J.R.M. and L.L.W.O. Supervision: J.R.M.
-
+\newpage
+\renewcommand{\refname}{Supplemental references}
 \bibliographystyle{apacite}

### 2023-02-28 paper/main.tex
typo fix in methods, added keywords, working on results section

@@
-\documentclass[english]{article}
+\documentclass[english, 11pt]{article}
 \usepackage{graphicx}
@@
 
+\bigskip
+\noindent
+\textbf{Keywords: information, compression, temporal decoding, dimensionality reduction, neuroimaging}
+
 \end{abstract}
@@
 continuum?
-
-% compression (redundancy) vs. complexity -- appear to balance each other, butnot necessarily.
 
@@
 
-The above prior studies are consistent with two potential descriptions of how
+The above studies are consistent with two potential descriptions of how
 cognitive processes are reflects in brain activity patterns. One possibility is
@@
 \section*{Results}
+
+We sought to understand whether higher-level cognition is reflected by more
+reliable and informative brain activity patterns, and how compressibility of
+brain activity patterns relates to cognitive complexity. We developed a
+computational framework for systematically assessing the informativeness and
+compressibility of brain activity patterns recorded under different cognitive
+circumstances. We used across-participant decoding accuracy (see
+\textit{Forward inference and decoding accuracy}) as a proxy for
+informativeness. To estimate the compressibility of the brain patterns, we used
+group principal components analysis (PCA) to project the brain patterns into
+$k$-dimensional spaces, for different values of $k$ (see \textit{Hierarchical
+Topographic Factor Analysis (HTFA)} and \textit{Principal components analysis
+(PCA)}). For more compressible brain patterns, decoding accuracy should be more
+robust to small values of $k$.
+
+We analyzed a dataset collected by \cite{SimoEtal16} that comprised four
+experimental conditions. These conditions exposed participants to stimuli that
+systematically varied in cognitive engagement. In the \textit{intact}
+experimental condition, participants listened to an audio recording of a
+10-minute story. In the \textit{paragraph}-scrambled experimental condition,
+partipants listened to a temporally scrambled version of the story, where the
+paragraphs occurred out of order, but where the same set of paragraphs was
+presented over the entire listening interval. All participants in this
+condition experienced the scrambled paragraphs in the same order. In the
+\textit{word}-scrambled experimental condition, participants listened to a
+temporally scrambled version of the story, where the words occurred in a random
+order. Again, all participants in this condition experienced the scrambled
+words in the same order. Finally, in the \textit{rest} experimental condition,
+participants lay in the scanner with no overt stimulus, while keeping their
+eyes open and blinking as needed. This public dataset provided a convenient
+means for testing our hypothesis that different levels of cognitive processing
+and engagement affect how informative and compressible the associated brain
+patterns are.
 
@@
 and~\ref{fig:inflection-thirds} we report the decoding accuracy for each
-condition and value of $k$, averaged across $n = 1,000$ cross validation folds.
+condition and value of $k$, averaged across $n = 100$ cross validation folds.
 

### 2023-02-28 paper/main.tex
added some notes
checking in results writeup for fig 2

@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{figs/decode_pcs_network}
+To evaluate the relation between informativeness and compression for brain
+activity from each experimental condition, we trained a series of
+across-participant temporal decoders on compressed representations of the data.
+Figure~\ref{fig:inflection}A displays the decoding accuracy as a function of
+the number of principal components used to represent the data. Several patterns
+were revealed by the analysis. First, in general (i.e., across experimental
+conditions), decoding accuracy improves as the number of components increases.
+However, decoding accuracy peaked at a higher level for experimental conditions
+that exposed participants to cognitively richer stimuli. The peak decoding
+accuracy was highest for the ``intact'' condition (versus paragraph: $t(XXX) =
+XXX, p = XXX$; versus word: $t(XXX) = XXX, p = XXX$; versus rest: $t(XXX) =
+XXX, p = XXX$), next highest for the ``paragraph'' condition (versus word:
+$t(XXX) = XXX, p = XXX$; versus rest: $t(XXX) = XXX, p = XXX$), and next
+highest for the ``word'' condition (versus rest: $t(XXX) = XXX, p = XXX$). 
+This ordering implies that cognitively richer conditions evoke more stable brain
+activity patterns across people.
 
-  \caption{\textbf{Network-specific decoding accuracy and compression.}
-  \textbf{A. Decoding accuracy and number of components for network-specific
-  inflection points.} We considered the seven networks identified by
-  \cite{YeoEtal11}. We computed each network's inflection point, for each
-  experimental condition, using the procedure described in
-  Figure~\ref{fig:inflection}C. \textbf{B. Network-specific decoding
-  accuracy.} Each of the seven networks are colored according to the decoding
-  accuracy at the network's inflection point for the ``intact'' experimental
-  condition (corresponding to the dark blue curve in Panel A). \textbf{C.
-  Network-specific compression.} Each of the seven networks are colored
-  according to the number of components at the network's inflection point for
-  the intact experimental condition. Larger numbers of components reflect lower
-  compressibility.}
+The cognitively richer conditions also displayed steeper initial slopes. For
+example, the intact condition decoders reached an arbitrarily chosen threshold
+of 5\% accuracy using fewer components than the paragraph condition decoders
+($t(XXX) = XXX, p = XXX$) or word condition decoders ($t(XXX) = XXX, p = XXX$),
+and decoding accuracy never exceeded 5\% for the rest condition. This suggests
+that brain activity patterns evoked by cognitively richer conditions are more
+compressible, such that representing the data using the same number of
+principal components provides more information to the temporal decoders
+(Fig.~\ref{fig:inflection}B).
 
-  \label{fig:networks}
-\end{figure}
+In every experimental condition, decoding accuracy appeared to asymptote (i.e.,
+hit an upper limit) beyond some characteristic number of components that
+differed across conditions. To quantify the ``inflection points'' at which the
+decoding curves in Figure~\ref{fig:inflection}A flattened out, we fit a sigmoid
+function to the average decoding curve for each condition. We defined the
+inflection point for each condition as the point on the fitted sigmoid where
+the second derivative was both positive and less than a threshold value of
+0.0001 (i.e., approaching 0 from the right). These inflection points reflect a
+``balance'' between higher decoding accuracy (which tends to be better when
+more components are used) and compression (which is better for fewer
+components). Plotting each condition's inflection point
+(Fig.~\ref{fig:inflection}D) reveals that both the number of components and the
+decoding accuracy at each inflection point increase systematically across
+conditions in proportion to cognitive richness.
 
@@
 
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figs/decode_pcs_network}
+
+  \caption{\textbf{Network-specific decoding accuracy and compression.}
+  \textbf{A. Decoding accuracy and number of components for network-specific
+  inflection points.} We considered the seven networks identified by
+  \cite{YeoEtal11}. We computed each network's inflection point, for each
+  experimental condition, using the procedure described in
+  Figure~\ref{fig:inflection}C. \textbf{B. Network-specific decoding
+  accuracy.} Each of the seven networks are colored according to the decoding
+  accuracy at the network's inflection point for the ``intact'' experimental
+  condition (corresponding to the dark blue curve in Panel A). \textbf{C.
+  Network-specific compression.} Each of the seven networks are colored
+  according to the number of components at the network's inflection point for
+  the intact experimental condition. Larger numbers of components reflect lower
+  compressibility.}
+
+  \label{fig:networks}
 \end{figure}

### 2023-02-28 paper/main.tex
checking in results writeup for fig 3

@@
 
-\begin{figure}
+\begin{figure}[tp]
   \centering
@@
 
-\begin{figure}
+If informativeness (to the temporal decoders) and compressibility vary with the
+cognitive richness of the stimulus, might these measures also vary over time
+\textit{within} a given condition? For example, participants in the intact
+condition might process the ongoing story more deeply later on in the story
+(compared with earlier in the story) given the additional narrative background
+and context they had been exposed to by that point. To examine this
+possibility, we divided each condition into three successive time segments. We
+computed decoding curves (Fig.~\ref{fig:inflection-thirds}A) and inflection
+points (Fig.~\ref{fig:inflection-thirds}B) for each segment and condition. We
+found that, in the two most cognitively rich conditions (intact and paragraph),
+both decoding accuracy and compressibility, as reflected by the change in
+decoding curves, increased with listening time (intact: $t(XXX) = XXX, p =
+XXX$; paragraph: $t(XXX) = XXX, p = XXX$). These changes may reflect an
+increase in comprehension or depth of processing with listening time. In
+contrast, the decoding accuracy and compressibility \textit{decreased} with
+listening time in the word condition ($t(XXX) = XXX, p = XXX$) and rest
+condition ($t(XXX) = XXX, p = XXX$). This might reflect the depletion of
+attentional resources in the less-engaging word and rest conditions.
+
+\begin{figure}[tp]
   \centering
-  \includegraphics[width=0.8\textwidth]{figs/decode_pcs_network}
+  \includegraphics[width=0.7\textwidth]{figs/decode_pcs_network}
 

### 2023-02-28 paper/main.tex
checking in draft of results section, added new discussion figure

@@
 \newcommand{\synthThirds}{S1}
+\newcommand{\accuracyNetworks}{S2}
+\newcommand{\componentsNetworks}{S3}
 
@@
 
+We also wondered how informativeness and compressibility in the different
+experimental conditions might vary across brain networks. We used a network
+parcellation identified by \cite{YeoEtal11} to segment the brain into seven
+distinct networks. The networks can be sorted (roughly) in order from
+lower-level to higher-level cortex as follows (Fig.~\ref{fig:networks}A):
+visual, somatomotor, dorsal attention, ventral attention, limbic,
+frontoparietal, and default mode. Next, we computed decoding curves separately
+for the activity patterns within each network and identified each network's
+inflection point, for each experimental condition. Moving from low-order
+networks to higher-order networks, we found that decoding accuracy (for the
+intact condition) tended to increase
+(Figs.~\ref{fig:networks}B,~\accuracyNetworks). This suggests that higher-order
+networks may carry more content-relevant or stimulus-driven ``information.'' We
+found no clear trends in the numbers of components at each network's inflection
+point across networks or conditions
+(Figs.~\ref{fig:networks}C,~\componentsNetworks). \textbf{JRM NOTE: expand on
+this once we have the new figures.}
+
 \begin{figure}[tp]
@@
 
+In addition to examining different networks in isolation, we wondered about the
+general structure of the full-brain (i.e., potentially multi-network) activity
+patterns reflected by different principal components across different
+experimental conditions. Figure~\ref{fig:neurosynth-pca} displays inflated
+brain maps of the top five highest-weighted components, for each experimental
+condition. We also used Neurosynth~\citep{RubiEtal17} to identify, for each
+component, the top five terms associated with each map (see \textit{Reverse
+inference}). We noticed (by inspection) several common themes across the sets
+of terms associated with each component and condition. Memory-related
+components included terms like ``middle temporal,'' ``memory retrieval,'' and
+``working memory.'' Sensory processing related components included terms like
+``primary visual,'' ``auditory cortex,'' ``v1,'' and so on. Other components
+were associated with sensory integration (e.g., ``thalamic,'' ``cingulate,''
+etc.), sentence comprehension (e.g., ``sentence,'' ``comprehension''), emotion
+and valence (e.g., ``amygdala,'' ``valance''), or the default mode network
+(e.g., ``default mode'').  The components we identified were relatively stable
+across story segments (Fig.~\synthThirds).
+
 \section*{Discussion}
+
+\begin{figure}[tp]
+  \centering
+  \includegraphics[width=\textwidth]{figs/discussion}
+
+\caption{\textbf{Summary of functions associated with top-weighted components
+by condition.} Each column corresponds to an experimental condition. Brain maps
+in the top row are reproduced from Figure~\ref{fig:neurosynth-pca}, for
+reference. Cognitive functions summarized from the top Neurosynth-derived terms
+in Figure~\ref{fig:neurosynth-pca} are listed in the rows on the left.  Shaded cells
+denote which experimental conditions were associated with one or more top-weighted
+principal components associated with the given function.}
+
+\label{fig:discussion}
+\end{figure}
 

### 2023-02-28 paper/main.tex
backing up progress on discussion section

@@
 
+We examined fMRI data collected as participants listened to an auditory
+recording of a story, scrambled recordings of the story, or underwent a resting
+state scan. We found that cognitively richer stimuli evoked more reliable
+(i.e., consistent across people) and information rich brain activity patterns.
+The brain patterns are also more compressible, in that each individual
+component provided more ``signal'' to temporal decoders relative to components
+of data from less cognitively rich conditions (Fig.~\ref{fig:inflection}). Over
+time (e.g., as the experiment progressed), these phenomena were strengthened.
+Specifically, across story segments, data from more cognitively rich conditions
+became more informative and compressible, and data from less cognitively rich
+conditions became \textit{less} informative and compressible
+(Fig.~\ref{fig:inflection-thirds}). We also repeated these analyses separately
+for different brain networks. We found that networks traditionally associated
+with higher-level cognitive functions tended to provide more informative brain
+patterns than networks traditionally associated with lower level cognitive
+functions (Fig.~\ref{fig:networks}). Finally, we examined the most dominant
+components of the brain activity patterns from each experimental condition. We
+used a reverse inference approach~\citep{RubiEtal17} to identify the terms in
+the neuroimaging literature most commonly associated with the corresponding
+maps. As summarized in Figure~\ref{fig:discussion}, we found that terms
+associated with memory and sensory processing were associated with the
+strongest components in all three story listening conditions. Terms associated
+with sensory integration were associated with the strongest components in the
+intact and paragraph-scrambled conditions. Terms associated with sentence
+comprehension, emotion, and valence were associated with the strongest
+components in the intact condition. Finally, terms associated with the default
+mode network were associated with the strongest components in the
+word-scrambled and resting state conditions. Taken together, our findings
+indicate that the informativeness and compressibility of our brain activity
+patterns are task-dependent, and these properties change systematically with
+factors like cognitive richness and depth of processing.
+
 \begin{figure}[tp]

### 2023-03-01 paper/main.tex
added some notes, updated decoding/inflection figure, backing up progress

@@
 
+other themes:
+- predictability implies redundancy
+
+- the structure of our brain patterns changes with cognitive task. maybe
+something about flexibility or adaptability? why might some states benefit from
+more compressibiilty or more informativeness?
+
+- for a fixed set of "symbols" the information rate decreases as the signal is
+compressed further. but our finding that each brain pattern component (symbol)
+becomes more informative as cognitive richness increases suggests that the
+"alphabet" of brain patterns is task dependent.
+
+- representations themselves may also change with task demands (e.g., Mack et al, 2020)
+
+- different brain networks appear to optimize in different ways, perhaps for different
+cognitive tasks or operations.
+
+\subsection*{Concluding remarks}
+
+- something about the fundamental nature of brain activity and the neural basis of thought.
+
 \section*{Methods}

### 2023-03-01 paper/main.tex
backing up progress on discussion section

@@
 
+Correlations or causal associations between different brain regions
+simultaneously imply that full-brain activity patterns will be compressible and
+also that those activity patterns will contain redundancies. For example, the
+extent that activity patterns at one brain area can be inferred or predicted
+from activity patterns at other areas~\citep[e.g.,][]{OwenEtal20, ScanEtal21},
+reflects overlap in the information available in or represented by those brain
+areas. If brain patterns in one area are recoverable using brain patterns in
+another area, then the ``signal'' used to convey the activity patterns could be
+compressed by removing the recoverable activity. Predictable (and therefore
+redundant) brain activity patterns are also more robust to signal corruption.
+For example, even if the activity patterns at one region are unreadable or
+unreliable at a given moment, that unreliability could be compensated for by
+other region's activity patterns that are predictive of the unreliable region.
+
 other themes:
-- predictability implies redundancy
 

### 2023-03-02 paper/main.tex
checking in complete draft of discussion section

@@
 \usepackage{hyperref}
-\usepackage{natbib}
+\usepackage[sort]{natbib}
 \usepackage{pxfonts}
@@
 
-Correlations or causal associations between different brain regions
-simultaneously imply that full-brain activity patterns will be compressible and
-also that those activity patterns will contain redundancies. For example, the
-extent that activity patterns at one brain area can be inferred or predicted
-from activity patterns at other areas~\citep[e.g.,][]{OwenEtal20, ScanEtal21},
-reflects overlap in the information available in or represented by those brain
-areas. If brain patterns in one area are recoverable using brain patterns in
-another area, then the ``signal'' used to convey the activity patterns could be
-compressed by removing the recoverable activity. Predictable (and therefore
-redundant) brain activity patterns are also more robust to signal corruption.
-For example, even if the activity patterns at one region are unreadable or
-unreliable at a given moment, that unreliability could be compensated for by
-other region's activity patterns that are predictive of the unreliable region.
+Our explorations of informativeness and compressibility are related to a much
+broader literature on the correlational and causal structure of brain activity
+patterns~\citep{PretEtal17, OwenEtal21, RogeEtal07, RubiSpor10, SizeEtal18,
+SmitEtal13b, SmitEtal13c, SrinEtal07, TomaVolk11, YeoEtal11, AdacEtal12,
+BassSpor17, BullSpor09, SporHone06, SporBetz16, SporZwi04, DhamEtal08,
+KorzEtal08, BrovEtal04}. Correlations or causal associations between different
+brain regions simultaneously imply that full-brain activity patterns will be
+compressible and also that those activity patterns will contain redundancies.
+For example, the extent that activity patterns at one brain area can be
+inferred or predicted from activity patterns at other
+areas~\citep[e.g.,][]{OwenEtal20, ScanEtal21}, reflects overlap in the
+information available in or represented by those brain areas. If brain patterns
+in one area are recoverable using brain patterns in another area, then the
+``signal'' used to convey the activity patterns could be compressed by removing
+the recoverable activity. Predictable (and therefore redundant) brain activity
+patterns are also more robust to signal corruption. For example, even if the
+activity patterns at one region are unreadable or unreliable at a given moment,
+that unreliability could be compensated for by other region's activity patterns
+that are predictive of the unreliable region.
 
-other themes:
+Our findings that informativeness and compressibility change with task demands
+may follow from task-dependent changes in full-brain correlation patterns. A
+number of prior studies have found that so-called ``functional connectivity''
+(i.e., correlational) patterns vary reliably across tasks, events, and
+situations~\citep{SimoEtal16, ColeEtal14, SmitEtal09, OwenEtal21}. By examining
+how these task-dependent changes in correlations affect informativeness and
+compressibility, our work suggests a potential reason for why the statistical
+structure of brain activity patterns might vary with cognitive task or with
+cognitive demands. For lower-level tasks, or for tasks that require relatively
+little ``deep'' cognitive processing, our brains may optimize activity patterns
+for robustness and redundancy over expressiveness, for example to maximize
+reliability. For higher-level tasks, or for tasks that require deeper cognitive
+processing, our brains may sacrifice some redundancy in favor of greater
+expressiveness.
 
-- the structure of our brain patterns changes with cognitive task. maybe
-something about flexibility or adaptability? why might some states benefit from
-more compressibiilty or more informativeness?
+In the information theory sense~\citep{Shan48}, when a signal is transmitted
+using a fixed alphabet of ``symbols,'' the information rate decreases as the
+signal is compressed (e.g., fewer symbols transmitted per unit time, using an
+alphabet with fewer symbols, etc.). Our finding that each individual brain
+component (symbol) becomes more informative as cognitive richness increases
+suggests that the ``alphabet'' of brain activity patterns is also
+task-dependent. Other work suggests that the representations that are
+\textit{reflected} by brain activity patterns may also change with task
+demands. For example, our brains may represent the same perceptual stimulus
+differently depending on which aspects of the stimulus or which combinations of
+features are relevant~\citep{MackEtal20}.
 
-- for a fixed set of "symbols" the information rate decreases as the signal is
-compressed further. but our finding that each brain pattern component (symbol)
-becomes more informative as cognitive richness increases suggests that the
-"alphabet" of brain patterns is task dependent.
-
-- representations themselves may also change with task demands (e.g., Mack et al, 2020)
-
-- different brain networks appear to optimize in different ways, perhaps for different
-cognitive tasks or operations.
+Different brain networks also varied in how informative and compressible their
+activity patterns were across experimental conditions (e.g.,
+Fig.~\ref{fig:networks}). This might reflect evolutionary optimizations that
+reflect the relevant constraints or demands placed on those networks. One
+possibility is that cortex is organized in a hierarchy of networks concerned
+with different levels of processing or function. To the extent that different
+levels of processing (e.g., low-level sensory processing versus ``deeper''
+higher-level processing) reflect different stimulus
+timescales~\citep[e.g.,][]{Mann20}, the network differences we observed might
+also relate to the timescales ot which each network is maximally
+sensitive~\citep{RegeEtal18, BaldEtal17,LernEtal11, HassEtal08}.
 
@@
 
-- something about the fundamental nature of brain activity and the neural basis of thought.
+Cognitive neuroscientists are still grappling with basic questions about the
+fundamental ``rules'' describing how our brains respond, and about how brain
+activity patterns and the associated cognitive representations and computations
+are linked. We identified two aspects of brain activity patterns,
+informativeness and compressibiilty, that appear to change systematically with
+task demands and across brain networks. Our work helps to clarify how the
+``neural code'' might be structured, and how the code might vary across tasks
+and brain areas.
 

### 2023-03-02 paper/main.tex
applied a coat of polish to the text-- fixed a bunch of minor issues throughout (mostly typos).  i also corrected some errors in figure 1.

@@
 
-\title{High-order cognition is supported by complex but compressible brain activity patterns}
+\title{High-order cognition is supported by information-rich but compressible brain activity patterns}
 
@@
 engagement. We considered two primary aspects of the data. First, we treated
-the number of features (components) required to achieve a threshold decoding
-accuracy as a proxy for the ``compressibility'' of the neural patterns (where
-fewer components indicate greater compression). Second, we treated the
-maximum achievable decoding accuracy across participants as an indicator of the
-``stability'' of the recorded patterns. Overall, we found that neural patterns
-recorded as participants listened to the intact story required fewer features
-to achieve comparable classification accuracy to the other experimental
-conditions. However, the peak decoding accuracy (achievable with more features)
-was also highest during intact story listening. Taken together, our work
+the maximum achievable decoding accuracy across participants as an indicator of
+the ``informativeness'' of the recorded patterns. Second, we treated the number
+of features (components) required to achieve a threshold decoding accuracy as a
+proxy for the ``compressibility'' of the neural patterns (where fewer
+components indicate greater compression). Overall, we found that the peak
+decoding accuracy (achievable without restricting the numbers of features)
+was highest in the intact (unscrambled) story listening condition.  However, the
+number of features required to achieve comparable classification accuracy
+was also lowest in the intact story listening condition.  Taken together, our work
 suggests that our brain networks flexibly reconfigure according to ongoing task
 demands, and that the activity patterns associated with higher-order cognition
-and high engagement are both more complex and more compressible than the
+and high engagement are both more informative and more compressible than the
 activity patterns associated with lower-order tasks and lower levels of
@@
 \noindent
-\textbf{Keywords: information, compression, temporal decoding, dimensionality reduction, neuroimaging}
+\textbf{Keywords: information, compression, temporal decoding, dimensionality
+reduction, neuroimaging}
 
@@
 occupying one or more positions along on a continuum. At one extreme, every
-node is fully independent of every other node. At the other extreme, all nodes
-behave identically. Each extreme optimizes key properties of how the network
-functions. When every node is independent, the network is maximally
+node is fully independent from every other node. At the other extreme, all
+nodes behave identically. Each extreme optimizes key properties of how the
+network functions. When every node is independent, the network is maximally
 \textit{expressive}: if we define the network's ``state'' as the activity
@@
 flexibility and expressiveness), are more information-rich than simpler or more
-constrained systems. The redundancy of a signal denotes the difference how
-expressive the signal \textit{could} be (i.e., proportional to the number of
-unique states or symbols used to transmit the signal) and the actual
+constrained systems. The redundancy of a signal denotes the difference between
+how expressive the signal \textit{could} be (i.e., proportional to the number
+of unique states or symbols used to transmit the signal) and the actual
 information rate (i.e., the entropy of each individual state or symbol). If a
@@
 sufficient to recover the full-brain activity pattern. Highly redundant systems
-are also robust, since there is little information loss from losing any given
-observation.
+are also robust, since there is little total information loss due to removing
+any given observation.
 
@@
 
-  \caption{\textbf{Information content and compressibility.} \textbf{A. Variance
-  explained for two images.} We applied principal components analysis to a
-  photograph and drawing, treating each row of the images as ``observations.''
-  Across columns, we identified the number of components required to explain
-  100\%, 95\%, 75\%, or 50\% of the cumulative variance in each image (the 100\%
-  columns denote the original images). The numbers of components are indicate in
-  parentheses, and the resulting ``compressed'' images are displayed. \textbf{B.
-  Representing two images with different numbers of components.} Using the same
-  principal component decompositions as in Panel A, we computed the cumulative
-  proportion of variance explained with 380 (original images), 20, 4, or 2
-  components. \textbf{C. Cumulative variance explained versus number of
-  components.} For the images displayed in Panels A and B, we plot the cumulative
-  proportion of variance explained as a function of the number of components used
-  to represent each image. \textbf{D. Information rate and compressibility.}
-  Across multiple images, the information rate (i.e., the amount of information
-  contained in each image; horizontal axis) is high when each individual pixel
-  provides information that cannot be inferred from other pixels.
-  High-information rate images tend to be high-resolution, and low-information
-  rate images tend to be low-resolution. Compressibility is related to the
-  difference between the information required to specify the original versus
-  compressed images (vertical axis). Highly compressible images often contain
-  predictable structure (redundancies) that can be leveraged to represent the
-  images much more efficiently than in their original feature spaces.}
+  \caption{\textbf{Information content and compressibility.} \textbf{A.
+  Variance explained for two images.} We applied principal components analysis
+  to a photograph and drawing, treating the rows of the images as
+  ``observations.'' Across columns, we identified the numbers of components
+  required to explain 100\%, 95\%, 75\%, or 50\% of the cumulative variance in
+  each image (the 100\% columns denote the original images). The numbers of
+  components are indicate in parentheses, and the resulting ``compressed''
+  images are displayed. \textbf{B. Representing two images with different
+  numbers of components.} Using the same principal component decompositions as
+  in Panel A, we computed the cumulative proportion of variance explained with
+  380 (original images), 20, 4, or 2 components. \textbf{C. Cumulative variance
+  explained versus number of components.} For the images displayed in Panels A
+  and B, we plot the cumulative proportion of variance explained as a function
+  of the number of components used to represent each image. \textbf{D.
+  Information rate and compressibility.} Across multiple images, the
+  information rate (i.e., the amount of information contained in each image;
+  horizontal axis) is high when each individual pixel provides information that
+  cannot be inferred from other pixels. High-information rate images tend to be
+  high-resolution, and low-information rate images tend to be low-resolution.
+  Compressibility is related to the difference between the information required
+  to specify the original versus compressed images (vertical axis). Highly
+  compressible images often contain predictable structure (redundancies) that
+  can be leveraged to represent the images much more efficiently than in their
+  original feature spaces.}
 
@@
 The above studies are consistent with two potential descriptions of how
-cognitive processes are reflects in brain activity patterns. One possibility is
-that the information rate of brain activity increases during more complex or
+cognitive processes are reflected in brain activity patterns. One possibility
+is that the information rate of brain activity increases during more complex or
 higher-level cognitive processing. If so, then the ability to reliably decode
@@
 decoding curve asymptotes. We fit sigmoid functions to each decoding curve
-(left sub-panel) and then computed the minimum number of components where the
-second derivative of the sigmoid was both positive and less than a threshold
-value of 0.0001. \textbf{D. Inflection points by condition.} Each dot displays
-the number of components ($x$-axis) and decoding accuracy ($y$-axis) at one
-condition's inflection point. All error ribbons denote bootstrap-estimated 95\%
-confidence intervals.}
+(left sub-panel) and then computed the minimum number of components for which
+the second derivative of the sigmoid was both positive and less than a
+threshold value of 0.0001. \textbf{D. Inflection points by condition.} Each dot
+displays the number of components ($x$-axis) and decoding accuracy ($y$-axis)
+at one condition's inflection point. All error ribbons denote
+bootstrap-estimated 95\% confidence intervals.}
 
@@
 
-To evaluate the relation between informativeness and compression for brain
+To evaluate the relation between informativeness and compressibiilty for brain
 activity from each experimental condition, we trained a series of
@@
 conditions), decoding accuracy improves as the number of components increases.
-However, decoding accuracy peaked at a higher level for experimental conditions
+However, decoding accuracy peaked at higher levels for experimental conditions
 that exposed participants to cognitively richer stimuli. The peak decoding
@@
 $t(XXX) = XXX, p = XXX$; versus rest: $t(XXX) = XXX, p = XXX$), and next
-highest for the ``word'' condition (versus rest: $t(XXX) = XXX, p = XXX$). 
-This ordering implies that cognitively richer conditions evoke more stable brain
+highest for the ``word'' condition (versus rest: $t(XXX) = XXX, p = XXX$). This
+ordering implies that cognitively richer conditions evoke more stable brain
 activity patterns across people.
@@
 networks to higher-order networks, we found that decoding accuracy (for the
-intact condition) tended to increase
-(Figs.~\ref{fig:networks}B,~\accuracyNetworks). This suggests that higher-order
-networks may carry more content-relevant or stimulus-driven ``information.'' We
-found no clear trends in the numbers of components at each network's inflection
-point across networks or conditions
-(Figs.~\ref{fig:networks}C,~\componentsNetworks). \textbf{JRM NOTE: expand on
-this once we have the new figures.}
+intact condition) tended to increase (Fig.~\ref{fig:networks}B). This suggests
+that higher-order networks may carry more content-relevant or stimulus-driven
+``information.'' We found no clear trends in the numbers of components at each
+network's inflection point across networks or conditions
+(Fig.~\ref{fig:networks}C).
 
@@
 (i.e., consistent across people) and information rich brain activity patterns.
-The brain patterns are also more compressible, in that each individual
-component provided more ``signal'' to temporal decoders relative to components
-of data from less cognitively rich conditions (Fig.~\ref{fig:inflection}). Over
-time (e.g., as the experiment progressed), these phenomena were strengthened.
-Specifically, across story segments, data from more cognitively rich conditions
-became more informative and compressible, and data from less cognitively rich
-conditions became \textit{less} informative and compressible
-(Fig.~\ref{fig:inflection-thirds}). We also repeated these analyses separately
-for different brain networks. We found that networks traditionally associated
-with higher-level cognitive functions tended to provide more informative brain
-patterns than networks traditionally associated with lower level cognitive
-functions (Fig.~\ref{fig:networks}). Finally, we examined the most dominant
-components of the brain activity patterns from each experimental condition. We
-used a reverse inference approach~\citep{RubiEtal17} to identify the terms in
-the neuroimaging literature most commonly associated with the corresponding
-maps. As summarized in Figure~\ref{fig:discussion}, we found that terms
-associated with memory and sensory processing were associated with the
-strongest components in all three story listening conditions. Terms associated
-with sensory integration were associated with the strongest components in the
-intact and paragraph-scrambled conditions. Terms associated with sentence
-comprehension, emotion, and valence were associated with the strongest
-components in the intact condition. Finally, terms associated with the default
-mode network were associated with the strongest components in the
-word-scrambled and resting state conditions. Taken together, our findings
-indicate that the informativeness and compressibility of our brain activity
-patterns are task-dependent, and these properties change systematically with
-factors like cognitive richness and depth of processing.
+The brain patterns evoked by cognitively richer stimuli were also more
+compressible, in that each individual component provided more ``signal'' to
+temporal decoders relative to components of data from less cognitively rich
+conditions (Fig.~\ref{fig:inflection}). Over time (e.g., as the experiment
+progressed), these phenomena were strengthened. Specifically, across story
+segments, data from more cognitively rich conditions became more informative
+and compressible, and data from less cognitively rich conditions became
+\textit{less} informative and compressible (Fig.~\ref{fig:inflection-thirds}).
+We also repeated these analyses separately for different brain networks. We
+found that networks traditionally associated with higher-level cognitive
+functions tended to provide more informative brain patterns than networks
+traditionally associated with lower level cognitive functions
+(Fig.~\ref{fig:networks}). Finally, we examined the most dominant components of
+the brain activity patterns from each experimental condition. We used a reverse
+inference approach~\citep{RubiEtal17} to identify the terms in the neuroimaging
+literature most commonly associated with the corresponding maps. As summarized
+in Figure~\ref{fig:discussion}, we found that terms associated with memory and
+sensory processing were associated with the strongest components in all three
+story listening conditions. Terms associated with sensory integration were
+associated with the strongest components in the intact and paragraph-scrambled
+conditions. Terms associated with sentence comprehension, emotion, and valence
+were associated with the strongest components in the intact condition. Finally,
+terms associated with the default mode network were associated with the
+strongest components in the word-scrambled and resting state conditions. Taken
+together, our findings indicate that the informativeness and compressibility of
+our brain activity patterns are task-dependent, and these properties change
+systematically with factors like cognitive richness and depth of processing.
 
@@
 compressible and also that those activity patterns will contain redundancies.
-For example, the extent that activity patterns at one brain area can be
+For example, the extent ot which activity patterns at one brain area can be
 inferred or predicted from activity patterns at other
@@
 activity patterns at one region are unreadable or unreliable at a given moment,
-that unreliability could be compensated for by other region's activity patterns
-that are predictive of the unreliable region.
+that unreliability could be compensated for by other regions' activity patterns
+that were predictive of the unreliable region.
 
@@
 how these task-dependent changes in correlations affect informativeness and
-compressibility, our work suggests a potential reason for why the statistical
+compressibility, our work suggests a potential reason why the statistical
 structure of brain activity patterns might vary with cognitive task or with
@@
 differently depending on which aspects of the stimulus or which combinations of
-features are relevant~\citep{MackEtal20}.
+features are task-relevant~\citep{MackEtal20}.
 
@@
 activity patterns were across experimental conditions (e.g.,
-Fig.~\ref{fig:networks}). This might reflect evolutionary optimizations that
-reflect the relevant constraints or demands placed on those networks. One
-possibility is that cortex is organized in a hierarchy of networks concerned
-with different levels of processing or function. To the extent that different
-levels of processing (e.g., low-level sensory processing versus ``deeper''
-higher-level processing) reflect different stimulus
+Fig.~\ref{fig:networks}). This might follow from evolutionary optimizations
+that reflect the relevant constraints or demands placed on those networks. One
+possibility is that cortex is organized in a hierarchy of networks ``concerned
+with'' or selective to different levels of processing or function. To the
+extent that different levels of processing (e.g., low-level sensory processing
+versus ``deeper'' higher-level processing) reflect different stimulus
 timescales~\citep[e.g.,][]{Mann20}, the network differences we observed might
-also relate to the timescales ot which each network is maximally
+also relate to the timescales at which each network is maximally
 sensitive~\citep{RegeEtal18, BaldEtal17,LernEtal11, HassEtal08}.
@@
 fundamental ``rules'' describing how our brains respond, and about how brain
-activity patterns and the associated cognitive representations and computations
-are linked. We identified two aspects of brain activity patterns,
+activity patterns and the associated underlying cognitive representations and
+computations are linked. We identified two aspects of brain activity patterns,
 informativeness and compressibiilty, that appear to change systematically with
@@
 participants). Full neuroimaging details may be found in the original paper for
-which the data were collected~\cite{SimoEtal16}. Procedures were approved by
+which the data were collected~\citep{SimoEtal16}. Procedures were approved by
 the Princeton University Committee on Activities Involving Human Subjects, and
@@
 
-Following our prior related work, we used HTFA~\cite{MannEtal18} to derive a
+Following our prior related work, we used HTFA~\citep{MannEtal18} to derive a
 compact representation of the neuroimaging data. In brief, this approach
@@
 dataset were performed in this lower-dimensional space. In other words, each
-participant's data matrix, $\mathbf{X}$, was a number-of-timepoints ($T$) by 700
-matrix of HTFA-derived factor weights (where the row and column labels were
-matched across participants). Code for carrying out HTFA on fMRI data may be
-found as part of the BrainIAK toolbox~\cite{CapoEtal17, KumaEtal21}, which may
-be downloaded at \href{https://brainiak.org/}{brainiak.org}.
+participant's data matrix was a number-of-timepoints ($T$) by 700 matrix of
+HTFA-derived factor weights (where the row and column labels were matched
+across participants). Code for carrying out HTFA on fMRI data may be found as
+part of the BrainIAK toolbox~\citep{CapoEtal17, KumaEtal21}, which may be
+downloaded at \href{https://brainiak.org/}{brainiak.org}.
 
@@
 participants' $T$ by 700 factor weight matrices. We used group PCA to project
-these 700-dimensional matrices into a series of $k$-dimensional spaces, for $k
-\in \{3, 4, 5, ..., 700\}$. This yielded a set of number-of-participants
+these 700-dimensional matrices into a series of shared $k$-dimensional spaces,
+for $k \in \{3, 4, 5, ..., 700\}$. This yielded a set of number-of-participants
 matrices, each with $T$ rows and $k$ columns.
@@
 the corresponding cognitive representations are reflected in similar spatial
-patterns across people~\cite{SimoChan20}. Following this logic, we used an
+patterns across people~\citep{SimoChan20}. Following this logic, we used an
 across-participant temporal decoding test developed by \cite{MannEtal18} to
@@
 of $k$ (i.e., number of principal components), we first used group PCA to
-project the data from each condition into a $k$-dimensional space. Next, we
-divided the participants into two groups: a template group,
+project the data from each condition into a shared $k$-dimensional space. Next,
+we divided the participants into two groups: a template group,
 $\mathcal{G}_{\mathrm{template}}$ (i.e., training data), and a to-be-decoded
@@
 \synthThirds). We then carried out a meta analysis using
-Neurosynth~\cite{RubiEtal17} to identify the 5 terms most commonly associated
+Neurosynth~\citep{RubiEtal17} to identify the 5 terms most commonly associated
 with the given map.
@@
 All of the code used to produce the figures and results in this manuscript,
-along with links to the corresponding datasets, may be found at 
+along with links to the corresponding data, may be found at
 \href{https://github.com/ContextLab/pca_paper}{github.com/ContextLab/pca\_paper}.
@@
 
-We acknowledge discussions with Rick Betzel, Emily Finn, and Jim Haxby. Our
-work was supported in part by NSF CAREER Award Number 2145172 to J.R.M. The
-content is solely the responsibility of the authors and does not necessarily
-represent the official views of our supporting organizations. The funders had
-no role in study design, data collection and analysis, decision to publish, or
-preparation of the manuscript.
+We acknowledge discussions with Rick Betzel, Luke Chang, Emily Finn, and Jim
+Haxby. Our work was supported in part by NSF CAREER Award Number 2145172 to
+J.R.M. The content is solely the responsibility of the authors and does not
+necessarily represent the official views of our supporting organizations. The
+funders had no role in study design, data collection and analysis, decision to
+publish, or preparation of the manuscript.
 

### 2023-03-02 paper/admin/cover_letter.tex
started to work on cover letter (backing up progress since i need to step away)

\title{cover letter}
%
% See http://texblog.org/2013/11/11/latexs-alternative-letter-class-newlfm/
% and http://www.ctan.org/tex-archive/macros/latex/contrib/newlfm
% for more information.
%
\documentclass[11pt,stdletter,orderfromtodate,sigleft]{newlfm}
\usepackage{blindtext, xfrac, animate, hyperref, pxfonts, geometry}

  \setlength{\voffset}{0in}

\newlfmP{dateskipbefore=0pt}
\newlfmP{sigsize=20pt}
\newlfmP{sigskipbefore=10pt}

\newlfmP{Headlinewd=0pt,Footlinewd=0pt}

\newcommand{\journal}{Nature Neuroscience}
\newcommand{\articletype}{Article}
\newcommand{\myTitle}{High-order cognition is supported by information-rich but compressible brain activity patterns}
\newcommand{\corresponding}{Jeremy R. Manning}

\namefrom{\vspace{-0.3in}\corresponding}
\addrfrom{
  Dartmouth College\\
  Department of Psychological \& Brain Sciences\\
  Moore Hall\\
  Hanover, NH, 03755}
\addrto{}
\dateset{\today}

\greetto{To the editors of \textit{\journal}:}

\closeline{Sincerely,}
%\usepackage{setspace}
%\linespread{0.85}
% How will your work make others in the field think differently and move the field forward?
% How does your work relate to the current literature on the topic?
% Who do you consider to be the most relevant audience for this work?
% Have you made clear in the letter what the work has and has not achieved?

\begin{document}
\begin{newlfm}

  We have enclosed our manuscript entitled \textit{\myTitle} to be considered
  for publication as an \textit{\articletype}. Our manuscript uses
  dimensionality reduction algorithms and pattern classifiers to explore how
  informative and compressible (in the information theory senses of those
  words) brain activity patterns are, under different cognitive circumstances.

  We developed a computational framework for evaluating our two measures of
  interest (informativeness and compressibility), and we applied the framework
  to a public fMRI dataset. In the dataset's experiment, participants either
  listened to an auditory recording of a story, listened to temporally
  scrambled recordings of the story, or underwent a resting state scan. Each
  condition is designed to engage cognitive processing and engagement at
  different depths. For example, listening to the intact story leads
  participants to mentally engage with the narrative, and to deeply process
  and connect narrative events in a way that leads to a rich understanding
  of the story.  Listening to temporally scrambled versions of the story might
  lead to understanding invidual words....

Thank you for considering this manuscript, and I hope you will find it suitable
for publication in \textit{\journal}.

\end{newlfm}
\end{document}

### 2023-03-03 paper/admin/cover_letter.tex
finished draft of cover letter

@@
 %
-\documentclass[11pt,stdletter,orderfromtodate,sigleft]{newlfm}
-\usepackage{blindtext, xfrac, animate, hyperref, pxfonts, geometry}
+\documentclass[10pt,stdletter,orderfromtodate,sigleft]{newlfm}
+\usepackage{pxfonts, geometry}
 
@@
   listened to an auditory recording of a story, listened to temporally
-  scrambled recordings of the story, or underwent a resting state scan. Each
-  condition is designed to engage cognitive processing and engagement at
-  different depths. For example, listening to the intact story leads
-  participants to mentally engage with the narrative, and to deeply process
-  and connect narrative events in a way that leads to a rich understanding
-  of the story.  Listening to temporally scrambled versions of the story might
-  lead to understanding invidual words....
+  scrambled recordings of the story (randomizing the orders of paragraphs or
+  words), or underwent a resting state scan. Each condition is designed to
+  engage cognitive processing and engagement at different depths. For example,
+  listening to the intact story leads participants to mentally engage with the
+  narrative, and to deeply process and connect narrative events in a way that
+  leads to a rich understanding of the story. Listening to temporally scrambled
+  versions of the story might lead to understanding invidual paragraphs or
+  words. However, individual moments in the temporally scrambled stories lack
+  contextual elements that enable participants to gain a deeper understanding.
+  Across conditions, we found that both informativeness and compressibility of
+  the brain patterns changed systematically with the cognitive ``richness'' of
+  the stimulus across the different experimental conditions. We also traced out
+  the brain networks associated with these changes. We found that networks
+  traditionally associated with higher-level cognitive functions tended to
+  exhibit more information-rich brain patterns than networks traditionally
+  associated with lower-level cognitive functions.
 
-Thank you for considering this manuscript, and I hope you will find it suitable
+  Taken together, our work provides new insights into the fundamental ``rules''
+  describing how our brains respond to and represent specific stimuli and
+  cognitive processes. We connect our findings about how the informativeness
+  and compressibility of brain activity change across experimental conditions
+  with prior work on task-dependent changes in functional connectivity (i.e.,
+  full-brain correlations). Our work helps to clarify how the ``neural code''
+  might be structured, and how the neural code might vary across tasks and
+  brain areas. We expect that our work will be of interest to a broad audience
+  including neuroscientists, cognitive psychologists and cognitive scientists,
+  and others.
+
+Thank you for considering our manuscript, and I hope you will find it suitable
 for publication in \textit{\journal}.

### 2023-03-07 code/notebooks/helpers.py
starting to work on network analyses (currently debugging)

@@ -1,0 +1,114 @@
+import numpy as np
+import pandas as pd
+import nibabel as nib
+import nilearn as nl
+from nilearn.input_data import NiftiMasker
+
+import os
+import warnings
+
+
+def nii2cmu(nifti_file, mask_file=None):
+    '''
+    inputs:
+      nifti_file: a filename of a .nii or .nii.gz file to be converted into
+                  CMU format
+                  
+      mask_file: a filename of a .nii or .nii.gz file to be used as a mask; all
+                 zero-valued voxels in the mask will be ignored in the CMU-
+                 formatted output.  If ignored or set to None, no voxels will
+                 be masked out.
+    
+    outputs:
+      Y: a number-of-timepoints by number-of-voxels numpy array containing the
+         image data.  Each row of Y is an fMRI volume in the original nifti
+         file.
+      
+      R: a number-of-voxels by 3 numpy array containing the voxel locations.
+         Row indices of R match the column indices in Y.
+    '''
+    def fullfact(dims):
+        '''
+        Replicates MATLAB's fullfact function (behaves the same way)
+        '''
+        vals = np.asmatrix(range(1, dims[0] + 1)).T
+        if len(dims) == 1:
+            return vals
+        else:
+            aftervals = np.asmatrix(fullfact(dims[1:]))
+            inds = np.asmatrix(np.zeros((np.prod(dims), len(dims))))
+            row = 0
+            for i in range(aftervals.shape[0]):
+                inds[row:(row + len(vals)), 0] = vals
+                inds[row:(row + len(vals)), 1:] = np.tile(aftervals[i, :], (len(vals), 1))
+                row += len(vals)
+            return inds
+    
+    with warnings.catch_warnings():
+        warnings.simplefilter("ignore")
+        
+        img = nib.load(nifti_file)
+        mask = NiftiMasker(mask_strategy='background')
+        if mask_file is None:
+            mask.fit(nifti_file)
+        else:
+            mask.fit(mask_file)
+    
+    hdr = img.header
+    S = img.get_sform()
+    vox_size = hdr.get_zooms()
+    im_size = img.shape
+    
+    if len(img.shape) > 3:
+        N = img.shape[3]
+    else:
+        N = 1
+    
+    Y = np.float32(mask.transform(nifti_file)).copy()
+    vmask = np.nonzero(np.array(np.reshape(mask.mask_img_.dataobj, (1, np.prod(mask.mask_img_.shape)), order='C')))[1]
[…]

### 2023-03-08 code/notebooks/helpers.py
running decoding and compression analyses (backing up progress and decoding results files)
- also working on networks analyses (in *_voxel.ipynb and helpers.py); trying to debug voxel locations and typed out some notes and potential directions

@@ -2,7 +2,7 @@ import numpy as np
 import pandas as pd
 import nibabel as nib
 import nilearn as nl
-from nilearn.input_data import NiftiMasker
+from nilearn.maskers import NiftiMasker
 
 import os
 import warnings
@@ -69,6 +69,10 @@ def nii2cmu(nifti_file, mask_file=None):
     vox_coords = fullfact(img.shape[0:3])[vmask, ::-1]-1
     
     R = np.array(np.dot(vox_coords, S[0:3, 0:3])) + S[:3, 3]
+
+    # center on the MNI152 brain (hard code this in)
+    mni_center = np.array([0.55741881, -21.52140703, 9.83783098])
+    R = R - R.mean(axis=0) + mni_center
     
     return {'Y': Y, 'R': R}
 

### 2023-03-08 paper/main.tex
added stats, updated figures 2 and 3 (including captions)

@@
 
-  \includegraphics[width=0.8\textwidth]{figs/decoding_and_inflection}
+  \includegraphics[width=\textwidth]{figs/decoding_and_inflection}
 
@@
 word, and rest), as a function of the number of components (features) used to
-represent the data. The horizontal red line denotes chance performance, and the
+represent the data. For each condition, decoding accuracy has been normalized
+by subtracting expected ``chance'' decoding performance (estimated as
+$\frac{1}{T}$, where $T$ is the number of timepoints in the given condition).
+This normalization was intended to enable fairer comparisons across conditions;
+a relative decoding accuracy of 0.0 denotes ``chance'' performance. The
 horizontal black line denotes 5\% decoding accuracy (used as a reference in
@@
 inset in Panel A. Intersections between each condition's decoding accuracy
-curve and the 5\% decoding accuracy reference line are marked by arrows.
-\textbf{C. Estimating inflection points.} We sought to identify an ``inflection
-point'' for each decoding curve, denoting the number of components at which the
-decoding curve asymptotes. We fit sigmoid functions to each decoding curve
-(left sub-panel) and then computed the minimum number of components for which
-the second derivative of the sigmoid was both positive and less than a
-threshold value of 0.0001. \textbf{D. Inflection points by condition.} Each dot
-displays the number of components ($x$-axis) and decoding accuracy ($y$-axis)
-at one condition's inflection point. All error ribbons denote
-bootstrap-estimated 95\% confidence intervals.}
+curve and the 5\% decoding accuracy reference line are marked by arrows. All
+error ribbons in Panels A and B denote bootstrap-estimated 95\% confidence
+intervals. \textbf{C. Estimating ``compressibility'' for each condition.} The
+density plots display the numbers of components required to reach at least 5\%
+(corrected) decoding accuracy, across all cross validation runs. For the
+``rest'' condition, where decoding accuracy never reached 5\% accuracy, we
+display the distribution of the numbers of components required to achieve the
+peak decoding accuracy in each fold. (This distribution is nearly flat, as
+illustrated in Panel E.) \textbf{D. Estimating ``informativeness'' for each
+condition.} The density plots display the peak (corrected) decoding accuracies
+achieved in each condition, across all cross validation runs. \textbf{E.
+Informativeness versus compressibility.} Each dot displays the minimum number
+of components required to achieve at least 5\% corrected decoding accuracy
+($x$-coordinate; for the rest condition the numbers of components are estimated
+using the peak decoding accuracies as in Panel C) and the peak (corrected)
+decoding accuracy ($y$-coordinate). The smaller dots correspond to individual
+cross validation runs and the larger dots display the averages across runs. The
+inset displays a zoomed-in view of the indicated region in the main panel.}
 
@@
 that exposed participants to cognitively richer stimuli. The peak decoding
-accuracy was highest for the ``intact'' condition (versus paragraph: $t(XXX) =
-XXX, p = XXX$; versus word: $t(XXX) = XXX, p = XXX$; versus rest: $t(XXX) =
-XXX, p = XXX$), next highest for the ``paragraph'' condition (versus word:
-$t(XXX) = XXX, p = XXX$; versus rest: $t(XXX) = XXX, p = XXX$), and next
-highest for the ``word'' condition (versus rest: $t(XXX) = XXX, p = XXX$). This
-ordering implies that cognitively richer conditions evoke more stable brain
-activity patterns across people.
+accuracy was highest for the ``intact'' condition (versus paragraph: $t(99) =
+35.205, p < 0.001)$; versus word: $t(99) = 43.172, p < 0.001)$; versus rest:
+$t(99) = 81.361, p < 0.001)$), next highest for the ``paragraph'' condition
+(versus word: $t(99) = 6.243, p < 0.001)$; versus rest: $t(99) = 50.748, p <
+0.001)$), and next highest for the ``word'' condition (versus rest: $t(99) =
+48.791, p < 0.001)$). This ordering implies that cognitively richer conditions
+evoke more stable brain activity patterns across people.
 
@@
 of 5\% accuracy using fewer components than the paragraph condition decoders
-($t(XXX) = XXX, p = XXX$) or word condition decoders ($t(XXX) = XXX, p = XXX$),
-and decoding accuracy never exceeded 5\% for the rest condition. This suggests
-that brain activity patterns evoked by cognitively richer conditions are more
-compressible, such that representing the data using the same number of
+($t(99) = -7.429, p < 0.001)$) or word condition decoders ($t(99) = -7.300, p <
+0.001)$), and decoding accuracy never exceeded 5\% for the rest condition. This
+suggests that brain activity patterns evoked by cognitively richer conditions
+are more compressible, such that representing the data using the same number of
 principal components provides more information to the temporal decoders
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{figs/information_and_compression_over_time}
+  \includegraphics[width=\textwidth]{figs/information_and_compression_over_time}
 
@@
 family of curves is plotted in the same format as Figure~\ref{fig:inflection}A
-but reflects data only from one third of the dataset. \textbf{B. Inflection
-points by condition and segment.} The dots re-plot the inflection points from
-Figure~\ref{fig:inflection}D for reference. The numbers denote the inflection
-points for each third of the dataset (1: first third; 2: second third; 3: third
-third; colors denote experimental conditions). \textbf{C. Change in decoding
-accuracy over time, by number of components.} For each number of components
-($x$-axis) and condition (color), we fit a regression line to the decoding
-accuracies obtained for the first, second, and third thirds of the dataset
-(corresponding to the left, middle, and right columns of Panel A,
-respectively). The $y$-axis denotes the slopes of the regression lines. All
-error ribbons denote bootstrap-estimated 95\% confidence intervals.}
+but reflects data only from one quarter of the dataset. \textbf{B.
+Informativeness versus compressibility by condition and segment.} Each scatter
+plot is in the same format as Figure~\ref{fig:inflection}E, but reflects data
+only from one quarter of the dataset. \textbf{C. Change in decoding accuracy
+over time, by number of components.} For each number of components ($x$-axis)
+and condition (color), we fit a regression line to the decoding accuracies
+obtained for the first, second, third, and fourth quarters of the dataset
+(corresponding to the columns of Panels A and B). The $y$-axis denotes the
+slopes of the regression lines. All error ribbons denote bootstrap-estimated
+95\% confidence intervals.}
 
@@
 both decoding accuracy and compressibility, as reflected by the change in
-decoding curves, increased with listening time (intact: $t(XXX) = XXX, p =
-XXX$; paragraph: $t(XXX) = XXX, p = XXX$). These changes may reflect an
-increase in comprehension or depth of processing with listening time. In
-contrast, the decoding accuracy and compressibility \textit{decreased} with
-listening time in the word condition ($t(XXX) = XXX, p = XXX$) and rest
-condition ($t(XXX) = XXX, p = XXX$). This might reflect the depletion of
-attentional resources in the less-engaging word and rest conditions.
+decoding curves, increased with listening time (e.g., at the annotated
+reference point of $k = 20$ components: intact: $t(99) = 7.915, p < 0.001)$;
+paragraph: $t(99) = 2.354, p = 0.021)$). These changes may reflect an increase
+in comprehension or depth of processing with listening time. In contrast, the
+decoding accuracy and compressibility \textit{decreased} with listening time in
+the word condition ($t(99) = -10.747, p < 0.001)$) and rest condition ($t(99) =
+-22.081, p < 0.001)$). This might reflect the depletion of attentional
+resources in the less-engaging word and rest conditions.
 

### 2023-03-08 code/notebooks/helpers.py
working on network analyses

@@ -3,6 +3,7 @@ import pandas as pd
 import nibabel as nib
 import nilearn as nl
 from nilearn.maskers import NiftiMasker
+from matplotlib.colors import LinearSegmentedColormap
 
 import os
 import warnings
@@ -71,8 +72,8 @@ def nii2cmu(nifti_file, mask_file=None):
     R = np.array(np.dot(vox_coords, S[0:3, 0:3])) + S[:3, 3]
 
     # center on the MNI152 brain (hard code this in)
-    mni_center = np.array([0.55741881, -21.52140703, 9.83783098])
-    R = R - R.mean(axis=0) + mni_center
+    #mni_center = np.array([0.55741881, -21.52140703, 9.83783098])
+    #R = R - R.mean(axis=0) + mni_center
     
     return {'Y': Y, 'R': R}
 
@@ -116,3 +117,48 @@ def cmu2nii(Y, R, template=None):
             data[locs[j, 0], locs[j, 1], locs[j, 2], i] = Y[i, j]
     
     return nib.Nifti1Image(data, affine=img.affine)
+
+
+# colormap tools-- source: https://towardsdatascience.com/beautiful-custom-colormaps-with-matplotlib-5bab3d1f0e72
+def hex_to_rgb(value):
+    '''
+    Converts hex to rgb colours
+    value: string of 6 characters representing a hex colour.
+    Returns: list length 3 of RGB values'''
+    value = value.strip("#") # removes hash symbol if present
+    lv = len(value)
+    return tuple(int(value[i:i + lv // 3], 16) for i in range(0, lv, lv // 3))
+
+
+def rgb_to_dec(value):
+    '''
+    Converts rgb to decimal colours (i.e. divides each value by 256)
+    value: list (length 3) of RGB values
+    Returns: list (length 3) of decimal values'''
+    return [v/256 for v in value]
+
+def get_continuous_cmap(hex_list, float_list=None, name='cmap', N=256):
+    ''' creates and returns a color map that can be used in heat map figures.
+        If float_list is not provided, colour map graduates linearly between each color in hex_list.
+        If float_list is provided, each color in hex_list is mapped to the respective location in float_list. 
+        
+        Parameters
+        ----------
+        hex_list: list of hex code strings
+        float_list: list of floats between 0 and 1, same length as hex_list. Must start with 0 and end with 1.
+        
+        Returns
+        ----------
+        colour map'''
[…]

### 2023-03-09 code/notebooks/helpers.py
got htfa network labels working!

@@ -137,6 +137,14 @@ def rgb_to_dec(value):
     Returns: list (length 3) of decimal values'''
     return [v/256 for v in value]
 
+def get_cmap(hex_list, first_color=(0.5, 0.5, 0.5, 0.0), name='custom', N=256):
+    cmap_list = [first_color]
+    cmap_list.append([hex_to_rgb(c) for c in hex_list])
+
+    # create the new map
+    return LinearSegmentedColormap.from_list(name, cmap_list, N=N)
+
+
 def get_continuous_cmap(hex_list, float_list=None, name='cmap', N=256):
     ''' creates and returns a color map that can be used in heat map figures.
         If float_list is not provided, colour map graduates linearly between each color in hex_list.

### 2023-03-09 code/notebooks/helpers.py
code reorganization to facilitate re-use in network analyses
working on network analyses (running PCA by condition + network)

@@ -1,12 +1,33 @@
 import numpy as np
 import pandas as pd
 import nibabel as nib
-import nilearn as nl
+import datawrangler as dw
+import seaborn as sns
+
 from nilearn.maskers import NiftiMasker
-from matplotlib.colors import LinearSegmentedColormap
+from matplotlib import pyplot as plt
+from sklearn.decomposition import IncrementalPCA as PCA
 
 import os
 import warnings
+import pickle
+
+basedir = os.path.split(os.path.split(os.getcwd())[0])[0]
+datadir = os.path.join(basedir, 'data')
+figdir = os.path.join(basedir, 'paper', 'figs', 'source')
+
+scratch_dir = os.path.join(basedir, 'data', 'scratch')
+if not os.path.exists(scratch_dir):
+    os.makedirs(scratch_dir)
+
+condition_colors = {
+    'intact': '#21409A',
+    'paragraph': '#00A14B',
+    'word': '#FFDE17',
+    'rest': '#7F3F98'
+}
+
+conditions = list(condition_colors.keys())
 
 
 def nii2cmu(nifti_file, mask_file=None):
@@ -119,54 +140,104 @@ def cmu2nii(Y, R, template=None):
     return nib.Nifti1Image(data, affine=img.affine)
 
 
-# colormap tools-- source: https://towardsdatascience.com/beautiful-custom-colormaps-with-matplotlib-5bab3d1f0e72
-def hex_to_rgb(value):
-    '''
-    Converts hex to rgb colours
-    value: string of 6 characters representing a hex colour.
-    Returns: list length 3 of RGB values'''
-    value = value.strip("#") # removes hash symbol if present
-    lv = len(value)
-    return tuple(int(value[i:i + lv // 3], 16) for i in range(0, lv, lv // 3))
+def group_pca(data, n_components=None, fname=None):
+    if fname is not None:
+        if os.path.exists(fname):
+            with open(fname, 'rb') as f:
+                return pickle.load(f)
+    
+    pca = PCA(n_components=n_components)
+    
+    x = dw.stack(data)
+    y = pca.fit_transform(x)
+
+    y = dw.unstack(pd.DataFrame(index=x.index, data=y))
+
+    if fname is not None:
+        with open(fname, 'wb') as f:
+            pickle.dump((y, pca), f)
+    
+    return y, pca
 
 
-def rgb_to_dec(value):
-    '''
-    Converts rgb to decimal colours (i.e. divides each value by 256)
-    value: list (length 3) of RGB values
[…]

### 2023-03-09 paper/main.tex
did some code reorganization, added network figure (fig 4) + updated caption

@@
   \centering
-  \includegraphics[width=0.7\textwidth]{figs/decode_pcs_network}
+  \includegraphics[width=0.9\textwidth]{figs/network_results}
 
   \caption{\textbf{Network-specific decoding accuracy and compression.}
-  \textbf{A. Decoding accuracy and number of components for network-specific
-  inflection points.} We considered the seven networks identified by
-  \cite{YeoEtal11}. We computed each network's inflection point, for each
-  experimental condition, using the procedure described in
-  Figure~\ref{fig:inflection}C. \textbf{B. Network-specific decoding
-  accuracy.} Each of the seven networks are colored according to the decoding
-  accuracy at the network's inflection point for the ``intact'' experimental
-  condition (corresponding to the dark blue curve in Panel A). \textbf{C.
-  Network-specific compression.} Each of the seven networks are colored
-  according to the number of components at the network's inflection point for
-  the intact experimental condition. Larger numbers of components reflect lower
-  compressibility.}
+  \textbf{A. Network parcellation map.} The glass brain displays the location
+  of each of seven networks identified by \cite{YeoEtal11}. \textbf{B. Node
+  labels.} We assigned network labels to each of 700 Hierarchical Topographic
+  Factor Analysis-derived nodes (see \textit{Hierarchical topographic factor
+  analysis (HTFA)}). For each node, we evaluated its radial basis function
+  (RBF) at the locations of every voxel in the parcellation map displayed in
+  Panel A. We summed up these RBF weights separately for the voxels in each
+  network. We defined each node's network label as its highest-weighted network
+  across all voxels. \textbf{C. Per-network node counts.} The bar plot displays
+  the total number of HTFA-derived nodes assigned to each network. \textbf{D.
+  Informativeness versus compressibility by network and condition.} Each dot
+  corresponds to a single cross validation run. The dots' coordinates denote
+  the minimum proportion of components (relative to the number of nodes in the
+  given network) required to achieve at least 5\% corrected decoding accuracy
+  ($x$-coordinate; for the rest condition these proportions are estimated using
+  the peak decoding accuracies) and peak (corrected) decoding accuracy
+  ($y$-coordinate). We repeated the analysis separately for each network
+  (color) and experimental condition (column). \textbf{E. Informativeness by
+  network.} For each network, sorted (roughly) from lower-order networks on the
+  left to higher-order networks on the right, the curves display the peak
+  (corrected) decoding accuracy for each experimental condition (color).
+  \textbf{F. Compressibility by network.} For each network, the curves display
+  the proportion of components (relative to the total possible number of
+  components displayed in Panel C) required to achieve at least 5\% decoding
+  accuracy (or, for the rest condition, peak decoding accuracy, as described
+  above). Error ribbons in Panels E and F denote 95\% bootstrap-estimated
+  confidence intervals.}
 

### 2023-03-09 code/notebooks/helpers.py

@@ -7,6 +7,8 @@ import seaborn as sns
 from nilearn.maskers import NiftiMasker
 from matplotlib import pyplot as plt
 from sklearn.decomposition import IncrementalPCA as PCA
+from tqdm import tqdm
+from scipy.spatial.distance import cdist
 
 import os
 import warnings
@@ -169,7 +171,7 @@ def accuracy(train, test):
     return np.mean([i == d for i, d in enumerate(labels)]) - 1 / len(labels)
 
 
-def cross_validation(data, n_iter=10, fname=None):
+def cross_validation(data, n_iter=10, fname=None, max_components=700):
     if fname is not None:
         if os.path.exists(fname):
             with open(fname, 'rb') as f:
@@ -227,6 +229,8 @@ def ridge_plot(x, column='Number of components', fname=None, xlim=[-99, 700], hu
     if fname is not None:
         g.savefig(os.path.join(figdir, fname + '.pdf'), bbox_inches='tight')
 
+    return fig
+
 
 def get_data():
     url = 'https://www.dropbox.com/s/29a48lv3j5ybcvw/pieman2_htfa.pkl?dl=1'
@@ -241,3 +245,42 @@ def get_data():
         data = pickle.load(open(fname, 'rb'))
     
     return data
+
+
+def info_and_compressibility(d, target=0.05):
+    def closest(x, target):
+        dists = np.abs(x.values - target)
+        dists[x.values < target] += 10 * np.max(dists)
+        return int(x.index.values[np.argmin(dists)])
+
+    df = []
+    for c in conditions:
+        dc = d[c].astype(float).pivot(index='Iteration', columns='Number of components', values='Relative decoding accuracy')
+        i = pd.DataFrame()
+        i['Number of components'] = dc.apply(lambda x: closest(x, target), axis=1, raw=False)
+        i['Relative decoding accuracy'] = dc.max(axis=1)
+        i['Condition'] = c
+        i['Iteration'] = dc.index.values.astype(int)
+        df.append(i)
+    return pd.concat(df, ignore_index=True, axis=0)
+
+
+def plot_info_and_compressibility_scatter(x, fname=None):
+    fig = plt.figure(figsize=(4, 3))
+    ax = plt.gca()
+
+    x = info_and_compressibility(x)
[…]

### 2023-03-09 code/notebooks/helpers.py
working on neurosynth decoding analysis

@@ -284,3 +284,6 @@ def plot_info_and_compressibility_scatter(x, fname=None):
         fig.savefig(os.path.join(figdir, fname + '.pdf'), bbox_inches='tight')
     
     return fig
+
+def rbf(R, center, width):
+    return np.exp(-np.sum((R - center) ** 2, axis=1) / width)

### 2023-03-09 code/notebooks/helpers.py
backing up preliminary decoding results

@@ -71,13 +71,19 @@ def nii2cmu(nifti_file, mask_file=None):
     with warnings.catch_warnings():
         warnings.simplefilter("ignore")
         
-        img = nib.load(nifti_file)
+        if type(nifti_file) == str:
+            img = nib.load(nifti_file)            
+        elif type(nifti_file) == nib.nifti1.Nifti1Image:
+            img = nifti_file
+        else:
+            raise ValueError('nifti_file must be a filename or nibabel Nifti1Image object')
+        
         mask = NiftiMasker(mask_strategy='background')
         if mask_file is None:
             mask.fit(nifti_file)
         else:
             mask.fit(mask_file)
-    
+
     hdr = img.header
     S = img.get_sform()
     vox_size = hdr.get_zooms()

### 2023-03-12 code/decode_helpers/decode_helpers/__init__.py
got rid of unused code

@@ -1,4 +1,0 @@
-from pathlib import Path
-
-version_info = (0, 0, 1)
-__version__ = '.'.join(map(str, version_info))

### 2023-03-12 code/decode_helpers/decode_helpers/helpers.py

@@ -1,280 +1,0 @@
-# coding: utf-8
-from __future__ import division
-import numpy as np
-import scipy.spatial.distance as sd
-from scipy.special import gamma
-from scipy.linalg import toeplitz
-from sklearn import linear_model
-from sklearn.base import BaseEstimator
-from scipy.spatial.distance import pdist, cdist
-from scipy.stats import pearsonr
-from sklearn.feature_selection import RFE
-from sklearn.model_selection import StratifiedKFold
-from sklearn.feature_selection import RFECV
-from scipy.optimize import minimize
-from scipy.stats import ttest_1samp as ttest
-import hypertools as hyp
-import pandas as pd
-import warnings
-from matplotlib import pyplot as plt
-
-
-def r2z(r):
-    """
-    Function that calculates the Fisher z-transformation
-
-    Parameters
-    ----------
-    r : int or ndarray
-        Correlation value
-
-    Returns
-    ----------
-    result : int or ndarray
-        Fishers z transformed correlation value
-
-    """
-    return 0.5*(np.log(1+r) - np.log(1-r))
-
-
-def z2r(z):
-    """
-    Function that calculates the inverse Fisher z-transformation
-
-    Parameters
-    ----------
-    z : int or ndarray
-        Fishers z transformed correlation value
-
-    Returns
-    ----------
-    result : int or ndarray
-        Correlation value
-
-    """
-    r = np.divide((np.exp(2*z) - 1), (np.exp(2*z) + 1))
-    r[np.isnan(r)] = 0
-    r[np.isinf(r)] = np.sign(r)[np.isinf(r)]
-    return r
-
-
-def mean_combine(vals):
-    '''
-    Compute the element-wise mean across each matrix in a list.
-
-    :param vals: a matrix, or a list of matrices
-    :return: a mean matrix
-    '''
-    if not (type(vals) == list):
-        return vals
-    else:
-        return np.mean(np.stack(vals, axis=2), axis=2)
-
-def corrmean_combine(corrs):
-    '''
-    Compute the mean element-wise correlation across each matrix in a list.
-
-    :param corrs: a matrix of vectorized correlation matrices (output of mat2vec), or a list
-                  of such matrices
-    :return: a mean vectorized correlation matrix
-    '''
-    if not (type(corrs) == list):
-        return corrs
-
-    elif np.shape(corrs)[0] == 1:
-        return corrs
-
-    else:
-        return z2r(np.mean(r2z(np.stack(corrs, axis=2)), axis=2))
[…]

### 2023-03-12 code/decode_helpers/setup.py

@@ -1,6 +1,0 @@
-# -*- coding: utf-8 -*-
-from setuptools import setup
-
-
-# reads from setup.cfg
-setup()

### 2023-03-12 code/scripts/config.py

@@ -1,66 +1,0 @@
-import os
-import socket
-
-config = dict()
-
-config['template'] = 'run_job.sh'
-
-local = False
-
-# ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-
-if local:
-    config['datadir'] = '/Users/lucyowen/repos/pca_paper-1/data'
-    config['workingdir'] = '/Users/lucyowen/repos/pca_paper-1'
-    config['startdir'] = os.path.dirname(os.path.dirname(os.path.realpath(__file__)))  # directory to start the job in
-    config['template'] = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'run_job_local.sh')
-else:
-    config['datadir'] = '/dartfs-hpc/rc/home/2/f002s72/Discovery_home/repos/pca_paper-1/data'
-    config['workingdir'] = '/dartfs/rc/lab/D/DBIC/CDL/f002s72/pca_paper/pieman'
-    config['startdir'] = os.path.dirname(os.path.dirname(os.path.realpath(__file__)))
-    config['template'] = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'run_job.sh')
-
-# if (socket.gethostname() == 'discovery7.hpcc.dartmouth.edu') or (socket.gethostname() == 'ndoli.hpcc.dartmouth.edu'):
-#     config['datadir'] = '/dartfs-hpc/rc/home/2/f002s72/Discovery_home/repos/pca_paper-1/data'
-#     config['workingdir'] = '/dartfs-hpc/rc/home/2/f002s72/Discovery_home/repos/pca_paper-1/'
-#     config['startdir'] = os.path.dirname(os.path.dirname(os.path.realpath(__file__)))
-#     config['template'] = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'run_job.sh')
-# else:
-#     config['datadir'] = '/Users/lucyowen/repos/pca_paper-1/data'
-#     config['workingdir'] = '/Users/lucyowen/repos/pca_paper-1'
-#     config['startdir'] = os.path.dirname(os.path.dirname(os.path.realpath(__file__)))  # directory to start the job in
-#     config['template'] = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'run_job_local.sh')
-
-# local = True
-# if local:
[…]

### 2023-03-12 code/scripts/pieman_pca_analysis.py

@@ -1,96 +1,0 @@
-from decode_helpers.helpers import pca_decoder
-from scipy.io import loadmat
-import numpy as np
-import sys
-import os
-from config import config
-import pandas as pd
-
-
-cond = sys.argv[1]
-chunk = sys.argv[2]
-reps = sys.argv[3]
-rfun = sys.argv[4]
-ndims = sys.argv[5]
-
-if len(sys.argv) < 7:
-    debug = False
-else:
-    debug = eval(sys.argv[6])
-
-result_name = 'pca_decode'
-
-
-if debug:
-    results_dir = os.path.join(config['resultsdir'], result_name, rfun + '_debug', 'ndims_'+ ndims)
-
-else:
-    results_dir = os.path.join(config['resultsdir'], result_name, rfun , 'ndims_'+ ndims)
-
-try:
-    if not os.path.exists(results_dir):
-        os.makedirs(results_dir)
-except OSError as err:
-   print(err)
-
-
-factors = 700
-
-if factors == 100:
-    pieman_name = 'pieman_ica100.mat'
-else:
-    pieman_name = 'pieman_data.mat'
-
-pieman_data = loadmat(os.path.join(config['datadir'], pieman_name))
-pieman_conds = ['intact', 'paragraph', 'word', 'rest']
-
-
-
-if debug:
-    data = []
-    conds = []
-    for c in pieman_conds:
-        next_data = list(map(lambda i: pieman_data[c][:, i][0][:30, :10], np.where(np.arange(pieman_data[c].shape[1]) != 3)[0]))
-        data.extend(next_data)
-        conds.extend([c]*len(next_data))
-    del pieman_data
-
-else:
-
-    data = []
-    conds = []
-    for c in pieman_conds:
-        if c == 'paragraph':
-            if factors == 700:
-                next_data = list(map(lambda i: pieman_data[c][:, i][0], np.where(np.arange(pieman_data[c].shape[1]) != 3)[0]))
-            else:
-                next_data = list(map(lambda i: pieman_data[c][:, i][0], np.where(np.arange(pieman_data[c].shape[1]) != 0)[0]))
-        else:
-            next_data = list(map(lambda i: pieman_data[c][:, i][0], np.arange(pieman_data[c].shape[1])))
-        data.extend(next_data)
-        conds.extend([c]*len(next_data))
-    del pieman_data
-
-
-data = np.array(data)
-conds = np.array(conds)
-
[…]

### 2023-03-12 code/scripts/pieman_pca_analysis_submit.py

@@ -1,160 +1,0 @@
-#!/usr/bin/python
-
-# create a bunch of job scripts
-
-from config import config
-from subprocess import call
-import numpy as np
-import os
-import socket
-import getpass
-import datetime as dt
-
-
-# ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-
-try:
-    os.stat(config['resultsdir'])
-except:
-    os.makedirs(config['resultsdir'])
-
-# each job command should be formatted as a string
-job_script = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'pieman_pca_analysis.py')
-
-cond_type = ['intact', 'paragraph', 'word', 'rest']
-
-
-# options for reps: integer
-reps =  str('100')
-
-# options for reps: rfuns
-rfuns =  [str('PCA')]
-
-
-# options for dims: integers
-dims = ['700']
-
-# options for debug: True or False
-debug = str('False')
-
-job_commands = list(np.array([list(map(lambda x: x[0]+" "+str(x[1])+" "+'unchunked'+" "+str(r)+
-                                                 " "+rfuns[0]+" "+ dims[0]+ " "+debug,
-                                       zip([job_script]*len(cond_type), cond_type)))
-                              for r in range(int(reps))]).flat)
-
-# job_names should specify the file name of each script (as a list, of the same length as job_commands)
-job_names = list(np.array([list(map(lambda x: os.path.basename(os.path.splitext(x)[0])+'_'+'unchunked'+'_'+str(r)+'_'
-                                              '_'+rfuns[0]+'_'+ dims[0] + '_'+debug+'.sh', cond_type))
-                           for r in range(int(reps))]).flat)
-
-
-
-# ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-
-assert(len(job_commands) == len(job_names))
-
-
-# job_command is referenced in the run_job.sh script
-# noinspection PyBroadException,PyUnusedLocal
-def create_job(name, job_command):
-    # noinspection PyUnusedLocal,PyShadowingNames
-    def create_helper(s, job_command):
-        x = [i for i, char in enumerate(s) if char == '<']
-        y = [i for i, char in enumerate(s) if char == '>']
-        if len(x) == 0:
-            return s
-
-        q = ''
-        index = 0
-        for i in range(len(x)):
-            q += s[index:x[i]]
-            unpacked = eval(s[x[i]+1:y[i]])
-            q += str(unpacked)
[…]

### 2023-03-12 code/scripts/pieman_pca_chunk.py

@@ -1,121 +1,0 @@
-from decode_helpers.helpers import pca_decoder
-from scipy.io import loadmat
-import numpy as np
-import sys
-import os
-from config import config
-import pandas as pd
-
-
-cond = sys.argv[1]
-chunk = sys.argv[2]
-reps = sys.argv[3]
-rfun = sys.argv[4]
-ndims = sys.argv[5]
-
-if len(sys.argv) < 7:
-    debug = False
-else:
-    debug = eval(sys.argv[6])
-
-result_name = 'pca_decode_chunked'
-
-if debug:
-    results_dir = os.path.join(config['resultsdir'], result_name, rfun + '_debug', 'ndims_'+ ndims)
-
-else:
-    results_dir = os.path.join(config['resultsdir'], result_name, rfun , 'ndims_'+ ndims)
-
-try:
-    if not os.path.exists(results_dir):
-        os.makedirs(results_dir)
-except OSError as err:
-   print(err)
-
-
-factors = 700
-
-if factors == 100:
-    pieman_name = 'pieman_ica100.mat'
-else:
-    pieman_name = 'pieman_data.mat'
-
-pieman_data = loadmat(os.path.join(config['datadir'], pieman_name))
-pieman_conds = ['intact', 'paragraph', 'word', 'rest']
-
-
-if debug:
-    data_chunks = [0] * 3
-    conds_chunks = [0] * 3
-    divided = 0
-    for third in list(range(3)):
-        data = []
-        conds = []
-
-        for c in pieman_conds:
-            next_data = list(map(lambda i: pieman_data[c][:, i][0][divided:divided+10,:20], np.arange(4)))
-            data.extend(next_data)
-            conds.extend([c]*len(next_data))
-
-        conds_chunks[third] = conds
-        data_chunks[third] = data
-        divided += 10
-
-    del pieman_data
-
-else:
-
-    data_chunks = [0] * 3
-    conds_chunks = [0] * 3
-    for third in list(range(3)):
-        data = []
-        conds = []
-        for c in pieman_conds:
-            timechunk = int(np.round(pieman_data[c][0][0].shape[0]/3, 0))
-            divided = third * timechunk
-            if c == 'paragraph':
-                if factors == 700:
-                    next_data = list(
-                        map(lambda i: pieman_data[c][:, i][0][divided:divided+timechunk,:], np.where(np.arange(pieman_data[c].shape[1]) != 3)[0]))
-                else:
-                    next_data = list(
[…]

### 2023-03-12 code/scripts/pieman_pca_chunk_submit.py

@@ -1,157 +1,0 @@
-#!/usr/bin/python
-
-# create a bunch of job scripts
-from config import config
-from subprocess import call
-import numpy as np
-import os
-import socket
-import getpass
-import datetime as dt
-
-
-# ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-
-try:
-    os.stat(config['resultsdir'])
-except:
-    os.makedirs(config['resultsdir'])
-
-# each job command should be formatted as a string
-job_script = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'pieman_pca_chunk.py')
-
-cond_type = ['intact', 'paragraph', 'word', 'rest']
-
-# options for reps: integer
-reps =  str('1')
-
-# options for reps: rfuns
-rfuns =  [str('PCA')]
-
-# options for dims: integers
-dims = ['700']
-
-# options for debug: True or False
-debug = str('False')
-
-job_commands = list(np.array([list(map(lambda x: x[0]+" "+str(x[1])+" "+'chunked'+" "+str(r)+
-                                                 " "+rfuns[0]+" "+ dims[0]+ " "+debug,
-                                       zip([job_script]*len(cond_type), cond_type)))
-                              for r in range(int(reps))]).flat)
-
-# job_names should specify the file name of each script (as a list, of the same length as job_commands)
-job_names = list(np.array([list(map(lambda x: os.path.basename(os.path.splitext(x)[0])+'_'+'chunked'+'_'+str(r)+'_'
-                                              '_'+rfuns[0]+'_'+ dims[0] + '_'+debug+'.sh', cond_type))
-                           for r in range(int(reps))]).flat)
-
-
-
-# ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-
-assert(len(job_commands) == len(job_names))
-
-
-# job_command is referenced in the run_job.sh script
-# noinspection PyBroadException,PyUnusedLocal
-def create_job(name, job_command):
-    # noinspection PyUnusedLocal,PyShadowingNames
-    def create_helper(s, job_command):
-        x = [i for i, char in enumerate(s) if char == '<']
-        y = [i for i, char in enumerate(s) if char == '>']
-        if len(x) == 0:
-            return s
-
-        q = ''
-        index = 0
-        for i in range(len(x)):
-            q += s[index:x[i]]
-            unpacked = eval(s[x[i]+1:y[i]])
-            q += str(unpacked)
[…]

### 2023-03-12 code/scripts/pieman_pca_elimination_analysis.py

@@ -1,95 +1,0 @@
-from decode_helpers.helpers import pca_decoder_rfe
-from scipy.io import loadmat
-import numpy as np
-import sys
-import os
-from config import config
-import pandas as pd
-
-
-cond = sys.argv[1]
-chunk = sys.argv[2]
-reps = sys.argv[3]
-rfun = sys.argv[4]
-ndims = sys.argv[5]
-
-if len(sys.argv) < 7:
-    debug = False
-else:
-    debug = eval(sys.argv[6])
-
-result_name = 'pca_decode_elimination'
-
-
-if debug:
-    results_dir = os.path.join(config['resultsdir'], result_name, rfun + '_debug', 'ndims_'+ ndims)
-
-else:
-    results_dir = os.path.join(config['resultsdir'], result_name, rfun , 'ndims_'+ ndims)
-
-try:
-    if not os.path.exists(results_dir):
-        os.makedirs(results_dir)
-except OSError as err:
-   print(err)
-
-
-factors = 700
-
-if factors == 100:
-    pieman_name = 'pieman_ica100.mat'
-else:
-    pieman_name = 'pieman_data.mat'
-
-pieman_data = loadmat(os.path.join(config['datadir'], pieman_name))
-pieman_conds = ['intact', 'paragraph', 'word', 'rest']
-
-
-
-if debug:
-    data = []
-    conds = []
-    for c in pieman_conds:
-        next_data = list(map(lambda i: pieman_data[c][:, i][0][:30, :10], np.where(np.arange(pieman_data[c].shape[1]) != 3)[0]))
-        data.extend(next_data)
-        conds.extend([c]*len(next_data))
-    del pieman_data
-
-else:
-
-    data = []
-    conds = []
-    for c in pieman_conds:
-        if c == 'paragraph':
-            if factors == 700:
-                next_data = list(map(lambda i: pieman_data[c][:, i][0], np.where(np.arange(pieman_data[c].shape[1]) != 3)[0]))
-            else:
-                next_data = list(map(lambda i: pieman_data[c][:, i][0], np.where(np.arange(pieman_data[c].shape[1]) != 0)[0]))
-        else:
-            next_data = list(map(lambda i: pieman_data[c][:, i][0], np.arange(pieman_data[c].shape[1])))
-        data.extend(next_data)
-        conds.extend([c]*len(next_data))
-    del pieman_data
-
-
-data = np.array(data)
-conds = np.array(conds)
[…]

### 2023-03-12 code/scripts/pieman_pca_elimination_analysis_submit.py

@@ -1,160 +1,0 @@
-#!/usr/bin/python
-
-# create a bunch of job scripts
-
-from config import config
-from subprocess import call
-import numpy as np
-import os
-import socket
-import getpass
-import datetime as dt
-
-
-# ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-
-try:
-    os.stat(config['resultsdir'])
-except:
-    os.makedirs(config['resultsdir'])
-
-# each job command should be formatted as a string
-job_script = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'pieman_pca_elimination_analysis.py')
-
-cond_type = ['intact', 'paragraph', 'word', 'rest']
-
-
-# options for reps: integer
-reps =  str('10')
-
-# options for reps: rfuns
-rfuns =  [str('PCA')]
-
-
-# options for dims: integers
-dims = ['700']
-
-# options for debug: True or False
-debug = str('False')
-
-job_commands = list(np.array([list(map(lambda x: x[0]+" "+str(x[1])+" "+'elim_unchunked'+" "+str(r)+
-                                                 " "+rfuns[0]+" "+ dims[0]+ " "+debug,
-                                       zip([job_script]*len(cond_type), cond_type)))
-                              for r in range(int(reps))]).flat)
-
-# job_names should specify the file name of each script (as a list, of the same length as job_commands)
-job_names = list(np.array([list(map(lambda x: os.path.basename(os.path.splitext(x)[0])+'_'+'elim_unchunked'+'_'+str(r)+'_'
-                                              '_'+rfuns[0]+'_'+ dims[0] + '_'+debug+'.sh', cond_type))
-                           for r in range(int(reps))]).flat)
-
-
-
-# ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-
-assert(len(job_commands) == len(job_names))
-
-
-# job_command is referenced in the run_job.sh script
-# noinspection PyBroadException,PyUnusedLocal
-def create_job(name, job_command):
-    # noinspection PyUnusedLocal,PyShadowingNames
-    def create_helper(s, job_command):
-        x = [i for i, char in enumerate(s) if char == '<']
-        y = [i for i, char in enumerate(s) if char == '>']
-        if len(x) == 0:
-            return s
-
-        q = ''
-        index = 0
-        for i in range(len(x)):
-            q += s[index:x[i]]
-            unpacked = eval(s[x[i]+1:y[i]])
[…]

### 2023-03-12 code/scripts/pieman_pca_network_analysis.py

@@ -1,102 +1,0 @@
-from decode_helpers.helpers import pca_decoder
-from scipy.io import loadmat
-import numpy as np
-import sys
-import os
-import supereeg as se
-from config import config
-import pandas as pd
-
-
-cond = sys.argv[1]
-chunk = sys.argv[2]
-reps = sys.argv[3]
-rfun = sys.argv[4]
-ndims = sys.argv[5]
-
-if len(sys.argv) < 7:
-    debug = False
-else:
-    debug = eval(sys.argv[6])
-
-result_name = 'pca_decode_network'
-
-if debug:
-    results_dir = os.path.join(config['resultsdir'], result_name, rfun + '_debug', 'ndims_'+ ndims)
-
-else:
-    results_dir = os.path.join(config['resultsdir'], result_name, rfun , 'ndims_'+ ndims)
-
-try:
-    if not os.path.exists(results_dir):
-        os.makedirs(results_dir)
-except OSError as err:
-   print(err)
-
-
-
-n_nets = 7
-
-print(str(n_nets))
-
-factors = 700
-
-if factors == 100:
-    pieman_name = 'pieman_ica100.mat'
-else:
-    pieman_name = 'pieman_data.mat'
-
-pieman_data = loadmat(os.path.join(config['datadir'], pieman_name))
-pieman_conds = ['intact', 'paragraph', 'word', 'rest']
-
-network_file = os.path.join(config['datadir'],'yeo_networks', 'networks.npz')
-
-network_npz = np.load(network_file)
-network_locs = network_npz['locs']
-network_data = network_npz['data']
-
-network_list = ['Visual', 'Somatomotor', 'Dorsal Attention', 'Ventral Attention',
-                'Limbic ', 'Frontoparietal', 'Default']
-
-for n in np.arange(n_nets):
-    mask_n = network_data == n + 1
-    mask_n_flat = mask_n.ravel()
-    data = []
-    conds = []
-
-
-    for c in pieman_conds:
-        if c == 'paragraph':
-            if factors == 700:
-                next_data = list(map(lambda i: pieman_data[c][:, i][0][:, mask_n_flat], np.where(np.arange(pieman_data[c].shape[1]) != 3)[0]))
-            else:
-                next_data = list(map(lambda i: pieman_data[c][:, i][0][:, mask_n_flat], np.where(np.arange(pieman_data[c].shape[1]) != 0)[0]))
-        else:
[…]

### 2023-03-12 code/scripts/pieman_pca_network_analysis_submit.py

@@ -1,158 +1,0 @@
-#!/usr/bin/python
-
-# create a bunch of job scripts
-from config import config
-from subprocess import call
-import numpy as np
-import os
-import socket
-import getpass
-import datetime as dt
-
-
-# ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-
-try:
-    os.stat(config['resultsdir'])
-except:
-    os.makedirs(config['resultsdir'])
-
-# each job command should be formatted as a string
-job_script = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'pieman_pca_network_analysis.py')
-
-cond_type = ['intact', 'paragraph', 'word', 'rest']
-
-# options for reps: integer
-reps =  str('100')
-
-# options for reps: rfuns
-rfuns =  [str('PCA')]
-
-
-# options for dims: integers
-dims = ['700']
-
-# options for debug: True or False
-debug = str('False')
-
-job_commands = list(np.array([list(map(lambda x: x[0]+" "+str(x[1])+" "+'unchunked'+" "+str(r)+
-                                                 " "+rfuns[0]+" "+ dims[0]+ " "+debug,
-                                       zip([job_script]*len(cond_type), cond_type)))
-                              for r in range(int(reps))]).flat)
-
-# job_names should specify the file name of each script (as a list, of the same length as job_commands)
-job_names = list(np.array([list(map(lambda x: os.path.basename(os.path.splitext(x)[0])+'_'+'unchunked'+'_'+str(r)+'_'
-                                              +rfuns[0]+'_'+ dims[0] + '_'+debug+'.sh', cond_type))
-                           for r in range(int(reps))]).flat)
-
-
-
-# ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-
-assert(len(job_commands) == len(job_names))
-
-
-# job_command is referenced in the run_job.sh script
-# noinspection PyBroadException,PyUnusedLocal
-def create_job(name, job_command):
-    # noinspection PyUnusedLocal,PyShadowingNames
-    def create_helper(s, job_command):
-        x = [i for i, char in enumerate(s) if char == '<']
-        y = [i for i, char in enumerate(s) if char == '>']
-        if len(x) == 0:
-            return s
-
-        q = ''
-        index = 0
-        for i in range(len(x)):
-            q += s[index:x[i]]
-            unpacked = eval(s[x[i]+1:y[i]])
-            q += str(unpacked)
[…]

### 2023-03-12 code/scripts/pieman_pca_network_chunk.py

@@ -1,130 +1,0 @@
-
-import timecorr as tc
-from timecorr.helpers import isfc, wisfc, mean_combine, corrmean_combine
-from scipy.io import loadmat
-import numpy as np
-import sys
-import os
-from config import config
-import pandas as pd
-import supereeg as se
-
-cond = sys.argv[1]
-chunk = sys.argv[2]
-reps = sys.argv[3]
-cfun = sys.argv[4]
-rfun = sys.argv[5]
-width = int(sys.argv[6])
-wp = sys.argv[7]
-ndims = sys.argv[8]
-
-if len(sys.argv) < 10:
-    debug = False
-else:
-    debug = eval(sys.argv[9])
-
-result_name = 'pca_decode_network_chunked'
-
-if debug:
-    results_dir = os.path.join(config['resultsdir'], result_name, rfun + '_debug', 'ndims_'+ ndims)
-
-else:
-    results_dir = os.path.join(config['resultsdir'], result_name, rfun , 'ndims_'+ ndims)
-
-try:
-    if not os.path.exists(results_dir):
-        os.makedirs(results_dir)
-except OSError as err:
-   print(err)
-
-
-
-laplace = {'name': 'Laplace', 'weights': tc.laplace_weights, 'params': {'scale': width}}
-delta = {'name': '$\delta$', 'weights': tc.eye_weights, 'params': tc.eye_params}
-gaussian = {'name': 'Gaussian', 'weights': tc.gaussian_weights, 'params': {'var': width}}
-mexican_hat = {'name': 'Mexican hat', 'weights': tc.mexican_hat_weights, 'params': {'sigma': width}}
-
-n_nets = 7
-
-print(str(n_nets))
-
-factors = 700
-
-if factors == 100:
-    pieman_name = 'pieman_ica100.mat'
-else:
-    pieman_name = 'pieman_data.mat'
-
-pieman_data = loadmat(os.path.join(config['datadir'], pieman_name))
-pieman_conds = ['intact', 'paragraph', 'word', 'rest']
-
-network_bo_file = os.path.join(config['datadir'], 'networks.bo')
-network_bo = se.load(network_bo_file)
-network_list = ['Visual', 'Somatomotor', 'Dorsal Attention', 'Ventral Attention',
-                'Limbic ', 'Frontoparietal', 'Default']
-
-weights_paramter = eval(wp)
-
-chunks = 3
-
-for n in np.arange(n_nets):
[…]

### 2023-03-12 code/scripts/pieman_pca_network_chunk_submit.py

@@ -1,169 +1,0 @@
-#!/usr/bin/python
-
-# create a bunch of job scripts
-from config import config
-from subprocess import call
-import numpy as np
-import os
-import socket
-import getpass
-import datetime as dt
-
-
-# ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-
-try:
-    os.stat(config['resultsdir'])
-except:
-    os.makedirs(config['resultsdir'])
-
-# each job command should be formatted as a string
-job_script = os.path.join(os.path.dirname(os.path.realpath(__file__)), 'pieman_pca_network_chunk.py')
-
-cond_type = ['intact', 'paragraph', 'word', 'rest']
-
-# options for levels: integer
-
-# options for reps: integer
-reps =  str('100')
-
-# options for reps: cfuns
-cfuns =  [str('isfc')]
-
-# options for reps: rfuns
-#rfuns =  [str('eigenvector_centrality'), str('pagerank_centrality'), str('strength')]
-rfuns =  [str('PCA')]
-
-# options for widths: integer
-widths = [str(10)]
-
-# options for weight functions: laplace, gaussian, mexican hat, delta
-weights = ['gaussian']
-
-# options for dims: integers
-dims = ['100']
-
-# options for debug: True or False
-debug = str('False')
-
-job_commands = list(np.array([list(map(lambda x: x[0]+" "+str(x[1])+" "+'chunk'+" "+str(r)+
-                                                 " "+cfuns[0]+" "+rfuns[0]+" "+widths[0]+" "+weights[0]+" "+ dims[0]+ " "+debug,
-                                       zip([job_script]*len(cond_type), cond_type)))
-                              for r in range(int(reps))]).flat)
-
-# job_names should specify the file name of each script (as a list, of the same length as job_commands)
-job_names = list(np.array([list(map(lambda x: os.path.basename(os.path.splitext(x)[0])+'_'+'chunk'+'_'+str(r)+'_'
-                                              +cfuns[0]+'_'+rfuns[0]+'_'+widths[0]+'_'+weights[0]+'_'+ dims[0] + '_'+debug+'.sh', cond_type))
-                           for r in range(int(reps))]).flat)
-
-
-
-# ====== MODIFY ONLY THE CODE BETWEEN THESE LINES ======
-
-assert(len(job_commands) == len(job_names))
-
-
-# job_command is referenced in the run_job.sh script
-# noinspection PyBroadException,PyUnusedLocal
-def create_job(name, job_command):
[…]

### 2023-03-12 code/scripts/run_job.sh

@@ -1,31 +1,0 @@
-#!/bin/bash -l
-
-# DO NOT MODIFY THIS FILE!
-# MODIFY config.py AND create_and_submit_jobs.py AS NEEDED
-
-# SLURM lines begin with "#SBATCH".  to-be-replaced text is sandwiched between angled brackets
-
-# declare a name for this job
-#SBATCH --job-name=<config['jobname']>
-
-# specify the number of cores and nodes (estimate 4GB of RAM per core)
-#SBATCH -N <config['nnodes']>
-
-#SBATCH -n <config['ppn']>
-
-# set the working directory *of this script* to the directory from which the job was submitted
-
-# set the working directory *of the job* to the specified start directory
-cd <config['startdir']>
-
-echo ACTIVATING timecorr VIRTUAL ENVIRONMENT
-
-module load python
-
-source activate pca_env
-
-# run the job
-<config['cmd_wrapper']> <job_command> #note: job_command is reserved for the job command; it should not be specified in config.py
-
-source deactivate pca_env
-

### 2023-03-12 code/scripts/test_submit.py

@@ -1,19 +1,0 @@
-import sys
-import os
-#sys.path.append(sys.path.append(os.path.dirname(os.path.dirname(os.path.abspath(__file__)))))
-
-#sys.path.append(os.path.abspath('..'))
-from decode_helpers.helpers import pca_decoder
-from scipy.io import loadmat
-import numpy as np
-import sys
-import os
-from config import config
-import pandas as pd
-import hypertools as hyp
-
-cond = sys.argv[1]
-
-
-print(cond)
-print('done')

### 2023-03-12 code/scripts/visbrain_plots/acc_by_network.py

@@ -1,157 +1,0 @@
-import supereeg as se
-import os
-import numpy as np
-from visbrain.objects import BrainObj, SceneObj, SourceObj
-from visbrain.gui import Brain
-
-#cmap = "yeo_colors_7_l"
-#cmap = "yeo_colors_7"
-
-
-cmap = "blue_7"
-template_brain = 'B3'
-
-CBAR_STATE = dict(cbtxtsz=12, clim=[0, 7], txtsz=10., width=.1, cbtxtsh=3.,
-                  rect=(-.3, -2., 1., 4.))
-KW = dict(title_size=14., zoom=1)
-
-ddir = '../../../data/'
-
-results_dir = os.path.join(ddir, 'results')
-
-nii_bo_dir = '../../../data/niis/networks'
-
-fig_dir = '../../../paper/figs/source/networks'
-
-
-acc_fname = os.path.join(results_dir, 'accuracy_by_network.npz')
-acc_file = np.load(acc_fname)
-
-
-hills = acc_file['hills'][:, 0, 0]
-maxs = acc_file['maxs'][:, 0, 0]
-
-def norm_vals(vals):
-    return (vals-vals.min())/(vals.max() - vals.min())
-
-
-
-
-b_yeo = se.load(os.path.join(nii_bo_dir, 'yeo_bo_6mm.bo'))
-
-data_yeo = b_yeo.get_data().values.ravel()
-xyz_yeo = b_yeo.locs.values
-
-data1 = data_yeo[data_yeo==1]
-xyz1 = xyz_yeo[data_yeo==1]
-
-data2 = data_yeo[data_yeo==2]
-xyz2 = xyz_yeo[data_yeo==2]
-
-data3 = data_yeo[data_yeo==3]
-xyz3 = xyz_yeo[data_yeo==3]
-
-data4 = data_yeo[data_yeo==4]
-xyz4 = xyz_yeo[data_yeo==4]
-
-data5 = data_yeo[data_yeo==5]
-xyz5 = xyz_yeo[data_yeo==5]
-
-data6 = data_yeo[data_yeo==6]
-xyz6 = xyz_yeo[data_yeo==6]
-
-data7 = data_yeo[data_yeo==7]
-xyz7 = xyz_yeo[data_yeo==7]
-
-kwargs = {}
-kwargs['alpha'] = 0.2
-
-
-s_obj_1 = SourceObj('iEEG', xyz1, data=data1, cmap=cmap)
[…]

### 2023-03-12 code/scripts/visbrain_plots/network_plot.py

@@ -1,80 +1,0 @@
-import supereeg as se
-import os
-import numpy as np
-from visbrain.objects import BrainObj, SceneObj, SourceObj
-
-cmap = "yeo_colors_7"
-
-template_brain = 'B3'
-
-CBAR_STATE = dict(cbtxtsz=12, clim=[0, 7], txtsz=10., width=.1, cbtxtsh=3.,
-                  rect=(-.3, -2., 1., 4.))
-KW = dict(title_size=14., zoom=1)
-
-
-nii_bo_dir = '../../../data/niis/networks'
-
-fig_dir = '../../../paper/figs/source/networks'
-
-sc = SceneObj(bgcolor='white', size=(1000, 1000))
-
-b_yeo = se.load(os.path.join(nii_bo_dir, 'yeo_bo_6mm.bo'))
-
-data1 = b_yeo.get_data().values.ravel()
-xyz1 = b_yeo.locs.values
-
-s_obj_1 = SourceObj('iEEG', xyz1, data=data1, cmap=cmap)
-s_obj_1.color_sources(data=data1)
-
-b_obj_proj_left = BrainObj(template_brain, hemisphere='left', translucent=False)
-b_obj_proj_left.project_sources(s_obj_1, clim=(1, 7), cmap=cmap)
-sc.add_to_subplot(b_obj_proj_left, row=0, col=0, rotate='left', use_this_cam=True)
-
-b_obj_proj_left = BrainObj(template_brain, hemisphere='left', translucent=False)
-b_obj_proj_left.project_sources(s_obj_1, clim=(1, 7), cmap=cmap)
-sc.add_to_subplot(b_obj_proj_left, row=1, col=0, rotate='right', use_this_cam=True)
-
-b_obj_proj_right = BrainObj(template_brain, hemisphere='right', translucent=False)
-b_obj_proj_right.project_sources(s_obj_1, clim=(1, 7), cmap=cmap)
-sc.add_to_subplot(b_obj_proj_right, row=1, col=1, rotate='left', use_this_cam=True)
-
-b_obj_proj_right = BrainObj(template_brain, hemisphere='right', translucent=False)
-b_obj_proj_right.project_sources(s_obj_1, clim=(1, 7), cmap=cmap)
-sc.add_to_subplot(b_obj_proj_right, row=0, col=1, rotate='right', use_this_cam=True)
-
[…]

### 2023-03-12 code/scripts/visbrain_plots/pca_plot_bos.py

@@ -1,88 +1,0 @@
-
-import supereeg as se
-import os
-import glob as glob
-from visbrain.objects import BrainObj, ColorbarObj, SceneObj, SourceObj
-
-colors = ['Blues_r', 'Greens_r', 'Purples_r', 'Reds_r']
-#cmap = "Purples_r"
-
-nii_bo_dir = '../../../data/niis/pcas'
-fig_dir = '../../../paper/figs/source/pcas'
-
-r = 10
-
-CBAR_STATE = dict(cbtxtsz=20, txtsz=20., txtcolor='black', width=.1, cbtxtsh=3.,
-                  rect=(-.3, -2., 1., 4.))
-KW = dict(title_size=14., zoom=2)
-
-template_brain = 'B3'
-
-conditions = ['intact', 'paragraph', 'rest', 'word']
-
-for e, c in enumerate(conditions):
-
-    cmap = colors[e]
-
-    b1 = se.load(os.path.join(nii_bo_dir,f'{c}_1_pca.bo'))
-    b2 = se.load(os.path.join(nii_bo_dir,f'{c}_2_pca.bo'))
-    b3 = se.load(os.path.join(nii_bo_dir,f'{c}_3_pca.bo'))
-    b4 = se.load(os.path.join(nii_bo_dir,f'{c}_4_pca.bo'))
-    b5 = se.load(os.path.join(nii_bo_dir,f'{c}_5_pca.bo'))
-
-    data1 = b1.get_data().values.ravel()
-    xyz1 = b1.locs.values
-
-    data2 = b2.get_data().values.ravel()
-    xyz2 = b2.locs.values
-
-    data3 = b3.get_data().values.ravel()
-    xyz3 = b3.locs.values
-
-    data4 = b4.get_data().values.ravel()
-    xyz4 = b4.locs.values
-
-    data5 = b5.get_data().values.ravel()
-    xyz5 = b5.locs.values
-
-
-    template_brain = 'B3'
-
-    sc = SceneObj(bgcolor='white', size=(1000, 1000))
-
-    CBAR_STATE = dict(cbtxtsz=12, clim=[0, 6], txtsz=10., width=.1, cbtxtsh=3.,
-                      rect=(-.3, -2., 1., 4.))
-    KW = dict(title_size=14., zoom=1)
-
-    s_obj_1 = SourceObj('iEEG', xyz1, data=data1, cmap=cmap)
-    s_obj_1.color_sources(data=data1)
[…]

### 2023-03-12 code/scripts/visbrain_plots/pca_plot_bos_chunks.py

@@ -1,92 +1,0 @@
-
-import supereeg as se
-import os
-import glob as glob
-from visbrain.objects import BrainObj, ColorbarObj, SceneObj, SourceObj
-
-colors = ['Blues_r', 'Greens_r', 'Purples_r', 'Reds_r']
-#cmap = "Purples_r"
-
-nii_bo_dir = '../../../data/niis/pcas'
-fig_dir = '../../../paper/figs/source/pcas'
-
-r = 10
-
-CBAR_STATE = dict(cbtxtsz=20, txtsz=20., txtcolor='black', width=.1, cbtxtsh=3.,
-                  rect=(-.3, -2., 1., 4.))
-KW = dict(title_size=14., zoom=2)
-
-template_brain = 'B3'
-
-conditions = ['intact', 'paragraph', 'rest', 'word']
-
-chunks = 3
-
-
-for chunk in range(chunks):
-    for e, c in enumerate(conditions):
-
-        cmap = colors[e]
-
-        b1 = se.load(os.path.join(nii_bo_dir,f'{c}_1_pca_chunk_{chunk+1}.bo'))
-        b2 = se.load(os.path.join(nii_bo_dir,f'{c}_2_pca_chunk_{chunk+1}.bo'))
-        b3 = se.load(os.path.join(nii_bo_dir,f'{c}_3_pca_chunk_{chunk+1}.bo'))
-        b4 = se.load(os.path.join(nii_bo_dir,f'{c}_4_pca_chunk_{chunk+1}.bo'))
-        b5 = se.load(os.path.join(nii_bo_dir,f'{c}_5_pca_chunk_{chunk+1}.bo'))
-
-        data1 = b1.get_data().values.ravel()
-        xyz1 = b1.locs.values
-
-        data2 = b2.get_data().values.ravel()
-        xyz2 = b2.locs.values
-
-        data3 = b3.get_data().values.ravel()
-        xyz3 = b3.locs.values
-
-        data4 = b4.get_data().values.ravel()
-        xyz4 = b4.locs.values
-
-        data5 = b5.get_data().values.ravel()
-        xyz5 = b5.locs.values
-
-
-        template_brain = 'B3'
-
-        sc = SceneObj(bgcolor='white', size=(1000, 1000))
-
-        CBAR_STATE = dict(cbtxtsz=12, clim=[0, 6], txtsz=10., width=.1, cbtxtsh=3.,
-                          rect=(-.3, -2., 1., 4.))
-        KW = dict(title_size=14., zoom=1)
-
[…]

### 2023-03-12 code/scripts/visbrain_plots/pca_plot_changes.py

@@ -1,120 +1,0 @@
-
-import supereeg as se
-import os
-import numpy as np
-import glob as glob
-from visbrain.objects import BrainObj, ColorbarObj, SceneObj, SourceObj
-
-colors = 'RdBu_r'
-
-nii_bo_dir = '../../../data/niis/pcas'
-fig_dir = '../../../paper/figs/source/pcas'
-
-r = 10
-
-CBAR_STATE = dict(cbtxtsz=20, txtsz=20., txtcolor='black', width=.1, cbtxtsh=3.,
-                  rect=(-.3, -2., 1., 4.))
-KW = dict(title_size=14., zoom=2)
-
-template_brain = 'B3'
-
-conditions = ['intact', 'paragraph', 'rest', 'word']
-
-for e, c in enumerate(conditions):
-
-    cmap = colors
-
-    b_1 = se.load(os.path.join(nii_bo_dir,f'{c}_1_pca_chunk_smallest_change3.bo'))
-    b1 = se.load(os.path.join(nii_bo_dir, f'{c}_1_pca_chunk_largest_change3.bo'))
-    b_2 = se.load(os.path.join(nii_bo_dir,f'{c}_2_pca_chunk_smallest_change3.bo'))
-    b2 = se.load(os.path.join(nii_bo_dir, f'{c}_2_pca_chunk_largest_change3.bo'))
-    b_3 = se.load(os.path.join(nii_bo_dir,f'{c}_3_pca_chunk_smallest_change3.bo'))
-    b3 = se.load(os.path.join(nii_bo_dir, f'{c}_3_pca_chunk_largest_change3.bo'))
-    b_4 = se.load(os.path.join(nii_bo_dir,f'{c}_4_pca_chunk_smallest_change3.bo'))
-    b4 = se.load(os.path.join(nii_bo_dir, f'{c}_4_pca_chunk_largest_change3.bo'))
-    b_5= se.load(os.path.join(nii_bo_dir,f'{c}_5_pca_chunk_smallest_change3.bo'))
-    b5 = se.load(os.path.join(nii_bo_dir, f'{c}_5_pca_chunk_largest_change3.bo'))
-
-    data1 = b1.get_data().values.ravel() * .9
-    xyz1 = b1.locs.values
-
-    data2 = b2.get_data().values.ravel() * .7
-    xyz2 = b2.locs.values
-
-    data3 = b3.get_data().values.ravel() * .5
-    xyz3 = b3.locs.values
-
[…]

### 2023-03-15 paper/figs/source/neurosynth_topics.tex
starting to add material to supplement, also added new figure to main text and removed the old neurosynth figures

\begin{tabular}{rlllllllllll}
\toprule
                                       Topic Label &     Cognitive Label &           Term 1 &        Term 2 &          Term 3 &         Term 4 &      Term 5 &         Term 6 &        Term 7 &         Term 8 &        Term 9 &        Term 10 \\
\midrule
            Cognitive control and task performance &   Cognitive control &            tasks &       control &         network &     conditions &  comparison &      performed &        common &     correlates &    experiment &            pre \\
                Developmental aging and maturation &                   - &              age &        adults &        children &    development & adolescents &          aging & developmental &    adolescence &     childhood &          adult \\
                Eye movements and visual attention &           Attention &              eye &          gaze &            eyes &         visual &    saccades &      movements &       saccade &      direction &          gait &         target \\
                      Facial and voice recognition &  Sensory perception &      recognition &      familiar &        identity &     unfamiliar &       voice &    familiarity &         route &         voices &            fg &         facial \\
        Social interaction and contextual behavior &    Social cognition &          context &          game &           human &    interaction &         ppi &     contextual &      contexts &         agency &  interactions &        partner \\
        Language processing and semantic knowledge & Language processing &         semantic &         words &            word &        lexical &      verbal &       language &         tasks &         naming &       fluency &   phonological \\
    Experimental design and behavioral performance &                   - &           trials &      stimulus &       responses &          trial &    reaction &           time &         event &         target &        events &          times \\
            Genetic polymorphisms and risk factors &                   - &         carriers &        allele &            gene &       genotype &         met &        genetic &  polymorphism &            val &          comt &             rs \\
     Sensorimotor integration and movement control &       Motor control &            motor &      movement &       movements &   sensorimotor &     primary &         finger &       control &        imagery &       sensory &      execution \\
                Drug addiction and substance abuse &                   - &          cocaine &         users &            drug &            bpd &    controls &       cannabis &     addiction &        craving &     dependent &         heroin \\
          Music perception and auditory processing &  Sensory perception &            music &       musical &           pitch &       auditory &   musicians &      sequences &        rhythm &      listening &          beat &        singing \\
           Menstrual cycle and hormonal regulation &                   - &            phase &         women &           cycle &         phases &   menstrual &             hf &    expression &            sex &        luteal &     follicular \\
              Cognitive functions and role playing &   Cognitive control &             role &          play &           human &          plays &   cognitive &       evidence &      critical &       distinct &           key &     subregions \\
                 Inhibition and gender differences &                   - &       inhibition &         women &      inhibitory &            sex &     females &         gender &         males &           stop &          male &         female \\
       Somatosensory stimulation and motor control &       Motor control &      stimulation & somatosensory &             tms &        tactile &     primary &          touch &         motor &           rtms &  transcranial &        sensory \\
           Multisensory integration and perception &  Sensory perception &         auditory &        visual &         sensory &       modality &      sounds &    integration &         sound &       stimulus &       primary &     modalities \\
                     Social perception and empathy &    Social cognition &           social &       empathy &      experience &         people &      person &      responses &   perspective &    individuals &    attachment &       empathic \\
          Gesture recognition and visual attention &           Attention &           target &      gestures &         targets &    orientation &      visual &    distractors &       gesture &          grasp &    distractor &       location \\
                               Experimental design &                   - &           design &         block &          blocks &          event &       mixed &      condition &            ca &        writing &       blocked &           runs \\
                            Alcohol cue reactivity &              Reward &             cues &           cue &         alcohol &   anticipation &        cued &    preparation &  anticipatory &       exposure &    expectancy &    preparatory \\
                       Neuroimaging and metabolism &                   - &              pet &    tomography &        emission &       positron &        flow &        glucose &       binding &     metabolism &     metabolic &       receptor \\
                    Abnormalities in schizophrenia &                   - &    schizophrenia &      controls &         reduced &  abnormalities &    symptoms &       deficits &       matched &        control &      abnormal &  schizophrenic \\
                            Eating and body weight &                   - &             food &         taste &            body &         weight &      eating &          women &         obese &         reward &         foods &        caloric \\
                    Sleep and olfactory processing &  Sensory perception &            sleep &     olfactory &            odor &             sd & deprivation &          odors &           rem &    wakefulness &         night &           wake \\
 Alzheimer's disease and mild cognitive impairment &                   - &               ad &       disease &             mci &      alzheimer &   cognitive &     impairment &          mild &           amci &      controls &        atrophy \\
             Working memory and executive function &              Memory &           memory &          load &           tasks &         verbal & maintenance &    performance &          term &     difficulty &          dual &      cognitive \\
                 Moral decision making and phobias &                   - &            moral &        phobia &          phobic &             ec &       guilt &         spider &      decision &            art &       phobics &    intentional \\
                               Language laterality & Language processing &         language &     asymmetry &    organization &          human &   dominance &    asymmetries &   lateralized &         handed &       located & representation \\
                                         Attention &           Attention &        attention &   attentional &          visual &        spatial &         top &      selective &      stimulus &        control &     orienting &         target \\
           Resting-state brain activity in smokers &       Resting state &             reho &       smokers &         resting &        smoking &    nicotine &    homogeneity &           sci &       controls &            rs &    spontaneous \\
                         Social cognition/judgment &    Social cognition &           social &     judgments &            mind &         mental &      theory &    mentalizing &     cognition &       judgment &        person &         people \\
                        Reward and decision making &              Reward &           reward &      decision &          choice &       outcomes &     rewards &       monetary &     decisions &   anticipation &     responses &        outcome \\
                       ADHD and attention deficits &           Attention &             adhd &      disorder &       attention &        deficit &    children &  hyperactivity &      deficits &        control &      controls &        reduced \\
Neurobiological variability and individual diff... &                   - &       individual &  relationship &           local &      dependent &      change &         global &     responses &       neuronal &     magnitude &          lower \\
                                 Spatial cognition &   Spatial cognition &          spatial &         space &        location &      locations &  navigation &        virtual &        visual &   visuospatial &         visuo &       position \\
            Therapeutic interventions and training &                   - &         training &   acupuncture &         therapy &             cr &     control &        trained &   improvement &    mindfulness &   stimulation &        induced \\
                    Color perception and deception &  Sensory perception &            color &        search &         feature &      deception &    features &      responses &        colour &      dimension &         lying &    conjunction \\
          Neurodegenerative diseases and disorders &                   - &          disease &            pd &        controls &        atrophy &    clinical &          motor &      dementia &             sd &      multiple &      sclerosis \\
                Cognitive control and interference &   Cognitive control &         conflict &       control &    interference &         stroop & incongruent &      selection &     congruent &         trials &     cognitive &     monitoring \\
          Structural MRI and brain volume analysis &                   - &           volume &          gray &           voxel &             gm & morphometry &           grey &       volumes &            vbm &       density &            age \\
                  Fear conditioning and extinction &             Emotion &             fear &     switching &    conditioning &      responses &    stimulus &     extinction &            cs &         switch &        threat &    conditioned \\
                      Skill learning and expertise &              Memory &         learning &      practice &         learned &       sequence & performance &       training &     sequences &          skill &      implicit &          motor \\
                                   PTSD and trauma &             Emotion &             ptsd &        trauma &          stress &       disorder &   traumatic &  posttraumatic &     childhood &      survivors &      exposure &       controls \\
         Neural oscillations and electrophysiology &                   - &        frequency &        source &           alpha &      amplitude &        beta &          gamma &      recorded &    frequencies &     potential &   simultaneous \\
          Temporal dynamics of stimulus processing &  Sensory perception &             time &     sustained &        duration &          onset &      period &          stage &        timing &          delay &     transient &          event \\
                         Tinnitus and hearing loss &  Sensory perception &         tinnitus &          loss &         hearing &         status &     driving &     subjective &     objective &         unfair &        offers &      rejection \\
           Abstract categories and representations & Language processing &         category &    adaptation & representations & categorization &  categories &       abstract &      stimulus & representation &      features &      knowledge \\
           Pain perception and sensory stimulation &  Sensory perception &             pain &       painful &     stimulation &  somatosensory &   intensity &        noxious &          heat &        chronic &       sensory &    nociceptive \\
                                 Body and primates &                   - &             body &         human &          humans &        monkeys &        itch &       primates &       species &         monkey &        bodies &        macaque \\
                Phonological processing in reading & Language processing &          reading &       chinese &    phonological &         visual &    language &        readers &      dyslexia &     characters &      children &           word \\
             Rule-based performance and complexity &   Cognitive control &             rule &    complexity &            size &          force &       rules &         effect &          grip &         amount &      distance &     artificial \\
Autism Spectrum Disorder (ASD) and social impai... &    Social cognition &              asd &        autism &          social &       spectrum & individuals &       controls &     disorders &       children &            td &        reduced \\
            Major depression disorder and emotions &             Emotion &       depression &           mdd &       depressed &     depressive &       major &           mood &      disorder &            sad &      controls &        control \\
                              Blindness and vision &  Sensory perception &            blind &        visual &         sighted &    individuals &       humor &       laughter &  congenitally &      blindness &    plasticity &        braille \\
                        Deafness and sign language & Language processing &        condition &    conditions &            deaf &        hearing &        sign &          signs &   referential &        signers &            ci &            nss \\
    Genetic risk and familial factors in psychosis &                   - &             risk &       genetic &              sz &       siblings &   relatives &    individuals &    unaffected &        factors &        family &      psychosis \\
                  Action observation and imitation &       Motor control &           action &       actions &     observation &          motor &      mirror &           goal &     imitation &      execution &      directed &      movements \\
                 Cognitive performance and control &   Cognitive control &      performance &     cognitive &         control &         memory &   executive &           test &         tasks &    individuals &    behavioral &       deficits \\
                     Mental disorders and controls &                   - &         disorder &           ocd &         bipolar &             hc &          bd &       controls &            ts &        control & abnormalities &     compulsive \\
Pharmacological effects of placebo and drug adm... &                   - &          placebo &      dopamine &          effect &           drug &          ht & administration &         blind &   dopaminergic &        double &             mg \\
                           Personality and anxiety &    Social cognition &          anxiety &   personality &           trait &    individuals &      scores &         traits &        threat &      disorders &        social &      avoidance \\
                 Mental imagery and math abilities &   Cognitive control &           mental &       imagery &       numerical &     arithmetic &    rotation &    calculation &     magnitude &          digit &         tasks &   mathematical \\
                     Priming and repetition effect &              Memory &          priming &    repetition &     suppression &       repeated &      effect &       implicit &       literal &         target &         prime &      metaphors \\
               Working memory and error monitoring &              Memory &               wm &         error &          errors &     prediction & performance &        correct &    monitoring &            ltm &      feedback &      predicted \\
                 Sentence comprehension and syntax & Language processing &        sentences &      language &        sentence &  comprehension &   syntactic &       semantic &         verbs &           verb &          word &        meaning \\
                            Resting state networks &       Resting state &          network &       resting &         default &           mode &        rest &      intrinsic &     cognitive &   correlations &          seed &    spontaneous \\
            Episodic memory encoding and retrieval &              Memory &           memory &      encoding &       retrieval &    recognition &    episodic &          items &    successful &     subsequent &  recollection &         recall \\
                         Visual object recognition &  Sensory perception &           object &        visual &         objects &          shape &      images &         scenes &         scene &      selective &   recognition &         stream \\
      Effective causal modeling of neural networks &                   - &        effective &        causal &        modeling &        dynamic &     network &            top &     influence &     modulation &  interactions &      causality \\
       Relational reasoning and fluid intelligence &                   - &        reasoning &         FALSE &            TRUE &   intelligence &  relational &         belief &     relations &      nonverbal &         fluid &     analogical \\
                  Lesion and stroke rehabilitation &                   - &          lesions &        lesion &          stroke &         damage &      injury &       controls &      recovery &        patient &        tracts &      integrity \\
         Affective valence and feedback processing &                   - &         negative &      positive &        feedback &        valence &     arousal &           bias &    positively &     negatively &    evaluation &     swallowing \\
               Autobiographical memory in epilepsy &              Memory & autobiographical &        events &          future &       epilepsy &    episodic &         memory &      memories &           past &          mtle &       personal \\
         Evidence and effect in behavioral studies &                   - &         evidence &       provide &          effect &     behavioral &  underlying &   demonstrated & understanding &      potential &     responses &        support \\
                Stress and physiological responses &                   - &           stress &          tdcs &        cortisol &      autonomic &       heart &      responses &          rate &     regulation & physiological &        induced \\
                    Speech and language processing & Language processing &           speech &      language &        auditory &     production &  perception &  comprehension &     listening &       acoustic &    linguistic &        prosody \\
Network interactions and evidence in human systems &                   - &          network &      evidence &           human &        systems &     support &        process &      distinct &    integration &       provide &        engaged \\
                           Neuroimaging techniques &                   - &         standard &        images &      individual &           time &       image &          voxel &       spatial &           test &      clinical &        mapping \\
              Visual perception of motion and form &  Sensory perception &           motion &        visual &      perception &     perceptual &  biological &        dynamic &        moving &          human &        static &       illusion \\
               Emotional processing and regulation &             Emotion &        emotional &       emotion &         neutral &         facial & expressions &      affective &     responses &       negative &    regulation &       emotions \\
\bottomrule
\end{tabular}

### 2023-03-15 paper/main.tex
starting to add material to supplement, also added new figure to main text and removed the old neurosynth figures

@@
 \newcommand{\componentsNetworks}{S3}
+\newcommand{\componentBrains}{S4}
+\newcommand{\topTerms}{S5}
+
+\newcommand{\topics}{S1}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figs/pca_neurosynth}
+  \includegraphics[width=\textwidth]{figs/neurosynth_by_component}
 
-\caption{\textbf{Top terms associated with the highest-weighted components by
-condition.} Each row corresponds to an experimental condition, and the colors
-correspond to the component number (ranked by proportion of variance
-explained). The inflated brain plots display the top 20 highest-weighted hubs
-(see \textit{Topographic Factor Analysis}) for each components'. The lists on
-the right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
-each components' brain map. Analogous maps computed separately for each story
-segment may be found in Figure~\synthThirds.}
+\caption{\textbf{Neurosynth topic weightings by component.} We used a reverse
+inference procedure (see \textit{Reverse inference}) to estimate the
+correspondences between brain images and a set of 80 topics derived from the
+Neurosynth database of neuroimaging articles~\citep{RubiEtal17}. \textbf{A.
+Topic correlations by component.} For each of the top five highest-weighted
+principal components (columns) derived from each experimental condition
+(colors), the radar plots display the correlations between the component images
+and the per-topic images derived for each topic (topic identities are denoted
+by the blue dots; a legend for the topic labels is in the lower right of the
+figure). The top-weighted terms for each topic may be found in Table~\topics.
+\textbf{B. Topic correlations averaged across components.} The radar plot is in
+the same format as the plots in Panel A, but here we display the per-condition
+average correlations across the top five components (for each condition)
+reflected in Panel A. \textbf{C. Component images.} Each plot diplays a right
+sagittal view of a glass brain projection of the top five principal components
+(columns) for each experimental conditions (rows). Additional projections for
+each component may be found in Figure~\componentBrains.}
 
@@
 
+\begin{figure}[tp]
+  \centering
+  \includegraphics[width=\textwidth]{figs/components_neurosynth_summary}
+
+\caption{\textbf{Summary of functions associated with top-weighted components
+by condition.} \textbf{A. Top-weighted topics by condition.} Here we display
+per-condition (rows, indicated by colored dots) topic correlations, averaged
+across topics that pertain to each of several broad cognitive functions
+(columns within each sub-panel, indicated by colored dots). Each sub-panel
+reflects correlations for the components indicated in the panel titles. A
+legend for the condition and several core cognitive function classifications is
+displayed in the lower right of the figure. Table~\topics~provides a list of
+each topic's top-weighted terms, along with each topic's manually labeled
+cognitive classification. A full list of the topics most highly associated with
+each component may be found in Figure~\topTerms. \textbf{B. Associations
+between per-condition components and cognitive functions.} The network plots
+denote positive average correlations between the component images for each
+condition (gray-outlined dots on the left sides of each network; colors denote
+conditions) and topic-specific brain maps associated with each indicated
+cognitive function (black-outlined dots on the right sides of each network;
+colors denote cognitive functions). The line thicknesses are proportional to
+the correlation values (correlation coefficients are noted in the heatmaps in
+Panel A). \textbf{C. Correlations between each principal component, by
+condition.} The heatmaps display the correlations between the brain maps
+(Fig.~\componentBrains) for each principal component (sub-panel), across each
+pair of conditions (rows and columns of each sub-panel's matrix, indicated by
+colored dots). \textbf{D. Associations between per-condition components, by
+component.} Each sub-panels network plot summarizes the pattern of correlations
+between the $n$\textsuperscript{th} top-weighted principal components
+(sub-panel) for each experimental condition (gray-outlined dots). The line
+thicknesses are proportional to the correlation values (correlation
+coefficients are noted in the heatmaps in Panel C).}
+
+\label{fig:neurosynth-summary}
+\end{figure}
+
 \section*{Discussion}
@@
 literature most commonly associated with the corresponding maps. As summarized
-in Figure~\ref{fig:discussion}, we found that terms associated with memory and
+in Figure~\ref{fig:neurosynth-summary}, we found that terms associated with memory and
 sensory processing were associated with the strongest components in all three
@@
 systematically with factors like cognitive richness and depth of processing.
-
-\begin{figure}[tp]
-  \centering
-  \includegraphics[width=\textwidth]{figs/discussion}
-
-\caption{\textbf{Summary of functions associated with top-weighted components
-by condition.} Each column corresponds to an experimental condition. Brain maps
-in the top row are reproduced from Figure~\ref{fig:neurosynth-pca}, for
-reference. Cognitive functions summarized from the top Neurosynth-derived terms
-in Figure~\ref{fig:neurosynth-pca} are listed in the rows on the left.  Shaded cells
-denote which experimental conditions were associated with one or more top-weighted
-principal components associated with the given function.}
-
-\label{fig:discussion}
-\end{figure}
 

### 2023-03-15 paper/supplement.tex
applied a coat of polish to the text-- fixed a bunch of minor issues throughout (mostly typos).  i also corrected some errors in figure 1.
backing up progress on neurosynth analyses
starting to add material to supplement, also added new figure to main text and removed the old neurosynth figures

@@
 \usepackage{soul}
+\usepackage[table]{xcolor}
+\usepackage{colortbl}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{lscape}
+\usepackage{rotating}
+\usepackage{supertabular}
+
 \linenumbers
@@
 \renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 
-\title{\textit{Supplemental materials for:} High-order cognition is supported by complex but compressible brain activity patterns}
+\title{\textit{Supplemental materials for:} High-order cognition is supported by information-rich but compressible brain activity patterns}
 
@@
 
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=\textwidth]{figs/pca_neurosynth_thirds}
+
+% \caption{\textbf{Top terms associated with the highest-weighted components by
+% condition, broken down by story segment.} Each group of three rows corresponds
+% to an experimental condition, and the colors correspond to the component number
+% (ranked by proportion of variance explained). The rows' numbering denotes the
+% story segment used to compute each map (1: first third; 2: second third; 3:
+% third third). The inflated brain plots display the top 20 highest-weighted hubs
+% (see \textit{Topographic Factor Analysis}) for each components. The lists on
+% the right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
+% each components' brain map. Analogous maps computed for the entire story may be
+% found in Figure~\synth.}
+
+% \label{fig:neurosynth-thirds}
+
+% \end{figure}
+
+\begin{landscape}
+  \tablehead{\toprule
+  Topic label &     Cognitive label &           Term 1 &        Term 2 &          Term 3 &         Term 4 &      Term 5 &         Term 6 &        Term 7 &         Term 8 &        Term 9 &        Term 10 \\
+  \midrule
+  }
+  \tabletail{\bottomrule}
+  \tiny
+
+\begin{supertabular}{rlllllllllll}
+  Cognitive control and task performance &   Cognitive control &            tasks &       control &         network &     conditions &  comparison &      performed &        common &     correlates &    experiment &            pre \\
+  Developmental aging and maturation &                   - &              age &        adults &        children &    development & adolescents &          aging & developmental &    adolescence &     childhood &          adult \\
+  Eye movements and visual attention &           Attention &              eye &          gaze &            eyes &         visual &    saccades &      movements &       saccade &      direction &          gait &         target \\
+        Facial and voice recognition &  Sensory perception &      recognition &      familiar &        identity &     unfamiliar &       voice &    familiarity &         route &         voices &            fg &         facial \\
+Social interaction and contextual behavior &    Social cognition &          context &          game &           human &    interaction &         ppi &     contextual &      contexts &         agency &  interactions &        partner \\
+Language processing and semantic knowledge & Language processing &         semantic &         words &            word &        lexical &      verbal &       language &         tasks &         naming &       fluency &   phonological \\
+Experimental design and behavioral performance &                   - &           trials &      stimulus &       responses &          trial &    reaction &           time &         event &         target &        events &          times \\
+Genetic polymorphisms and risk factors &                   - &         carriers &        allele &            gene &       genotype &         met &        genetic &  polymorphism &            val &          comt &             rs \\
+Sensorimotor integration and movement control &       Motor control &            motor &      movement &       movements &   sensorimotor &     primary &         finger &       control &        imagery &       sensory &      execution \\
+  Drug addiction and substance abuse &                   - &          cocaine &         users &            drug &            bpd &    controls &       cannabis &     addiction &        craving &     dependent &         heroin \\
+Music perception and auditory processing &  Sensory perception &            music &       musical &           pitch &       auditory &   musicians &      sequences &        rhythm &      listening &          beat &        singing \\
+Menstrual cycle and hormonal regulation &                   - &            phase &         women &           cycle &         phases &   menstrual &             hf &    expression &            sex &        luteal &     follicular \\
+Cognitive functions and role playing &   Cognitive control &             role &          play &           human &          plays &   cognitive &       evidence &      critical &       distinct &           key &     subregions \\
+   Inhibition and gender differences &                   - &       inhibition &         women &      inhibitory &            sex &     females &         gender &         males &           stop &          male &         female \\
+Somatosensory stimulation and motor control &       Motor control &      stimulation & somatosensory &             tms &        tactile &     primary &          touch &         motor &           rtms &  transcranial &        sensory \\
+Multisensory integration and perception &  Sensory perception &         auditory &        visual &         sensory &       modality &      sounds &    integration &         sound &       stimulus &       primary &     modalities \\
+       Social perception and empathy &    Social cognition &           social &       empathy &      experience &         people &      person &      responses &   perspective &    individuals &    attachment &       empathic \\
+Gesture recognition and visual attention &           Attention &           target &      gestures &         targets &    orientation &      visual &    distractors &       gesture &          grasp &    distractor &       location \\
+                 Experimental design &                   - &           design &         block &          blocks &          event &       mixed &      condition &            ca &        writing &       blocked &           runs \\
+              Alcohol cue reactivity &              Reward &             cues &           cue &         alcohol &   anticipation &        cued &    preparation &  anticipatory &       exposure &    expectancy &    preparatory \\
+         Neuroimaging and metabolism &                   - &              pet &    tomography &        emission &       positron &        flow &        glucose &       binding &     metabolism &     metabolic &       receptor \\
+      Abnormalities in schizophrenia &                   - &    schizophrenia &      controls &         reduced &  abnormalities &    symptoms &       deficits &       matched &        control &      abnormal &  schizophrenic \\
+              Eating and body weight &                   - &             food &         taste &            body &         weight &      eating &          women &         obese &         reward &         foods &        caloric \\
+      Sleep and olfactory processing &  Sensory perception &            sleep &     olfactory &            odor &             sd & deprivation &          odors &           rem &    wakefulness &         night &           wake \\
+Alzheimer's disease and mild cognitive impairment &                   - &               ad &       disease &             mci &      alzheimer &   cognitive &     impairment &          mild &           amci &      controls &        atrophy \\
+Working memory and executive function &              Memory &           memory &          load &           tasks &         verbal & maintenance &    performance &          term &     difficulty &          dual &      cognitive \\
+   Moral decision making and phobias &                   - &            moral &        phobia &          phobic &             ec &       guilt &         spider &      decision &            art &       phobics &    intentional \\
+                 Language laterality & Language processing &         language &     asymmetry &    organization &          human &   dominance &    asymmetries &   lateralized &         handed &       located & representation \\
+                           Attention &           Attention &        attention &   attentional &          visual &        spatial &         top &      selective &      stimulus &        control &     orienting &         target \\
+Resting-state brain activity in smokers &       Resting state &             reho &       smokers &         resting &        smoking &    nicotine &    homogeneity &           sci &       controls &            rs &    spontaneous \\
+           Social cognition/judgment &    Social cognition &           social &     judgments &            mind &         mental &      theory &    mentalizing &     cognition &       judgment &        person &         people \\
+          Reward and decision making &              Reward &           reward &      decision &          choice &       outcomes &     rewards &       monetary &     decisions &   anticipation &     responses &        outcome \\
+         ADHD and attention deficits &           Attention &             adhd &      disorder &       attention &        deficit &    children &  hyperactivity &      deficits &        control &      controls &        reduced \\
+Neurobiological variability and individual diff...
+&                   - &       individual &  relationship &           local &      dependent &      change &         global &     responses &       neuronal &     magnitude &          lower \\
+                   Spatial cognition &   Spatial cognition &          spatial &         space &        location &      locations &  navigation &        virtual &        visual &   visuospatial &         visuo &       position \\
+Therapeutic interventions and training &                   - &         training &   acupuncture &         therapy &             cr &     control &        trained &   improvement &    mindfulness &   stimulation &        induced \\
+      Color perception and deception &  Sensory perception &            color &        search &         feature &      deception &    features &      responses &        colour &      dimension &         lying &    conjunction \\
+Neurodegenerative diseases and disorders &                   - &          disease &            pd &        controls &        atrophy &    clinical &          motor &      dementia &             sd &      multiple &      sclerosis \\
+  Cognitive control and interference &   Cognitive control &         conflict &       control &    interference &         stroop & incongruent &      selection &     congruent &         trials &     cognitive &     monitoring \\
+Structural MRI and brain volume analysis &                   - &           volume &          gray &           voxel &             gm & morphometry &           grey &       volumes &            vbm &       density &            age \\
+    Fear conditioning and extinction &             Emotion &             fear &     switching &    conditioning &      responses &    stimulus &     extinction &            cs &         switch &        threat &    conditioned \\
+        Skill learning and expertise &              Memory &         learning &      practice &         learned &       sequence & performance &       training &     sequences &          skill &      implicit &          motor \\
+                     PTSD and trauma &             Emotion &             ptsd &        trauma &          stress &       disorder &   traumatic &  posttraumatic &     childhood &      survivors &      exposure &       controls \\
+Neural oscillations and electrophysiology &                   - &        frequency &        source &           alpha &      amplitude &        beta &          gamma &      recorded &    frequencies &     potential &   simultaneous \\
+Temporal dynamics of stimulus processing &  Sensory perception &             time &     sustained &        duration &          onset &      period &          stage &        timing &          delay &     transient &          event \\
+           Tinnitus and hearing loss &  Sensory perception &         tinnitus &          loss &         hearing &         status &     driving &     subjective &     objective &         unfair &        offers &      rejection \\
+Abstract categories and representations & Language processing &         category &    adaptation & representations & categorization &  categories &       abstract &      stimulus & representation &      features &      knowledge \\
+Pain perception and sensory stimulation &  Sensory perception &             pain &       painful &     stimulation &  somatosensory &   intensity &        noxious &          heat &        chronic &       sensory &    nociceptive \\
+                   Body and primates &                   - &             body &         human &          humans &        monkeys &        itch &       primates &       species &         monkey &        bodies &        macaque \\
+  Phonological processing in reading & Language processing &          reading &       chinese &    phonological &         visual &    language &        readers &      dyslexia &     characters &      children &           word \\
+Rule-based performance and complexity &   Cognitive control &             rule &    complexity &            size &          force &       rules &         effect &          grip &         amount &      distance &     artificial \\
+Autism Spectrum Disorder (ASD) and social impai...
+&    Social cognition &              asd &        autism &          social &       spectrum & individuals &       controls &     disorders &       children &            td &        reduced \\
+Major depression disorder and emotions &             Emotion &       depression &           mdd &       depressed &     depressive &       major &           mood &      disorder &            sad &      controls &        control \\
+                Blindness and vision &  Sensory perception &            blind &        visual &         sighted &    individuals &       humor &       laughter &  congenitally &      blindness &    plasticity &        braille \\
+          Deafness and sign language & Language processing &        condition &    conditions &            deaf &        hearing &        sign &          signs &   referential &        signers &            ci &            nss \\
+Genetic risk and familial factors in psychosis &                   - &             risk &       genetic &              sz &       siblings &   relatives &    individuals &    unaffected &        factors &        family &      psychosis \\
+    Action observation and imitation &       Motor control &           action &       actions &     observation &          motor &      mirror &           goal &     imitation &      execution &      directed &      movements \\
+   Cognitive performance and control &   Cognitive control &      performance &     cognitive &         control &         memory &   executive &           test &         tasks &    individuals &    behavioral &       deficits \\
+       Mental disorders and controls &                   - &         disorder &           ocd &         bipolar &             hc &          bd &       controls &            ts &        control & abnormalities &     compulsive \\
+Pharmacological effects of placebo and drug adm...
+&                   - &          placebo &      dopamine &          effect &           drug &          ht & administration &         blind &   dopaminergic &        double &             mg \\
+             Personality and anxiety &    Social cognition &          anxiety &   personality &           trait &    individuals &      scores &         traits &        threat &      disorders &        social &      avoidance \\
+   Mental imagery and math abilities &   Cognitive control &           mental &       imagery &       numerical &     arithmetic &    rotation &    calculation &     magnitude &          digit &         tasks &   mathematical \\
+       Priming and repetition effect &              Memory &          priming &    repetition &     suppression &       repeated &      effect &       implicit &       literal &         target &         prime &      metaphors \\
+ Working memory and error monitoring &              Memory &               wm &         error &          errors &     prediction & performance &        correct &    monitoring &            ltm &      feedback &      predicted \\
+   Sentence comprehension and syntax & Language processing &        sentences &      language &        sentence &  comprehension &   syntactic &       semantic &         verbs &           verb &          word &        meaning \\
+              Resting state networks &       Resting state &          network &       resting &         default &           mode &        rest &      intrinsic &     cognitive &   correlations &          seed &    spontaneous \\
+Episodic memory encoding and retrieval &              Memory &           memory &      encoding &       retrieval &    recognition &    episodic &          items &    successful &     subsequent &  recollection &         recall \\
+           Visual object recognition &  Sensory perception &           object &        visual &         objects &          shape &      images &         scenes &         scene &      selective &   recognition &         stream \\
+Effective causal modeling of neural networks &                   - &        effective &        causal &        modeling &        dynamic &     network &            top &     influence &     modulation &  interactions &      causality \\
+Relational reasoning and fluid intelligence &                   - &        reasoning &         FALSE &            TRUE &   intelligence &  relational &         belief &     relations &      nonverbal &         fluid &     analogical \\
+    Lesion and stroke rehabilitation &                   - &          lesions &        lesion &          stroke &         damage &      injury &       controls &      recovery &        patient &        tracts &      integrity \\
+Affective valence and feedback processing &                   - &         negative &      positive &        feedback &        valence &     arousal &           bias &    positively &     negatively &    evaluation &     swallowing \\
+ Autobiographical memory in epilepsy &              Memory & autobiographical &        events &          future &       epilepsy &    episodic &         memory &      memories &           past &          mtle &       personal \\
+Evidence and effect in behavioral studies &                   - &         evidence &       provide &          effect &     behavioral &  underlying &   demonstrated & understanding &      potential &     responses &        support \\
+  Stress and physiological responses &                   - &           stress &          tdcs &        cortisol &      autonomic &       heart &      responses &          rate &     regulation & physiological &        induced \\
+      Speech and language processing & Language processing &           speech &      language &        auditory &     production &  perception &  comprehension &     listening &       acoustic &    linguistic &        prosody \\
+Network interactions and evidence in human systems &                   - &          network &      evidence &           human &        systems &     support &        process &      distinct &    integration &       provide &        engaged \\
+             Neuroimaging techniques &                   - &         standard &        images &      individual &           time &       image &          voxel &       spatial &           test &      clinical &        mapping \\
+Visual perception of motion and form &  Sensory perception &           motion &        visual &      perception &     perceptual &  biological &        dynamic &        moving &          human &        static &       illusion \\
+ Emotional processing and regulation &             Emotion &        emotional &       emotion &         neutral &         facial & expressions &      affective &     responses &       negative &    regulation &       emotions \\
+
+\end{supertabular}
+
+\caption{\vspace{0.2cm} \normalsize \noindent Table S1: \textbf{Neurosynth-derived topics.} We report the top-weighted terms
+for each of 80 topics identified using Latent Dirichlet
+Allocation~\citep{BleiEtal03} applied to 9,204 functional neuroimaging articles
+in the Neurosynth database~\citep{RubiEtal17}. The topics, as well as
+associated brain maps identified using Neurosynth, were identified and reported
+in several prior studies~\citep{FoxEtal14, SulEtal17, ChenEtal20}. The topic
+labels for each topic were generated automatically with the following ChatGPT
+(chat.openai.com) prompt: ``Please help me come up with intuitive labels for
+topic topics I found by fitting a topic model to thousands of neuroscience and
+psychology articles. I'll paste in the top 10 highest-weighted words for each
+topic, and I'd like you to respond with a suggested label. For each topic,
+please respond with just the topic label and no other formatting or text. Here
+are the next topic's top words:'' followed by a comma-separated list of the
+given topic's top-weighted words reflected in the table. For some topics,
+ChatGPT responded with a longer-form response rather than a concise topic
+label. In these instances, on a case-by-case basis, we used a second follow-up
+prompt to achieve the given topic's label: ``could you please come up with a
+more concise label for that topic?''. We then manually identified a set of 11
+cognitive labels that were intended to encapsulate a representative range of
+widely studied low-level and high-level cognitive functions. In choosing the
+set of cognitive labels, we jointly considered each topic's ChatGPT-derived
+topic label, along with the top-weighted words for the topic. We attempted to
+generate a concise set of labels that still spanned the full set of cognitive
+functions reflected across the 80 topics. Topics that appeared unrelated to
+specific cognitive functions (e.g., topics related to specific methods or
+clinical themes) are designated with dashes.
+
+\label{tab:topics}}
+
+\end{landscape}
+
 \begin{figure}
   \centering
-  \includegraphics[width=\textwidth]{figs/pca_neurosynth_thirds}
+  \includegraphics[width=\textwidth]{figs/top_terms_by_component}
 
 \caption{\textbf{Top terms associated with the highest-weighted components by
-condition, broken down by story segment.} Each group of three rows corresponds
-to an experimental condition, and the colors correspond to the component number
-(ranked by proportion of variance explained). The rows' numbering denotes the
-story segment used to compute each map (1: first third; 2: second third; 3:
-third third). The inflated brain plots display the top 20 highest-weighted hubs
-(see \textit{Topographic Factor Analysis}) for each components. The lists on
-the right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
-each components' brain map. Analogous maps computed for the entire story may be
-found in Figure~\synth.}
+condition, broken down by story segment.} Each group of five rows corresponds
+to an experimental condition (denoted by color, as indicated in the legend in
+the lower right), and the columns and shading correspond to the component
+number (ranked by proportion of variance explained). The colored squares in
+front of many of the topics denote manually identified cognitive labels
+(Tab.~S1).}
 
-\label{fig:neurosynth-thirds}
+\label{fig:top-terms}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{figs/topic_contrasts}
+
+\caption{\textbf{Fill this in...}}
+
+\label{fig:contrasts}
 

### 2023-03-15 paper/main.tex
finished creating and documenting remaining figures

@@
 \newcommand{\topTerms}{S5}
+\newcommand{\neurosynthFull}{S6}
 
@@
 thicknesses are proportional to the correlation values (correlation
-coefficients are noted in the heatmaps in Panel C).}
+coefficients are noted in the heatmaps in Panel C). Also see
+Fig.~\neurosynthFull~for additional information.}
 

### 2023-03-15 paper/supplement.tex
finished creating and documenting remaining figures

@@
 \usepackage{rotating}
-\usepackage{supertabular}
+\usepackage{xtab}
 
-\linenumbers
+%\linenumbers
 
 \newcommand{\synth}{1}
+\newcommand{\neurosynth}{5}
 \renewcommand{\thefigure}{S\arabic{figure}}
@@
 
-\begin{supertabular}{rlllllllllll}
+\begin{xtabular}{rlllllllllll}
   Cognitive control and task performance &   Cognitive control &            tasks &       control &         network &     conditions &  comparison &      performed &        common &     correlates &    experiment &            pre \\
@@
 
-\end{supertabular}
+ \label{tab:topics}}
+\end{xtabular}
 
@@
 
-\label{tab:topics}}
-
 \end{landscape}
 
-\begin{figure}
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{figs/variance_explained}
+
+\caption{\textbf{Cumulative variance explained by component, condition, and part.} 
+Each panel displays the cumulative variance explained in the neuroimaging data as a function of the number of principal components.
+Colors denote experimental conditions.  The left panel displays results for all data, and the right panels display results separated
+by story segment (Q1: first quarter; Q2: second quarter; Q3: third quarter; Q4: fourth quarter).}
+
+\label{fig:var-explained}
+\end{figure}
+
+\begin{figure}[t]
   \centering
@@
 
-\begin{figure}
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{figs/component_brain_maps}
+
+\caption{\textbf{Brain maps by component and condition.} For the top five
+highest-weighted principal components (rows), from each experimental condition
+(columns), the components' brain maps are projected onto four views: left
+sagittal, axial, right sagittal, and coronal.  The color scale is the same for all panels and matches the coloring in Fig.~\neurosynth C.}
+
+\label{fig:components}
+\end{figure}
+
+\begin{figure}[t]
   \centering
@@
 
-\caption{\textbf{Fill this in...}}
+\caption{\textbf{Comparisons between per-component topic correlations across conditions.}  Each sub-panel displays a scatterplot comparing the per-topic correlations for two or more experimental conditions.
+Each dot donotes the correlations for a single topic (indicated by the legend on the right).  The topics are colored according to the ranked order of the correlations
+between the topic's brain maps and the brain map for the first principal component in the intact condition.  \textbf{A. Comparisons between correlations for each pair of experimental conditions.}  The
+conditions being compared are marked on the $x$ and $y$ axes.
+Each sub-panel (column) reflects the correlations for one principal component.
+\textbf{B.
+Comparisons between \textit{differences} in correlations for pairs of experimental conditions.}
+In these sub-panels, the $x$ and $y$ coordinates reflect differences in correlations for the indicated experimental conditions, for the given component (column).
+All panels: the across-topic correlations reported in each panel are between each topic's $x$ and $y$ coordinates.}
 
@@
 
+\end{figure}
+
+\begin{figure}[tp]
+  \centering
+  \includegraphics[width=\textwidth]{figs/components_neurosynth_full}
+
+\caption{\textbf{Functions associated with top-weighted components by
+condition.} \textbf{A. Top-weighted topics by condition.} Here we display
+per-condition (rows, indicated by colored dots) topic correlations, averaged
+across topics that pertain to each of several broad cognitive functions
+(columns within each sub-panel, indicated by colored dots). Each sub-panel
+reflects correlations for the components indicated in the panel titles. A
+legend for the condition and cognitive function classifications is displayed in
+the lower right of the figure. Table~S1 provides a list of each topic's
+top-weighted terms, along with each topic's manually labeled cognitive
+classification. A full list of the topics most highly associated with each
+component may be found in Figure~\ref{fig:top-terms}. \textbf{B. Associations
+between per-condition components and cognitive functions.} The network plots
+denote positive average correlations between the component images for each
+condition (gray-outlined dots on the left sides of each network; colors denote
+conditions) and topic-specific brain maps associated with each indicated
+cognitive function (black-outlined dots on the right sides of each network;
+colors denote cognitive functions). The line thicknesses are proportional to
+the correlation values (correlation coefficients are noted in the heatmaps in
+Panel A). \textbf{C. Correlations between each principal component, by
+condition.} The heatmaps display the correlations between the brain maps
+(Fig.~\componentBrains) for each principal component (sub-panel), across each
+pair of conditions (rows and columns of each sub-panel's matrix, indicated by
+colored dots). \textbf{D. Associations between per-condition topic weights, by
+component.} Each sub-panels network plot summarizes the pattern of correlations
+between the topic correlations from each of the $n$\textsuperscript{th} top-weighted principal components
+(sub-panel), for each experimental condition (gray-outlined dots). The line
+thicknesses are proportional to the correlation values (correlation
+coefficients are noted in the heatmaps in Panel C).}
+
+\label{fig:neurosynth-full}
 \end{figure}

### 2023-03-15 paper/main.tex
updating main text

@@
 
-\newcommand{\synthThirds}{S1}
-\newcommand{\accuracyNetworks}{S2}
-\newcommand{\componentsNetworks}{S3}
-\newcommand{\componentBrains}{S4}
-\newcommand{\topTerms}{S5}
-\newcommand{\neurosynthFull}{S6}
+\newcommand{\varExplained}{S1}
+\newcommand{\topTerms}{S2}
+\newcommand{\componentBrains}{S3}
+\newcommand{\topicContrasts}{S4}
+\newcommand{\neurosynthFull}{S5}
 
@@
 We take as a given that brain activity is highly flexible: our brains can
-exhibit nearly infinite activity patterns. This flexibility implies that our
+exhibit nearly infinite varieties of activity patterns. This flexibility implies that our
 brains activity patterns are highly information rich. However, brain activity
@@
 listening times in each condition, as a proxy for how ``informative'' the
-task-specific activity patterns were~\citep{SimoChan20}. We also use principle
+task-specific activity patterns were~\citep{SimoChan20}. We also used principle
 components analysis to generate lower-dimensional representations of the
@@
 Figure~\ref{fig:inflection}A displays the decoding accuracy as a function of
-the number of principal components used to represent the data. Several patterns
-were revealed by the analysis. First, in general (i.e., across experimental
-conditions), decoding accuracy improves as the number of components increases.
-However, decoding accuracy peaked at higher levels for experimental conditions
-that exposed participants to cognitively richer stimuli. The peak decoding
+the number of principal components used to represent the data (also see
+Fig.~\varExplained). Several patterns were revealed by the analysis. First, in
+general (i.e., across experimental conditions), decoding accuracy tends to
+improve as the number of components are increased. However, decoding accuracy
+peaked at higher levels for experimental conditions that exposed participants
+to cognitively richer stimuli (Fig.~\ref{fig:inflection}D). The peak decoding
 accuracy was highest for the ``intact'' condition (versus paragraph: $t(99) =
@@
 principal components provides more information to the temporal decoders
-(Fig.~\ref{fig:inflection}B).
-
-In every experimental condition, decoding accuracy appeared to asymptote (i.e.,
-hit an upper limit) beyond some characteristic number of components that
-differed across conditions. To quantify the ``inflection points'' at which the
-decoding curves in Figure~\ref{fig:inflection}A flattened out, we fit a sigmoid
-function to the average decoding curve for each condition. We defined the
-inflection point for each condition as the point on the fitted sigmoid where
-the second derivative was both positive and less than a threshold value of
-0.0001 (i.e., approaching 0 from the right). These inflection points reflect a
-``balance'' between higher decoding accuracy (which tends to be better when
-more components are used) and compression (which is better for fewer
-components). Plotting each condition's inflection point
-(Fig.~\ref{fig:inflection}D) reveals that both the number of components and the
-decoding accuracy at each inflection point increase systematically across
-conditions in proportion to cognitive richness.
+(Figs.~\ref{fig:inflection}B, C). Taken together, as shown in
+Figure~\ref{fig:inflection}E, we found that brain activity patterns evoked by
+cognitively richer conditions tended to be both more informative (i.e.,
+associated with higher decoding accuracies) \textit{and} more compressible
+(i.e., requiring fewer components to achieve the 5\% accuracy threshold).
 
@@
 
-\label{fig:inflection-thirds}
+\label{fig:inflection-quarters}
 
@@
 and context they had been exposed to by that point. To examine this
-possibility, we divided each condition into three successive time segments. We
-computed decoding curves (Fig.~\ref{fig:inflection-thirds}A) and inflection
-points (Fig.~\ref{fig:inflection-thirds}B) for each segment and condition. We
-found that, in the two most cognitively rich conditions (intact and paragraph),
-both decoding accuracy and compressibility, as reflected by the change in
-decoding curves, increased with listening time (e.g., at the annotated
-reference point of $k = 20$ components: intact: $t(99) = 7.915, p < 0.001)$;
-paragraph: $t(99) = 2.354, p = 0.021)$). These changes may reflect an increase
-in comprehension or depth of processing with listening time. In contrast, the
-decoding accuracy and compressibility \textit{decreased} with listening time in
-the word condition ($t(99) = -10.747, p < 0.001)$) and rest condition ($t(99) =
--22.081, p < 0.001)$). This might reflect the depletion of attentional
-resources in the less-engaging word and rest conditions.
+possibility, we divided each condition into four successive time segments. We
+computed decoding curves (Fig.~\ref{fig:inflection-quarters}A) and the numbers
+of components required to achieve 5\% decoding accuracy
+(Fig.~\ref{fig:inflection-quarters}B) for each segment and condition. We found
+that, in the two most cognitively rich conditions (intact and paragraph), both
+decoding accuracy and compressibility, as reflected by the change in decoding
+curves, increased with listening time (e.g., at the annotated reference point
+of $k = 20$ components in Fig.~\ref{fig:inflection-quarters}C: intact: $t(99) =
+7.915, p < 0.001)$; paragraph: $t(99) = 2.354, p = 0.021)$). These changes may
+reflect an increase in comprehension or depth of processing with listening
+time. In contrast, the decoding accuracy and compressibility \textit{decreased}
+with listening time in the word condition ($t(99) = -10.747, p < 0.001)$) and
+rest condition ($t(99) = -22.081, p < 0.001)$). This might reflect the
+depletion of attentional resources in the less-engaging word and rest
+conditions.
 
@@
 distinct networks. The networks can be sorted (roughly) in order from
-lower-level to higher-level cortex as follows (Fig.~\ref{fig:networks}A):
+lower-level to higher-level cortex as follows (Figs.~\ref{fig:networks}A--C):
 visual, somatomotor, dorsal attention, ventral attention, limbic,
@@
 inflection point, for each experimental condition. Moving from low-order
-networks to higher-order networks, we found that decoding accuracy (for the
-intact condition) tended to increase (Fig.~\ref{fig:networks}B). This suggests
-that higher-order networks may carry more content-relevant or stimulus-driven
-``information.'' We found no clear trends in the numbers of components at each
-network's inflection point across networks or conditions
-(Fig.~\ref{fig:networks}C).
+networks to higher-order networks, we found that decoding accuracy tended to
+increase, particularly in the higer-order experimental conditions
+(Fig.~\ref{fig:networks}D, E). This suggests that higher-order networks may
+carry more content-relevant or stimulus-driven ``information.'' We found no
+clear trends in the proportions of components required to achieve 5\% decoding
+accuracy across networks or conditions (Fig.~\ref{fig:networks}F).
 
@@
 by the blue dots; a legend for the topic labels is in the lower right of the
-figure). The top-weighted terms for each topic may be found in Table~\topics.
-\textbf{B. Topic correlations averaged across components.} The radar plot is in
-the same format as the plots in Panel A, but here we display the per-condition
-average correlations across the top five components (for each condition)
-reflected in Panel A. \textbf{C. Component images.} Each plot diplays a right
-sagittal view of a glass brain projection of the top five principal components
-(columns) for each experimental conditions (rows). Additional projections for
-each component may be found in Figure~\componentBrains.}
+figure). An annotated list of the top-weighted topics for each component and
+condition may be found in Figure~\topTerms~and the top-weighted terms for each
+topic may be found in Table~\topics. \textbf{B. Topic correlations averaged
+across components.} The radar plot is in the same format as the plots in Panel
+A, but here we display the per-condition average correlations across the top
+five components (for each condition) reflected in Panel A. \textbf{C. Component
+images.} Each plot diplays a right sagittal view of a glass brain projection of
+the top five principal components (columns) for each experimental conditions
+(rows). Additional projections for each component may be found in
+Figure~\componentBrains.}
 
@@
 patterns reflected by different principal components across different
-experimental conditions. Figure~\ref{fig:neurosynth-pca} displays inflated
-brain maps of the top five highest-weighted components, for each experimental
-condition. We also used Neurosynth~\citep{RubiEtal17} to identify, for each
-component, the top five terms associated with each map (see \textit{Reverse
-inference}). We noticed (by inspection) several common themes across the sets
-of terms associated with each component and condition. Memory-related
-components included terms like ``middle temporal,'' ``memory retrieval,'' and
-``working memory.'' Sensory processing related components included terms like
-``primary visual,'' ``auditory cortex,'' ``v1,'' and so on. Other components
-were associated with sensory integration (e.g., ``thalamic,'' ``cingulate,''
-etc.), sentence comprehension (e.g., ``sentence,'' ``comprehension''), emotion
-and valence (e.g., ``amygdala,'' ``valance''), or the default mode network
-(e.g., ``default mode'').  The components we identified were relatively stable
-across story segments (Fig.~\synthThirds).
+experimental conditions. As shown in Figure~\ref{fig:neurosynth-pca}, we used
+Neurosynth~\citep{RubiEtal17} to identify, for each component, the associations
+with each of 80 themes (see \textit{Reverse inference}).  In general, the first
+principal components across all of the experimental conditions tended to weight
+most heavily on themes related to cognitive control, memory,
+language processing, attention, and perception.
+
+To elucidate....\textbf{JRM STOPPED HERE}
+
+We noticed (by
+inspection) several common themes across the sets of terms associated with each
+component and condition. Memory-related components included terms like ``middle
+temporal,'' ``memory retrieval,'' and ``working memory.'' Sensory processing
+related components included terms like ``primary visual,'' ``auditory cortex,''
+``v1,'' and so on. Other components were associated with sensory integration
+(e.g., ``thalamic,'' ``cingulate,'' etc.), sentence comprehension (e.g.,
+``sentence,'' ``comprehension''), emotion and valence (e.g., ``amygdala,''
+``valance''), or the default mode network (e.g., ``default mode''). The
+components we identified were relatively stable across story segments
+(Fig.~\synthThirds).
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figs/components_neurosynth_summary}
+  \includegraphics[width=0.9\textwidth]{figs/components_neurosynth_summary}
 
@@
 thicknesses are proportional to the correlation values (correlation
-coefficients are noted in the heatmaps in Panel C). Also see
+coefficients are noted in the heatmaps in Panel C). \textbf{E. Change in
+weights by condition.} The bar heights reflect the slopes of regression lines,
+fit separately to the top five components from each condition, between the
+ChatGPT-derived ``rank'' of each cognitive classification and the correlations
+between the component and topic maps associated with cognitive processes at the
+given rank (see \textit{Ranking cognitive processes}).  Also see
 Fig.~\neurosynthFull~for additional information.}

### 2023-03-16 paper/main.tex
working on main text; almost done revising/updating

@@
 \newcommand{\topics}{S1}
+\newcommand{\topicTags}{S2}
 
@@
 most heavily on themes related to cognitive control, memory,
-language processing, attention, and perception.
+language processing, attention, and perception.  Other components appeared to
+vary more across conditions.
 
-To elucidate....\textbf{JRM STOPPED HERE}
-
-We noticed (by
-inspection) several common themes across the sets of terms associated with each
-component and condition. Memory-related components included terms like ``middle
-temporal,'' ``memory retrieval,'' and ``working memory.'' Sensory processing
-related components included terms like ``primary visual,'' ``auditory cortex,''
-``v1,'' and so on. Other components were associated with sensory integration
-(e.g., ``thalamic,'' ``cingulate,'' etc.), sentence comprehension (e.g.,
-``sentence,'' ``comprehension''), emotion and valence (e.g., ``amygdala,''
-``valance''), or the default mode network (e.g., ``default mode''). The
-components we identified were relatively stable across story segments
-(Fig.~\synthThirds).
+To gain further insights into which brain functions might be most closely
+associated with the top-weighted components from each experimental condition,
+we manually grouped each Neurosynth-derived topic into a smaller set of core
+cognitive functions. Separately for each component, we computed the average
+weightings across all topics that were tagged as being associated with each of
+these cognitive functions (Figs.~\ref{fig:neurosynth-summary}A,~\neurosynthFull
+A). To help visualize these associations, we used the patterns of associations
+for each component to construct graphs whose nodes were experimental conditions
+and cognitive functions (Figs.~\ref{fig:neurosynth-summary}B,~\neurosynthFull
+B). We also computed correlations between the sets of per-topic weightings from
+each of the top-weighted components from each experimental condition
+(Fig.~\ref{fig:neurosynth-summary}C, D) and between the brain maps for each
+condition's components (Fig.~\neurosynthFull C, D). Taken together, we found
+that each component appeared to weight on a fundamental set of cognitive
+functions that varied by experimental condition. For example, the top principal
+components from every condition weighted similarly on the full set of
+Neurosynth topics (Fig.~\ref{fig:neurosynth-pca}A) and cognitive functions
+(Figs.~\ref{fig:neurosynth-summary}A, B and~\neurosynthFull~A, B), suggesting
+that these components might reflect a set of functions or activity patterns
+that are common across all conditions. The second components' weightings were
+similar across the intact, paragraph, and rest conditions (highest-weighted
+functions: cognitive control, memory, social cognition, and resting state), but
+different for the word condition (highest-weighted functions: sensory
+perception and cognitive control). The fourth components' weighting were
+grouped the paragraph and word conditions (highest-weighted functions: memory,
+language processing, and cognitive control) and the intact and rest conditions
+(highest-weighted functions: emotion, social cognition). We also used
+ChatGPT~\citep{ChatGPT} to sort the list of manually tagged cognitive functions
+from lowest-level to highest-level (Tab.~\topicTags,
+Fig.~\ref{fig:neurosynth-summary}E; also see \textit{Ranking cognititive
+processes}). We found that higher-level functions tended to be weighted on more
+heavily by top components from the intact and paragraph conditions than
+lower-level functions. The top components from the word condition showed the
+opposite tendancy, whereby \textit{lower}-level functions tended to be weighted
+on more heavily than higher-level functions. The components from the rest
+condition showed almost no differences in the weights associated with
+high-level versus low-level functions.
 
@@
 
+\textbf{JRM NOTE: UPDATE THE NEXT TWO SUB-SECTIONS}
 \subsection*{Reverse inference}
@@
 with the given map.
+
+\subsection*{Ranking cognitive processes}
 

### 2023-03-16 paper/main.tex
checking in edits

@@
 and compressible, and data from less cognitively rich conditions became
-\textit{less} informative and compressible (Fig.~\ref{fig:inflection-thirds}).
+\textit{less} informative and compressible (Fig.~\ref{fig:inflection-quarters}).
 We also repeated these analyses separately for different brain networks. We
@@
 chance perfomance is $\frac{1}{T}$). In Figures~\ref{fig:inflection}
-and~\ref{fig:inflection-thirds} we report the decoding accuracy for each
+and~\ref{fig:inflection-quarters} we report the decoding accuracy for each
 condition and value of $k$, averaged across $n = 100$ cross validation folds.
 
-\textbf{JRM NOTE: UPDATE THE NEXT TWO SUB-SECTIONS}
 \subsection*{Reverse inference}
@@
 other studies, we created summary maps of each principal component, for each
-experimental condition, by summing together the 20 HTFA-derived RBF nodes (see
-\textit{Hierarchical Topographic Factor Analysis}) with the highest absolute
-value weights for each of the top 5 components (Figs.~\ref{fig:neurosynth-pca},
-\synthThirds). We then carried out a meta analysis using
-Neurosynth~\citep{RubiEtal17} to identify the 5 terms most commonly associated
-with the given map.
+experimental condition. Each principal component comprises 700 ``weights'' on
+each of the HTFA-derived RBF nodes (see \textit{Hierarchical Topographic Factor
+Analysis}). For each node, we evaluated its RBF at the locations of every voxel
+in the standard 2~mm MNI152 template brain and multiplied the RBF by the node's
+weight. The sum of these weighted RBF activation maps provides a full-brain
+image, in MNI152 space, of the given principal component.
+
+Next, we considered 80 topics estimated using Latent Dirichlet
+Allocation~\citep{BleiEtal03} applied to 9,204 functional neuroimaging articles
+in the Neurosynth database~\citep{RubiEtal17}. The topics, as well as
+associated brain maps identified using Neurosynth, were identified and reported
+in several prior studies~\citep{FoxEtal14, SulEtal17, ChenEtal20}. The topic
+labels for each topic were generated automatically with the following ChatGPT
+(chat.openai.com) prompt: ``Please help me come up with intuitive labels for
+topic topics I found by fitting a topic model to thousands of neuroscience and
+psychology articles. I'll paste in the top 10 highest-weighted words for each
+topic, and I'd like you to respond with a suggested label. For each topic,
+please respond with just the topic label and no other formatting or text. Here
+are the next topic's top words:'' followed by a comma-separated list of the
+given topic's top-weighted words reflected in the Table~\topics. For some
+topics, ChatGPT responded with a longer-form response rather than a concise
+topic label. In these instances, on a case-by-case basis, we used a second
+follow-up prompt to achieve the given topic's label: ``could you please come up
+with a more concise label for that topic?''. We then manually identified a set
+of 11 cognitive labels that were intended to encapsulate a representative range
+of widely studied low-level and high-level cognitive functions. In choosing the
+set of cognitive labels, we jointly considered each topic's ChatGPT-derived
+topic label, along with the top-weighted words for the topic. We attempted to
+generate a concise set of labels that still spanned the full set of cognitive
+functions reflected across the 80 topics. Topics that appeared unrelated to
+specific cognitive functions (e.g., topics related to specific methods or
+clinical themes) are designated with dashes in Table~\topics.
+
+Finally, following an approach used in several prior studies~\citep{FoxEtal14,
+SulEtal17, ChenEtal20} we treated the correlation between a given component's
+brain map and each topic's brain map as a reflection of how much the component
+was reflective of the given topic. This resulted in a set of 80 ``weights''
+(correlation coefficients) for each component's brain map, with one weight per
+Neurosynth-derived topic.
 
 \subsection*{Ranking cognitive processes}
+
+We manually identified 11 cognitive labels
+spanning the set of 80 Neurosynth-derived topics: cognitive control, language
+processing, memory, emotion, social cognition, spatial cognition, attention,
+reward, sensory perception, motor control, and resting state. We then used
+ChatGPT to automatically ``rank'' the processes from high-level to low-level
+using the following prompt: ``Please rank these cognitive processes from
+highest-level to lowest-level, where higher values indicate higher-order or
+higher-level processes. return the result as a csv file with a header row and
+two columns: `Cognitive label' and `Rank'. Here are the processes: cognitive
+control, language processing, memory, emotion, social cognition, spatial
+cognition, attention, reward, sensory perception, motor control, resting
+state''.  Table~\topicTags displays the output.  We used these labels in
+the analysis presented in Figure~\ref{fig:neurosynth-summary}E to help summarize
+difference in topic weightings across experimental conditions.
 

### 2023-03-16 paper/supplement.tex
checking in edits

@@
 \usepackage{soul}
-\usepackage[table]{xcolor}
-\usepackage{colortbl}
 \usepackage{booktabs}
-\usepackage{longtable}
+\usepackage{supertabular}
 \usepackage{lscape}
-\usepackage{rotating}
-\usepackage{xtab}
 
@@
 
+%\begin{landscape}
+
   \begin{landscape}
-  \tablehead{\toprule
+  \begin{center}
+\tiny
+  \tablehead{
+  \toprule
   Topic label &     Cognitive label &           Term 1 &        Term 2 &          Term 3 &         Term 4 &      Term 5 &         Term 6 &        Term 7 &         Term 8 &        Term 9 &        Term 10 \\
-  \midrule
-  }
+  \midrule}
+  %\midrule
+  %}
   \tabletail{\bottomrule}
-  \tiny
 
-\begin{xtabular}{rlllllllllll}
+  \bottomcaption{\textbf{Neurosynth-derived topics.} We report the top-weighted terms for each
+  of 80 topics identified using Latent Dirichlet Allocation~\citep{BleiEtal03}
+  applied to 9,204 functional neuroimaging articles in the Neurosynth
+  database~\citep{RubiEtal17}. See \textit{Reverse inference} for additional
+  information.}
+%\begin{xtabular}{rlllllllllll}
+
+  \begin{supertabular}{rlllllllllll}
   Cognitive control and task performance &   Cognitive control &            tasks &       control &         network &     conditions &  comparison &      performed &        common &     correlates &    experiment &            pre \\
@@
 
- \label{tab:topics}}
-\end{xtabular}
+ \label{tab:topics}
 
-\caption{\vspace{0.2cm} \normalsize \noindent Table S1: \textbf{Neurosynth-derived topics.} We report the top-weighted terms
-for each of 80 topics identified using Latent Dirichlet
-Allocation~\citep{BleiEtal03} applied to 9,204 functional neuroimaging articles
-in the Neurosynth database~\citep{RubiEtal17}. The topics, as well as
-associated brain maps identified using Neurosynth, were identified and reported
-in several prior studies~\citep{FoxEtal14, SulEtal17, ChenEtal20}. The topic
-labels for each topic were generated automatically with the following ChatGPT
-(chat.openai.com) prompt: ``Please help me come up with intuitive labels for
-topic topics I found by fitting a topic model to thousands of neuroscience and
-psychology articles. I'll paste in the top 10 highest-weighted words for each
-topic, and I'd like you to respond with a suggested label. For each topic,
-please respond with just the topic label and no other formatting or text. Here
-are the next topic's top words:'' followed by a comma-separated list of the
-given topic's top-weighted words reflected in the table. For some topics,
-ChatGPT responded with a longer-form response rather than a concise topic
-label. In these instances, on a case-by-case basis, we used a second follow-up
-prompt to achieve the given topic's label: ``could you please come up with a
-more concise label for that topic?''. We then manually identified a set of 11
-cognitive labels that were intended to encapsulate a representative range of
-widely studied low-level and high-level cognitive functions. In choosing the
-set of cognitive labels, we jointly considered each topic's ChatGPT-derived
-topic label, along with the top-weighted words for the topic. We attempted to
-generate a concise set of labels that still spanned the full set of cognitive
-functions reflected across the 80 topics. Topics that appeared unrelated to
-specific cognitive functions (e.g., topics related to specific methods or
-clinical themes) are designated with dashes.
+\end{supertabular}
+\end{center}
+\end{landscape}
 
-\end{landscape}
+\newpage
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{lr}
+  \toprule
+  Cognitive label &  Rank \\
+  \midrule
+  Cognitive control   &    10 \\
+  Language processing &     9 \\
+  Memory              &     8 \\
+  Emotion             &     7 \\
+  Social cognition    &     6 \\
+  Spatial cognition   &     5 \\
+  Attention           &     4 \\
+  Reward              &     3 \\
+  Sensory perception  &     2 \\
+  Motor control       &     1 \\
+  Resting state       &     0 \\
+  \bottomrule
+\end{tabular}
+
+\caption{\textbf{Ranking cognitive processes.} The table displays the output of
+a ChatGPT~\citep{ChatGPT} prompt asking for a ranking of the cognitive
+processes reflected in the labels from Table~\ref{tab:topics}. See \textit{Ranking cognitive
+processes} for additional detail.}
+\label{tab:topic-tags}
+
+\end{center}
+\end{table}
+
+\newpage
 
@@
 front of many of the topics denote manually identified cognitive labels
-(Tab.~S1).}
+(Tab.~\ref{tab:topics}).}
 
@@
 legend for the condition and cognitive function classifications is displayed in
-the lower right of the figure. Table~S1 provides a list of each topic's
+the lower right of the figure. Table~\ref{tab:topics} provides a list of each topic's
 top-weighted terms, along with each topic's manually labeled cognitive
@@
 condition.} The heatmaps display the correlations between the brain maps
-(Fig.~\componentBrains) for each principal component (sub-panel), across each
+(Fig.~\ref{fig:components}) for each principal component (sub-panel), across each
 pair of conditions (rows and columns of each sub-panel's matrix, indicated by
@@
 component.} Each sub-panels network plot summarizes the pattern of correlations
-between the topic correlations from each of the $n$\textsuperscript{th} top-weighted principal components
-(sub-panel), for each experimental condition (gray-outlined dots). The line
-thicknesses are proportional to the correlation values (correlation
-coefficients are noted in the heatmaps in Panel C).}
+between the topic correlations from each of the $n$\textsuperscript{th}
+top-weighted principal components (sub-panel), for each experimental condition
+(gray-outlined dots). The line thicknesses are proportional to the correlation
+values (correlation coefficients are noted in the heatmaps in Panel C).}
 

### 2023-03-16 paper/compile.sh

@@ -1,0 +1,20 @@
+#!/bin/bash
+
+latex -interaction=nonstopmode main
+bibtex main
+latex -interaction=nonstopmode main
+latex -interaction=nonstopmode main
+latex -interaction=nonstopmode main
+latex -interaction=nonstopmode main
+pdflatex -interaction=nonstopmode main
+
+latex -interaction=nonstopmode supplement
+bibtex supplement
+latex -interaction=nonstopmode supplement
+latex -interaction=nonstopmode supplement
+latex -interaction=nonstopmode supplement
+latex -interaction=nonstopmode supplement
+pdflatex -interaction=nonstopmode supplement
+
+rm *.cb* *.dvi *.log *.blg *.aux *.fff *.out
+

### 2023-03-16 paper/main.tex
finished complete drafts of main text and supplement

@@
 information, we should be able to use the activity patterns to accurately
-differentiate between different aspects of the task~\citep[e.g., using pattern
+differentiate between different aspects of a task~\citep[e.g., using pattern
 classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
@@
 And if the signal is compressible, a low-dimensional embedding of the signal
-will be similarly informative to the original signal
+will be similarly informative as the original signal
 (Fig.~\ref{fig:information-compression}D).
@@
 peak decoding accuracy in each fold. (This distribution is nearly flat, as
-illustrated in Panel E.) \textbf{D. Estimating ``informativeness'' for each
+also illustrated in Panel E.) \textbf{D. Estimating ``informativeness'' for each
 condition.} The density plots display the peak (corrected) decoding accuracies
@@
 of 5\% accuracy using fewer components than the paragraph condition decoders
-($t(99) = -7.429, p < 0.001)$) or word condition decoders ($t(99) = -7.300, p <
-0.001)$), and decoding accuracy never exceeded 5\% for the rest condition. This
+($t(99) = -7.429, p < 0.001$) or word condition decoders ($t(99) = -7.300, p <
+0.001$), and decoding accuracy never exceeded 5\% for the rest condition. This
 suggests that brain activity patterns evoked by cognitively richer conditions
@@
 of $k = 20$ components in Fig.~\ref{fig:inflection-quarters}C: intact: $t(99) =
-7.915, p < 0.001)$; paragraph: $t(99) = 2.354, p = 0.021)$). These changes may
+7.915, p < 0.001$; paragraph: $t(99) = 2.354, p = 0.021$). These changes may
 reflect an increase in comprehension or depth of processing with listening
 time. In contrast, the decoding accuracy and compressibility \textit{decreased}
-with listening time in the word condition ($t(99) = -10.747, p < 0.001)$) and
-rest condition ($t(99) = -22.081, p < 0.001)$). This might reflect the
+with listening time in the word condition ($t(99) = -10.747, p < 0.001$) and
+rest condition ($t(99) = -22.081, p < 0.001$). This might reflect the
 depletion of attentional resources in the less-engaging word and rest
@@
 different for the word condition (highest-weighted functions: sensory
-perception and cognitive control). The fourth components' weighting were
+perception and cognitive control). The fourth components' weighting 
 grouped the paragraph and word conditions (highest-weighted functions: memory,
@@
 colored dots). \textbf{D. Associations between per-condition components, by
-component.} Each sub-panels network plot summarizes the pattern of correlations
+component.} Each sub-panel's network plot summarizes the pattern of correlations
 between the $n$\textsuperscript{th} top-weighted principal components
@@
 and compressible, and data from less cognitively rich conditions became
-\textit{less} informative and compressible (Fig.~\ref{fig:inflection-quarters}).
-We also repeated these analyses separately for different brain networks. We
-found that networks traditionally associated with higher-level cognitive
-functions tended to provide more informative brain patterns than networks
-traditionally associated with lower level cognitive functions
-(Fig.~\ref{fig:networks}). Finally, we examined the most dominant components of
-the brain activity patterns from each experimental condition. We used a reverse
-inference approach~\citep{RubiEtal17} to identify the terms in the neuroimaging
-literature most commonly associated with the corresponding maps. As summarized
-in Figure~\ref{fig:neurosynth-summary}, we found that terms associated with memory and
-sensory processing were associated with the strongest components in all three
-story listening conditions. Terms associated with sensory integration were
-associated with the strongest components in the intact and paragraph-scrambled
-conditions. Terms associated with sentence comprehension, emotion, and valence
-were associated with the strongest components in the intact condition. Finally,
-terms associated with the default mode network were associated with the
-strongest components in the word-scrambled and resting state conditions. Taken
-together, our findings indicate that the informativeness and compressibility of
-our brain activity patterns are task-dependent, and these properties change
-systematically with factors like cognitive richness and depth of processing.
+\textit{less} informative and compressible
+(Fig.~\ref{fig:inflection-quarters}). We also repeated these analyses
+separately for different brain networks. We found that networks traditionally
+associated with higher-level cognitive functions tended to provide more
+informative brain patterns than networks traditionally associated with lower
+level cognitive functions (Fig.~\ref{fig:networks}). Finally, we examined the
+most dominant components of the brain activity patterns from each experimental
+condition. We used a reverse inference approach~\citep{RubiEtal17} to identify
+the terms in the neuroimaging literature most commonly associated with the
+corresponding maps. As summarized in Figure~\ref{fig:neurosynth-summary}E, we
+found that the intact and paragraph conditions tended to weight on higher-level
+cognitive processes more than lower-level cognitive processes, whereas the word
+condition weighted on lower-level processes more than higher-level processes
+and the rest condition showed no difference in high-level versus low-level
+weighting. Taken together, our findings indicate that the informativeness and
+compressibility of our brain activity patterns are task-dependent, and these
+properties change systematically with factors like cognitive richness and depth
+of processing.
 
@@
 compressible and also that those activity patterns will contain redundancies.
-For example, the extent ot which activity patterns at one brain area can be
+For example, the extent to which activity patterns at one brain area can be
 inferred or predicted from activity patterns at other
@@
 information available in or represented by those brain areas. If brain patterns
-in one area are recoverable using brain patterns in another area, then the
+in one area are recoverable using brain patterns in another area, then a
 ``signal'' used to convey the activity patterns could be compressed by removing
@@
 smaller number of radial basis function (RBF) nodes~\citep[in this case, 700
-nodes, as determined by an optimization procedure;][]{MannEtal18}). This
+nodes, as determined by an optimization procedure;][]{MannEtal18}. This
 provides a convenient representation for examining full-brain activity patterns
@@
 weight. The sum of these weighted RBF activation maps provides a full-brain
-image, in MNI152 space, of the given principal component.
+image, in MNI152 space, of the given principal component (Fig.~\componentBrains).
 
@@
 topic label. In these instances, on a case-by-case basis, we used a second
-follow-up prompt to achieve the given topic's label: ``could you please come up
+follow-up prompt to achieve the given topic's label: ``Could you please come up
 with a more concise label for that topic?''. We then manually identified a set
@@
 highest-level to lowest-level, where higher values indicate higher-order or
-higher-level processes. return the result as a csv file with a header row and
+higher-level processes. Return the result as a csv file with a header row and
 two columns: `Cognitive label' and `Rank'. Here are the processes: cognitive
@@
 cognition, attention, reward, sensory perception, motor control, resting
-state''.  Table~\topicTags displays the output.  We used these labels in
+state''.  Table~\topicTags~displays the output.  We used these labels in
 the analysis presented in Figure~\ref{fig:neurosynth-summary}E to help summarize

### 2023-03-16 paper/supplement.tex
finished complete drafts of main text and supplement

@@
 
-\caption{\textbf{Top terms associated with the highest-weighted components by
-condition, broken down by story segment.} Each group of five rows corresponds
-to an experimental condition (denoted by color, as indicated in the legend in
-the lower right), and the columns and shading correspond to the component
-number (ranked by proportion of variance explained). The colored squares in
-front of many of the topics denote manually identified cognitive labels
-(Tab.~\ref{tab:topics}).}
+\caption{\textbf{Highest-weighted topics associated with the highest-weighted
+components by condition, broken down by story segment.} Each group of five rows
+corresponds to an experimental condition (denoted by color, as indicated in the
+legend in the lower right), and the columns and shading correspond to the
+component number (ranked by proportion of variance explained). The colored
+squares in front of many of the topics denote manually identified cognitive
+labels (Tab.~\ref{tab:topics}).}
 
@@
 (columns), the components' brain maps are projected onto four views: left
-sagittal, axial, right sagittal, and coronal.  The color scale is the same for all panels and matches the coloring in Fig.~\neurosynth C.}
+sagittal, axial, right sagittal, and coronal. The color scale is the same for
+all panels and matches the coloring in Figure~\neurosynth C.}
 
@@
 colored dots). \textbf{D. Associations between per-condition topic weights, by
-component.} Each sub-panels network plot summarizes the pattern of correlations
+component.} Each sub-panel's network plot summarizes the pattern of correlations
 between the topic correlations from each of the $n$\textsuperscript{th}

### 2023-03-16 code/notebooks/helpers.py
updated code after verifying in a clean environment

@@ -13,6 +13,7 @@ from scipy.spatial.distance import cdist
 import os
 import warnings
 import pickle
+import requests
 
 basedir = os.path.split(os.path.split(os.getcwd())[0])[0]
 datadir = os.path.join(basedir, 'data')
@@ -239,7 +240,7 @@ def ridge_plot(x, column='Number of components', fname=None, xlim=[-99, 700], hu
 
 
 def get_data():
-    url = 'https://www.dropbox.com/s/29a48lv3j5ybcvw/pieman2_htfa.pkl?dl=1'
+    url = 'https://www.dropbox.com/s/fjntg498tl6gfvz/pieman2_htfa.pkl?dl=1'
 
     fname = os.path.join(datadir, 'pieman2_htfa.pkl')
     if not os.path.exists(fname):

### 2023-03-17 paper/main.tex
final round of edits before submission

@@
 We take as a given that brain activity is highly flexible: our brains can
-exhibit nearly infinite varieties of activity patterns. This flexibility implies that our
-brains activity patterns are highly information rich. However, brain activity
-patterns are also highly structured. For example, full-brain correlation
-matrices are stable within~\citep{FinnEtal15, FinnEtal17, GratEtal18} and
-across~\citep{YeoEtal11, GlerEtal12, GratEtal18, ColeEtal14} individuals. This
-stability suggests that our brains' activity patterns are at least partially
-constrained, for example by anatomical, external, or internal factors.
-Constraints on brain activity that limit its flexibility decrease
+exhibit nearly infinite varieties of activity patterns. This flexibility
+implies that our brains' activity patterns are highly information rich. However,
+brain activity patterns are also highly structured. For example, full-brain
+correlation matrices are stable within~\citep{FinnEtal15, FinnEtal17,
+GratEtal18} and across~\citep{YeoEtal11, GlerEtal12, GratEtal18, ColeEtal14}
+individuals. This stability suggests that our brains' activity patterns are at
+least partially constrained, for example by anatomical, external, or internal
+factors. Constraints on brain activity that limit its flexibility decrease
 expressiveness (i.e., its information rate). However, constraints on brain
@@
   each image (the 100\% columns denote the original images). The numbers of
-  components are indicate in parentheses, and the resulting ``compressed''
+  components are indicated in parentheses, and the resulting ``compressed''
   images are displayed. \textbf{B. Representing two images with different
@@
 experimental condition, participants listened to an audio recording of a
-10-minute story. In the \textit{paragraph}-scrambled experimental condition,
-partipants listened to a temporally scrambled version of the story, where the
-paragraphs occurred out of order, but where the same set of paragraphs was
-presented over the entire listening interval. All participants in this
-condition experienced the scrambled paragraphs in the same order. In the
-\textit{word}-scrambled experimental condition, participants listened to a
-temporally scrambled version of the story, where the words occurred in a random
-order. Again, all participants in this condition experienced the scrambled
-words in the same order. Finally, in the \textit{rest} experimental condition,
-participants lay in the scanner with no overt stimulus, while keeping their
-eyes open and blinking as needed. This public dataset provided a convenient
-means for testing our hypothesis that different levels of cognitive processing
-and engagement affect how informative and compressible the associated brain
-patterns are.
+10-minute Moth Radio Hour story, \textit{Pie Man}, by Jim O'Grady. In the
+\textit{paragraph}-scrambled experimental condition, participants listened to a
+temporally scrambled version of the story, where the paragraphs occurred out of
+order, but where the same set of paragraphs was presented over the entire
+listening interval. All participants in this condition experienced the
+scrambled paragraphs in the same order. In the \textit{word}-scrambled
+experimental condition, participants listened to a temporally scrambled version
+of the story, where the words occurred in a random order. Again, all
+participants in this condition experienced the scrambled words in the same
+order. Finally, in the \textit{rest} experimental condition, participants lay
+in the scanner with no overt stimulus, while keeping their eyes open and
+blinking as needed. This public dataset provided a convenient means for testing
+our hypothesis that different levels of cognitive processing and engagement
+affect how informative and compressible the associated brain patterns are.
 
@@
 
-To evaluate the relation between informativeness and compressibiilty for brain
+To evaluate the relation between informativeness and compressibility for brain
 activity from each experimental condition, we trained a series of
@@
 cognitively richer conditions tended to be both more informative (i.e.,
-associated with higher decoding accuracies) \textit{and} more compressible
+associated with higher peak decoding accuracies) \textit{and} more compressible
 (i.e., requiring fewer components to achieve the 5\% accuracy threshold).
@@
 (corresponding to the columns of Panels A and B). The $y$-axis denotes the
-slopes of the regression lines. All error ribbons denote bootstrap-estimated
+slopes of the regression lines. The black vertical line marks $k = 20$ components,
+as referenced in the main text.  All error ribbons denote bootstrap-estimated
 95\% confidence intervals.}
@@
   accuracy (or, for the rest condition, peak decoding accuracy, as described
-  above). Error ribbons in Panels E and F denote 95\% bootstrap-estimated
-  confidence intervals.}
+  above). Error ribbons in Panels E and F denote bootstrap-estimated
+  95\% confidence intervals.}
 
@@
 networks to higher-order networks, we found that decoding accuracy tended to
-increase, particularly in the higer-order experimental conditions
+increase, particularly in the higher-level experimental conditions
 (Fig.~\ref{fig:networks}D, E). This suggests that higher-order networks may
@@
 five components (for each condition) reflected in Panel A. \textbf{C. Component
-images.} Each plot diplays a right sagittal view of a glass brain projection of
-the top five principal components (columns) for each experimental conditions
+images.} Each plot displays a right sagittal view of a glass brain projection of
+the top five principal components (columns) for each experimental condition
 (rows). Additional projections for each component may be found in
@@
 functions that varied by experimental condition. For example, the top principal
-components from every condition weighted similarly on the full set of
-Neurosynth topics (Fig.~\ref{fig:neurosynth-pca}A) and cognitive functions
-(Figs.~\ref{fig:neurosynth-summary}A, B and~\neurosynthFull~A, B), suggesting
-that these components might reflect a set of functions or activity patterns
-that are common across all conditions. The second components' weightings were
-similar across the intact, paragraph, and rest conditions (highest-weighted
-functions: cognitive control, memory, social cognition, and resting state), but
-different for the word condition (highest-weighted functions: sensory
-perception and cognitive control). The fourth components' weighting 
-grouped the paragraph and word conditions (highest-weighted functions: memory,
-language processing, and cognitive control) and the intact and rest conditions
-(highest-weighted functions: emotion, social cognition). We also used
-ChatGPT~\citep{ChatGPT} to sort the list of manually tagged cognitive functions
-from lowest-level to highest-level (Tab.~\topicTags,
-Fig.~\ref{fig:neurosynth-summary}E; also see \textit{Ranking cognititive
+components from every condition weighted similarly (across conditions) on the
+full set of Neurosynth topics (Fig.~\ref{fig:neurosynth-pca}A) and cognitive
+functions (Figs.~\ref{fig:neurosynth-summary}A, B and~\neurosynthFull~A, B),
+suggesting that these components might reflect a set of functions or activity
+patterns that are common across all conditions. The second components'
+weightings were similar across the intact, paragraph, and rest conditions
+(highest-weighted functions: cognitive control, memory, social cognition, and
+resting state), but different for the word condition (highest-weighted
+functions: sensory perception and cognitive control). The fourth components'
+weighting grouped the paragraph and word conditions (highest-weighted
+functions: memory, language processing, and cognitive control) and the intact
+and rest conditions (highest-weighted functions: emotion, social cognition). We
+also used ChatGPT~\citep{ChatGPT} to sort the list of manually tagged cognitive
+functions from lowest-level to highest-level (Tab.~\topicTags,
+Fig.~\ref{fig:neurosynth-summary}E; also see \textit{Ranking cognitive
 processes}). We found that higher-level functions tended to be weighted on more
@@
 lower-level functions. The top components from the word condition showed the
-opposite tendancy, whereby \textit{lower}-level functions tended to be weighted
+opposite tendency, whereby \textit{lower}-level functions tended to be weighted
 on more heavily than higher-level functions. The components from the rest
@@
 colors denote cognitive functions). The line thicknesses are proportional to
-the correlation values (correlation coefficients are noted in the heatmaps in
+the correlation values (correlation coefficients are noted in the heat maps in
 Panel A). \textbf{C. Correlations between each principal component, by
-condition.} The heatmaps display the correlations between the brain maps
+condition.} The heat maps display the correlations between the brain maps
 (Fig.~\componentBrains) for each principal component (sub-panel), across each
@@
 thicknesses are proportional to the correlation values (correlation
-coefficients are noted in the heatmaps in Panel C). \textbf{E. Change in
+coefficients are noted in the heat maps in Panel C). \textbf{E. Change in
 weights by condition.} The bar heights reflect the slopes of regression lines,
@@
 broader literature on the correlational and causal structure of brain activity
-patterns~\citep{PretEtal17, OwenEtal21, RogeEtal07, RubiSpor10, SizeEtal18,
-SmitEtal13b, SmitEtal13c, SrinEtal07, TomaVolk11, YeoEtal11, AdacEtal12,
-BassSpor17, BullSpor09, SporHone06, SporBetz16, SporZwi04, DhamEtal08,
-KorzEtal08, BrovEtal04}. Correlations or causal associations between different
-brain regions simultaneously imply that full-brain activity patterns will be
-compressible and also that those activity patterns will contain redundancies.
-For example, the extent to which activity patterns at one brain area can be
-inferred or predicted from activity patterns at other
-areas~\citep[e.g.,][]{OwenEtal20, ScanEtal21}, reflects overlap in the
+patterns and networks~\citep{PretEtal17, OwenEtal21, RogeEtal07, RubiSpor10,
+SizeEtal18, SmitEtal13b, SmitEtal13c, SrinEtal07, TomaVolk11, YeoEtal11,
+AdacEtal12, BassSpor17, BullSpor09, SporHone06, SporBetz16, SporZwi04,
+DhamEtal08, KorzEtal08, BrovEtal04, LynnBass21}. Correlations or causal
+associations between different brain regions simultaneously imply that
+full-brain activity patterns will be compressible and also that those activity
+patterns will contain redundancies. For example, the extent to which activity
+patterns at one brain area can be inferred or predicted from activity patterns
+at other areas~\citep[e.g.,][]{OwenEtal20, ScanEtal21}, reflects overlap in the
 information available in or represented by those brain areas. If brain patterns
@@
 computations are linked. We identified two aspects of brain activity patterns,
-informativeness and compressibiilty, that appear to change systematically with
+informativeness and compressibility, that appear to change systematically with
 task demands and across brain networks. Our work helps to clarify how the
@@
 in several prior studies~\citep{FoxEtal14, SulEtal17, ChenEtal20}. The topic
-labels for each topic were generated automatically with the following ChatGPT
-(chat.openai.com) prompt: ``Please help me come up with intuitive labels for
-topic topics I found by fitting a topic model to thousands of neuroscience and
+labels for each topic were generated automatically with the following
+ChatGPT~\citep{ChatGPT} prompt: ``Please help me come up with intuitive labels
+for topics I found by fitting a topic model to thousands of neuroscience and
 psychology articles. I'll paste in the top 10 highest-weighted words for each
@@
 SulEtal17, ChenEtal20} we treated the correlation between a given component's
-brain map and each topic's brain map as a reflection of how much the component
-was reflective of the given topic. This resulted in a set of 80 ``weights''
-(correlation coefficients) for each component's brain map, with one weight per
-Neurosynth-derived topic.
+brain map and each topic's brain map as an approximate measure of how much the
+component was reflective of the given topic. This resulted in a set of 80
+``weights'' (correlation coefficients) for each component's brain map, with one
+weight per Neurosynth-derived topic.
 
@@
 
-Conceptulization: J.R.M. and L.L.W.O. Methodology: J.R.M. and L.L.W.O.
-Implementation: L.L.W.O. Analysis: L.L.W.O. Writing, Reviewing, and Editing:
-J.R.M. and L.L.W.O. Supervision: J.R.M.
+Conceptualization: J.R.M. and L.L.W.O. Methodology: J.R.M. and L.L.W.O.
+Implementation: J.R.M. and L.L.W.O. Analysis: J.R.M. and L.L.W.O. Writing,
+Reviewing, and Editing: J.R.M. and L.L.W.O. Funding acquisition: J.R.M.
+Supervision: J.R.M.
 

### 2023-03-17 paper/supplement.tex
final round of edits before submission

@@
 \maketitle
-
-% \begin{figure}
-%   \centering
-%   \includegraphics[width=\textwidth]{figs/pca_neurosynth_thirds}
-
-% \caption{\textbf{Top terms associated with the highest-weighted components by
-% condition, broken down by story segment.} Each group of three rows corresponds
-% to an experimental condition, and the colors correspond to the component number
-% (ranked by proportion of variance explained). The rows' numbering denotes the
-% story segment used to compute each map (1: first third; 2: second third; 3:
-% third third). The inflated brain plots display the top 20 highest-weighted hubs
-% (see \textit{Topographic Factor Analysis}) for each components. The lists on
-% the right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
-% each components' brain map. Analogous maps computed for the entire story may be
-% found in Figure~\synth.}
-
-% \label{fig:neurosynth-thirds}
-
-% \end{figure}
-
-%\begin{landscape}
 
@@
 \caption{\textbf{Comparisons between per-component topic correlations across conditions.}  Each sub-panel displays a scatterplot comparing the per-topic correlations for two or more experimental conditions.
-Each dot donotes the correlations for a single topic (indicated by the legend on the right).  The topics are colored according to the ranked order of the correlations
+Each dot denotes the correlations for a single topic (indicated by the legend on the right).  The topics are colored according to the ranked order of the correlations
 between the topic's brain maps and the brain map for the first principal component in the intact condition.  \textbf{A. Comparisons between correlations for each pair of experimental conditions.}  The
@@
 colors denote cognitive functions). The line thicknesses are proportional to
-the correlation values (correlation coefficients are noted in the heatmaps in
+the correlation values (correlation coefficients are noted in the heat maps in
 Panel A). \textbf{C. Correlations between each principal component, by
-condition.} The heatmaps display the correlations between the brain maps
+condition.} The heat maps display the correlations between the brain maps
 (Fig.~\ref{fig:components}) for each principal component (sub-panel), across each
@@
 (gray-outlined dots). The line thicknesses are proportional to the correlation
-values (correlation coefficients are noted in the heatmaps in Panel C).}
+values (correlation coefficients are noted in the heat maps in Panel C).}
 

### 2023-03-17 paper/admin/cover_letter.tex
final round of edits before submission
title change

@@
 \newcommand{\articletype}{Article}
-\newcommand{\myTitle}{High-order cognition is supported by information-rich but compressible brain activity patterns}
+\newcommand{\myTitle}{High-level cognition is supported by information-rich but compressible brain activity patterns}
 \newcommand{\corresponding}{Jeremy R. Manning}
@@
   leads to a rich understanding of the story. Listening to temporally scrambled
-  versions of the story might lead to understanding invidual paragraphs or
+  versions of the story might lead to understanding individual paragraphs or
   words. However, individual moments in the temporally scrambled stories lack

### 2023-12-18 code/notebooks/helpers.py
working on new simulations, also moved plot_accuracy function to helpers

@@ -294,3 +294,25 @@ def plot_info_and_compressibility_scatter(x, fname=None):
 
 def rbf(R, center, width):
     return np.exp(-np.sum((R - center) ** 2, axis=1) / width)
+
+
+def plot_accuracy(x, figdir=None, fname=None, conditions=['intact', 'paragraph', 'word', 'rest'], condition_colors=condition_colors, ylim=[-0.01, 0.35], xlim=[3, 700]):
+    if figdir is not None and not os.path.exists(figdir):
+        os.makedirs(figdir)
+
+    fig = plt.figure(figsize=(4, 3))
+    ax = plt.gca()
+
+    for c in conditions:
+        sns.lineplot(x[c], x='Number of components', y='Relative decoding accuracy', label=c.capitalize(), color=condition_colors[c], legend=False)
+    
+    ax.set_xlabel('Number of components', fontsize=12)
+    ax.set_ylabel('Relative decoding accuracy', fontsize=12)
+    ax.set_ylim(ylim)
+    ax.set_xlim(xlim)
+    ax.spines[['right', 'top']].set_visible(False)
+
+    if fname is not None:
+        fig.savefig(os.path.join(figdir, fname + '.pdf'), bbox_inches='tight')
+
+    return fig

### 2023-12-19 code/notebooks/helpers.py
updated simulations and information and compressibility figure

@@ -296,15 +296,18 @@ def rbf(R, center, width):
     return np.exp(-np.sum((R - center) ** 2, axis=1) / width)
 
 
-def plot_accuracy(x, figdir=None, fname=None, conditions=['intact', 'paragraph', 'word', 'rest'], condition_colors=condition_colors, ylim=[-0.01, 0.35], xlim=[3, 700]):
+def plot_accuracy(x, figdir=None, fname=None, conditions=['intact', 'paragraph', 'word', 'rest'], condition_colors=condition_colors, ylim=[-0.01, 0.35], xlim=[3, 700], ax=None):
     if figdir is not None and not os.path.exists(figdir):
         os.makedirs(figdir)
 
-    fig = plt.figure(figsize=(4, 3))
-    ax = plt.gca()
+    if ax is None:
+        fig = plt.figure(figsize=(4, 3))
+        ax = plt.gca()
+    else:
+        fig = plt.gcf()
 
     for c in conditions:
-        sns.lineplot(x[c], x='Number of components', y='Relative decoding accuracy', label=c.capitalize(), color=condition_colors[c], legend=False)
+        sns.lineplot(x[c], x='Number of components', y='Relative decoding accuracy', label=c.capitalize(), color=condition_colors[c], legend=False, ax=ax)
     
     ax.set_xlabel('Number of components', fontsize=12)
     ax.set_ylabel('Relative decoding accuracy', fontsize=12)

### 2023-12-19 paper/main.tex
updated stats (correct for multiple comparisons) in decoding_and_compression notebook   - note: still need to update stats in paper

@@
 
-\title{High-order cognition is supported by information-rich but compressible brain activity patterns}
+\title{High-level cognition is supported by information-rich but compressible brain activity patterns}
 
@@
   \caption{\textbf{Information content and compressibility.} \textbf{A.
-  Variance explained for two images.} We applied principal components analysis
-  to a photograph and drawing, treating the rows of the images as
-  ``observations.'' Across columns, we identified the numbers of components
-  required to explain 100\%, 95\%, 75\%, or 50\% of the cumulative variance in
-  each image (the 100\% columns denote the original images). The numbers of
-  components are indicated in parentheses, and the resulting ``compressed''
-  images are displayed. \textbf{B. Representing two images with different
-  numbers of components.} Using the same principal component decompositions as
-  in Panel A, we computed the cumulative proportion of variance explained with
-  380 (original images), 20, 4, or 2 components. \textbf{C. Cumulative variance
-  explained versus number of components.} For the images displayed in Panels A
-  and B, we plot the cumulative proportion of variance explained as a function
-  of the number of components used to represent each image. \textbf{D.
-  Information rate and compressibility.} Across multiple images, the
-  information rate (i.e., the amount of information contained in each image;
-  horizontal axis) is high when each individual pixel provides information that
-  cannot be inferred from other pixels. High-information rate images tend to be
-  high-resolution, and low-information rate images tend to be low-resolution.
-  Compressibility is related to the difference between the information required
-  to specify the original versus compressed images (vertical axis). Highly
-  compressible images often contain predictable structure (redundancies) that
-  can be leveraged to represent the images much more efficiently than in their
-  original feature spaces.}
-
+  Variance explained for two images.} Consider two example images: a photograph
+  and simple drawing. The images have the same resolutions, but their content
+  is very different. The photograph is more information rich: when we compare
+  the original photograph and drawing (leftmost column), we can see that the
+  photograph captures much more detail. We can also apply principal components
+  analysis to the images, treating the rows of the images as ``observations.''
+  Across columns, we identified the numbers of components required to explain
+  100\%, 95\%, 75\%, or 50\% of the cumulative variance in each image (the
+  100\% columns denote the original images). The numbers of components are
+  indicated in parentheses, and the resulting ``compressed'' images are
+  displayed. This analysis reveals that the drawing is more compressible: just
+  16 components can explain 95\% of the variance in teh drawing, whereas 50
+  components are required to explain 95\% of the variance in the photograph.
+  \textbf{B. Representing two images with different numbers of components.}
+  Using the same principal component decompositions as in Panel A, we computed
+  the cumulative proportion of variance explained with 380 (original images),
+  20, 4, or 2 components. This analysis provides another way of characterizing
+  the compressibility of each image by showing that the same number of
+  components can explain more variance in the drawing than in the photograph.
+  \textbf{C. Template data from four synthetic datasets.} We constructed four
+  synthetic datasets, each comprising 25 features (columns) observed across 100
+  samples (rows) from each of 10 simulated ``participants.'' The datasets were
+  constructed to contain different levels of informativeness and
+  compressibility (see \textit{Synthetic data}). \textbf{D. Decoding accuracy
+  by number of components.} For each synthetic dataset, we trained
+  across-participant classifiers to decode timepoint labels.  Each panel displays the
+  decoding accuracy as a function of the number of components used to represent
+  the data. Error ribbons denote bootstrap-estimated 95\% confidence intervals.}
 \label{fig:information-compression} 
@@
 
+\subsection*{Synthetic data}
+
+To help illustrate the relationship between informativeness and compressibility
+(Fig.~\ref{fig:information-compression}), we generated four synthetic datasets,
+varying in informativeness and compressibility. Each dataset comprised
+simulated observations of $K = 25$ features across $N = 100$ timepoints, from each of $S = 10$
+participants. To create each dataset, we first constructed a ``template''
+matrix of $N$ timepoints by $K$ features. We then generated participant-specific
+data by adding independent noise to each entry in template matrix, drawn from
+the unit normal distribution (i.e., with a mean of 0 and a variance of 1). We
+repeated this process for each participant, yielding $S$ participant-specific
+matrices for each dataset.
+
+Since we estimate informativeness using the temporal decoding accuracy across
+participants, highly informative data will tend to have observations that are
+highly timepoint specific. Relatively uninformative data, in contrast, will
+tend to have more similar observations across timepoints. To generate data with
+``high informativeness,'' we constructed template matrices whose rows
+(observations) were drawn independently from zero-mean multivariate normal
+distributions. The covariances of these distributions were determined according
+to the desired compressibility of the data, as described below. We used a
+multi-step process to generate data with ``low informativeness.'' First we
+generated new template matrices using the same procedure as for the ``high
+informativeness'' datasets. We then multiplied each matrix by a constant ($\rho
+= 0.1$) and computed the cumulative sum of each matrix's rows. This yielded
+matrices whose rows were highly similar across observations.
+
+Compressibility reflects the extent to which decoding accuracy is affected by
+reducing the number of components used to represent the data. Highly
+compressible data will tend to exhibit more similarities across features,
+whereas less compressible data will tend to show greater independence across
+features. To generate data with ``high compressibility,'' we set the covariance
+matrix of the multivariate normal distribution to a toeplitz matrix whose first
+row was given by $\left[K, K - 1, ..., 1\right]$. To generate data with ``low
+compressibility,'' we set the covariance matrix to the identity matrix.
+
+Template matrices for datasets with high informativeness and high
+compressibility, high informativeness and low compressibility, low
+informativeness and high compressibility, and low informativeness and low
+compressibility are displayed in Figure~\ref{fig:information-compression}C. The
+corresponding decoding curves are displayed in
+Figure~\ref{fig:information-compression}D.
+
 \section*{Data and code availability}

### 2023-12-20 paper/main.tex
working on introduction text related to fig 1.

@@
 important aspects of the original signal. In general, information content and
-compressibility are related but are partially dissociable
+compressibility are often related but are also dissociable
 (Fig.~\ref{fig:information-compression}). If a given signal (e.g., a
 representation of brain activity patterns) contains more information about
-ongoing cognitive processes, then the peak decoding accuracy should be high.
-And if the signal is compressible, a low-dimensional embedding of the signal
-will be similarly informative as the original signal
-(Fig.~\ref{fig:information-compression}D).
+ongoing cognitive processes, then the peak decoding accuracy should be high. In
+the simulations shown in Figure~\ref{fig:information-compression}C we construct
+synthetic datasets that have high or low levels of informativeness by varying
+temporal autocorrelations in the data (see \textit{Synthetic data}).
+If a signal is compressible, then a low-dimensional embedding of the signal
+will be similarly informative as the original signal.  In the simulations
+shown in Figure~\ref{fig:information-compression}C we construct synthetic
+datasets that have high or low levels of compressibility by varying
+the covariance structure across features (see \textit{Synthetic data}).  As shown
+in Figure~\ref{fig:information-compression}D, highly informative datasets
+yield higher decoding accuracies than less informative datasets (i.e., the peaks of the curves in
+the top panels of Fig.~\ref{fig:information-compression}D are higher than the
+peaks of the curves in the bottom panels). Highly compressible datasets show
+steeper slopes when we plot decoding accuracy as a function of the number of
+components used to represent the data (i.e., the slopes of the curves in the
+left panels of Fig.~\ref{fig:information-compression}D are steeper than the
+slopes of the curves in the right panels).  Whereas characterizing
+the informativeness and compressibility of synthetic data can be instructive,
+we are ultimately interested in understanding how these properties relate to
+brain activity patterns recorded under different cognitive circumstances.
 
@@
 
+Unless otherwise noted, all statistical tests are two-sided, all error bars and
+error ribbons denote bootstrap-estimated 95\% confidence intervals across
+participants, and all reported $p$-values are corrected for multiple
+comparisons using the Benjamini-Hochberg procedure~\citep{BenjHoch95}.
+
 \subsection*{Functional neuroimaging data collected during story

### 2023-12-21 paper/main.tex
adding stats to Fig. 4 [WIP]

@@
 conditions.
+
+These results make some intuitive sense. As the contextual information
+available to participants increases (i.e., over time in the cognitively rich
+``intact'' and ``paragraph'' conditions), it makes sense that this might
+constrain neural responses to a greater extent. While this pattern may not
+necessarily hold for \textit{every} possible story or stimulus, we suspect that
+it is generally the case that our knowledge about what is happening in a story
+tends to increase as we experience more about it. In turn, this could lead to
+greater consistency in different people's interpretations of and neural
+responses to the stimulus. Similarly, as participants are left to ``mind
+wander,'' or as they experience mental fatigue (i.e., over time in the less
+cognitively rich ``word'' and ``rest'' conditions), we suggest that this might
+lead to greater variability in neural responses across people, resulting in
+lower decoding accuracy. Again, it is not necessarily the case that every
+possible ``unengaging'' stimulus will lead to greater neural variability as
+time progresses, but we suspect this phenomenon is likely to hold for a variety
+of such stimuli. These findings replicate at least to a limited extent (e.g.,
+across both cognitively rich conditions and across both cognitively
+impoverished conditions, and for the different groups of participants in each
+of those conditions). However, determining whether these patterns generalize to
+other stimuli would require additional study (with new stimuli).
 

### 2023-12-21 code/notebooks/helpers.py

@@ -319,3 +319,10 @@ def plot_accuracy(x, figdir=None, fname=None, conditions=['intact', 'paragraph',
         fig.savefig(os.path.join(figdir, fname + '.pdf'), bbox_inches='tight')
 
     return fig
+
+
+def pstring(pval):
+    if pval < 0.001:
+        return 'p < 0.001'
+    else:
+        return f'p = {pval:.3f}'

### 2023-12-21 paper/main.tex
working on addressing R1 comments [WIP]
still not working: decoding_and_compression notebook is currently in disrepair-- i need to re-write the ridge_plot function to include a "kdeplot" replacement where the area sums to 1.

@@
 \usepackage{soul}
+\usepackage{rotating}
 \linenumbers
@@
 
-We applied dimensionality reduction algorithms and pattern classifiers to
-functional neuroimaging data collected as participants listened to a story,
-temporally scrambled versions of the story, or underwent a resting state
-scanning session. These experimental conditions were intended to require
-different depths of processing and inspire different levels of cognitive
-engagement. We considered two primary aspects of the data. First, we treated
-the maximum achievable decoding accuracy across participants as an indicator of
-the ``informativeness'' of the recorded patterns. Second, we treated the number
-of features (components) required to achieve a threshold decoding accuracy as a
+Brain activity patterns are highly flexible and often complex, but also highly
+structured. Here we examined how fundamental properties of brain activity
+patterns relate to ongoing cognitive processes. To this end, we applied
+dimensionality reduction algorithms and pattern classifiers to functional
+neuroimaging data collected as participants listened to a story, temporally
+scrambled versions of the story, or underwent a resting state scanning session.
+These experimental conditions were intended to require different depths of
+processing and inspire different levels of cognitive engagement. We considered
+two primary aspects of the data. First, we treated the maximum achievable
+decoding accuracy across participants as an indicator of the
+``informativeness'' of the recorded patterns. Second, we treated the number of
+features (components) required to achieve a threshold decoding accuracy as a
 proxy for the ``compressibility'' of the neural patterns (where fewer
 components indicate greater compression). Overall, we found that the peak
-decoding accuracy (achievable without restricting the numbers of features)
-was highest in the intact (unscrambled) story listening condition.  However, the
-number of features required to achieve comparable classification accuracy
-was also lowest in the intact story listening condition.  Taken together, our work
+decoding accuracy (achievable without restricting the numbers of features) was
+highest in the intact (unscrambled) story listening condition. However, the
+number of features required to achieve comparable classification accuracy was
+also lowest in the intact story listening condition. Taken together, our work
 suggests that our brain networks flexibly reconfigure according to ongoing task
@@
 available to participants increases (i.e., over time in the cognitively rich
-``intact'' and ``paragraph'' conditions), it makes sense that this might
+intact and paragraph conditions), it makes sense that this might
 constrain neural responses to a greater extent. While this pattern may not
@@
 wander,'' or as they experience mental fatigue (i.e., over time in the less
-cognitively rich ``word'' and ``rest'' conditions), we suggest that this might
+cognitively rich word and rest conditions), we suggest that this might
 lead to greater variability in neural responses across people, resulting in
@@
 
-We also wondered how informativeness and compressibility in the different
-experimental conditions might vary across brain networks. We used a network
-parcellation identified by \cite{YeoEtal11} to segment the brain into seven
-distinct networks. The networks can be sorted (roughly) in order from
+If the informativeness and compressibility of brain activity patterns vary over
+time, do these properties might also vary across brain networks? We used a
+network parcellation identified by \cite{YeoEtal11} to segment the brain into
+seven distinct networks. The networks can be sorted (roughly) in order from
 lower-level to higher-level cortex as follows (Figs.~\ref{fig:networks}A--C):
@@
 networks to higher-order networks, we found that decoding accuracy tended to
-increase, particularly in the higher-level experimental conditions
-(Fig.~\ref{fig:networks}D, E). This suggests that higher-order networks may
-carry more content-relevant or stimulus-driven ``information.'' We found no
-clear trends in the proportions of components required to achieve 5\% decoding
-accuracy across networks or conditions (Fig.~\ref{fig:networks}F).
+increase in the higher-level experimental conditions and decrease (slightly) in
+the lower-level experimental conditions (Fig.~\ref{fig:networks}D, E;
+Spearman's rank correlation between decoding accuracy and network order:
+intact: $\rho = 0.362, p < 0.001$; paragraph: $\rho = 0.441, p < 0.001$; word:
+$\rho = -0.102, p = 0.007$; rest: $\rho = -0.354, p < 0.001$). This suggests
+that higher-order networks may carry more content-relevant or stimulus-driven
+``information.'' We found no clear trends in the proportions of components
+required to achieve 5\% decoding accuracy across networks or conditions
+(Fig.~\ref{fig:networks}F).
 
-\begin{figure}[tp]
+\begin{sidewaysfigure}[tp]
   \centering
@@
 
-\end{figure}
+\end{sidewaysfigure}
 
-In addition to examining different networks in isolation, we wondered about the
-general structure of the full-brain (i.e., potentially multi-network) activity
-patterns reflected by different principal components across different
-experimental conditions. As shown in Figure~\ref{fig:neurosynth-pca}, we used
+Whereas the above analyses examinine different networks in isolation, how does
+full-brain (i.e., potentially multi-network) activity patterns reflected by
+different principal components vary across different experimental conditions?
+As shown in Figure~\ref{fig:neurosynth-pca}, we used
 Neurosynth~\citep{RubiEtal17} to identify, for each component, the associations
@@
 principal components across all of the experimental conditions tended to weight
-most heavily on themes related to cognitive control, memory,
-language processing, attention, and perception.  Other components appeared to
-vary more across conditions.
+most heavily on themes related to cognitive control, memory, language
+processing, attention, and perception. Other components appeared to vary more
+across conditions.
 
@@
 heavily by top components from the intact and paragraph conditions than
-lower-level functions. The top components from the word condition showed the
-opposite tendency, whereby \textit{lower}-level functions tended to be weighted
-on more heavily than higher-level functions. The components from the rest
-condition showed almost no differences in the weights associated with
-high-level versus low-level functions.
+lower-level functions (intact vs. word: $t(198) = 11.059, p < 0.001$; intact
+vs. rest: $t(198) = 3.699, p < 0.001$; paragraph vs. word: $t(198) = 13.504, p
+< 0.001$; paragraph vs. rest: $t(198) = 4.812, p < 0.001$; also see
+\textit{Ranking cognitive processes}). The top components from the word
+condition showed the opposite tendency, whereby \textit{lower}-level functions
+tended to be weighted on more heavily than higher-level functions (word vs.
+rest: $t(198) = -7.315, p < 0.001$). The weighting trends for the intact and
+paragraph conditions were not reliably different ($t(198) = -0.479, p =
+0.633$). The components from the rest condition showed almost no differences in
+the weights associated with high-level versus low-level functions (rest vs. 0:
+$t(99) = 1.836, p = 0.081$). These findings suggest that when participants were
+engaged more strongly (in the more engaging intact and paragraph conditions),
+their dominant neural patterns reflected higher-level cognitive functions. In
+contrast, when participants were engaged less strongly (in the less engaging
+word and rest conditions), their dominant neural patterns reflected lower-level
+cognitive functions. Although they were highly statistically reliable, it is
+also important to note that these latter effects are also relatively small
+(e.g., the slopes for \textit{all} of the experimental conditions are
+numerically close to zero; Fig.~\ref{fig:neurosynth-summary}E). We suggest that
+this phenomenon may merit further investigation in future work.
 
@@
 colored dots). \textbf{D. Associations between per-condition components, by
-component.} Each sub-panel's network plot summarizes the pattern of correlations
-between the $n$\textsuperscript{th} top-weighted principal components
-(sub-panel) for each experimental condition (gray-outlined dots). The line
-thicknesses are proportional to the correlation values (correlation
+component.} Each sub-panel's network plot summarizes the pattern of
+correlations between the $n$\textsuperscript{th} top-weighted principal
+components (sub-panel) for each experimental condition (gray-outlined dots).
+The line thicknesses are proportional to the correlation values (correlation
 coefficients are noted in the heat maps in Panel C). \textbf{E. Change in
@@
 between the component and topic maps associated with cognitive processes at the
-given rank (see \textit{Ranking cognitive processes}).  Also see
+given rank (see \textit{Ranking cognitive processes}) Error bars denote
+bootstrap-estimated 95\% confidence intervals. Also see
 Fig.~\neurosynthFull~for additional information.}
@@
 
-We manually identified 11 cognitive labels
-spanning the set of 80 Neurosynth-derived topics: cognitive control, language
-processing, memory, emotion, social cognition, spatial cognition, attention,
-reward, sensory perception, motor control, and resting state. We then used
-ChatGPT to automatically ``rank'' the processes from high-level to low-level
-using the following prompt: ``Please rank these cognitive processes from
-highest-level to lowest-level, where higher values indicate higher-order or
-higher-level processes. Return the result as a csv file with a header row and
-two columns: `Cognitive label' and `Rank'. Here are the processes: cognitive
-control, language processing, memory, emotion, social cognition, spatial
-cognition, attention, reward, sensory perception, motor control, resting
-state''.  Table~\topicTags~displays the output.  We used these labels in
-the analysis presented in Figure~\ref{fig:neurosynth-summary}E to help summarize
-difference in topic weightings across experimental conditions.
+We manually identified 11 cognitive labels spanning the set of 80
+Neurosynth-derived topics: cognitive control, language processing, memory,
+emotion, social cognition, spatial cognition, attention, reward, sensory
+perception, motor control, and resting state. We then used ChatGPT to
+automatically ``rank'' the processes from high-level to low-level using the
+following prompt: ``Please rank these cognitive processes from highest-level to
+lowest-level, where higher values indicate higher-order or higher-level
+processes. Return the result as a csv file with a header row and two columns:
+`Cognitive label' and `Rank'. Here are the processes: cognitive control,
+language processing, memory, emotion, social cognition, spatial cognition,
+attention, reward, sensory perception, motor control, resting state''.
+Table~\topicTags~displays the output.
+
+In the analysis presented in Figure~\ref{fig:neurosynth-summary}E, we summarize
+difference in topic weightings across experimental conditions. In particular,
+we sought to characterize how the dominant neural patterns evoked by each
+experimental condition weighted on different cognitive functions. For each of
+the top five principal components from each experimental condition
+(Fig.~\ref{fig:neurosynth-pca}), we computed the average weightings for each of
+the 11 manually identified (and ChatGPT-ranked) cognitive labels described
+above (Tab.~\topicTags). We then fit a line separately for each experiment
+condition ($x$-values: cognitive rank; $y$-values: weights). In carrying out
+this analysis, we used a bootstrap procedure to estimate the variability in the
+slopes of the regression lines, whereby we repeated this process for each of $n
+= 100$ iterations, each time resampling (with replacement) the set of observed
+ranks and weights. This procedure yielded distributions of 100 estimated slopes
+for each experimental condition. We used these distributions to compare the
+slopes across experimental conditions and to estimate 95\% confidence intervals.
 

### 2023-12-21 code/notebooks/helpers.py

@@ -100,10 +100,6 @@ def nii2cmu(nifti_file, mask_file=None):
     vox_coords = fullfact(img.shape[0:3])[vmask, ::-1]-1
     
     R = np.array(np.dot(vox_coords, S[0:3, 0:3])) + S[:3, 3]
-
-    # center on the MNI152 brain (hard code this in)
-    #mni_center = np.array([0.55741881, -21.52140703, 9.83783098])
-    #R = R - R.mean(axis=0) + mni_center
     
     return {'Y': Y, 'R': R}
 
@@ -208,12 +204,18 @@ def cross_validation(data, n_iter=10, fname=None, max_components=700):
     return results
 
 
-def ridge_plot(x, column='Number of components', fname=None, xlim=[-99, 700], hue='Condition', palette=[condition_colors[c] for c in conditions]):
+def ridge_plot(x, column='Number of components', fname=None, xlim=[-99, 700], hue='Condition', palette=[condition_colors[c] for c in conditions], scale_start=0.25, scale_height=0.1):
 
     sns.set_theme(style="white", rc={"axes.facecolor": (0, 0, 0, 0)})
     g = sns.FacetGrid(x, row=hue, hue=hue, palette=palette, height=1, aspect=6)
     g.map(sns.kdeplot, column, bw_adjust=1, clip_on=True, fill=True, alpha=1, common_norm=True, linewidth=1.5)
     g.refline(y=0, linewidth=1.5, linestyle='-', color=None, clip_on=False)
+
+    # plot a scale bar in the upper right
+    if scale_height is not None:        
+        # compute the x position of the scale bar -- 98% of the way to the right
+        x = xlim[0] + 0.98 * (xlim[1] - xlim[0])    
+        g.axes[0][0].plot([x, x], [scale_start, scale_start + scale_height], color='k', linewidth=1.5)
 
     def label(x, color, label):
         ax = plt.gca()
@@ -231,7 +233,7 @@ def ridge_plot(x, column='Number of components', fname=None, xlim=[-99, 700], hu
 
     ax = plt.gca()
     ax.set_xlim(xlim[0], xlim[1])
-    ax.set_xlabel(column, fontsize=12)
+    ax.set_xlabel(column, fontsize=12)    
 
     if fname is not None:
         g.savefig(os.path.join(figdir, fname + '.pdf'), bbox_inches='tight')
@@ -254,7 +256,7 @@ def get_data():
     return data
 
 
[…]

### 2023-12-22 code/notebooks/helpers.py
udpated figures-- added scale bars to ridge plots

@@ -9,6 +9,7 @@ from matplotlib import pyplot as plt
 from sklearn.decomposition import IncrementalPCA as PCA
 from tqdm import tqdm
 from scipy.spatial.distance import cdist
+import statsmodels.api as sm
 
 import os
 import warnings
@@ -205,10 +206,30 @@ def cross_validation(data, n_iter=10, fname=None, max_components=700):
 
 
 def ridge_plot(x, column='Number of components', fname=None, xlim=[-99, 700], hue='Condition', palette=[condition_colors[c] for c in conditions], scale_start=0.25, scale_height=0.1):
+    def pdf_plot(x, ax=None, xlim=[0.0, 1.0], resolution=1000, **kwargs):
+        if ax is None:
+            ax = plt.gca()
+        
+        if 'color' in kwargs:
+            color = kwargs['color']
+            kwargs.pop('color')
+        else:
+            color = 'k'
+        
+        density = sm.nonparametric.KDEUnivariate(x)
+        density.fit()
+
+        xs = np.linspace(xlim[0], xlim[1], resolution)
+        ys = density.evaluate(xs)
+        ys = ys / np.sum(ys)
+
+        ax.fill(xs, ys, color=color, **kwargs)
+        return ax
 
     sns.set_theme(style="white", rc={"axes.facecolor": (0, 0, 0, 0)})
     g = sns.FacetGrid(x, row=hue, hue=hue, palette=palette, height=1, aspect=6)
-    g.map(sns.kdeplot, column, bw_adjust=1, clip_on=True, fill=True, alpha=1, common_norm=True, linewidth=1.5)
+    # g.map(sns.kdeplot, column, bw_adjust=1, clip_on=True, fill=True, alpha=1, common_norm=True, linewidth=1.5)  # REPLACE WITH A NEW FUNCTION THAT NORMALIZES AREA TO 1
+    g.map(pdf_plot, column, xlim=xlim, linewidth=1.5)
     g.refline(y=0, linewidth=1.5, linestyle='-', color=None, clip_on=False)
 
     # plot a scale bar in the upper right

### 2023-12-22 paper/main.tex
added a significance statement, color bars to figs 6 and S5

@@
 reduction, neuroimaging}
+
+\bigskip 
+\noindent
+
+\textbf{Significance Statement:} How our brains respond to ongoing experiences
+depends on what we are doing and thinking about, how engaged we are, and our
+past experiences, among other factors. To gain insights into how brain activity
+patterns reflect ongoing cognitive processes, we examined two fundamental
+aspects of brain activity under different cognitive circumstances:
+informativeness and compressibility. Informativeness refers to the extent to
+which brain patterns are both temporally specific and consistent across
+different people. Compressibility refers to the extent to how robust the
+informativeness of brain patterns is to dimensionality reduction. We found that
+brain activity patterns evoked by higher-level cognitive tasks are both more
+informative \textit{and} more compressible than activity evoked by lower-level
+tasks. Our findings suggest that our brains flexibly reconfigure themselves to
+optimize different aspects of how they function according to ongoing cognitive
+demands.
 
@@
 intervals. \textbf{C. Estimating ``compressibility'' for each condition.} The
-density plots display the numbers of components required to reach at least 5\%
-(corrected) decoding accuracy, across all cross validation runs. For the
-``rest'' condition, where decoding accuracy never reached 5\% accuracy, we
-display the distribution of the numbers of components required to achieve the
-peak decoding accuracy in each fold. (This distribution is nearly flat, as
-also illustrated in Panel E.) \textbf{D. Estimating ``informativeness'' for each
-condition.} The density plots display the peak (corrected) decoding accuracies
-achieved in each condition, across all cross validation runs. \textbf{E.
-Informativeness versus compressibility.} Each dot displays the minimum number
-of components required to achieve at least 5\% corrected decoding accuracy
-($x$-coordinate; for the rest condition the numbers of components are estimated
-using the peak decoding accuracies as in Panel C) and the peak (corrected)
-decoding accuracy ($y$-coordinate). The smaller dots correspond to individual
-cross validation runs and the larger dots display the averages across runs. The
-inset displays a zoomed-in view of the indicated region in the main panel.}
+probability density plots display the numbers of components required to reach
+at least 5\% (corrected) decoding accuracy, across all cross validation runs.
+For the ``rest'' condition, where decoding accuracy never reached 5\% accuracy,
+we display the distribution of the numbers of components required to achieve
+the peak decoding accuracy in each fold. (This distribution is nearly flat, as
+also illustrated in Panel E.) The scale bar denotes a height of 0.01.
+\textbf{D. Estimating ``informativeness'' for each condition.} The probability
+density plots display the peak (corrected) decoding accuracies achieved in each
+condition, across all cross validation runs. The scale bar denotes a height of
+0.01. \textbf{E. Informativeness versus compressibility.} Each dot displays the
+minimum number of components required to achieve at least 5\% corrected
+decoding accuracy ($x$-coordinate; for the rest condition the numbers of
+components are estimated using the peak decoding accuracies as in Panel C) and
+the peak (corrected) decoding accuracy ($y$-coordinate). The smaller dots
+correspond to individual cross validation runs and the larger dots display the
+averages across runs. The inset displays a zoomed-in view of the indicated
+region in the main panel.}
 

### 2023-12-22 paper/main.tex
starting to work on network shuffling analysis requested by R2 [WIP]

@@
 that unreliability could be compensated for by other regions' activity patterns
-that were predictive of the unreliable region.
+that were predictive of the unreliable region. Whereas compressible brain
+patterns a robust to spatial signal corruption, high versus low informativeness
+reflects a similar (though dissociable; e.g.,
+Fig.~\ref{fig:information-compression}) tradeoff between expressiveness and
+robustness of \textit{temporal} patterns. Highly informative brain patterns (by
+our measure; i.e., patterns that yield greater temporal decoding accuracy) are
+expressive about ongoing experiences or cognitive states, since each moment's
+pattern is reliably distinguishable from other moments' patterns. However, when
+each moment's pattern is unique, brain activity becomes less robust to temporal
+signal corruption. Our finding that brain activity patterns becomes more
+informative (i.e., less robust to temporal signal corruption) and compressible
+(i.e., more robust to spatial signal corruption) when cognitive engagement is
+higher suggests that our brain may optimize its activity patterns to prioritize
+either temporal or spatial robustness, according to task demands.
 
@@
 
-Different brain networks also varied in how informative and compressible their
-activity patterns were across experimental conditions (e.g.,
-Fig.~\ref{fig:networks}). This might follow from evolutionary optimizations
-that reflect the relevant constraints or demands placed on those networks. One
-possibility is that cortex is organized in a hierarchy of networks ``concerned
-with'' or selective to different levels of processing or function. To the
-extent that different levels of processing (e.g., low-level sensory processing
-versus ``deeper'' higher-level processing) reflect different stimulus
-timescales~\citep[e.g.,][]{Mann20}, the network differences we observed might
-also relate to the timescales at which each network is maximally
-sensitive~\citep{RegeEtal18, BaldEtal17,LernEtal11, HassEtal08}.
+We found that different brain networks varied in how informative and
+compressible their activity patterns were across experimental conditions (e.g.,
+Fig.~\ref{fig:networks}). This might
+follow from evolutionary optimizations that reflect the relevant constraints or
+demands placed on those networks. One possibility is that cortex is organized
+in a hierarchy of networks ``concerned with'' or selective to different levels
+of processing or function. To the extent that different levels of processing
+(e.g., low-level sensory processing versus ``deeper'' higher-level processing)
+reflect different stimulus timescales~\citep[e.g.,][]{Mann20}, the network
+differences we observed might also relate to the timescales at which each
+network is maximally sensitive~\citep{RegeEtal18, BaldEtal17,LernEtal11,
+HassEtal08}.
+
+Our reverse inference analyses (Figs.~\ref{fig:neurosynth-pca},
+\ref{fig:neurosynth-summary}) also provide some insights into how neural
+activity patterns change with cognitive engagement or task demands. Prior work
+has shown that the components and network ``parcels'' identified through
+covarying activity patterns can be highly similar even across different
+tasks~\citep[including ``rest,'' e.g.,][]{SmitEtal09, LairEtal11}. We
+replicated this basic finding in that the first principal components from all
+four experimental conditions were strikingly similar (e.g., see the leftmost
+columns of Figs.~\ref{fig:neurosynth-pca}A and~\ref{fig:neurosynth-summary}C,
+D). We also found some small, though statistically reliable, systematic changes
+in the weights associated with different cognitive functions across conditions
+(Fig.~\ref{fig:neurosynth-summary}E). This result provided an additional way of
+characterizing network-level differences across conditions
+(Fig.~\ref{fig:networks}E).  Taken together, these findings suggest that although
+similar networks may be involved in different tasks, the ways in which those
+networks are engaged may vary systematically with task demands.
 
@@
 informativeness and compressibility, that appear to change systematically with
-task demands and across brain networks. Our work helps to clarify how the
-``neural code'' might be structured, and how the code might vary across tasks
-and brain areas.
+task demands and across brain networks. We speculate that these changes may
+reflect ongoing tradeoffs between how robust to signal corruption versus how
+expressive about ongoing cognitive states our brains' activity patterns are.
+Our work also provides a new framework for evaluating these tradeoffs in other
+datasets, or in future studies.
 
@@
 
+We recognize that ChatGPT is not omniscient, nor should it be treated as an
+expert cognitive neuroscientist. We therefore reviewed ChatGPT's responses
+carefully by hand to verify that they seemed reasonable to us. Whereas prior
+work has often constructed such rankings by hand, we see our use of ChatGPT in
+this case as a small additional ``sanity check'' on our rankings that helped us
+to be slightly more objective than if we had simply created the rankings
+ourselves manually.
+
 In the analysis presented in Figure~\ref{fig:neurosynth-summary}E, we summarize

### 2023-12-22 paper/main.tex
working on addressing R2 and R3 concerns.
i've also added some notes on the network shuffling/permuation analysis, which still needs to be implemented and integrated into the paper... [WIP]

@@
 comprised functional magnetic resonance imaging (fMRI) data collected as
-participants listened to an audio recording of a 10-minute story, temporally
+participants listened to an audio recording of a 7-minute story, temporally
 scrambled recordings of the story, or underwent a resting state
@@
 experimental condition, participants listened to an audio recording of a
-10-minute Moth Radio Hour story, \textit{Pie Man}, by Jim O'Grady. In the
+7-minute Moth Radio Hour story, \textit{Pie Man}, by Jim O'Grady. In the
 \textit{paragraph}-scrambled experimental condition, participants listened to a
@@
 required to achieve 5\% decoding accuracy across networks or conditions
-(Fig.~\ref{fig:networks}F).
+(Fig.~\ref{fig:networks}F). We note that the limbic network we considered here
+often overlaps with low (imaging) signal regions, and therefore it may be
+difficult to draw strong conclusions about this network's informativeness or
+compressibility.
 
@@
 native English speakers with normal hearing and provided written informed
-consent.
+consent. We have excerpted the relevant portions of the dataset documention
+here to provide information about the scanning parameters and preprocessing
+steps used to generate the data we analyzed (the original descriptions may be
+found at the above link):
+
+\begin{quote}
+Subjects were scanned in a 3T full-body MRI scanner (Skyra; Siemens) with a
+sixteen-channel head coil. For functional scans, images were acquired using a
+T2* weighted echo planer imaging (EPI) pulse sequence [repetition time (TR),
+1500 ms; echo time (TE), 28 ms; flip angle, $64^\circ$], each volume comprising
+27 slices of 4 mm thickness with 0 mm gap; slice acquisition order was
+interleaved. In-plane resolution was $3 \times 3$ mm\textsuperscript{2} [field
+of view (FOV), 192 $\times$ 192 mm\textsuperscript{2}]. Anatomical images were
+acquired using a T1-weighted magnetization-prepared rapid-acquisition gradient
+echo (MPRAGE) pulse sequence (TR, 2300 ms; TE, 3.08 ms; flip angle 9$^\circ$;
+0.89 mm\textsuperscript{3} resolution; FOV, 256 mm\textsuperscript{2}). To
+minimize head movement, subjects' heads were stabilized with foam padding.
+Stimuli were presented using the Psychophysics toolbox~\citep{Brai97,Pell97}.
+Subjects were provided with an MRI compatible in-ear mono earbuds (Sensimetrics
+Model S14), which provided the same audio input to each ear. MRI-safe passive
+noise-canceling headphones were placed over the earbuds, for noise removal and
+safety.
+
+Functional data were preprocessed and analyzed using FSL
+(www.fmrib.ox.ac.uk/fsl), including correction for head motion and
+slice-acquisition time, spatial smoothing (6 mm FWHM Gaussian kernel), and
+high-pass temporal filtering (140 s period). Preprocessed data were aligned to
+coplanar and high-resolution anatomicals and the standard MNI152 brain, and
+interpolated to 3-mm isotropic voxels.
+\end{quote}
+
+The intact and word conditions each comprised 300 TRs (7.5 minutes) per participant.
+The paragraph condition comprised 272 TRs (6.8 minutes) per participant. The rest
+condition comprised 400 TRs (10 minutes) per participant.
 
@@
 downloaded at \href{https://brainiak.org/}{brainiak.org}.
+
+We also considered alternative approaches to obtaining compact representations
+of the neuroimaging data, including network
+parcellations~\citep[e.g.,][]{SchaEtal18, GordEtal16}. Whereas network
+parcellations are typically derived from large resting state datasets, HTFA may
+be applied to much smaller datasets. In our prior work, we showed that HTFA
+applied to the same dataset used here can explain full-brain activity to within
+a maximum of 0.25 standard deviations of each voxel's observed activity in the
+original dataset, taken across all voxels, images, and participants, using the
+700-node representation we also employed here~\citep{MannEtal18}. Some of the
+explanatory power of HTFA comes from the fact that each node's explanatory
+power falls off smoothly with distance to its center. Intuitively, the result
+is a representation that looks like a lightly spatially smoothed version of the
+original data, but where the degree of smoothing varies across the brain
+according to how spatially autocorrelated the local activity patterns are.
 

### 2023-12-23 paper/main.tex
finished addressing "writing-based" comments from reviewers.  still to do: control analysis with network labels...i'm trying to figure out how to do this in a way that will actually run in a reasonable amount of time...

@@
 less compressible brain activity patterns~\citep{MackEtal20, OwenEtal21}. These
-findings complement other work suggesting that functional connectivity
-(correlation) patterns are task-dependent~\citep{FinnEtal17, OwenEtal20,
-ColeEtal14}, although see~\cite{GratEtal18}. Higher-order cognitive processing
-of a common stimulus also appears to drive more stereotyped task-related
-activity and functional connectivity across individuals~\citep{HassEtal08,
-LernEtal11, SimoChan20, SimoEtal16}.
+patterns hold even when the stimulus itself is held
+constant~\citep{MackEtal20}. These findings complement other work suggesting
+that functional connectivity (correlation) patterns are
+task-dependent~\citep{FinnEtal17, OwenEtal20, ColeEtal14}, although
+see~\cite{GratEtal18}. Higher-order cognitive processing of a common stimulus
+also appears to drive more stereotyped task-related activity and functional
+connectivity across individuals~\citep{HassEtal08, LernEtal11, SimoChan20,
+SimoEtal16}.
 
@@
 complexity or with the level (or ``depth'') of cognitive processing. A second
-possibility is that the compressibility of brain activity patterns increases
+possibility is that the compressibility of brain activity patterns decreases
 during more complex or higher-level cognitive processing. If so, then
-individual features of brain recordings, or compressed representations of brain
-recordings, should carry more information during complex or high-level (versus
-simple or low-level) cognitive tasks.
+individual features of brain recordings should carry more information (over and
+above the information carried by other features) during complex or high-level
+(versus simple or low-level) cognitive tasks. The tradeoffs between these two
+aspects of brain activity may also vary across brain regions or networks, for
+example according to each region's functional role.
 
@@
 
+One potential limitation of our work concerns how our measure of
+informativeness might generalize across tasks or cognitive representations and
+processes. Our use of across-participant temporal decoding accuracy as a proxy
+for informativeness is motivated in part by prior work that introduced
+across-participant similarity (in time-varying response to a stimulus) as a
+means of identifying stimulus-driven brain activity
+patterns~\citep{SimoEtal16}. Intuitively, only activity patterns are are driven
+by the stimulus would be expected to synchronize (i.e., be time-locked to the
+stimulus) across participants. This approach implicitly removes idiosyncratic
+responses (e.g., neural patterns that are \textit{not} similar across people).
+However, there are also some published examples, including in our own prior
+work, that indicate that some types of stimulus-evoked activity will be missed
+by across-participant comparisons. For example we have reported how brain
+regions like the ventromedial prefrontal cortex (vmPFC) show stimulus-driven
+responses that are, for the most part, not similar across people. In that paper
+(and drawing on other work), we suggest that the vmPFC seems to represent or
+support highly idiosyncratic internal states, like affective responses.
+Although we would consider the vmPFC to be a ``high-level'' region (e.g., we
+consider affect to be a relatively high-level aspect of cognition), the measure
+of informativeness that we used in our current study would identify regions
+like the vmPFC as having \textit{low} informativeness. This is because
+across-participant decoding accuracy (our proxy measure for informativeness)
+will only be high for representations or responses that are common across
+people.
+
+Relatedly, even in the experimental conditions we describe as ``less
+cognitively engaging,'' we do not think it likely that high-level thought or
+cognitive processing is absent. Rather, we suggest that these high-level
+representations tend to be more idiosyncratic when the stimulus is less
+engaging, and therefore less constraining on people's thoughts. Nonetheless,
+even during highly engaging tasks, people may engage in idiosyncratic
+stimulus-driven processes. For example, people might retrieve personal
+information as they listened to the story. Those retrievals could happen at
+different times for different people according to each individual's prior
+experiences, and even when those sorts of retrievals happen to be temporally
+synchronized across people, the specific memories or information being
+retrieved might still be idiosyncratic. Our measure of informativeness is
+insensitive to these processes. Further, even in response to an identical
+stimulus, task instructions or participants' interal goals could change the
+relationship between compressibility and informativeness. Some work has shown
+that the ``dimensionality'' of neural representations can change systematically
+with task complexity, even in response to an identical
+stimulus~\citep{MackEtal20}. Taken together, we expect that the way we have
+defined informativeness in this paper, and the specific dataset we examined,
+are likely to have influenced our findings. While we see our approach as a
+reasonable first step, we also suggest that future work should explore
+alternative measures of informativeness and compressibility, and should examine
+how these measures vary across different tasks and datasets.
+
 In the information theory sense~\citep{Shan48}, when a signal is transmitted

### 2023-12-24 paper/main.tex
backing up network shuffling (permutation) analyses

@@
 difficult to draw strong conclusions about this network's informativeness or
-compressibility.
+compressibility. We also considered the possibility that the correlations with
+network order might be influenced by the numbers of nodes in each network. We
+designed a permutation-based procedure to address this possibility, whereby we
+repeated the above analyses using shuffled network labels (see \textit{Network
+permutation tests}).  The correlations between decoding accuracy and network
+order... <INSERT RESULTS>
 

### 2023-12-24 paper/main.tex
added network permutation analyses + t-tests, report results in paper

@@
 permutation tests}). The correlations between decoding accuracy and network
-order... <INSERT RESULTS>
+order were reliably more positive than the shuffled correlations for the intact
+($t(1998) = 276.431, p < 0.001$) and paragraph ($t(1998) = 330.334, p < 0.001$)
+conditions, and reliably more negative for the word ($t(1998) = -16.386, p <
+0.001$) and rest ($t(1998) = -318.631, p < 0.001$) conditions. These results
+suggest that the correlations between decoding accuracy and network order were
+not driven solely by the numbers of nodes in each network.
 
@@
 
+\subsection*{Network permutation tests}
+
+In our analyses of how informativeness varied across brain networks, we
+considered the possibility that the correlations with network order might be
+influenced by the numbers of nodes in each network. We designed a
+permutation-based procedure to address this possibility, whereby we repeated
+the above analyses using shuffled network labels. Specifically, for each of $n_1 =
+10$ iterations, we randomly shuffled (without replacement) the network labels of
+the HTFA nodes, and then we re-ran our entire decoding analysis pipeline,
+including applying PCA with $3...m$ features for each condition (where $m$ is the
+number of nodes in the given network), and then running 100 cross-validation
+runs of the decoding procedure for each condition and number of components.
+This resulted in 10 sets of shuffled data where each network had the same
+numbers of nodes, but where the decoding results no longer maintained the
+fidelity of each individual network.
+
+We sampled the original and shuffled datasets (with replacement) to create $n_2
+= 1000$ bootstrap samples. For each bootstrap sample, we computed the
+correlations between the decoding accuracies and network order for each
+condition and number of components. This yielded a distribution of $n_2$
+correlation values for each condition, for both the original and shuffled
+datasets. We then compared the distributions of Spearman's $\rho$ values for
+the original and shuffled datasets using a two-sided independent samples
+Welch's $t$-tests.
+
 \subsection*{Principal components analysis (PCA)}

### 2023-12-26 paper/old.tex
submitted version (PNAS R2)

--- 2023-12-23 paper/main.tex
@@
 \usepackage{soul}
-\usepackage{rotating}
 \linenumbers
@@
 
-Brain activity patterns are highly flexible and often complex, but also highly
-structured. Here we examined how fundamental properties of brain activity
-patterns relate to ongoing cognitive processes. To this end, we applied
-dimensionality reduction algorithms and pattern classifiers to functional
-neuroimaging data collected as participants listened to a story, temporally
-scrambled versions of the story, or underwent a resting state scanning session.
-These experimental conditions were intended to require different depths of
-processing and inspire different levels of cognitive engagement. We considered
-two primary aspects of the data. First, we treated the maximum achievable
-decoding accuracy across participants as an indicator of the
-``informativeness'' of the recorded patterns. Second, we treated the number of
-features (components) required to achieve a threshold decoding accuracy as a
+We applied dimensionality reduction algorithms and pattern classifiers to
+functional neuroimaging data collected as participants listened to a story,
+temporally scrambled versions of the story, or underwent a resting state
+scanning session. These experimental conditions were intended to require
+different depths of processing and inspire different levels of cognitive
+engagement. We considered two primary aspects of the data. First, we treated
+the maximum achievable decoding accuracy across participants as an indicator of
+the ``informativeness'' of the recorded patterns. Second, we treated the number
+of features (components) required to achieve a threshold decoding accuracy as a
 proxy for the ``compressibility'' of the neural patterns (where fewer
 components indicate greater compression). Overall, we found that the peak
-decoding accuracy (achievable without restricting the numbers of features) was
-highest in the intact (unscrambled) story listening condition. However, the
-number of features required to achieve comparable classification accuracy was
-also lowest in the intact story listening condition. Taken together, our work
+decoding accuracy (achievable without restricting the numbers of features)
+was highest in the intact (unscrambled) story listening condition.  However, the
+number of features required to achieve comparable classification accuracy
+was also lowest in the intact story listening condition.  Taken together, our work
 suggests that our brain networks flexibly reconfigure according to ongoing task
@@
 reduction, neuroimaging}
-
-\bigskip 
-\noindent
-
-\textbf{Significance Statement:} How our brains respond to ongoing experiences
-depends on what we are doing and thinking about, how engaged we are, and our
-past experiences, among other factors. To gain insights into how brain activity
-patterns reflect ongoing cognitive processes, we examined two fundamental
-aspects of brain activity under different cognitive circumstances:
-informativeness and compressibility. Informativeness refers to the extent to
-which brain patterns are both temporally specific and consistent across
-different people. Compressibility refers to the extent to how robust the
-informativeness of brain patterns is to dimensionality reduction. We found that
-brain activity patterns evoked by higher-level cognitive tasks are both more
-informative \textit{and} more compressible than activity evoked by lower-level
-tasks. Our findings suggest that our brains flexibly reconfigure themselves to
-optimize different aspects of how they function according to ongoing cognitive
-demands.
 
@@
   \caption{\textbf{Information content and compressibility.} \textbf{A.
-  Variance explained for two images.} Consider two example images: a photograph
-  and simple drawing. The images have the same resolutions, but their content
-  is very different. The photograph is more information rich: when we compare
-  the original photograph and drawing (leftmost column), we can see that the
-  photograph captures much more detail. We can also apply principal components
-  analysis to the images, treating the rows of the images as ``observations.''
-  Across columns, we identified the numbers of components required to explain
-  100\%, 95\%, 75\%, or 50\% of the cumulative variance in each image (the
-  100\% columns denote the original images). The numbers of components are
-  indicated in parentheses, and the resulting ``compressed'' images are
-  displayed. This analysis reveals that the drawing is more compressible: just
-  16 components can explain 95\% of the variance in teh drawing, whereas 50
-  components are required to explain 95\% of the variance in the photograph.
-  \textbf{B. Representing two images with different numbers of components.}
-  Using the same principal component decompositions as in Panel A, we computed
-  the cumulative proportion of variance explained with 380 (original images),
-  20, 4, or 2 components. This analysis provides another way of characterizing
-  the compressibility of each image by showing that the same number of
-  components can explain more variance in the drawing than in the photograph.
-  \textbf{C. Template data from four synthetic datasets.} We constructed four
-  synthetic datasets, each comprising 25 features (columns) observed across 100
-  samples (rows) from each of 10 simulated ``participants.'' The datasets were
-  constructed to contain different levels of informativeness and
-  compressibility (see \textit{Synthetic data}). \textbf{D. Decoding accuracy
-  by number of components.} For each synthetic dataset, we trained
-  across-participant classifiers to decode timepoint labels.  Each panel displays the
-  decoding accuracy as a function of the number of components used to represent
-  the data. Error ribbons denote bootstrap-estimated 95\% confidence intervals.}
+  Variance explained for two images.} We applied principal components analysis
+  to a photograph and drawing, treating the rows of the images as
+  ``observations.'' Across columns, we identified the numbers of components
+  required to explain 100\%, 95\%, 75\%, or 50\% of the cumulative variance in
+  each image (the 100\% columns denote the original images). The numbers of
+  components are indicated in parentheses, and the resulting ``compressed''
+  images are displayed. \textbf{B. Representing two images with different
+  numbers of components.} Using the same principal component decompositions as
+  in Panel A, we computed the cumulative proportion of variance explained with
+  380 (original images), 20, 4, or 2 components. \textbf{C. Cumulative variance
+  explained versus number of components.} For the images displayed in Panels A
+  and B, we plot the cumulative proportion of variance explained as a function
+  of the number of components used to represent each image. \textbf{D.
+  Information rate and compressibility.} Across multiple images, the
+  information rate (i.e., the amount of information contained in each image;
+  horizontal axis) is high when each individual pixel provides information that
+  cannot be inferred from other pixels. High-information rate images tend to be
+  high-resolution, and low-information rate images tend to be low-resolution.
+  Compressibility is related to the difference between the information required
+  to specify the original versus compressed images (vertical axis). Highly
+  compressible images often contain predictable structure (redundancies) that
+  can be leveraged to represent the images much more efficiently than in their
+  original feature spaces.}
+
 \label{fig:information-compression} 
@@
 important aspects of the original signal. In general, information content and
-compressibility are often related but are also dissociable
+compressibility are related but are partially dissociable
 (Fig.~\ref{fig:information-compression}). If a given signal (e.g., a
 representation of brain activity patterns) contains more information about
-ongoing cognitive processes, then the peak decoding accuracy should be high. In
-the simulations shown in Figure~\ref{fig:information-compression}C we construct
-synthetic datasets that have high or low levels of informativeness by varying
-temporal autocorrelations in the data (see \textit{Synthetic data}).
-If a signal is compressible, then a low-dimensional embedding of the signal
-will be similarly informative as the original signal.  In the simulations
-shown in Figure~\ref{fig:information-compression}C we construct synthetic
-datasets that have high or low levels of compressibility by varying
-the covariance structure across features (see \textit{Synthetic data}).  As shown
-in Figure~\ref{fig:information-compression}D, highly informative datasets
-yield higher decoding accuracies than less informative datasets (i.e., the peaks of the curves in
-the top panels of Fig.~\ref{fig:information-compression}D are higher than the
-peaks of the curves in the bottom panels). Highly compressible datasets show
-steeper slopes when we plot decoding accuracy as a function of the number of
-components used to represent the data (i.e., the slopes of the curves in the
-left panels of Fig.~\ref{fig:information-compression}D are steeper than the
-slopes of the curves in the right panels).  Whereas characterizing
-the informativeness and compressibility of synthetic data can be instructive,
-we are ultimately interested in understanding how these properties relate to
-brain activity patterns recorded under different cognitive circumstances.
+ongoing cognitive processes, then the peak decoding accuracy should be high.
+And if the signal is compressible, a low-dimensional embedding of the signal
+will be similarly informative as the original signal
+(Fig.~\ref{fig:information-compression}D).
 
@@
 less compressible brain activity patterns~\citep{MackEtal20, OwenEtal21}. These
-patterns hold even when the stimulus itself is held
-constant~\citep{MackEtal20}. These findings complement other work suggesting
-that functional connectivity (correlation) patterns are
-task-dependent~\citep{FinnEtal17, OwenEtal20, ColeEtal14}, although
-see~\cite{GratEtal18}. Higher-order cognitive processing of a common stimulus
-also appears to drive more stereotyped task-related activity and functional
-connectivity across individuals~\citep{HassEtal08, LernEtal11, SimoChan20,
-SimoEtal16}.
+findings complement other work suggesting that functional connectivity
+(correlation) patterns are task-dependent~\citep{FinnEtal17, OwenEtal20,
+ColeEtal14}, although see~\cite{GratEtal18}. Higher-order cognitive processing
+of a common stimulus also appears to drive more stereotyped task-related
+activity and functional connectivity across individuals~\citep{HassEtal08,
+LernEtal11, SimoChan20, SimoEtal16}.
 
@@
 complexity or with the level (or ``depth'') of cognitive processing. A second
-possibility is that the compressibility of brain activity patterns decreases
+possibility is that the compressibility of brain activity patterns increases
 during more complex or higher-level cognitive processing. If so, then
-individual features of brain recordings should carry more information (over and
-above the information carried by other features) during complex or high-level
-(versus simple or low-level) cognitive tasks. The tradeoffs between these two
-aspects of brain activity may also vary across brain regions or networks, for
-example according to each region's functional role.
+individual features of brain recordings, or compressed representations of brain
+recordings, should carry more information during complex or high-level (versus
+simple or low-level) cognitive tasks.
 
@@
 comprised functional magnetic resonance imaging (fMRI) data collected as
-participants listened to an audio recording of a 7-minute story, temporally
+participants listened to an audio recording of a 10-minute story, temporally
 scrambled recordings of the story, or underwent a resting state
@@
 experimental condition, participants listened to an audio recording of a
-7-minute Moth Radio Hour story, \textit{Pie Man}, by Jim O'Grady. In the
+10-minute Moth Radio Hour story, \textit{Pie Man}, by Jim O'Grady. In the
 \textit{paragraph}-scrambled experimental condition, participants listened to a
@@
 intervals. \textbf{C. Estimating ``compressibility'' for each condition.} The
-probability density plots display the numbers of components required to reach
-at least 5\% (corrected) decoding accuracy, across all cross validation runs.
-For the ``rest'' condition, where decoding accuracy never reached 5\% accuracy,
-we display the distribution of the numbers of components required to achieve
-the peak decoding accuracy in each fold. (This distribution is nearly flat, as
-also illustrated in Panel E.) The scale bar denotes a height of 0.01.
-\textbf{D. Estimating ``informativeness'' for each condition.} The probability
-density plots display the peak (corrected) decoding accuracies achieved in each
-condition, across all cross validation runs. The scale bar denotes a height of
-0.01. \textbf{E. Informativeness versus compressibility.} Each dot displays the
-minimum number of components required to achieve at least 5\% corrected
-decoding accuracy ($x$-coordinate; for the rest condition the numbers of
-components are estimated using the peak decoding accuracies as in Panel C) and
-the peak (corrected) decoding accuracy ($y$-coordinate). The smaller dots
-correspond to individual cross validation runs and the larger dots display the
-averages across runs. The inset displays a zoomed-in view of the indicated
-region in the main panel.}
+density plots display the numbers of components required to reach at least 5\%
+(corrected) decoding accuracy, across all cross validation runs. For the
+``rest'' condition, where decoding accuracy never reached 5\% accuracy, we
+display the distribution of the numbers of components required to achieve the
+peak decoding accuracy in each fold. (This distribution is nearly flat, as
+also illustrated in Panel E.) \textbf{D. Estimating ``informativeness'' for each
+condition.} The density plots display the peak (corrected) decoding accuracies
+achieved in each condition, across all cross validation runs. \textbf{E.
+Informativeness versus compressibility.} Each dot displays the minimum number
+of components required to achieve at least 5\% corrected decoding accuracy
+($x$-coordinate; for the rest condition the numbers of components are estimated
+using the peak decoding accuracies as in Panel C) and the peak (corrected)
+decoding accuracy ($y$-coordinate). The smaller dots correspond to individual
+cross validation runs and the larger dots display the averages across runs. The
+inset displays a zoomed-in view of the indicated region in the main panel.}
 
@@
 
-These results make some intuitive sense. As the contextual information
-available to participants increases (i.e., over time in the cognitively rich
-intact and paragraph conditions), it makes sense that this might
-constrain neural responses to a greater extent. While this pattern may not
-necessarily hold for \textit{every} possible story or stimulus, we suspect that
-it is generally the case that our knowledge about what is happening in a story
-tends to increase as we experience more about it. In turn, this could lead to
-greater consistency in different people's interpretations of and neural
-responses to the stimulus. Similarly, as participants are left to ``mind
-wander,'' or as they experience mental fatigue (i.e., over time in the less
-cognitively rich word and rest conditions), we suggest that this might
-lead to greater variability in neural responses across people, resulting in
-lower decoding accuracy. Again, it is not necessarily the case that every
-possible ``unengaging'' stimulus will lead to greater neural variability as
-time progresses, but we suspect this phenomenon is likely to hold for a variety
-of such stimuli. These findings replicate at least to a limited extent (e.g.,
-across both cognitively rich conditions and across both cognitively
-impoverished conditions, and for the different groups of participants in each
-of those conditions). However, determining whether these patterns generalize to
-other stimuli would require additional study (with new stimuli).
-
 \begin{figure}[tp]
@@
 
-If the informativeness and compressibility of brain activity patterns vary over
-time, do these properties might also vary across brain networks? We used a
-network parcellation identified by \cite{YeoEtal11} to segment the brain into
-seven distinct networks. The networks can be sorted (roughly) in order from
+We also wondered how informativeness and compressibility in the different
+experimental conditions might vary across brain networks. We used a network
+parcellation identified by \cite{YeoEtal11} to segment the brain into seven
+distinct networks. The networks can be sorted (roughly) in order from
 lower-level to higher-level cortex as follows (Figs.~\ref{fig:networks}A--C):
@@
 networks to higher-order networks, we found that decoding accuracy tended to
-increase in the higher-level experimental conditions and decrease (slightly) in
-the lower-level experimental conditions (Fig.~\ref{fig:networks}D, E;
-Spearman's rank correlation between decoding accuracy and network order:
-intact: $\rho = 0.362, p < 0.001$; paragraph: $\rho = 0.441, p < 0.001$; word:
-$\rho = -0.102, p = 0.007$; rest: $\rho = -0.354, p < 0.001$). This suggests
-that higher-order networks may carry more content-relevant or stimulus-driven
-``information.'' We found no clear trends in the proportions of components
-required to achieve 5\% decoding accuracy across networks or conditions
-(Fig.~\ref{fig:networks}F). We note that the limbic network we considered here
-often overlaps with low (imaging) signal regions, and therefore it may be
-difficult to draw strong conclusions about this network's informativeness or
-compressibility.
+increase, particularly in the higher-level experimental conditions
+(Fig.~\ref{fig:networks}D, E). This suggests that higher-order networks may
+carry more content-relevant or stimulus-driven ``information.'' We found no
+clear trends in the proportions of components required to achieve 5\% decoding
+accuracy across networks or conditions (Fig.~\ref{fig:networks}F).
 
-\begin{sidewaysfigure}[tp]
+\begin{figure}[tp]
   \centering
@@
 
-\end{sidewaysfigure}
+\end{figure}
 
-Whereas the above analyses examinine different networks in isolation, how does
-full-brain (i.e., potentially multi-network) activity patterns reflected by
-different principal components vary across different experimental conditions?
-As shown in Figure~\ref{fig:neurosynth-pca}, we used
+In addition to examining different networks in isolation, we wondered about the
+general structure of the full-brain (i.e., potentially multi-network) activity
+patterns reflected by different principal components across different
+experimental conditions. As shown in Figure~\ref{fig:neurosynth-pca}, we used
 Neurosynth~\citep{RubiEtal17} to identify, for each component, the associations
@@
 principal components across all of the experimental conditions tended to weight
-most heavily on themes related to cognitive control, memory, language
-processing, attention, and perception. Other components appeared to vary more
-across conditions.
+most heavily on themes related to cognitive control, memory,
+language processing, attention, and perception.  Other components appeared to
+vary more across conditions.
 
@@
 heavily by top components from the intact and paragraph conditions than
-lower-level functions (intact vs. word: $t(198) = 11.059, p < 0.001$; intact
-vs. rest: $t(198) = 3.699, p < 0.001$; paragraph vs. word: $t(198) = 13.504, p
-< 0.001$; paragraph vs. rest: $t(198) = 4.812, p < 0.001$; also see
-\textit{Ranking cognitive processes}). The top components from the word
-condition showed the opposite tendency, whereby \textit{lower}-level functions
-tended to be weighted on more heavily than higher-level functions (word vs.
-rest: $t(198) = -7.315, p < 0.001$). The weighting trends for the intact and
-paragraph conditions were not reliably different ($t(198) = -0.479, p =
-0.633$). The components from the rest condition showed almost no differences in
-the weights associated with high-level versus low-level functions (rest vs. 0:
-$t(99) = 1.836, p = 0.081$). These findings suggest that when participants were
-engaged more strongly (in the more engaging intact and paragraph conditions),
-their dominant neural patterns reflected higher-level cognitive functions. In
-contrast, when participants were engaged less strongly (in the less engaging
-word and rest conditions), their dominant neural patterns reflected lower-level
-cognitive functions. Although they were highly statistically reliable, it is
-also important to note that these latter effects are also relatively small
-(e.g., the slopes for \textit{all} of the experimental conditions are
-numerically close to zero; Fig.~\ref{fig:neurosynth-summary}E). We suggest that
-this phenomenon may merit further investigation in future work.
+lower-level functions. The top components from the word condition showed the
+opposite tendency, whereby \textit{lower}-level functions tended to be weighted
+on more heavily than higher-level functions. The components from the rest
+condition showed almost no differences in the weights associated with
+high-level versus low-level functions.
 
@@
 colored dots). \textbf{D. Associations between per-condition components, by
-component.} Each sub-panel's network plot summarizes the pattern of
-correlations between the $n$\textsuperscript{th} top-weighted principal
-components (sub-panel) for each experimental condition (gray-outlined dots).
-The line thicknesses are proportional to the correlation values (correlation
+component.} Each sub-panel's network plot summarizes the pattern of correlations
+between the $n$\textsuperscript{th} top-weighted principal components
+(sub-panel) for each experimental condition (gray-outlined dots). The line
+thicknesses are proportional to the correlation values (correlation
 coefficients are noted in the heat maps in Panel C). \textbf{E. Change in
@@
 between the component and topic maps associated with cognitive processes at the
-given rank (see \textit{Ranking cognitive processes}) Error bars denote
-bootstrap-estimated 95\% confidence intervals. Also see
+given rank (see \textit{Ranking cognitive processes}).  Also see
 Fig.~\neurosynthFull~for additional information.}
@@
 that unreliability could be compensated for by other regions' activity patterns
-that were predictive of the unreliable region. Whereas compressible brain
-patterns a robust to spatial signal corruption, high versus low informativeness
-reflects a similar (though dissociable; e.g.,
-Fig.~\ref{fig:information-compression}) tradeoff between expressiveness and
-robustness of \textit{temporal} patterns. Highly informative brain patterns (by
-our measure; i.e., patterns that yield greater temporal decoding accuracy) are
-expressive about ongoing experiences or cognitive states, since each moment's
-pattern is reliably distinguishable from other moments' patterns. However, when
-each moment's pattern is unique, brain activity becomes less robust to temporal
-signal corruption. Our finding that brain activity patterns becomes more
-informative (i.e., less robust to temporal signal corruption) and compressible
-(i.e., more robust to spatial signal corruption) when cognitive engagement is
-higher suggests that our brain may optimize its activity patterns to prioritize
-either temporal or spatial robustness, according to task demands.
+that were predictive of the unreliable region.
 
@@
 
-One potential limitation of our work concerns how our measure of
-informativeness might generalize across tasks or cognitive representations and
-processes. Our use of across-participant temporal decoding accuracy as a proxy
-for informativeness is motivated in part by prior work that introduced
-across-participant similarity (in time-varying response to a stimulus) as a
-means of identifying stimulus-driven brain activity
-patterns~\citep{SimoEtal16}. Intuitively, only activity patterns are are driven
-by the stimulus would be expected to synchronize (i.e., be time-locked to the
-stimulus) across participants. This approach implicitly removes idiosyncratic
-responses (e.g., neural patterns that are \textit{not} similar across people).
-However, there are also some published examples, including in our own prior
-work, that indicate that some types of stimulus-evoked activity will be missed
-by across-participant comparisons. For example we have reported how brain
-regions like the ventromedial prefrontal cortex (vmPFC) show stimulus-driven
-responses that are, for the most part, not similar across people. In that paper
-(and drawing on other work), we suggest that the vmPFC seems to represent or
-support highly idiosyncratic internal states, like affective responses.
-Although we would consider the vmPFC to be a ``high-level'' region (e.g., we
-consider affect to be a relatively high-level aspect of cognition), the measure
-of informativeness that we used in our current study would identify regions
-like the vmPFC as having \textit{low} informativeness. This is because
-across-participant decoding accuracy (our proxy measure for informativeness)
-will only be high for representations or responses that are common across
-people.
-
-Relatedly, even in the experimental conditions we describe as ``less
-cognitively engaging,'' we do not think it likely that high-level thought or
-cognitive processing is absent. Rather, we suggest that these high-level
-representations tend to be more idiosyncratic when the stimulus is less
-engaging, and therefore less constraining on people's thoughts. Nonetheless,
-even during highly engaging tasks, people may engage in idiosyncratic
-stimulus-driven processes. For example, people might retrieve personal
-information as they listened to the story. Those retrievals could happen at
-different times for different people according to each individual's prior
-experiences, and even when those sorts of retrievals happen to be temporally
-synchronized across people, the specific memories or information being
-retrieved might still be idiosyncratic. Our measure of informativeness is
-insensitive to these processes. Further, even in response to an identical
-stimulus, task instructions or participants' interal goals could change the
-relationship between compressibility and informativeness. Some work has shown
-that the ``dimensionality'' of neural representations can change systematically
-with task complexity, even in response to an identical
-stimulus~\citep{MackEtal20}. Taken together, we expect that the way we have
-defined informativeness in this paper, and the specific dataset we examined,
-are likely to have influenced our findings. While we see our approach as a
-reasonable first step, we also suggest that future work should explore
-alternative measures of informativeness and compressibility, and should examine
-how these measures vary across different tasks and datasets.
-
 In the information theory sense~\citep{Shan48}, when a signal is transmitted
@@
 
-We found that different brain networks varied in how informative and
-compressible their activity patterns were across experimental conditions (e.g.,
-Fig.~\ref{fig:networks}). This might
-follow from evolutionary optimizations that reflect the relevant constraints or
-demands placed on those networks. One possibility is that cortex is organized
-in a hierarchy of networks ``concerned with'' or selective to different levels
-of processing or function. To the extent that different levels of processing
-(e.g., low-level sensory processing versus ``deeper'' higher-level processing)
-reflect different stimulus timescales~\citep[e.g.,][]{Mann20}, the network
-differences we observed might also relate to the timescales at which each
-network is maximally sensitive~\citep{RegeEtal18, BaldEtal17,LernEtal11,
-HassEtal08}.
-
-Our reverse inference analyses (Figs.~\ref{fig:neurosynth-pca},
-\ref{fig:neurosynth-summary}) also provide some insights into how neural
-activity patterns change with cognitive engagement or task demands. Prior work
-has shown that the components and network ``parcels'' identified through
-covarying activity patterns can be highly similar even across different
-tasks~\citep[including ``rest,'' e.g.,][]{SmitEtal09, LairEtal11}. We
-replicated this basic finding in that the first principal components from all
-four experimental conditions were strikingly similar (e.g., see the leftmost
-columns of Figs.~\ref{fig:neurosynth-pca}A and~\ref{fig:neurosynth-summary}C,
-D). We also found some small, though statistically reliable, systematic changes
-in the weights associated with different cognitive functions across conditions
-(Fig.~\ref{fig:neurosynth-summary}E). This result provided an additional way of
-characterizing network-level differences across conditions
-(Fig.~\ref{fig:networks}E).  Taken together, these findings suggest that although
-similar networks may be involved in different tasks, the ways in which those
-networks are engaged may vary systematically with task demands.
+Different brain networks also varied in how informative and compressible their
+activity patterns were across experimental conditions (e.g.,
+Fig.~\ref{fig:networks}). This might follow from evolutionary optimizations
+that reflect the relevant constraints or demands placed on those networks. One
+possibility is that cortex is organized in a hierarchy of networks ``concerned
+with'' or selective to different levels of processing or function. To the
+extent that different levels of processing (e.g., low-level sensory processing
+versus ``deeper'' higher-level processing) reflect different stimulus
+timescales~\citep[e.g.,][]{Mann20}, the network differences we observed might
+also relate to the timescales at which each network is maximally
+sensitive~\citep{RegeEtal18, BaldEtal17,LernEtal11, HassEtal08}.
 
@@
 informativeness and compressibility, that appear to change systematically with
-task demands and across brain networks. We speculate that these changes may
-reflect ongoing tradeoffs between how robust to signal corruption versus how
-expressive about ongoing cognitive states our brains' activity patterns are.
-Our work also provides a new framework for evaluating these tradeoffs in other
-datasets, or in future studies.
+task demands and across brain networks. Our work helps to clarify how the
+``neural code'' might be structured, and how the code might vary across tasks
+and brain areas.
 
@@
 conditions (Fig.~\ref{fig:information-compression}).
-
-Unless otherwise noted, all statistical tests are two-sided, all error bars and
-error ribbons denote bootstrap-estimated 95\% confidence intervals across
-participants, and all reported $p$-values are corrected for multiple
-comparisons using the Benjamini-Hochberg procedure~\citep{BenjHoch95}.
 
@@
 native English speakers with normal hearing and provided written informed
-consent. We have excerpted the relevant portions of the dataset documention
-here to provide information about the scanning parameters and preprocessing
-steps used to generate the data we analyzed (the original descriptions may be
-found at the above link):
-
-\begin{quote}
-Subjects were scanned in a 3T full-body MRI scanner (Skyra; Siemens) with a
-sixteen-channel head coil. For functional scans, images were acquired using a
-T2* weighted echo planer imaging (EPI) pulse sequence [repetition time (TR),
-1500 ms; echo time (TE), 28 ms; flip angle, $64^\circ$], each volume comprising
-27 slices of 4 mm thickness with 0 mm gap; slice acquisition order was
-interleaved. In-plane resolution was $3 \times 3$ mm\textsuperscript{2} [field
-of view (FOV), 192 $\times$ 192 mm\textsuperscript{2}]. Anatomical images were
-acquired using a T1-weighted magnetization-prepared rapid-acquisition gradient
-echo (MPRAGE) pulse sequence (TR, 2300 ms; TE, 3.08 ms; flip angle 9$^\circ$;
-0.89 mm\textsuperscript{3} resolution; FOV, 256 mm\textsuperscript{2}). To
-minimize head movement, subjects' heads were stabilized with foam padding.
-Stimuli were presented using the Psychophysics toolbox~\citep{Brai97,Pell97}.
-Subjects were provided with an MRI compatible in-ear mono earbuds (Sensimetrics
-Model S14), which provided the same audio input to each ear. MRI-safe passive
-noise-canceling headphones were placed over the earbuds, for noise removal and
-safety.
-
-Functional data were preprocessed and analyzed using FSL
-(www.fmrib.ox.ac.uk/fsl), including correction for head motion and
-slice-acquisition time, spatial smoothing (6 mm FWHM Gaussian kernel), and
-high-pass temporal filtering (140 s period). Preprocessed data were aligned to
-coplanar and high-resolution anatomicals and the standard MNI152 brain, and
-interpolated to 3-mm isotropic voxels.
-\end{quote}
-
-The intact and word conditions each comprised 300 TRs (7.5 minutes) per participant.
-The paragraph condition comprised 272 TRs (6.8 minutes) per participant. The rest
-condition comprised 400 TRs (10 minutes) per participant.
+consent.
 
@@
 downloaded at \href{https://brainiak.org/}{brainiak.org}.
-
-We also considered alternative approaches to obtaining compact representations
-of the neuroimaging data, including network
-parcellations~\citep[e.g.,][]{SchaEtal18, GordEtal16}. Whereas network
-parcellations are typically derived from large resting state datasets, HTFA may
-be applied to much smaller datasets. In our prior work, we showed that HTFA
-applied to the same dataset used here can explain full-brain activity to within
-a maximum of 0.25 standard deviations of each voxel's observed activity in the
-original dataset, taken across all voxels, images, and participants, using the
-700-node representation we also employed here~\citep{MannEtal18}. Some of the
-explanatory power of HTFA comes from the fact that each node's explanatory
-power falls off smoothly with distance to its center. Intuitively, the result
-is a representation that looks like a lightly spatially smoothed version of the
-original data, but where the degree of smoothing varies across the brain
-according to how spatially autocorrelated the local activity patterns are.
 
@@
 
-We manually identified 11 cognitive labels spanning the set of 80
-Neurosynth-derived topics: cognitive control, language processing, memory,
-emotion, social cognition, spatial cognition, attention, reward, sensory
-perception, motor control, and resting state. We then used ChatGPT to
-automatically ``rank'' the processes from high-level to low-level using the
-following prompt: ``Please rank these cognitive processes from highest-level to
-lowest-level, where higher values indicate higher-order or higher-level
-processes. Return the result as a csv file with a header row and two columns:
-`Cognitive label' and `Rank'. Here are the processes: cognitive control,
-language processing, memory, emotion, social cognition, spatial cognition,
-attention, reward, sensory perception, motor control, resting state''.
-Table~\topicTags~displays the output.
-
-We recognize that ChatGPT is not omniscient, nor should it be treated as an
-expert cognitive neuroscientist. We therefore reviewed ChatGPT's responses
-carefully by hand to verify that they seemed reasonable to us. Whereas prior
-work has often constructed such rankings by hand, we see our use of ChatGPT in
-this case as a small additional ``sanity check'' on our rankings that helped us
-to be slightly more objective than if we had simply created the rankings
-ourselves manually.
-
-In the analysis presented in Figure~\ref{fig:neurosynth-summary}E, we summarize
-difference in topic weightings across experimental conditions. In particular,
-we sought to characterize how the dominant neural patterns evoked by each
-experimental condition weighted on different cognitive functions. For each of
-the top five principal components from each experimental condition
-(Fig.~\ref{fig:neurosynth-pca}), we computed the average weightings for each of
-the 11 manually identified (and ChatGPT-ranked) cognitive labels described
-above (Tab.~\topicTags). We then fit a line separately for each experiment
-condition ($x$-values: cognitive rank; $y$-values: weights). In carrying out
-this analysis, we used a bootstrap procedure to estimate the variability in the
-slopes of the regression lines, whereby we repeated this process for each of $n
-= 100$ iterations, each time resampling (with replacement) the set of observed
-ranks and weights. This procedure yielded distributions of 100 estimated slopes
-for each experimental condition. We used these distributions to compare the
-slopes across experimental conditions and to estimate 95\% confidence intervals.
-
-\subsection*{Synthetic data}
-
-To help illustrate the relationship between informativeness and compressibility
-(Fig.~\ref{fig:information-compression}), we generated four synthetic datasets,
-varying in informativeness and compressibility. Each dataset comprised
-simulated observations of $K = 25$ features across $N = 100$ timepoints, from each of $S = 10$
-participants. To create each dataset, we first constructed a ``template''
-matrix of $N$ timepoints by $K$ features. We then generated participant-specific
-data by adding independent noise to each entry in template matrix, drawn from
-the unit normal distribution (i.e., with a mean of 0 and a variance of 1). We
-repeated this process for each participant, yielding $S$ participant-specific
-matrices for each dataset.
-
-Since we estimate informativeness using the temporal decoding accuracy across
-participants, highly informative data will tend to have observations that are
-highly timepoint specific. Relatively uninformative data, in contrast, will
-tend to have more similar observations across timepoints. To generate data with
-``high informativeness,'' we constructed template matrices whose rows
-(observations) were drawn independently from zero-mean multivariate normal
-distributions. The covariances of these distributions were determined according
-to the desired compressibility of the data, as described below. We used a
-multi-step process to generate data with ``low informativeness.'' First we
-generated new template matrices using the same procedure as for the ``high
-informativeness'' datasets. We then multiplied each matrix by a constant ($\rho
-= 0.1$) and computed the cumulative sum of each matrix's rows. This yielded
-matrices whose rows were highly similar across observations.
-
-Compressibility reflects the extent to which decoding accuracy is affected by
-reducing the number of components used to represent the data. Highly
-compressible data will tend to exhibit more similarities across features,
-whereas less compressible data will tend to show greater independence across
-features. To generate data with ``high compressibility,'' we set the covariance
-matrix of the multivariate normal distribution to a toeplitz matrix whose first
-row was given by $\left[K, K - 1, ..., 1\right]$. To generate data with ``low
-compressibility,'' we set the covariance matrix to the identity matrix.
-
-Template matrices for datasets with high informativeness and high
-compressibility, high informativeness and low compressibility, low
-informativeness and high compressibility, and low informativeness and low
-compressibility are displayed in Figure~\ref{fig:information-compression}C. The
-corresponding decoding curves are displayed in
-Figure~\ref{fig:information-compression}D.
+We manually identified 11 cognitive labels
+spanning the set of 80 Neurosynth-derived topics: cognitive control, language
+processing, memory, emotion, social cognition, spatial cognition, attention,
+reward, sensory perception, motor control, and resting state. We then used
+ChatGPT to automatically ``rank'' the processes from high-level to low-level
+using the following prompt: ``Please rank these cognitive processes from
+highest-level to lowest-level, where higher values indicate higher-order or
+higher-level processes. Return the result as a csv file with a header row and
+two columns: `Cognitive label' and `Rank'. Here are the processes: cognitive
+control, language processing, memory, emotion, social cognition, spatial
+cognition, attention, reward, sensory perception, motor control, resting
+state''.  Table~\topicTags~displays the output.  We used these labels in
+the analysis presented in Figure~\ref{fig:neurosynth-summary}E to help summarize
+difference in topic weightings across experimental conditions.
 

### 2023-12-26 paper/main.tex

@@
 
-\textbf{Significance Statement:} How our brains respond to ongoing experiences
-depends on what we are doing and thinking about, how engaged we are, and our
-past experiences, among other factors. To gain insights into how brain activity
-patterns reflect ongoing cognitive processes, we examined two fundamental
-aspects of brain activity under different cognitive circumstances:
-informativeness and compressibility. Informativeness refers to the extent to
-which brain patterns are both temporally specific and consistent across
-different people. Compressibility refers to the extent to how robust the
-informativeness of brain patterns is to dimensionality reduction. We found that
-brain activity patterns evoked by higher-level cognitive tasks are both more
-informative \textit{and} more compressible than activity evoked by lower-level
-tasks. Our findings suggest that our brains flexibly reconfigure themselves to
-optimize different aspects of how they function according to ongoing cognitive
+\textbf{Significance Statement:} How our brains respond to ongoing
+experiences depends on what we are doing and thinking about, among
+other factors. To study how brain activity reflects ongoing cognition,
+we examined two fundamental aspects of brain activity under different
+cognitive circumstances: informativeness and
+compressibility. Informativeness refers to the extent to which brain
+patterns are both temporally specific and consistent across different
+people. Compressibility refers to how robust the informativeness of
+brain patterns is to dimensionality reduction. Brain activity evoked
+by higher-level cognitive tasks are both more informative \textit{and} more
+compressible than activity evoked by lower-level tasks. Our findings
+suggest that our brains flexibly reconfigure themselves to optimize
+different aspects of how they function according to ongoing cognitive
 demands.
@@
 from the network without any loss of function or expressive power, as long as
-any single node remains. Presumably, most natural systems tend to occupy
-positions between these extremes. We wondered: might the human brain
-reconfigure itself to be more flexible or more robust according to ongoing
-demands? In other words, might the brain reconfigure its connections or
-behaviors under different circumstances to change its position along this
-continuum?
+any single node remains. In addition to considering flexibility across space
+(nodes), these properties may also vary, largely independently, across time. A
+network is maximally expressive when its nodes' activity patterns vary in
+meaningful ways from moment to moment, whereas it is maximally robust to signal
+corruption when its activity is constant over time. Presumably, most natural
+systems tend to occupy positions between these temporal and spatial extremes. Under
+different circumstances, it may even prove benefitial for systems to make
+different tradeoffs between expressiveness and robustness along the temporal
+and spatial dimensions.  We wondered: might the human brain reconfigure itself to
+be more flexible or more robust according to ongoing demands? In other words,
+might the brain reconfigure its connections or behaviors under different
+circumstances to change its position along these continuums?
 
@@
   displayed. This analysis reveals that the drawing is more compressible: just
-  16 components can explain 95\% of the variance in teh drawing, whereas 50
+  16 components can explain 95\% of the variance in the drawing, whereas 50
   components are required to explain 95\% of the variance in the photograph.
@@
 accuracy was highest for the ``intact'' condition (versus paragraph: $t(99) =
-35.205, p < 0.001)$; versus word: $t(99) = 43.172, p < 0.001)$; versus rest:
-$t(99) = 81.361, p < 0.001)$), next highest for the ``paragraph'' condition
-(versus word: $t(99) = 6.243, p < 0.001)$; versus rest: $t(99) = 50.748, p <
-0.001)$), and next highest for the ``word'' condition (versus rest: $t(99) =
-48.791, p < 0.001)$). This ordering implies that cognitively richer conditions
+35.205, p < 0.001$; versus word: $t(99) = 43.172, p < 0.001$; versus rest:
+$t(99) = 81.361, p < 0.001$), next highest for the ``paragraph'' condition
+(versus word: $t(99) = 6.243, p < 0.001$; versus rest: $t(99) = 50.748, p <
+0.001$), and next highest for the ``word'' condition (versus rest: $t(99) =
+48.791, p < 0.001$). This ordering implies that cognitively richer conditions
 evoke more stable brain activity patterns across people.
@@
 it is generally the case that our knowledge about what is happening in a story
-tends to increase as we experience more about it. In turn, this could lead to
+tends to increase as we experience more of it. In turn, this could lead to
 greater consistency in different people's interpretations of and neural
@@
 If the informativeness and compressibility of brain activity patterns vary over
-time, do these properties might also vary across brain networks? We used a
+time, might these properties also vary across brain networks? We used a
 network parcellation identified by \cite{YeoEtal11} to segment the brain into
@@
 
-Whereas the above analyses examinine different networks in isolation, how does
+Whereas the above analyses examined different networks in isolation, how does
 full-brain (i.e., potentially multi-network) activity patterns reflected by
@@
 with each of 80 themes (see \textit{Reverse inference}). In general, the first
-principal components across all of the experimental conditions tended to weight
+principal components across all of the experimental conditions tended to weigh
 most heavily on themes related to cognitive control, memory, language
@@
 condition's components (Fig.~\neurosynthFull C, D). Taken together, we found
-that each component appeared to weight on a fundamental set of cognitive
+that each component appeared to weigh on a fundamental set of cognitive
 functions that varied by experimental condition. For example, the top principal
-components from every condition weighted similarly (across conditions) on the
+components from every condition weighed similarly (across conditions) on the
 full set of Neurosynth topics (Fig.~\ref{fig:neurosynth-pca}A) and cognitive
@@
 Fig.~\ref{fig:neurosynth-summary}E; also see \textit{Ranking cognitive
-processes}). We found that higher-level functions tended to be weighted on more
+processes}). We found that higher-level functions tended to be weighted more
 heavily by top components from the intact and paragraph conditions than
@@
 condition showed the opposite tendency, whereby \textit{lower}-level functions
-tended to be weighted on more heavily than higher-level functions (word vs.
-rest: $t(198) = -7.315, p < 0.001$). The weighting trends for the intact and
+tended to be weighted more heavily than higher-level functions (word vs. rest:
+$t(198) = -7.315, p < 0.001$). The weighting trends for the intact and
 paragraph conditions were not reliably different ($t(198) = -0.479, p =
-0.633$). The components from the rest condition showed almost no differences in
-the weights associated with high-level versus low-level functions (rest vs. 0:
-$t(99) = 1.836, p = 0.081$). These findings suggest that when participants were
-engaged more strongly (in the more engaging intact and paragraph conditions),
-their dominant neural patterns reflected higher-level cognitive functions. In
-contrast, when participants were engaged less strongly (in the less engaging
-word and rest conditions), their dominant neural patterns reflected lower-level
-cognitive functions. Although they were highly statistically reliable, it is
-also important to note that these latter effects are also relatively small
-(e.g., the slopes for \textit{all} of the experimental conditions are
-numerically close to zero; Fig.~\ref{fig:neurosynth-summary}E). We suggest that
-this phenomenon may merit further investigation in future work.
+0.633$). The components from the rest condition showed only a small trending
+difference in the weights associated with high-level versus low-level functions
+(rest vs. 0: $t(99) = 1.836, p = 0.081$). These findings suggest that when
+participants were engaged more strongly (in the more engaging intact and
+paragraph conditions), their dominant neural patterns reflected higher-level
+cognitive functions. In contrast, when participants were engaged less strongly
+(in the less engaging word and rest conditions), their dominant neural patterns
+reflected lower-level cognitive functions. Although they were highly
+statistically reliable, it is also important to note that these latter effects
+are also relatively small (e.g., the slopes for \textit{all} of the
+experimental conditions are numerically close to zero;
+Fig.~\ref{fig:neurosynth-summary}E). We suggest that this phenomenon may merit
+further investigation in future work.
 
@@
 (columns within each sub-panel, indicated by colored dots). Each sub-panel
-reflects correlations for the components indicated in the panel titles. A
-legend for the condition and several core cognitive function classifications is
-displayed in the lower right of the figure. Table~\topics~provides a list of
-each topic's top-weighted terms, along with each topic's manually labeled
-cognitive classification. A full list of the topics most highly associated with
-each component may be found in Figure~\topTerms. \textbf{B. Associations
-between per-condition components and cognitive functions.} The network plots
-denote positive average correlations between the component images for each
-condition (gray-outlined dots on the left sides of each network; colors denote
-conditions) and topic-specific brain maps associated with each indicated
-cognitive function (black-outlined dots on the right sides of each network;
-colors denote cognitive functions). The line thicknesses are proportional to
-the correlation values (correlation coefficients are noted in the heat maps in
-Panel A). \textbf{C. Correlations between each principal component, by
-condition.} The heat maps display the correlations between the brain maps
-(Fig.~\componentBrains) for each principal component (sub-panel), across each
-pair of conditions (rows and columns of each sub-panel's matrix, indicated by
-colored dots). \textbf{D. Associations between per-condition components, by
-component.} Each sub-panel's network plot summarizes the pattern of
-correlations between the $n$\textsuperscript{th} top-weighted principal
-components (sub-panel) for each experimental condition (gray-outlined dots).
-The line thicknesses are proportional to the correlation values (correlation
-coefficients are noted in the heat maps in Panel C). \textbf{E. Change in
-weights by condition.} The bar heights reflect the slopes of regression lines,
-fit separately to the top five components from each condition, between the
-ChatGPT-derived ``rank'' of each cognitive classification and the correlations
-between the component and topic maps associated with cognitive processes at the
-given rank (see \textit{Ranking cognitive processes}) Error bars denote
-bootstrap-estimated 95\% confidence intervals. Also see
-Fig.~\neurosynthFull~for additional information.}
+reflects correlations for the components indicated in the panel titles.
+Table~\topics~provides a list of each topic's top-weighted terms, along with
+each topic's manually labeled cognitive classification. A full list of the
+topics most highly associated with each component may be found in
+Figure~\topTerms. \textbf{B. Associations between per-condition components and
+cognitive functions.} The network plots denote positive average correlations
+between the component images for each condition (gray-outlined dots on the left
+sides of each network; colors denote conditions) and topic-specific brain maps
+associated with each indicated cognitive function (black-outlined dots on the
+right sides of each network; colors denote cognitive functions). The line
+thicknesses are proportional to the correlation values (correlation
+coefficients are noted in the heat maps in Panel A). \textbf{C. Correlations
+between each principal component, by condition.} The heat maps display the
+correlations between the brain maps (Fig.~\componentBrains) for each principal
+component (sub-panel), across each pair of conditions (rows and columns of each
+sub-panel's matrix, indicated by colored dots). \textbf{D. Associations between
+per-condition components, by component.} Each sub-panel's network plot
+summarizes the pattern of correlations between the $n$\textsuperscript{th}
+top-weighted principal components (sub-panel) for each experimental condition
+(gray-outlined dots). The line thicknesses are proportional to the correlation
+values (correlation coefficients are noted in the heat maps in Panel C).
+\textbf{E. Change in weights by condition.} The bar heights reflect the slopes
+of regression lines, fit separately to the top five components from each
+condition, between the ChatGPT-derived ``rank'' of each cognitive
+classification and the correlations between the component and topic maps
+associated with cognitive processes at the given rank (see \textit{Ranking
+cognitive processes}) Error bars denote bootstrap-estimated 95\% confidence
+intervals. Also see Fig.~\neurosynthFull~for additional information.}
 
@@
 corresponding maps. As summarized in Figure~\ref{fig:neurosynth-summary}E, we
-found that the intact and paragraph conditions tended to weight on higher-level
+found that the intact and paragraph conditions tended to weight higher-level
 cognitive processes more than lower-level cognitive processes, whereas the word
-condition weighted on lower-level processes more than higher-level processes
-and the rest condition showed no difference in high-level versus low-level
-weighting. Taken together, our findings indicate that the informativeness and
+condition weighted lower-level processes more than higher-level processes
+and the rest condition showed no reliable difference in high-level versus low-level
+weightings. Taken together, our findings indicate that the informativeness and
 compressibility of our brain activity patterns are task-dependent, and these
@@
 that were predictive of the unreliable region. Whereas compressible brain
-patterns a robust to spatial signal corruption, high versus low informativeness
-reflects a similar (though dissociable; e.g.,
+patterns are robust to spatial signal corruption, high versus low
+informativeness reflects a similar (though dissociable; e.g.,
 Fig.~\ref{fig:information-compression}) tradeoff between expressiveness and
@@
 One potential limitation of our work concerns how our measure of
-informativeness might generalize across tasks or cognitive representations and
-processes. Our use of across-participant temporal decoding accuracy as a proxy
-for informativeness is motivated in part by prior work that introduced
-across-participant similarity (in time-varying response to a stimulus) as a
-means of identifying stimulus-driven brain activity
-patterns~\citep{SimoEtal16}. Intuitively, only activity patterns are are driven
-by the stimulus would be expected to synchronize (i.e., be time-locked to the
-stimulus) across participants. This approach implicitly removes idiosyncratic
-responses (e.g., neural patterns that are \textit{not} similar across people).
-However, there are also some published examples, including in our own prior
-work, that indicate that some types of stimulus-evoked activity will be missed
-by across-participant comparisons. For example we have reported how brain
-regions like the ventromedial prefrontal cortex (vmPFC) show stimulus-driven
-responses that are, for the most part, not similar across people. In that paper
-(and drawing on other work), we suggest that the vmPFC seems to represent or
-support highly idiosyncratic internal states, like affective responses.
-Although we would consider the vmPFC to be a ``high-level'' region (e.g., we
-consider affect to be a relatively high-level aspect of cognition), the measure
-of informativeness that we used in our current study would identify regions
-like the vmPFC as having \textit{low} informativeness. This is because
-across-participant decoding accuracy (our proxy measure for informativeness)
-will only be high for representations or responses that are common across
-people.
+informativeness might generalize across different tasks, cognitive
+representations, and processes. Our use of across-participant temporal decoding
+accuracy as a proxy for informativeness is motivated in part by prior work that
+introduced across-participant similarity (in time-varying response to a
+stimulus) as a means of identifying stimulus-driven brain activity
+patterns~\citep{SimoEtal16}. Intuitively, only activity patterns that are
+driven by the stimulus would be expected to synchronize (i.e., be time-locked
+to the stimulus) across participants. This approach implicitly removes
+idiosyncratic responses (e.g., neural patterns that are \textit{not} similar
+across people). However, there are also some published examples, including in
+our own prior work, that indicate that some types of stimulus-evoked activity
+will be missed by across-participant comparisons. For example we have reported
+how brain regions like the ventromedial prefrontal cortex (vmPFC) show
+stimulus-driven responses that are, for the most part, \textit{not} similar
+across people. In that paper (and drawing on other work), we suggest that the
+vmPFC seems to represent or support highly idiosyncratic internal states, like
+affective responses. Although we would consider the vmPFC to be a
+``high-level'' region (e.g., we consider affect to be a relatively high-level
+aspect of cognition), the measure of informativeness that we used in our
+current study would identify regions like the vmPFC as having \textit{low}
+informativeness. This is because across-participant decoding accuracy (our
+proxy measure for informativeness) will only be high for representations or
+responses that are common across people.
 
 Relatedly, even in the experimental conditions we describe as ``less
-cognitively engaging,'' we do not think it likely that high-level thought or
-cognitive processing is absent. Rather, we suggest that these high-level
-representations tend to be more idiosyncratic when the stimulus is less
+cognitively engaging,'' we think it likely that high-level thought or cognitive
+processing is still present. Rather, we suggest that these high-level
+representations will tend to be more idiosyncratic when the stimulus is less
 engaging, and therefore less constraining on people's thoughts. Nonetheless,
@@
 different times for different people according to each individual's prior
-experiences, and even when those sorts of retrievals happen to be temporally
+experiences. Even when those sorts of retrievals happen to be temporally
 synchronized across people, the specific memories or information being
@@
 insensitive to these processes. Further, even in response to an identical
-stimulus, task instructions or participants' interal goals could change the
+stimulus, task instructions or participants' internal goals could change the
 relationship between compressibility and informativeness. Some work has shown
@@
 (e.g., low-level sensory processing versus ``deeper'' higher-level processing)
-reflect different stimulus timescales~\citep[e.g.,][]{Mann20}, the network
+reflect different stimulus timescales~\citep[e.g.,][]{Mann23b}, the network
 differences we observed might also relate to the timescales at which each
@@
 native English speakers with normal hearing and provided written informed
-consent. We have excerpted the relevant portions of the dataset documention
+consent. We have excerpted the relevant portions of the dataset documentation
 here to provide information about the scanning parameters and preprocessing
@@
 700-node representation we also employed here~\citep{MannEtal18}. Some of the
-explanatory power of HTFA comes from the fact that each node's explanatory
-power falls off smoothly with distance to its center. Intuitively, the result
-is a representation that looks like a lightly spatially smoothed version of the
+explanatory power of HTFA comes from the fact that each node's influence falls
+off smoothly with distance to its center. Intuitively, the result is a
+representation that looks like a lightly spatially smoothed version of the
 original data, but where the degree of smoothing varies across the brain
@@
 
-In our analyses of how informativeness varied across brain networks, we
-considered the possibility that the correlations with network order might be
-influenced by the numbers of nodes in each network. We designed a
-permutation-based procedure to address this possibility, whereby we repeated
-the above analyses using shuffled network labels. Specifically, for each of $n_1 =
-10$ iterations, we randomly shuffled (without replacement) the network labels of
-the HTFA nodes, and then we re-ran our entire decoding analysis pipeline,
-including applying PCA with $3...m$ features for each condition (where $m$ is the
-number of nodes in the given network), and then running 100 cross-validation
-runs of the decoding procedure for each condition and number of components.
-This resulted in 10 sets of shuffled data where each network had the same
-numbers of nodes, but where the decoding results no longer maintained the
-fidelity of each individual network.
+In our analyses of how informativeness varied across brain networks
+(Fig.~\ref{fig:networks}), we considered the possibility that the correlations
+with network order might be influenced by the numbers of nodes in each network.
+We designed a permutation-based procedure to address this possibility, whereby
+we repeated the above analyses using shuffled network labels. Specifically, for
+each of $n_1 = 10$ iterations, we randomly shuffled (without replacement) the
+network labels of the HTFA nodes, and then we re-ran our entire decoding
+analysis pipeline, including applying PCA with $3...m$ features for each
+condition (where $m$ is the number of nodes in the given network), and then
+running 100 cross-validation runs of the decoding procedure for each condition
+and number of components. This resulted in 10 sets of shuffled data, where each
+network had the same numbers of nodes, but where the decoding results no longer
+maintained the fidelity of each individual network.
 
@@
 datasets. We then compared the distributions of Spearman's $\rho$ values for
-the original and shuffled datasets using a two-sided independent samples
-Welch's $t$-tests.
+the original and shuffled datasets using two-sided independent samples Welch's
+$t$-tests.
 
@@
 the top five principal components from each experimental condition
-(Fig.~\ref{fig:neurosynth-pca}), we computed the average weightings for each of
+(Fig.~\ref{fig:neurosynth-pca}), we computed the average weights for each of
 the 11 manually identified (and ChatGPT-ranked) cognitive labels described
@@
 for each experimental condition. We used these distributions to compare the
-slopes across experimental conditions and to estimate 95\% confidence intervals.
+slopes across experimental conditions and to estimate 95\% confidence
+intervals.
 

### 2024-02-13 paper/main.tex
adding autocorrelation analysis (will remove later)

@@
 stimulus-driven responses that are, for the most part, \textit{not} similar
-across people. In that paper (and drawing on other work), we suggest that the
+across people~\citep{ChanEtal21a}. In that paper (and drawing on other work), we suggest that the
 vmPFC seems to represent or support highly idiosyncratic internal states, like

### 2024-03-11 paper/supplement.tex
title change
submitted version (PNAS R2)
adding autocorrelation analysis (will remove later)
added autocorrelation analysis

@@
 
-\title{\textit{Supplemental materials for:} High-order cognition is supported by information-rich but compressible brain activity patterns}
+\title{\textit{Supplemental materials for:} High-level cognition is supported by information-rich but compressible brain activity patterns}
 
@@
   \centering
+  \includegraphics[width=\textwidth]{figs/autocorrelations}
+
+\caption{\textbf{Autocorrelations.} \textbf{A. Temporal autocorrelations.} The
+panel displays the average correlation (taken across all nodes and
+participants) for each experimental condition (color), as a function of offset
+lag (in TRs). \textbf{B. Spatial autocorrelations.} The panel displays the
+average correlation between the activation timeseries for each pair of nodes,
+for each experimental condition (color), as a function of distance (in mm).  Error ribbons in
+both panels denote bootstrap-estimated across-participant 95\% confidence intervals.}
+\label{fig:autocorrs}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
   \includegraphics[width=\textwidth]{figs/variance_explained}
@@
 
+\clearpage
 \newpage
 \renewcommand{\refname}{Supplemental references}
+
+\newcounter{bibcount}
+\makeatletter
+\patchcmd{\@lbibitem}{\item[}{\item[\hfil\stepcounter{bibcount}{\thebibcount.}}{}{}
+\setlength{\bibhang}{2\parindent}
+\renewcommand\NAT@bibsetup%
+   [1]{\setlength{\leftmargin}{\bibhang}\setlength{\itemindent}{-\parindent}%
+       \setlength{\itemsep}{\bibsep}\setlength{\parsep}{\z@}}
+\makeatother
+
 \bibliographystyle{apacite}

### 2024-03-12 paper/main.tex
added autocorrelation analysis
add spatial_autocorrelations.pkl to dropbox:   - load locally if possible   - else download from dropbox   - if failure, compute and store locally
the pkl file is 1.2 GB, so it's too large to store directly on GitHub

--- 2023-12-26 paper/main.tex
@@
 
-\newcommand{\varExplained}{S1}
-\newcommand{\topTerms}{S2}
-\newcommand{\componentBrains}{S3}
-\newcommand{\topicContrasts}{S4}
-\newcommand{\neurosynthFull}{S5}
+\newcommand{\autocorr}{S1}
+\newcommand{\varExplained}{S2}
+\newcommand{\topTerms}{S3}
+\newcommand{\componentBrains}{S4}
+\newcommand{\topicContrasts}{S5}
+\newcommand{\neurosynthFull}{S6}
 
@@
 AdacEtal12, BassSpor17, BullSpor09, SporHone06, SporBetz16, SporZwi04,
-DhamEtal08, KorzEtal08, BrovEtal04, LynnBass21}. Correlations or causal
+DhamEtal08, KorzEtal08, BrovEtal04, LynnBass21, ShinnEtal23}. Correlations or causal
 associations between different brain regions simultaneously imply that
@@
 informativeness might generalize across different tasks, cognitive
-representations, and processes. Our use of across-participant temporal decoding
-accuracy as a proxy for informativeness is motivated in part by prior work that
-introduced across-participant similarity (in time-varying response to a
-stimulus) as a means of identifying stimulus-driven brain activity
-patterns~\citep{SimoEtal16}. Intuitively, only activity patterns that are
-driven by the stimulus would be expected to synchronize (i.e., be time-locked
-to the stimulus) across participants. This approach implicitly removes
-idiosyncratic responses (e.g., neural patterns that are \textit{not} similar
-across people). However, there are also some published examples, including in
-our own prior work, that indicate that some types of stimulus-evoked activity
-will be missed by across-participant comparisons. For example we have reported
-how brain regions like the ventromedial prefrontal cortex (vmPFC) show
-stimulus-driven responses that are, for the most part, \textit{not} similar
-across people. In that paper (and drawing on other work), we suggest that the
-vmPFC seems to represent or support highly idiosyncratic internal states, like
-affective responses. Although we would consider the vmPFC to be a
-``high-level'' region (e.g., we consider affect to be a relatively high-level
-aspect of cognition), the measure of informativeness that we used in our
-current study would identify regions like the vmPFC as having \textit{low}
-informativeness. This is because across-participant decoding accuracy (our
-proxy measure for informativeness) will only be high for representations or
-responses that are common across people.
+representations, and processes. As in many neuroimaing studies, our findings
+may also be influenced by our preprocessing decisions, which could in turn
+affect temporal and spatial correlations in the data. Our use of
+across-participant temporal decoding accuracy as a proxy for informativeness is
+motivated in part by prior work that introduced across-participant similarity
+(in time-varying response to a stimulus) as a means of identifying
+stimulus-driven brain activity patterns~\citep{SimoEtal16}. Intuitively, only
+activity patterns that are driven by the stimulus would be expected to
+synchronize (i.e., be time-locked to the stimulus) across participants. This
+approach implicitly removes idiosyncratic responses (e.g., neural patterns that
+are \textit{not} similar across people). However, there are also some published
+examples, including in our own prior work, that indicate that some types of
+stimulus-evoked activity will be missed by across-participant comparisons. For
+example we have reported how brain regions like the ventromedial prefrontal
+cortex (vmPFC) show stimulus-driven responses that are, for the most part,
+\textit{not} similar across people~\citep{ChanEtal21a}. In that paper (and
+drawing on other work), we suggest that the vmPFC seems to represent or support
+highly idiosyncratic internal states, like affective responses. Although we
+would consider the vmPFC to be a ``high-level'' region (e.g., we consider
+affect to be a relatively high-level aspect of cognition), the measure of
+informativeness that we used in our current study would identify regions like
+the vmPFC as having \textit{low} informativeness. This is because
+across-participant decoding accuracy (our proxy measure for informativeness)
+will only be high for representations or responses that are common across
+people.
 
